---
title: "Untitled"
format: html
---


```{r}
# BiocManager::install("mygene")
# library(mygene)
library(dplyr)
library(parseMetab)
```

# Run differential analysis on genes within each metabolic system
```{r}
OUTDIR = "../results/CellFie_Hu2025/"
taskInfo2 = readRDS(paste0(OUTDIR, "/CellFieOut/taskInfo.RDS"))
systems = taskInfo2$Depth1 %>% unique()
rslt_lists = lapply(systems, function(s) {
    print(s)
    name = gsub(" .*", "", s)
    outDirHu = paste0("../results/FeatureDifferential_CellFie/", name, "Differential/Hu/")
    dir.create(outDirHu, recursive=TRUE, showWarnings=FALSE)
    outDirZhou = paste0("../results/FeatureDifferential_CellFie/", name, "Differential/Zhou/")
    dir.create(outDirZhou, recursive=TRUE, showWarnings=FALSE)

    rslt_list = systemDiffAnalysis(System=s, outDirHu, outDirZhou)

    visualizeDiffFeatures(limma_rslt=rslt_list$limma_rslt_listZhou, 
        outdir=outDirZhou, 
        taskInfo=rslt_list$systemZhou,
        max_logFC=4)

    visualizeDiffFeatures(limma_rslt=rslt_list$limma_rslt_listHu, 
        outdir=outDirHu, 
        taskInfo=rslt_list$systemHu,max_logFC=4)

    visualizeDiffFeatures(limma_rslt=rslt_list$limma_rslt_listZhou, 
        outdir=outDirZhou, 
        taskInfo=rslt_list$systemZhou,
        max_logFC=4, depth="Depth2")

    visualizeDiffFeatures(limma_rslt=rslt_list$limma_rslt_listHu, 
        outdir=outDirHu, 
        taskInfo=rslt_list$systemHu, 
        max_logFC=4, depth="Depth2")

    return(rslt_list)
}) %>% setNames(systems)


# names(rslt_lists) = systems
saveRDS(rslt_lists, "../results/FeatureDifferential_CellFie/rslt_list.RDS")
rslt_lists = readRDS("../results/FeatureDifferential_CellFie/rslt_list.RDS")
```

# Compare effect size among systems
```{r}
outDirSummary = "../results/FeatureDifferential_CellFie/summary/"
outDirSummary %>% dir.create(., recursive=TRUE, showWarnings=FALSE)

makeEffectSizeBoxplot_system(
    rslt_lists, dataset="Hu", 
    winsorize.quantil=0.05)
makeEffectSizeBoxplot_system(
    rslt_lists, dataset="Zhou", 
    winsorize.quantil=0.05)

makeEffectSizeBoxplot_system(
    rslt_lists, 
    dataset="Hu", effect.statistic="t",
    winsorize.quantil=0.02)
makeEffectSizeBoxplot_system(
    rslt_lists, dataset="Zhou", 
    effect.statistic="t",
    winsorize.quantil=0.02) 

```

# Compare effect size among cancer types
```{r}
outDirSummary = "../results/FeatureDifferential_CellFie/summary/"
outDirSummary %>% dir.create(., recursive=TRUE, showWarnings=FALSE)

makeEffectSizeBoxplot_cancer(
    rslt_lists, dataset="Hu", winsorize.quantil=0.01)
makeEffectSizeBoxplot_cancer(
    rslt_lists, dataset="Zhou", winsorize.quantil=0.01)

makeEffectSizeBoxplot_cancer(
    rslt_lists, 
    dataset="Hu", effect.statistic="t",
    winsorize.quantil=0)
makeEffectSizeBoxplot_cancer(
    rslt_lists, dataset="Zhou", 
    effect.statistic="t",
    winsorize.quantil=0) 

```



```{r}
# GBM is not included here
OUTDIR = "../results/CellFie_Zhou2020/"
NdatZhou = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
makeExprBoxplot_system(
    NormExpr=NdatZhou, 
    rslt_lists,
    dataset="Zhou", 
    winsorize.quantil=0.05,
    outDirSummary)


OUTDIR = "../results/CellFie_Hu2025/"
NdatHu = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
rm(OUTDIR)
gc()
makeExprBoxplot_system(
    NormExpr=NdatHu, 
    rslt_lists,
    dataset="Hu", 
    winsorize.quantil=0.05,
    outDirSummary)

```


# Network analysis
Here, we will construct a co-expression network and identify hub features, which enables better detection of metabolic rewiring in the context of cancer versus control comparisons. However, this approach requires a sufficiently large sample size (typically >20 samples), so we will focus on the Hu2025 dataset for this analysis.

```{r}
# Proteomics data
## Hu 2025
OUTDIR = "../results/CellFie_Hu2025/"
NdatHu = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
NdatHu = NdatHu[!duplicated(NdatHu$genes),]
rownames(NdatHu) = NULL
NdatHu = NdatHu %>% tibble::column_to_rownames(var="genes")
rm(OUTDIR)
OUTDIR = "../results/CellFie_Hu2025/"
taskInfo2 = readRDS(paste0(OUTDIR, "/CellFieOut/taskInfo.RDS"))

lapply(names(rslt_lists), function(s) {
    getSystemNetwork(
        dat=NdatHu, 
        System=s, 
        suffix="Hu",
        outdir="../results/FeatureDifferential_CellFie/")
})
# getSystemNetwork(
#         dat=NdatHu, 
#         System=names(rslt_lists), 
#         outdir="../results/FeatureDifferential_CellFie/")

## Zhou 2020
OUTDIR = "../results/CellFie_Zhou2020/"
NdatHu = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
NdatHu = NdatHu[!duplicated(NdatHu$genes),]
rownames(NdatHu) = NULL
NdatHu = NdatHu %>% tibble::column_to_rownames(var="genes")
rm(OUTDIR)
OUTDIR = "../results/CellFie_Zhou2020/"
taskInfo2 = readRDS(paste0(OUTDIR, "/CellFieOut/taskInfo.RDS"))
lapply(names(rslt_lists), function(s) {
    getSystemNetwork(
        dat=NdatHu, 
        System=s, 
        suffix="Zhou",
        outdir="../results/FeatureDifferential_CellFie/")
})



# Transcriptomics data
NdatTCGA = 
    read.csv2("../results/TCGA/Limma_GSVA/NormalizedCount.csv", 
    sep="\t")
NdatTCGA = NdatTCGA %>% tibble::column_to_rownames(var="Features") %>%
    tibble::rownames_to_column(var="symbol")
NdatTCGA[,2:ncol(NdatTCGA)] = 
    lapply(NdatTCGA[,2:ncol(NdatTCGA)], function(x) as.numeric(x))
databaseDIR="../results/TCGA/KEGG_metabolism/"
metab_db_table = 
    readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS"))
kWithin = lapply(unique(metab_db_table$class), function(s) {
    getSystemNetwork(
        dat=NdatTCGA, 
        System=s, 
        suffix="TCGA",
        outdir="../results/FeatureDifferential_CellFie/",
        size.index=1,
        framework = "TCGA", 
        keggTable=metab_db_table)
}) %>% setNames(unique(metab_db_table$class))
kWithin2 = lapply(names(kWithin), function(k) {
    dat = as.data.frame(sort(kWithin[[k]], decreasing=T))
    colnames(dat) = k
    dat %>% tibble::rownames_to_column(var="Gene")
}) #%>% setNames(names(kWithin))
openxlsx::write.xlsx(kWithin2, 
    "../results/FeatureDifferential_CellFie/Network/Hubness.xlsx")


lapply(unique(metab_db_table$class), function(s) {
    getSystemNetwork(
        dat=NdatTCGA, 
        System=s, 
        suffix="TCGA_larger",
        outdir="../results/FeatureDifferential_CellFie/",
        size.index=2,
        vertex.size.index=5,
        framework = "TCGA", 
        keggTable=metab_db_table)
})

lapply(unique(metab_db_table$class), function(s) {
    getSystemNetwork(
        dat=NdatTCGA, 
        System=s, 
        suffix="TCGA_larger2",
        outdir="../results/FeatureDifferential_CellFie/",
        size.index=5,
        vertex.size.index=3,
        framework = "TCGA", 
        keggTable=metab_db_table)
})

```

```{r}
out = capture.output(sessionInfo())
writeLines(out, 
    "../results/FeatureDifferential_CellFie/sessionInfo.txt")

```
