---
title: "Untitled"
format: html
---


```{r}
# BiocManager::install("mygene")
# library(mygene)
library(dplyr)
library(parseMetab)
```

```{r}
#' Convert gene entrez ID to gene symbol using mygene, then extract the LIPIDS METABOLISM.
#' 
#' @param taskInfo2 A cellfie output tabel generated with the processCellFieOutput() function. It should have a column named GeneAssociatedToEssentialRxnsTask.
#' @examples 
#' # getSystem(taskInfo2)
#' 
getSystem = function(taskInfo2, System) {
    entrezID = unique(taskInfo2$GeneAssociatedToEssentialRxnsTask)
    res = mygene::queryMany(
        entrezID,
        scopes = "entrezgene",
        fields = "symbol",
        species = "human"
    )
    mapping = setNames(res$symbol, res$query)
    taskInfo2$GeneAssociatedToEssentialRxnsTask_symbol = 
        mapping[as.character(taskInfo2$GeneAssociatedToEssentialRxnsTask)]
    head(taskInfo2)
    system = taskInfo2 %>% filter(Depth1%in%System)
    return(system)
}


visualizeDiffFeatures = function(
    limma_rslt=limma_rslt_list, 
    outdir=outDirZhou, 
    taskInfo=systemZhou,
    h=NULL, w=NULL,
    max_logFC=3,
    depth="Depth3") {

    logFCdat = lapply(names(limma_rslt), function(l) {
        lr = limma_rslt[[l]] %>% dplyr::select(Feature, logFC) %>%
            setNames(c("Feature", l))
    }) %>% Reduce(dplyr::full_join, .) %>% as.data.frame() %>%
        tibble::column_to_rownames(var="Feature")
    logFCdat = logFCdat[rev(order(rowSums(logFCdat))), ]

    padj_dat = lapply(names(limma_rslt), function(l) {
        lr = limma_rslt[[l]] %>% dplyr::select(Feature, adj.P.Val) %>%
            setNames(c("Feature", l))
    }) %>% Reduce(dplyr::full_join, .) %>% as.data.frame() %>%
        tibble::column_to_rownames(var="Feature")
    padj_dat = padj_dat[rownames(logFCdat), ]

    sig_mat = matrix("", nrow=nrow(logFCdat), ncol=ncol(logFCdat))
    sig_mat[padj_dat<0.05] = "*"

    taskInfo$Depth = taskInfo[[depth]]
    ann_row = taskInfo %>% dplyr::select(Depth, GeneAssociatedToEssentialRxnsTask_symbol) %>%
        na.omit() %>% unique() %>% 
        filter(GeneAssociatedToEssentialRxnsTask_symbol %in% as.character(rownames(logFCdat))) %>%
        mutate(value = 1) %>%  # mark presence with 1
        tidyr::pivot_wider(names_from = Depth, 
            values_from = value,
            values_fill = list(value = 0)) %>%
        tibble::column_to_rownames(var="GeneAssociatedToEssentialRxnsTask_symbol")
    ann_row = ann_row[rownames(logFCdat), ]
    ann_colors = 
        replicate(ncol(ann_row), c("0" = "snow2", "1" = "#d95f02"), 
        simplify = FALSE) %>% setNames(colnames(ann_row))
    
    logFCdat[abs(logFCdat)>max_logFC] = max_logFC
    max_abs = max(abs(logFCdat), na.rm = TRUE)
    # Define breaks centered at 0
    breaks = seq(-max_abs, max_abs, length.out = 101)
    p1 = pheatmap::pheatmap(logFCdat,
            color = colorRampPalette(c("blue", "white", "red"))(100),
            breaks = breaks,
            # annotation_col = ann_row,   # top bar
            display_numbers = sig_mat,
            annotation_row = ann_row,
            annotation_colors = ann_colors,
            show_rownames=TRUE, 
            show_colnames=TRUE,
            fontsize = 10,
            border_color = "white",
            annotation_legend = FALSE,
            main=paste0("logFC (*: padj<0.05)"))
    if (is.null(w)) {
        w = 0.165*(ncol(logFCdat)+length(ann_colors))+2
    }
    if (is.null(h)) {
        h = 0.12*nrow(logFCdat)+0.05*max(nchar(colnames(ann_row))) +2.5
    }
    pdf(paste0(outdir, "/logFC_tumorVSnormal", depth, ".pdf"),
        h=h, w=w)
    print(p1)
    dev.off()
}

systemDiffAnalysis = function(
    System="LIPIDS METABOLISM",
    outDirHu, outDirZhou){
    print("Extract the specific METABOLISM genes ...")
    # Hu2025
    OUTDIR = "../results/CellFie_Hu2025/"
    taskInfo2 = readRDS(paste0(OUTDIR, "/CellFieOut/taskInfo.RDS"))
    systemHu = getSystem(taskInfo2, System=System)
    systemHu = systemHu %>% unique()
    saveRDS(systemHu, paste0(outDirHu, "/systemTaskInfoHu.RDS"))
    openxlsx::write.xlsx(list(Hu2025=systemHu),
        paste0(outDirHu, "systemTaskInfoHu.xlsx"))
    rm(taskInfo2, OUTDIR)
    gc()

    # Zhou2020
    OUTDIR = "../results/CellFie_Zhou2020/"
    taskInfo2 = 
        read.csv(paste0(OUTDIR, "/CellFieOut/all/detailScoring_new.csv"), sep="\t", header=TRUE)
    head(taskInfo2)
    systemZhou = getSystem(taskInfo2, System=System)
    systemZhou = systemZhou %>% unique()
    saveRDS(systemZhou, paste0(outDirZhou, "/systemTaskInfoZhou.RDS"))
    openxlsx::write.xlsx(list(Zhou2020=systemZhou),
        paste0(outDirZhou, "systemTaskInfoZhou.xlsx"))
    rm(taskInfo2, OUTDIR)
    gc()

    print("Paired differential analysis ... ")
    OUTDIR = "../results/CellFie_Hu2025/"
    NdatHu = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
    systemHu = readRDS(paste0(outDirHu, "/systemTaskInfoHu.RDS"))
    metaHu = readRDS(paste0(OUTDIR, "/data/Hu2025/Meta_hu2025.RDS"))
    saveRDS(metaHu, paste0(outDirHu, "/metaHu.RDS"))

    NdatHu = NdatHu[!duplicated(NdatHu$genes),]
    rownames(NdatHu) = NULL
    NdatHu = NdatHu %>% tibble::column_to_rownames(var="genes")
    NdatHu[] = lapply(NdatHu, as.numeric)
    sharedGenes = 
        intersect(as.character(unique(systemHu$GeneAssociatedToEssentialRxnsTask)), rownames(NdatHu))
    patientPair = which(table(metaHu$Patient)==2) %>% names()
    limma_rslt_listHu = 
        lapply(unique(metaHu$Cancer_type), function(c) {
            print(c)
            meta_sub = metaHu %>% filter(Patient%in%patientPair) %>% filter(Cancer_type==c)
            if ((nrow(meta_sub)>4) & length(unique(meta_sub$Condition))==2) {
                limmaTest_CellFie(
                    dat=NdatHu[sharedGenes,], 
                    paired=TRUE, 
                    currentCovariate=NULL, 
                    checkCovariate=FALSE, 
                    meta=meta_sub, 
                    paired_variable="Patient")  
            } else {NA}
        }) %>%setNames(unique(metaHu$Cancer_type))
    limma_rslt_listHu = limma_rslt_listHu[!is.na(limma_rslt_listHu)]
    rm(NdatHu, metaHu, sharedGenes, OUTDIR, patientPair)
    gc()

    mappingHu = setNames(systemHu$GeneAssociatedToEssentialRxnsTask_symbol,
        systemHu$GeneAssociatedToEssentialRxnsTask)
    limma_rslt_listHu = lapply(limma_rslt_listHu, function(x) {
            x$rslt_of_interest %>% 
            tibble::rownames_to_column(var="EntrezID")%>%
            mutate(Feature=as.character(mappingHu[EntrezID])) %>%
            dplyr::select(Feature, everything())})
    saveRDS(limma_rslt_listHu, paste0(outDirHu, "limma_rslt_listHu.RDS"))
    openxlsx::write.xlsx(
        limma_rslt_listHu, 
        file = paste0(outDirHu, "limma_reslut.xlsx"))

    OUTDIR = "../results/CellFie_Zhou2020/"
    NdatZhou = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
    systemZhou = readRDS(paste0(outDirZhou, "/systemTaskInfoZhou.RDS"))
    metaZhou = readRDS(paste0(OUTDIR,"/data/Zhou2020/meta.RDS"))
    saveRDS(metaZhou, paste0(outDirZhou, "/metaZhou.RDS"))

    sum(metaZhou$Sample %in% colnames(NdatZhou))
    sum(colnames(NdatZhou) %in% metaZhou$Sample)

    NdatZhou = NdatZhou[!duplicated(NdatZhou$genes),]
    rownames(NdatZhou) = NULL
    NdatZhou = NdatZhou %>% tibble::column_to_rownames(var="genes")
    NdatZhou[] = lapply(NdatZhou, as.numeric)
    sharedGenes = 
        intersect(as.character(unique(systemZhou$GeneAssociatedToEssentialRxnsTask)), rownames(NdatZhou))
    patientPair = which(table(metaZhou$Patient)==2) %>% names()
    limma_rslt_listZhou = 
        lapply(unique(metaZhou$Cancer_type), function(c) {
            print(c)
            meta_sub = metaZhou %>% filter(Patient%in%patientPair) %>% filter(Cancer_type==c)
            if ((nrow(meta_sub)>4) & length(unique(meta_sub$Condition))==2) {
                limmaTest_CellFie(
                    dat=NdatZhou[sharedGenes,], 
                    paired=TRUE, 
                    currentCovariate=NULL, 
                    checkCovariate=FALSE, 
                    meta=meta_sub, 
                    paired_variable="Patient")  
            } else {NA}
        }) %>%setNames(unique(metaZhou$Cancer_type))
    limma_rslt_listZhou = limma_rslt_listZhou[!is.na(limma_rslt_listZhou)]
    rm(NdatZhou, metaZhou, sharedGenes, OUTDIR, patientPair)
    gc()

    print("Save differential analysis results ...")
    mappingZhou = 
        setNames(systemZhou$GeneAssociatedToEssentialRxnsTask_symbol,
        systemZhou$GeneAssociatedToEssentialRxnsTask)
    limma_rslt_listZhou = lapply(limma_rslt_listZhou,function(x) {
            x$rslt_of_interest %>% 
            tibble::rownames_to_column(var="EntrezID")%>%
            mutate(Feature=as.character(mappingZhou[EntrezID])) %>%
            dplyr::select(Feature, everything())})
    saveRDS(limma_rslt_listZhou, paste0(outDirZhou, "limma_rslt_listZhou.RDS"))
    openxlsx::write.xlsx(
        limma_rslt_listZhou, 
        file = paste0(outDirZhou, "limma_reslut.xlsx"))

    return(list(limma_rslt_listZhou = limma_rslt_listZhou,
            limma_rslt_listHu = limma_rslt_listHu,
            systemHu = systemHu,
            systemZhou = systemZhou))
}

```

# Run differential analysis on genes within each metabolic system
```{r}
OUTDIR = "../results/CellFie_Hu2025/"
taskInfo2 = readRDS(paste0(OUTDIR, "/CellFieOut/taskInfo.RDS"))
systems = taskInfo2$Depth1 %>% unique()
rslt_lists = lapply(systems, function(s) {
    print(s)
    name = gsub(" .*", "", s)
    outDirHu = paste0("../results/FeatureDifferential_CellFie/", name, "Differential/Hu/")
    dir.create(outDirHu, recursive=TRUE, showWarnings=FALSE)
    outDirZhou = paste0("../results/FeatureDifferential_CellFie/", name, "Differential/Zhou/")
    dir.create(outDirZhou, recursive=TRUE, showWarnings=FALSE)

    rslt_list = systemDiffAnalysis(System=s, outDirHu, outDirZhou)

    visualizeDiffFeatures(limma_rslt=rslt_list$limma_rslt_listZhou, 
        outdir=outDirZhou, 
        taskInfo=rslt_list$systemZhou,
        max_logFC=4)

    visualizeDiffFeatures(limma_rslt=rslt_list$limma_rslt_listHu, 
        outdir=outDirHu, 
        taskInfo=rslt_list$systemHu,max_logFC=4)

    visualizeDiffFeatures(limma_rslt=rslt_list$limma_rslt_listZhou, 
        outdir=outDirZhou, 
        taskInfo=rslt_list$systemZhou,
        max_logFC=4, depth="Depth2")

    visualizeDiffFeatures(limma_rslt=rslt_list$limma_rslt_listHu, 
        outdir=outDirHu, 
        taskInfo=rslt_list$systemHu, 
        max_logFC=4, depth="Depth2")

    return(rslt_list)
}) %>% setNames(systems)


# names(rslt_lists) = systems
saveRDS(rslt_lists, "../results/FeatureDifferential_CellFie/rslt_list.RDS")
rslt_lists = readRDS("../results/FeatureDifferential_CellFie/rslt_list.RDS")
```

# Compare effect size among systems
```{r}
outDirSummary = "../results/FeatureDifferential_CellFie/summary/"
outDirSummary %>% dir.create(., recursive=TRUE, showWarnings=FALSE)

# Hu2025
makeEffectSizeBoxplot_system = function(
    rslt_lists, dataset="Hu", 
    effect.statistic="logFC",
    winsorize.quantil=0.05) {
    if (dataset=="Hu") {
        limma_rslt = "limma_rslt_listHu"
    } else if (dataset=="Zhou") {
        limma_rslt = "limma_rslt_listZhou"
    } else {stop("dataset must be 'Hu' or 'Zhou'")}

    dat4plot = lapply(names(rslt_lists), function(sname) {
        s = rslt_lists[[sname]]
        lapply(s[[limma_rslt]], function(c) {
            c$logFC = c[[effect.statistic]]
            c %>% dplyr::select(Feature, logFC)
        }) %>% Reduce(rbind, .) %>%
        mutate(System = sname)
    }) %>% Reduce(rbind, .)

    dat4plot$System = gsub("METABOLISM", "M.", dat4plot$System)
    order = dat4plot %>% group_by(System) %>% 
        summarize(median=median(logFC)) %>% 
        arrange(median) %>% pull(System) %>% unique()
    dat4plot$System = factor(dat4plot$System, levels=order)
    quantileLower = quantile(dat4plot$logFC, probs=winsorize.quantil)
    quantileUpper = quantile(dat4plot$logFC, probs=1-winsorize.quantil)
    dat4plot$logFC[dat4plot$logFC < quantileLower] = quantileLower
    dat4plot$logFC[dat4plot$logFC > quantileUpper] = quantileUpper

    pdf(paste0(outDirSummary, "/effectSize_",effect.statistic,"_system_", dataset, ".pdf"), h=4, w= 3)
    print(dat4plot %>% ggplot2::ggplot(., ggplot2::aes(x=System, y=logFC)) +
        ggplot2::geom_violin(trim=FALSE, fill="#E69F00", color="#E69F00")+
        ggplot2::geom_boxplot(width=0.4, fill="grey", alpha = 0.4, color="black", linewidth = 0.1, outlier.size = 0.5) + 
        ggplot2::labs(title = "",
            x="System", y=effect.statistic)+
        ggplot2::geom_hline(yintercept = 0, linetype = "dashed", color = "red", linewidth = 0.5) +
        ggplot2::theme_minimal() +
        ggplot2::theme(legend.position="none",
            axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0)))
    dev.off()
}

makeEffectSizeBoxplot_system(
    rslt_lists, dataset="Hu", 
    winsorize.quantil=0.05)
makeEffectSizeBoxplot_system(
    rslt_lists, dataset="Zhou", 
    winsorize.quantil=0.05)

makeEffectSizeBoxplot_system(
    rslt_lists, 
    dataset="Hu", effect.statistic="t",
    winsorize.quantil=0.02)
makeEffectSizeBoxplot_system(
    rslt_lists, dataset="Zhou", 
    effect.statistic="t",
    winsorize.quantil=0.02) 

```

# Compare effect size among cancer types
```{r}
outDirSummary = "../results/FeatureDifferential_CellFie/summary/"
outDirSummary %>% dir.create(., recursive=TRUE, showWarnings=FALSE)

makeEffectSizeBoxplot_cancer = function(
    rslt_lists, dataset="Hu", 
    effect.statistic="logFC",
    winsorize.quantil=0.05) {
    if (dataset=="Hu") {
        limma_rslt = "limma_rslt_listHu"
    } else if (dataset=="Zhou") {
        limma_rslt = "limma_rslt_listZhou"
    } else {stop("dataset must be 'Hu' or 'Zhou'")}

    cancers = names(rslt_lists[[1]][[limma_rslt]])
    dat4plot = lapply(cancers, function(c) {
        lapply(names(rslt_lists), function(sname) {
            cdat = rslt_lists[[sname]][[limma_rslt]][[c]]
            cdat$logFC = cdat[[effect.statistic]]
            cdat %>% dplyr::select(Feature, logFC)
        }) %>% Reduce(rbind, .) %>%
        mutate(Cancer = c)
    }) %>% Reduce(rbind, .)

    order = dat4plot %>% group_by(Cancer) %>% 
        summarize(median=median(logFC)) %>% 
        arrange(median) %>% pull(Cancer) %>% unique()
    dat4plot$Cancer = factor(dat4plot$Cancer, levels=order)
    quantileLower = quantile(dat4plot$logFC, probs=winsorize.quantil)
    quantileUpper = quantile(dat4plot$logFC, probs=1-winsorize.quantil)
    dat4plot$logFC[dat4plot$logFC < quantileLower] = quantileLower
    dat4plot$logFC[dat4plot$logFC > quantileUpper] = quantileUpper

    w = length(order)*0.15 + 1
    pdf(paste0(outDirSummary, "/effectSize_",effect.statistic,"_Cancer_", dataset, ".pdf"), h=4, w= w)
    print(dat4plot %>% ggplot2::ggplot(., ggplot2::aes(x=Cancer, y=logFC)) +
        # ggplot2::geom_violin(trim=FALSE, fill="#E69F00", color="white")+
        ggplot2::geom_boxplot(width=0.75, fill="#E69F00", outlier.size = 0.5, line.width=0.3) + 
        ggplot2::labs(title = "",
            x="Cancer", y=effect.statistic)+
        ggplot2::geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
        ggplot2::theme_minimal() +
        ggplot2::theme(legend.position="none",
            axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0)))
    dev.off()
}

makeEffectSizeBoxplot_cancer(
    rslt_lists, dataset="Hu", winsorize.quantil=0.01)
makeEffectSizeBoxplot_cancer(
    rslt_lists, dataset="Zhou", winsorize.quantil=0.01)

makeEffectSizeBoxplot_cancer(
    rslt_lists, 
    dataset="Hu", effect.statistic="t",
    winsorize.quantil=0)
makeEffectSizeBoxplot_cancer(
    rslt_lists, dataset="Zhou", 
    effect.statistic="t",
    winsorize.quantil=0) 

```



```{r}
makeExprBoxplot_system = function(
    NormExpr=NdatHu, 
    rslt_lists,
    dataset="Hu", 
    winsorize.quantil=0.05,
    outDirSummary) {

    if (dataset=="Hu") {
        limma_rslt = "limma_rslt_listHu"
    } else if (dataset=="Zhou") {
        limma_rslt = "limma_rslt_listZhou"
    } else {stop("dataset must be 'Hu' or 'Zhou'")}
    NormExpr = NormExpr[!duplicated(NormExpr$genes), ] 
    rownames(NormExpr) = NULL
    NormExpr = NormExpr %>% 
        tibble::column_to_rownames(var="genes")
    dat4plot = lapply(names(rslt_lists), function(sname) {
        cancers = names(rslt_lists[[sname]][[limma_rslt]])
        lapply(cancers, function(c) {
            features = rslt_lists[[sname]][[limma_rslt]][[c]][["EntrezID"]]
            dat = NormExpr[features, ]
            rownames(dat) = NULL
            rownames(dat) = rslt_lists[[sname]][[limma_rslt]][[c]][["Feature"]]
            data.frame(Feature=rownames(dat), meanExpr = rowMeans(dat), System=sname)
        }) %>% Reduce(rbind, .)
    }) %>% Reduce(rbind, .)
    dat4plot$System = gsub("METABOLISM", "M.", dat4plot$System)
    order = dat4plot %>% group_by(System) %>% 
        summarize(median=median(meanExpr)) %>% 
        arrange(median) %>% pull(System) %>% unique()
    dat4plot$System = factor(dat4plot$System, levels=order)
    quantileLower = quantile(dat4plot$meanExpr, probs=winsorize.quantil)
    quantileUpper = quantile(dat4plot$meanExpr, probs=1-winsorize.quantil)
    dat4plot$meanExpr[dat4plot$meanExpr < quantileLower] = quantileLower
    dat4plot$meanExpr[dat4plot$meanExpr > quantileUpper] = quantileUpper

    pdf(paste0(outDirSummary, "/FeatureAbundance_system_", dataset, ".pdf"), h=4, w= 2)
    print(dat4plot %>% ggplot2::ggplot(., ggplot2::aes(x=System, y=meanExpr)) +
        ggplot2::geom_violin(trim=FALSE, fill="#E69F00", color="white", width=1)+
        ggplot2::geom_boxplot(width=0.1, line.width=0.1) + 
        ggplot2::labs(title = "",
            x="System", y="meanExpr")+
        # ggplot2::geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
        ggplot2::theme_minimal() +
        ggplot2::theme(legend.position="none",
            axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0)))
    dev.off()
}

OUTDIR = "../results/CellFie_Zhou2020/"
NdatZhou = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
makeExprBoxplot_system(
    NormExpr=NdatZhou, 
    rslt_lists,
    dataset="Zhou", 
    winsorize.quantil=0.05,
    outDirSummary)


OUTDIR = "../results/CellFie_Hu2025/"
NdatHu = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
rm(OUTDIR)
gc()
makeExprBoxplot_system(
    NormExpr=NdatHu, 
    rslt_lists,
    dataset="Hu", 
    winsorize.quantil=0.05,
    outDirSummary)

```


# Network analysis
Here, we will construct a co-expression network and identify hub features, which enables better detection of metabolic rewiring in the context of cancer versus control comparisons. However, this approach requires a sufficiently large sample size (typically >20 samples), so we will focus on the Hu2025 dataset for this analysis.

```{r}
#' @param framework this can be either "CellFie" or "GSVA"
#' 
getSystemNetwork = function(
    dat=NdatHu, taskInfo=taskInfo2, System, outdir, suffix="",
    framework="CellFie", keggTable=NULL, size.index=1,
    vertex.size.index=10) {
    if (framework == "CellFie") {
        systemHu = getSystem(taskInfo, System=System)
        sub = dat[rownames(dat) %in% systemHu$GeneAssociatedToEssentialRxnsTask,]
        mapping = setNames(
            systemHu$GeneAssociatedToEssentialRxnsTask_symbol, 
            as.character(systemHu$GeneAssociatedToEssentialRxnsTask))
        sub$symbol = as.character(mapping[rownames(sub)])
        sub = sub[!duplicated(sub$symbol), ]
        rownames(sub) = NULL
        sub = sub %>% tibble::column_to_rownames(var="symbol")
    } else if (!is.null(keggTable)){
            keggTable_sub = filter(keggTable, class==System)
            sub = dat[dat$symbol %in% keggTable_sub$gene_symbol,]
            sub = sub[!duplicated(sub$symbol), ]
            rownames(sub) = NULL
            sub = sub %>% tibble::column_to_rownames(var="symbol")
    } else {
            kegg_metab_dbs = 
                getdb_metabolism(databaseDIR=paste0(outdir, "/KEGG_metabolism/"))
            keggTable_sub = filter(kegg_metab_dbs$kegg_metab_db, class==System)
            sub = dat[dat$symbol %in% keggTable_sub$gene_symbol,]
            sub = sub[!duplicated(sub$symbol), ]
            rownames(sub) = NULL
            sub = sub %>% tibble::column_to_rownames(var="symbol")
    }

    # Calcaulte co-expression matrix: genes x samples and find hub features
    cor_mat = cor(t(sub), method = "spearman")
    kWithin = rowSums(abs(cor_mat)) - 1 # hubness
    hub.nr = max(3, 0.03*length(kWithin))
    hubs = names(sort(kWithin, decreasing = TRUE))[1:hub.nr]

    library(igraph)
    # Filter the cor matrix to only keep those cor >0.6
    # correlation threshold
    thr = 0.6
    cor_mat[is.na(cor_mat)] = 0
    edges = which(
        abs(cor_mat) > thr & lower.tri(cor_mat),
        arr.ind = TRUE
    )
    # You’re extracting the pairs of genes whose absolute correlation is above your threshold, 
    # but only from the lower triangle to avoid duplicates, and returning their positions as row/column indices. 
    # These pairs represent the edges in your co-expression network.
    edge_df = data.frame(
    from   = rownames(cor_mat)[edges[,1]],
    to     = colnames(cor_mat)[edges[,2]],
    weight = cor_mat[edges]
    )
    # Build a filtered graph
    g = graph_from_data_frame(edge_df, directed = FALSE)

    # add node attributes
    V(g)$kWithin = kWithin[V(g)$name] #hubness
    V(g)$isHub   = V(g)$name %in% hubs
    # # You’re making a smaller graph containing only your hub genes plus their direct neighbors
    # g_viz = induced_subgraph(
    #   g,
    #   vids = unique(unlist(ego(g, nodes = hubs, order = 1)))
    # )

    set.seed(42)
    paste0(outdir, "/Network/") %>% dir.create(recursive=T, showWarnings=F)
    h = w = size.index*length(V(g))/10 + 5
    pdf(paste0(outdir, "/Network/Network_hub_", gsub(" ", "", System),suffix,".pdf"), h=h,w=w)
    library(scales)
    edge.color = ifelse(E(g)$weight > 0, "snow2", "skyblue")
    E(g)$weight = abs(E(g)$weight)
    plot(
        g,
        layout = layout_with_fr,
        
        vertex.size = vertex.size.index * (V(g)$kWithin - min(V(g)$kWithin)) / (max(V(g)$kWithin) - min(V(g)$kWithin)),  
        # normalize size nicely
        vertex.color = ifelse(V(g)$isHub, alpha("#D73027",0.6), alpha("orange", 0.6)),  # nicer red and blue shades
        vertex.frame.color = "gold",   # subtle border around nodes
        
        vertex.label = V(g)$name,
        vertex.label.cex = 1,
        vertex.label.color = "black",
        vertex.label.family = "Helvetica",
        
        edge.width = 1 + 4 * abs(E(g)$weight),  # base width plus scaled by weight
        edge.color = edge.color,  # clean blue/red colors
        edge.curved = 0.1,  # slight curve to edges for better visibility
        
        margin = c(0, 0, 0, 0),  # reduce plot margin
        main = "Co-expression Network: Hubs and Neighbors"
    )
    dev.off()

    # # Only keep postively correlated edges
    # edge_df <- subset(edge_df, weight > thr)
    # V(g)$betweenness <- betweenness(g)
    # V(g)$eigen <- eigen_centrality(g)$vector
    # library(ggraph)
    # set.seed(1)
    # ggraph(g_viz, layout = "fr") +
    #   geom_edge_link(aes(width = abs(weight), color = weight > 0), alpha = 0.7) +
    #   geom_node_point(aes(size = kWithin, color = isHub)) +
    #   geom_node_text(aes(label = name), repel = TRUE) +
    #   theme_void()
    return(kWithin)
}

# Proteomics data
## Hu 2025
OUTDIR = "../results/CellFie_Hu2025/"
NdatHu = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
NdatHu = NdatHu[!duplicated(NdatHu$genes),]
rownames(NdatHu) = NULL
NdatHu = NdatHu %>% tibble::column_to_rownames(var="genes")
rm(OUTDIR)
OUTDIR = "../results/CellFie_Hu2025/"
taskInfo2 = readRDS(paste0(OUTDIR, "/CellFieOut/taskInfo.RDS"))

lapply(names(rslt_lists), function(s) {
    getSystemNetwork(
        dat=NdatHu, 
        System=s, 
        suffix="Hu",
        outdir="../results/FeatureDifferential_CellFie/")
})
# getSystemNetwork(
#         dat=NdatHu, 
#         System=names(rslt_lists), 
#         outdir="../results/FeatureDifferential_CellFie/")

## Zhou 2020
OUTDIR = "../results/CellFie_Zhou2020/"
NdatHu = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
NdatHu = NdatHu[!duplicated(NdatHu$genes),]
rownames(NdatHu) = NULL
NdatHu = NdatHu %>% tibble::column_to_rownames(var="genes")
rm(OUTDIR)
OUTDIR = "../results/CellFie_Zhou2020/"
taskInfo2 = readRDS(paste0(OUTDIR, "/CellFieOut/taskInfo.RDS"))
lapply(names(rslt_lists), function(s) {
    getSystemNetwork(
        dat=NdatHu, 
        System=s, 
        suffix="Zhou",
        outdir="../results/FeatureDifferential_CellFie/")
})



# Transcriptomics data
NdatTCGA = 
    read.csv2("../results/TCGA/Limma_GSVA/NormalizedCount.csv", 
    sep="\t")
NdatTCGA = NdatTCGA %>% tibble::column_to_rownames(var="Features") %>%
    tibble::rownames_to_column(var="symbol")
NdatTCGA[,2:ncol(NdatTCGA)] = 
    lapply(NdatTCGA[,2:ncol(NdatTCGA)], function(x) as.numeric(x))
databaseDIR="../results/TCGA/KEGG_metabolism/"
metab_db_table = 
    readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS"))
kWithin = lapply(unique(metab_db_table$class), function(s) {
    getSystemNetwork(
        dat=NdatTCGA, 
        System=s, 
        suffix="TCGA",
        outdir="../results/FeatureDifferential_CellFie/",
        size.index=1,
        framework = "TCGA", 
        keggTable=metab_db_table)
}) %>% setNames(unique(metab_db_table$class))
kWithin2 = lapply(names(kWithin), function(k) {
    dat = as.data.frame(sort(kWithin[[k]], decreasing=T))
    colnames(dat) = k
    dat %>% tibble::rownames_to_column(var="Gene")
}) #%>% setNames(names(kWithin))
openxlsx::write.xlsx(kWithin2, 
    "../results/FeatureDifferential_CellFie/Network/Hubness.xlsx")


lapply(unique(metab_db_table$class), function(s) {
    getSystemNetwork(
        dat=NdatTCGA, 
        System=s, 
        suffix="TCGA_larger",
        outdir="../results/FeatureDifferential_CellFie/",
        size.index=2,
        vertex.size.index=5,
        framework = "TCGA", 
        keggTable=metab_db_table)
})

lapply(unique(metab_db_table$class), function(s) {
    getSystemNetwork(
        dat=NdatTCGA, 
        System=s, 
        suffix="TCGA_larger2",
        outdir="../results/FeatureDifferential_CellFie/",
        size.index=5,
        vertex.size.index=3,
        framework = "TCGA", 
        keggTable=metab_db_table)
})

```

```{r}
out = capture.output(sessionInfo())
writeLines(out, 
    "../results/FeatureDifferential_CellFie/sessionInfo.txt")

```
