---
title: Cancer metabolism
subtitle: Hu20025
author: Yuan Li
toc: true
number-sections: true
format: 
  html:
    #embed-resources: true #This works from command line not from Rstudio
    code-fold: true
    code-tools: true
    code-link: true
    theme: pandoc
    df-print: paged
    html-math-method: katex #use katex for math text
  pdf:
    code-line-numbers: true
    geometry: 
      - top=30mm
      - left=30mm
  docx: default
language:
  code-summary: Click to show code
editor_options:
  chunk_output_type: inline
editor: visual
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 7,
                      results    = "hold",
                      message    = FALSE,
                      cache.lazy = FALSE,
                      warning    = FALSE)

```

> Within the conda env SMS7546, run dpkg -i quarto-linux-arm64.deb (one needs to download quarto-linux-arm64.deb first to the work folder)

> To fun this notebook: quarto render Hu2025.qmd –to html

# Import and Prepare the data for CellFie

```{r}
library(dplyr)
library(ggplot2)
library(parseMetab)
```

```{r}
Pdat = 
  read.table("../data/Hu2025/ALL_PRO_PROJECT_RATIO_maxquant_20220316.txt", sep="\t", header=T)
Pdat[,9:ncol(Pdat)] = lapply(Pdat[,9:ncol(Pdat)], function(x) as.numeric(as.character(x)))
dim(Pdat)
# There are negative values, are those log2 transformed or scaled, if scaled, the mean should be 0
rowMeans(Pdat[, 9:ncol(Pdat)], na.rm=T) %>% mean(na.rm=T)
# -0.837735
colMeans(Pdat[, 9:ncol(Pdat)], na.rm=T) %>% mean(na.rm=T)
# -0.2027342
# The data probably has been log transformed, so we will transform them back
Pdat[,9:ncol(Pdat)] = 2^Pdat[,9:ncol(Pdat)]
rowMeans(Pdat[, 9:ncol(Pdat)], na.rm=T) %>% mean(na.rm=T)
colMeans(Pdat[, 9:ncol(Pdat)], na.rm=T) %>% mean(na.rm=T)

Pdat = Pdat[, c(3, 9:ncol(Pdat))] %>% as.data.frame() %>%
  mutate(genes=gene_symbol) %>% 
  dplyr::select(genes, everything(), -gene_symbol)

```



## Check whether the data has been normalized or not

```{r}
#' @import ggplot2
overviewPlot = function(data, variable2plot="TotalAbundance", plot="barplot", 
                        unit="Sample", log=F, ylim=F, title="") {
  if (variable2plot=="TotalAbundance") {
    temp = colSums(data, na.rm=T)
    if (log) temp=log2(temp+0.0001)
  } else if (variable2plot=="Nr.MissingValues") {
    temp = apply(data, 2, function(col) {
            sum(is.na(col))})
  } else if (variable2plot=="Nr.NonMissingValues") {
    temp = apply(data, 2, function(col) {
            sum(!is.na(col))})
  } else {
    temp = colSums(data, na.rm=T)
    if (log) temp=log2(temp+0.0001)
  }
  if (variable2plot %in% c("TotalAbundance","Nr.MissingValues","Nr.NonMissingValues")) {
    temp2 = data.frame(varaible2plot=as.numeric(temp),
                       Sample=names(temp),
                       Group=gsub("_([^_]+)$", "", names(temp)))
    if (log) {
      variable2plot = paste0(variable2plot, "_log2")
    }
    colnames(temp2)[1] = variable2plot
  }
  if (plot == "boxplot") {
    if (unit=="Sample") {
      if (log) {
        temp3 = tidyr::gather(log2(data+0.0001), key="Sample", value="Abundance")
        temp3$Group = gsub(".*_","",temp3$Sample) %>% gsub("NAT","Normal",.)
      } else {
        temp3 = tidyr::gather(data, key="Sample", value="Abundance")
        temp3$Group = gsub(".*_","",temp3$Sample) %>% gsub("NAT","Normal",.)
      }
      if (log) {
        v2plot = "Abundance_log2"
        colnames(temp3)[2]=v2plot} else {v2plot="Abundance"}
      p = ggplot2::ggplot(temp3, ggplot2::aes_string(x="Sample", y=v2plot)) +
        ggplot2::geom_boxplot(ggplot2::aes(fill=Group),
                    outlier.shape = 16,
                    outlier.size = 1,
                    outlier.colour = "grey") +
      ggplot2::labs(x="", y=v2plot)
    } else {
      p = ggplot2::ggplot(temp2, ggplot2::aes_string(x="Group", y=variable2plot)) +
        ggplot2::geom_boxplot(ggplot2::aes(fill=Group),
                     outlier.shape = 16,
                     outlier.size = 1,
                     outlier.colour = "grey") +
        ggplot2::labs(x="", y=variable2plot) 
    }
  } else if (plot == "barplot") {
    p = ggplot2::ggplot(temp2, ggplot2::aes_string(x="Sample", y=variable2plot)) +
      ggplot2::geom_bar(aes(fill=Group), stat = "identity") +
      ggplot2::theme(legend.position = "none")
  }
  if (ylim) {p = p + ylim(-250, 250)}
  p + ggplot2::theme_bw() +
    ggplot2::theme(axis.text.x=ggplot2::element_text(angle=270, vjust=0.5,hjust=1)) +
    #ggplot2::theme(legend.position = "none") +
    ggtitle(title)
}

OUTDIR = "../results/CellFie_Hu2025/"
QCDIR = paste0(OUTDIR, "/data/QC/")
dir.create(QCDIR, recursive=T, showWarnings=F)

cancers = gsub("_.*", "", colnames(Pdat)) %>% unique() %>% setdiff(., "genes")

pdf(paste0(QCDIR, "RawDataBoxplot_Hu2025.pdf"), h=5, w=36.5)
  lapply(cancers, function(c) {
    overviewPlot(Pdat[, grep(c, colnames(Pdat), value=T)], 
        variable2plot="TotalAbundance", plot="boxplot", 
        unit="Sample", log=T, ylim=F,
        title = paste0("Others_", c))
})
dev.off()

```

```{r}
library(AnnotationDbi)
library(org.Hs.eg.db)
library(dplyr)
Pdat2 = AnnotationDbi::select(
  org.Hs.eg.db,
  keys = Pdat$genes,
  columns = c("ENTREZID"),
  keytype = "SYMBOL"
) %>% na.omit() %>%
  .[!duplicated(.$SYMBOL),] %>%
  setNames(c("genes", "ENTREZID")) %>%
  right_join(Pdat) %>%
  na.omit() %>%
  dplyr::select(-genes)
names(Pdat2)[1] = "genes"

# Log transform the data
OUTDIR = "../results/CellFie_Hu2025/"
INDIR = paste0(OUTDIR, "data/Hu2025/")
dir.create(INDIR, recursive = T, showWarnings = F)

```

```{r}
#| eval: false
# Save the data table
# Save the input data to a csv file
lapply(cancers, function(c) {
  write.table(Pdat2[, c("genes", grep(c,colnames(Pdat2),value=T))], 
    paste0(INDIR, "CancerProteomicsHu2025_",c,".csv"), sep=",", 
    quote=F, row.names=F)
})

```

```{r}
# Check the number of samples for each cancer
lapply(cancers, function(c) {
  cat(paste0(c, ": ",
  ncol(Pdat2[, c(grep(c,colnames(Pdat2),value=T))]),
  "; "
  ))
}) 
# ESCC: 248; LIHC: 330; STAD: 160; GBM: 109; HNSC: 175; KIRC: 194; LUAD: 211; LUSC: 208; PAAD: 207; UCEC: 120; BRCA: 142; COAD: 197; OV: 103
rm(Pdat)

OUTDIR = "../results/CellFie_Hu2025/"
saveRDS(Pdat2, paste0(OUTDIR, "/Pdat2.RDS"))
# Pdat2 = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
SampleNames = colnames(Pdat2)
cancers = gsub("_.*", "", SampleNames) %>% unique() %>% setdiff(., "genes")
rm(Pdat2)
gc()

```

# Run CellFie

  Submit the non-negative normalized data file to the [genePatten webserver](https://cloud.genepattern.org/gp/pages/index.jsf), and run it with the following settings:

  -   Data="PsoPproteomics.csv",
  -   Sample.Number=,
  -   Reference.Model= Human_iHsa,
  -   Thresholding.Approach="local",
  -   Percentile.Or.Value="percentile",
  -   Local.Threshold.Type="minmaxmean",
  -   Local.Lower.Bound="25",
  -   Local.Upper.Bound="75"
  -   Uses gene-specific thresholds, customized per gene (based on its expression distribution)
  -   For each gene, CellFie computes its 25th and 75th percentile and considers it ON if expression > mean of [25th, 75th].

# Analyze CellFie output

## Sort out meta data

```{r}
library(dplyr)
OUTDIR = "../results/CellFie_Hu2025/"
QCDIR = paste0(OUTDIR, "/data/QC/")
cancers = c("ESCC", "LIHC", "STAD", "GBM", "HNSC", "KIRC", "LUAD", "LUSC", "PAAD", "UCEC", "BRCA", "COAD", "OV")

meta =  
  read.table("../data/Hu2025/ALL_PRO_PROJECT_RATIO_maxquant_20220316.sample.txt", 
    sep="\t", header=T) %>% mutate(Sample=Sample_ID) %>% 
  mutate(Patient=Patient_ID) %>% mutate(Cancer_type=Tumor_code) %>%
  mutate(Condition=Sample_type) %>% 
  mutate(Batch=gsub(".*Proteome_", "", Run_ID)) %>%
  mutate(Cancer_Condition = paste0(Cancer_type, "_", Condition))
meta$Condition[meta$Condition=="NAT"] = "Normal"
table(meta$Condition)
head(meta)

test = meta %>% mutate(Batch=gsub(".*Proteome_", "", Run_ID)) %>% 
  dplyr::select(Tumor_code, Sample_type, Batch) %>% 
  dplyr::count(Tumor_code, Sample_type, Batch) %>% 
  arrange(Batch)
# Confounding batch: 
Cbatch = names(table(test$Batch))[table(test$Batch)==1]
test[test$Batch%in%Cbatch, ]
table(meta$Tumor_code)
test[test$Batch%in%Cbatch, ] %>% filter(Tumor_code=="BRCA") %>% pull(n) %>% sum()
# We can see that for all BRCA batches, 2 batches of KIRC, 1 batch of HNSC, and 1 batch of UCEC, 
# Conditions are completely confounded with Batch, we will remove those batches

# We have in total 2404 sampels to start with
dim(meta)
rm(test)
gc()

# We will remove 189 samples:
meta %>% filter(Batch %in% Cbatch) %>% dim()

# There will be 2215 samples left
meta %>% filter(!Batch %in% Cbatch) %>% dim()

meta_all = meta %>% filter(!Batch %in% Cbatch)
dim(meta_all)

openxlsx::write.xlsx(list(
    meta = meta_all,
    meta_summary = meta_all %>% dplyr::count(Tumor_code, Paired),
    SampleSummary = meta_all %>% dplyr::count(Cancer_type, Condition)
  ), 
  file=paste0(OUTDIR, "/data/Hu2025/Meta_hu2025.xlsx"))

OUTDIR = "../results/CellFie_Hu2025/"
saveRDS(meta_all, paste0(OUTDIR, "/data/Hu2025/Meta_hu2025.RDS"))
# meta_all = readRDS(paste0(OUTDIR, "/data/Hu2025/Meta_hu2025.RDS"))

```


## Data summary

### Sample size summary

```{r}
meta_all = readRDS(paste0(OUTDIR, "/data/Hu2025/Meta_hu2025.RDS"))
SampleSizeCircularBarplot(meta=meta_all, 
  title="Hu2025", OUT_DIR=paste0(OUTDIR, "/Summary/"), h=4, w=4.5)

```


## Process CellFie output

```{r}
Pdat2 = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
SampleNames = colnames(Pdat2)
cancers = gsub("_.*", "", SampleNames) %>% unique() %>% setdiff(., "genes")
rm(Pdat2)
gc()
cfs = processCellFieOutput(combined=FALSE,
    outdir = paste0(OUTDIR, "/CellFieOut/"), 
    SampleNames, meta=meta, samples2keep=meta_all$Sample)
saveRDS(cfs, paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
rm(cfs, meta)
gc()

# Boxplot of task size (unique gene nr) summary
# plotTaskSizes(taskInfo = lapply(cancers, function(cancer) {
#     read.csv(paste0(OUTDIR, "/CellFieOut/", cancer, "/detailScoring_new.csv"),
#           header = TRUE, sep="\t", stringsAsFactors = FALSE) 
#   }) %>% Reduce(rbind,.), 
#               OUTDIR = paste0(OUTDIR, "/CellFieOut/../Summary/"))

taskInfo2 = lapply(cancers, function(c) {
    read.csv(paste0(OUTDIR, "/CellFieOut/", c, "/detailScoring_new.csv"),
          header = TRUE, sep="\t", stringsAsFactors = FALSE) %>%
          select(Depth1, Depth2, Depth3, GeneAssociatedToEssentialRxnsTask)
  }) %>% Reduce(rbind,.) %>% unique()

saveRDS(taskInfo2, paste0(OUTDIR, "/CellFieOut/taskInfo.RDS"))
head(taskInfo2)

```

### Metabolic activity boxplot by cancer

```{r}
meta_all = readRDS(paste0(OUTDIR, "/data/Hu2025/Meta_hu2025.RDS"))
cfs = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
ActivityBoxplot_byCancer(meta=meta_all, 
  cf=cfs, OUT_DIR=paste0(OUTDIR, "/Summary/"), title="Hu2025")

```

### Metabolic activity boxplot by system

```{r}
cfs = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
ActivityBoxplot_bySystem(
  cf=cfs%>%mutate(Depth1=gsub("METABOLISM","M.", Depth1)), 
  OUT_DIR=paste0(OUTDIR, "/Summary/"), w=2, h=3.5, title="Hu2025")

```


## Limma analysis (paired)

```{r}
cfs = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
meta_all = readRDS(paste0(OUTDIR, "/data/Hu2025/Meta_hu2025.RDS"))
depth = "Depth3"
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")
result3 = runLimmaCellFie(depth, cfs=cfs, 
  meta=meta_all, OUTDIRV, w=9.5, height=23)
saveRDS(result3, paste0(OUTDIRV, "/result3.RDS"))
# result3 = readRDS(paste0(OUTDIRV, "/result3.RDS"))

adj.P.Val.cutoff = 0.05
makeTaskScoreHeatmap_CellFie(rlst_list=result3$rlst_list, 
  cf5_Depth=result3$cf5_Depth, OUTDIRV=OUTDIRV,#  w=9.5,h=23,
  adj.P.Val.cutoff, meta=meta_all)

makeLimmaDotplot_CellFie(
  rlst_list=result3$rlst_list,
  adj.P.Val.cutoff, # w=9, h=19,  
  OUTDIRV=OUTDIRV, Depth=depth, lowfigure=F, title = "", 
  suffix="Hu2025_tasks_all", taskHierarchy=cfs, nested=T, filter=F)

makeLimmaDotplot_CellFie(
  rlst_list=result3$rlst_list,
  adj.P.Val.cutoff, # w=9, h=13.5,  
  OUTDIRV=OUTDIRV, Depth=depth, lowfigure=F, title = "", 
  suffix="Hu2025_tasks", taskHierarchy=cfs, nested=T, filter=T)

rm(result3)
gc()

```

```{r}
meta_all = readRDS(paste0(OUTDIR, "/data/Hu2025/Meta_hu2025.RDS"))
depth = "Depth2"
Depth2 = readRDS(paste0(OUTDIR, "Summary/TaskScores_", depth,".RDS"))
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth2/")
result2 = runLimmaCellFie(depth, cfs=Depth2, 
  meta=meta_all, OUTDIRV, w=9.5, height=17)
rm(Depth2)
gc()
saveRDS(result2, paste0(OUTDIRV, "/result2.RDS"))
# result2 = readRDS(paste0(OUTDIRV, "/result2.RDS"))

adj.P.Val.cutoff = 0.05
makeTaskScoreHeatmap_CellFie(rlst_list=result2$rlst_list, 
  cf5_Depth=result2$cf5_Depth, OUTDIRV=OUTDIRV, # w=9.5,h=23,
  adj.P.Val.cutoff, meta=meta_all)
cfs = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
makeLimmaDotplot_CellFie(
  rlst_list=result2$rlst_list,
  adj.P.Val.cutoff, w=8.5, #w=15, h=10,  
  OUTDIRV=OUTDIRV, Depth=depth, lowfigure=F, title = "", 
  suffix="Hu2025_subsystem_all", taskHierarchy=cfs, nested=T, filter=F)

rm(result2, cfs)
gc()

```

```{r}
meta_all = readRDS(paste0(OUTDIR, "/data/Hu2025/Meta_hu2025.RDS"))
depth = "Depth1"
Depth1 = readRDS(paste0(OUTDIR, "Summary/TaskScores_", depth,".RDS"))
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth1/")
result1 = runLimmaCellFie(depth, cfs=Depth1, 
  meta=meta_all, OUTDIRV, w=9.5, height=3.3)
saveRDS(result1, paste0(OUTDIRV, "/result1.RDS"))
rm(Depth1)
gc()

adj.P.Val.cutoff = 0.05
makeTaskScoreHeatmap_CellFie(rlst_list=result1$rlst_list, 
  cf5_Depth=result1$cf5_Depth, OUTDIRV, w=9.5,h=6.5,
  adj.P.Val.cutoff, meta=meta_all)

cfs = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
makeLimmaDotplot_CellFie(
  rlst_list=result1$rlst_list,
  adj.P.Val.cutoff, h=5, # w=9,  
  OUTDIRV=OUTDIRV, Depth=depth, lowfigure=T, title = "", 
  suffix="Hu2025_system_all", taskHierarchy=cfs, nested=F, filter=F)

makeLimmaDotplot_CellFie(
  rlst_list=result1$rlst_list,
  adj.P.Val.cutoff, h=3.5, # w=9,  
  OUTDIRV=OUTDIRV, Depth=depth, lowfigure=F, title = "", 
  suffix="Hu2025_system_all2", taskHierarchy=cfs, nested=F, filter=F)

rm(result1, cfs)
gc()
```


### Volcano plots

```{r}
adj.P.Val.cutoff = 0.05
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth1/")
result1 = readRDS(paste0(OUTDIRV, "result1.RDS"))
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth2/")
result2 = readRDS(paste0(OUTDIRV, "result2.RDS"))
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")
result3 = readRDS(paste0(OUTDIRV, "/result3.RDS"))
plots = lapply(names(result1$rlst_list), function(cancer) {
  dat2plot = rbind(result1$rlst_list[[cancer]]$rslt_of_interest %>% 
                      tibble::rownames_to_column(var="Task") %>%
                      #filter(adj.P.Val<adj.P.Val.cutoff) %>%
                      mutate(Depth = "Depth1"),
                   result2$rlst_list[[cancer]]$rslt_of_interest %>% 
                      tibble::rownames_to_column(var="Task") %>%
                      #filter(adj.P.Val<adj.P.Val.cutoff)%>%
                      mutate(Depth = "Depth2"),
                   result3$rlst_list[[cancer]]$rslt_of_interest%>% 
                      tibble::rownames_to_column(var="Task") %>%
                      #filter(adj.P.Val<adj.P.Val.cutoff)%>%
                      mutate(Depth = "Depth3"))
  #print(head(dat2plot))
  p = doVolcano_CellFie(resi=dat2plot, NAME=cancer, adj.P.Val.cutoff=0.05, 
                logFC.cutoff=0, lfcShrink=F)
  p
}) 

pdf(paste0(OUTDIR, "/Limma/Volcano_plots.pdf"),h=4,w=6)
  plots
dev.off()

pdf(paste0(OUTDIR, "/Limma/Volcano_plots2.pdf"),h=20,w=18)
  cowplot::plot_grid(plots[[1]],plots[[2]],plots[[3]],plots[[4]], 
                     plots[[5]],plots[[6]],plots[[7]],plots[[8]],
                     plots[[9]],plots[[10]],plots[[11]],plots[[12]],
                     ncol=3,nrow=5)
dev.off()
rm(plots)
rm (result1, result2, restul3)
gc()
```


### Limma result summary

#### Sig task nr barplot _ by cancer

```{r}
OUTDIR = "../results/CellFie_Hu2025/"
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")
result3 = readRDS(paste0(OUTDIRV, "/result3.RDS"))
SigNrBarplot(rlst_list = result3$rlst_list, 
  sig.cutoff=0.05, w=2.75, h=3.0, OUT_DIR=paste0(OUTDIR, "/Limma/"))
```

#### Differential effect boxplot by system

```{r}
OUTDIR = "../results/CellFie_Hu2025/"
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")
result3 = readRDS(paste0(OUTDIRV, "/result3.RDS"))
cfs = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
diffEffectBoxplot_bySystem(rlst_list = result3$rlst_list, 
  cf=cfs%>%mutate(Depth1=gsub("METABOLISM","M.", Depth1)), 
  OUT_DIR=paste0(OUTDIR, "/Limma/"), 
  w=2, h=3.5, title="Hu2025")

```

#### Differential effect boxplot by cancer

```{r}
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")
result3 = readRDS(paste0(OUTDIRV, "/result3.RDS"))
diffEffectBoxplot_byCancer(
  rlst_list = result3$rlst_list, 
  OUT_DIR=paste0(OUTDIR, "/Limma/"), w=2.75, h=3.0, title="Hu2025")
```


# SessionInfo
```{r}
out = capture.output(
  sessionInfo()
)
writeLines(out, paste0(OUTDIR,"/sessionInfo.txt"))
out
```

```{r}

```
