---
title: "Regulators"
format: html
---

```{r}
miRNAmRNAinteractions = function(targets, outdir, suffix) {
    # BiocManager::install("multiMiR")
    library(multiMiR)
    multiMiR_results <- get_multimir(
        predicted.cutoff = NULL, # Turns off predicted targets (TargetScan, etc.)
        # database = "validated",  # Restricts data to validated tables (miRTarBase)
        target = targets # Add your target hubgenes here
    )

    library(igraph)
    library(ggraph)
    library(tidygraph)
    library(dplyr)

    # 1. Clean the edge list using your exact multiMiR column names
    network_edges <- multiMiR_results@data %>%
        dplyr::select(mature_mirna_id, target_symbol) %>%
        distinct() # Removes the duplicate validation entries for a cleaner plot

    # 2. Build explicit node categories (miRNA vs mRNA)
    nodes_mirna <- data.frame(name = unique(network_edges$mature_mirna_id), type = "miRNA")
    nodes_mrna  <- data.frame(name = unique(network_edges$target_symbol), type = "mRNA Target")
    network_nodes <- rbind(nodes_mirna, nodes_mrna)

    # 3. Assemble the tidygraph network object
    graph_obj <- tbl_graph(nodes = network_nodes, edges = network_edges, directed = FALSE)

    # 4. Generate the PDF plot
    # Just change pdf(...) to cairo_pdf(...)
    h = nrow(multiMiR_results@data)/37.5
    w = (nrow(multiMiR_results@data)/37.5) + 1
    cairo_pdf(paste0(outdir, "/Interaction_miRNA_mRNA_Network", suffix,".pdf"), 
        width = w, height = h)
    print(ggraph(graph_obj, layout = "stress") + 
        geom_edge_link(alpha = 0.25, color = "gray60", linewidth = 0.4) +
        geom_node_point(aes(color = type, shape = type), size = 5) +
        geom_node_text(aes(label = name), repel = TRUE, size = 2.8, fontface = "bold", max.overlaps = 20) +
        scale_color_manual(values = c("miRNA" = "#2E9FDF", "mRNA Target" = "#E7B800")) +
        scale_shape_manual(values = c("miRNA" = 16, "mRNA Target" = 17)) +
        theme_graph() +
        labs(title = "miRNA-mRNA Regulatory Network Map",
            subtitle = "Experimentally Validated miRTarBase Targets"))
    dev.off()

    return(multiMiR_results@data)
}

```

```{r}
outdir = "../results/Regulators/"
dir.create(outdir, recursive=TRUE, showWarnings=FALSE)
```

# miRNA-mRNA interactions: multiMiR

## Ntworks for visualizing validated interacting miRNA for hubgenes

```{r}
outdir_mi = paste0(outdir, "/miRNA_mRNA/")
dir.create(outdir_mi, recursive=T, showWarnings=F)
```

```{r}
targets = c("RPN2", "ALG3", "DDOST") 
hubs1 = miRNAmRNAinteractions(targets, outdir_mi, suffix="_GlycanHubGenes")
targets = c("DTYMK", "RRM2",  "NME1" ) 
hubs2 = miRNAmRNAinteractions(targets, outdir_mi, suffix="_NucleotideHubGenes")

rbind(hubs1, hubs2)  %>% openxlsx::write.xlsx(.,
    paste0(outdir_mi, "miRNA_mRNA_interactions_GlycanHubGens_NucleotideHubGenes.xlsx")
)
```

## Find all valdiated interacting miRNA each of glycan/nucleotide pathways

```{r}
databaseDIR="../results/TCGA/KEGG_metabolism/"
metab_db_table = 
    readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS"))

unique(metab_db_table$class)
list_lists = lapply(c("Glycan biosynthesis and metabolism", "Nucleotide metabolism"), function(current_class){
    
    out_dir = paste0(outdir_mi, "/", gsub(" ", ".", current_class), "/")
    dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
    
    class_p = filter(metab_db_table, class == current_class)
    pathways = pull(class_p, gs_name) %>% unique()
    
    lapply(pathways, function(pathway) {
        print(pathway)
        if (!file.exists(paste0(out_dir, "/", gsub(" ", ".", pathway), ".xlsx"))) {
            all_targets = filter(class_p, gs_name == pathway) %>% pull(gene_symbol) %>% unique()
            if (length(all_targets) == 0) return(NULL)
            
            # Divide targets into batches of 10
            batch_size <- 10
            batches <- split(all_targets, ceiling(seq_along(all_targets) / batch_size))
            
            batch_results <- lapply(seq_along(batches), function(i) {
                batch_genes <- batches[[i]]
                
                # --- INFINTIE RETRY LOOP START ---
                repeat {
                    cat(paste0("Processing ", pathway, ": Batch ", i, "/", length(batches), "...\n"))
                    
                    # Attempt to run the multiMiR function
                    res <- tryCatch({
                        miRNAmRNAinteractions(
                            batch_genes, 
                            outdir = out_dir, 
                            suffix = paste0("_", gsub(" ", ".", gsub(" ", ".", pathway)), "_batch", i)
                        )
                    }, error = function(e) {
                        # Capture server drop/timeout string message
                        message(paste("❌ Server Error:", e$message))
                        return("SERVER_DOWN") 
                    })
                    
                    # If the function ran successfully (did not return our error flag), break the repeat loop
                    if (!identical(res, "SERVER_DOWN")) {
                        return(res)
                    }
                    
                    # If it failed, wait 60 seconds before retrying the exact same batch automatically
                    message("⏳ Server disconnected or sent an empty reply. Waiting 60 seconds to retry...")
                    Sys.sleep(60)
                }
                # --- INFINITE RETRY LOOP END ---
                
            })
            
            # Combine the successful batches together
            batch_results <- batch_results[!sapply(batch_results, is.null)]
            if (length(batch_results) > 0) {
                pathway_rslt <- do.call(rbind, batch_results)
                pathway_rslt$Class <- current_class
                openxlsx::write.xlsx(pathway_rslt, paste0(out_dir, "/", gsub(" ", ".", pathway), ".xlsx"))
                return(pathway_rslt)
            } else {
                return(NULL)
            }
        }
    }) %>% setNames(gsub(" ", ".", pathways))
}) %>% setNames(gsub(" ", ".", c("Glycan biosynthesis and metabolism", "Nucleotide metabolism")))

```


## Find enriched miRNA for glycan/nucleotide classes

```{r}
# Extract miRNAs interacting with at least 10 genes in each glycan/nucleotide classes
miRNAs = lapply(list_lists, function(l) {
    l ???
})

# Query all at once
multi_query <- get_multimir(
  mirna    = miRNAs
)

# Create a clean, unique lookup table of miRNA-to-mRNA pairs
mirna_target_lookup <- multi_query@data %>%
  dplyr::select(mature_mirna_id, target_symbol) %>%
  dplyr::distinct()

# Run ORA enrichment analysis of miRNA targets set within each of Glycan/Nucleotide classes

```



# lnRNA-mRNA interactions: EVLncRNAs3_human_data

```{r}
LnRNAmRNAnetwork = function(lnRNAmRNAinteractions, Targets, outdir, suffix) {
    # 1. Clean the edge list, keeping the Regulation column for edge aesthetics
    network_edges <- lnRNAmRNAinteractions %>%
        filter(Target %in% Targets) %>%
        dplyr::select(LncRNA, Target, Regulation) %>%
        distinct() # Removes absolute duplicate rows for a cleaner plot
    
    if (nrow(network_edges)>0) {
        openxlsx::write.xlsx(network_edges, paste0(outdir, "lnRNAmRNAInteractions", suffix, ".xlsx"))

        # 2. Build explicit node categories (lncRNA vs mRNA)
        nodes_lnc  <- data.frame(name = unique(network_edges$LncRNA), type = "lncRNA")
        nodes_mrna <- data.frame(name = unique(network_edges$Target), type = "mRNA Target")
        network_nodes <- rbind(nodes_lnc, nodes_mrna)

        # 3. Assemble the tidygraph network object
        graph_obj <- tbl_graph(nodes = network_nodes, edges = network_edges, directed = FALSE)

        # 4. Generate the PDF plot with your dynamic sizing logic
        h = nrow(lnRNAmRNAinteractions) / 37.5
        w = (nrow(lnRNAmRNAinteractions) / 37.5) + 1

        # Safety threshold: Ensure the canvas remains readable for small datasets
        if (h < 7) h <- 7
        if (w < 8) w <- 8

        cairo_pdf(paste0(outdir, "/ncRNA_mRNA_inteaction_Network", suffix, ".pdf"), 
                width = w, height = h)

        print(
            ggraph(graph_obj, layout = "stress") + 
                # Color edges dynamically based on target regulation pattern
                geom_edge_link(aes(color = Regulation), alpha = 0.5, linewidth = 0.5) +
                
                # Color and shape nodes based on molecule type
                geom_node_point(aes(color = type, shape = type), size = 5) +
                geom_node_text(aes(label = name), repel = TRUE, size = 2.8, fontface = "bold", max.overlaps = 20) +
                
                # Custom color styling for nodes and edge paths
                scale_color_manual(values = c("lncRNA" = "#8E44AD", "mRNA Target" = "#E7B800")) + 
                scale_shape_manual(values = c("lncRNA" = 16, "mRNA Target" = 17)) +             
                
                # Color mapping for regulation trends (Adjust strings to match your table exactly)
                scale_edge_color_manual(values = c(
                    "up-regulated"   = "#E74C3C", # Soft Crimson
                    "down-regulated" = "#3498DB", # Soft Cerulean
                    "unknown"        = "gray60",  # Neutral gray
                    "N/A"            = "gray60"
                ), na.value = "gray60") +
                
                theme_graph() +
                theme(legend.position = "right") +
                labs(
                    title = "lncRNA-mRNA Regulatory Network Map",
                    subtitle = "Experimentally Validated EVLncRNAs3 Targets",
                    color = "Molecule Class",
                    shape = "Molecule Class",
                    edge_color = "Target Regulation"
                )
        )

        dev.off()
    } else {network_edges = NULL}

    return(network_edges)
}

```


```{r}
outdir_ln = paste0(outdir, "/LncRNA_mRNA/")
dir.create(outdir_ln, recursive=T, showWarnings=F)
library(readxl)
EVLncRNAs3 = 
    read_excel(paste0(outdir_ln, "/EVLncRNAs3_human_data/function_human.xls"), sheet = 1)

lnRNAmRNAinteractions = 
    EVLncRNAs3[, c("LncRNA name", "Interaction target", 
        "Expression pattern of Interaction target", "Level of interaction")] %>%
    setNames(c("LncRNA", "Target", "Regulation", "InteractionType"))  %>%
    filter(InteractionType == "lncRNA-mRNA")

# No lnRNA interaction found for:
targets = c("RPN2", "ALG3", "DDOST") 
targets = c("DTYMK", "RRM2",  "NME1" ) 

```


## Get all the lnRNAs interacting with pathway memembers for the glycan/nucleotide pathways

```{r}
ln_lists = lapply(c("Glycan biosynthesis and metabolism", "Nucleotide metabolism"), function(current_class){
    
    out_dir = paste0(outdir_ln, "/", gsub(" ", ".", current_class), "/")
    dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
    
    class_p = filter(metab_db_table, class == current_class)
    pathways = pull(class_p, gs_name) %>% unique()
    genes = pull(class_p, gene_symbol)
    sum(lnRNAmRNAinteractions$Target %in% genes)
    # 1
})

```

! There are not much shared genes!


```{r}
out = capture.output(sessionInfo())
writeLines(out, paste0(outdir, "/sessionInfo.txt"))

```
