---
title: "TCGA"
author: Yuan Li (yuan.li@nbis.se)
format: html
---



```{r}
# BiocManager::install("TCGAbiolinks")
library(TCGAbiolinks)
library(SummarizedExperiment)
library(dplyr)
library(dplyr)
library(ggplot2)
library(parseMetab)
# remove.packages("parseMetab")
# detach("package:KinaseDownstream", unload = TRUE)
# remotes::install_github("Maj18/parseMetab", dependencies = TRUE)
# library(parseMetab)
```

# Download SMTdb cancer sample data

```{bash}
#!/bin/bash

(
  echo -e "cancer\ttissue\tpaper\tpaper_html\tdatasetID\tspot_num\tcell_num\tbac_num\tslice_type\tsample"
  jq -r '
    .data[] | [
      .cancer,
      .tissue,
      .paper,
      .paper_html,
      .datasetID,
      .spot_num,
      .cell_num,
      .bac_num,
      .slice_type,
      .sample
    ] | @tsv
  ' ../data/SMTdb/download_file.json
) > ../data/SMTdb/output.tsv

# First get the header to know column numbers
header=$(head -1 ../data/SMTdb/output.tsv)

# Example columns based on your JSON (adjust if different):
# cancer = column 1
# datasetID = column 5
# slice_type = column 9

# List of files to download
# "sc_celltype_file.txt", "sc_gene_exp_count.txt"
files=("st_cluster.txt" "st_neb.txt" "st_dev_rctd.txt" "st_gene_exp_count.txt")

tail -n +2 ../data/SMTdb/output.tsv | awk -F '\t' '{ print $1 "\t" $5 }' | while IFS=$'\t' read -r cancer dataset_id; do
  mkdir -p "../data/SMTdb/$cancer"/${dataset_id}

  for file in "../data/SMTdb/${files[@]}"; do
    output_file="../data/SMTdb/${cancer}/${dataset_id}/${file}"
    if [ -f "$output_file" ]; then
      echo "File $output_file already exists, skipping download."
    else
      echo "Downloading $file for $dataset_id into folder $cancer"
      curl -L -o "$output_file" "http://bio-bigdata.hrbmu.edu.cn/SMTdb/downloadCopy.jsp?filename=${file}&loc=${dataset_id}"
    fi
  done

done

```

# Process the data

```{r}
INDIR = "../data/SMTdb/"
meta = read.table(
  paste0(INDIR, "/output.tsv"),
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE,
  fill = TRUE,
  quote = "",
  comment.char = ""
)

# We will only keep cancers that have >5 samples:
print("We will only keep cancer samples that have >5 cancer samples...")
cancers2keep = meta %>% filter(sample_type=="cancer") %>%
    group_by(cancer) %>% 
    summarise(count = n()) %>%
    filter(count > 5) %>% pull(cancer)
meta = meta %>% 
    filter(cancer %in% cancers2keep) %>% filter(sample_type=="cancer")
OUTDIR = "../results/SMTdb/"
OUTDIR %>% dir.create(recursive=T, showWarnings=FALSE)
write.table(meta, paste0(OUTDIR, "/meta.csv"),
                    quote=F, sep="\t", row.names=F)
saveRDS(meta, paste0(OUTDIR, "/meta.RDS"))
# meta = readRDS(paste0(OUTDIR, "/meta.RDS"))
head(meta)
dim(meta)
```

## Import data

```{r}
#' Import SMTdb files
#' 
#' Import SMTdb files for all cancer types and all samples.
#' 
#' @param INDIR Directory where the SMTdb files are placed, it shoud be organized as <Cancer_type>/<Sample_ID>/*.txt.
#' @param pattern What files to import, e.g. "st_neb.txt".
#' 
#' @return A list of list of tables:
#'  The first level of lists contains all cancer types.
#'  The second level of lists contains all samples within each cancer type.
#'
#' @examples 
#' # importFiles_SMTdb(INDIR, pattern="st_gene_exp_count.txt")
#' 
importFiles_SMTdb = function(INDIR, pattern="st_gene_exp_count.txt", meta) {
    cancers = meta$cancer %>% unique()
    # cancers = list.dirs(INDIR, full.names=FALSE, recursive=FALSE)
    lapply(cancers, function(c) {
        cancer_dir = paste0(INDIR, "/", c, "/")
        # file_paths = 
        #     list.files(cancer_dir, 
        #     pattern=pattern, full.names=TRUE, recursive=TRUE)
        # file_samples = list.dirs(cancer_dir, full.names=F, recursive=FALSE)
        if (pattern=="st_neb.txt") {
            file_samples = 
                meta %>% 
                filter(cancer==c&sample_type=="cancer") %>%
                pull(datasetID) %>% unique()
        } else {
            file_samples = 
                meta %>% filter(cancer==c) %>% pull(datasetID) %>% unique()
        }
        file_paths = paste0(INDIR, "/", c, "/", file_samples, "/", pattern)

        lapply(file_paths, function(file) {
            tryCatch(read.table(file, header = TRUE, 
                sep = "\t", stringsAsFactors = FALSE, row.names = NULL) %>%
            as.data.frame(), error=function(e) NA)
        }) %>% setNames(file_samples)
    }) %>% setNames(cancers)
}

INDIR = "../data/SMTdb/"
SpatialNeighborhood = importFiles_SMTdb(INDIR, pattern="st_neb.txt", meta=meta)
STexpr_all = importFiles_SMTdb(INDIR, pattern="st_gene_exp_count.txt", meta=meta)

# Save the objects
OUTDIR = "../results/SMTdb/"
OUTDIR %>% dir.create(., recursive=TRUE, showWarnings=FALSE)
saveRDS(SpatialNeighborhood, paste0(OUTDIR, "/SpatialNeighborhood.RDS"))
# SpatialNeighborhood = readRDS(paste0(OUTDIR, "/SpatialNeighborhood.RDS"))
saveRDS(STexpr_all, paste0(OUTDIR, "/STexpr_all.RDS"))
# STexpr_all = readRDS(paste0(OUTDIR, "/STexpr_all.RDS"))
lapply(STexpr_all, function(STexpr) names(STexpr))
lapply(SpatialNeighborhood, function(STexpr) names(STexpr))
lapply(SpatialNeighborhood, function(x) length(x))
lapply(STexpr_all, function(x) length(x))

# Get ST data for only cancer
STexpr_cancer = lapply(names(STexpr_all), function(c) {
    cancer_samples = names(SpatialNeighborhood[[c]])
    STexpr_all[[c]][cancer_samples]
}) %>% setNames(names(STexpr_all))
saveRDS(STexpr_cancer, paste0(OUTDIR, "/STexpr_cancer.RDS"))
lapply(STexpr_cancer, function(STexpr) names(STexpr))

rm(STexpr_all)
gc()

```

## Make pseudo-bulk per sample

### Malignant spots on cacner samples

```{r}
#' Make pseudobulk data from spatial transcriptomics data.
#' 
#' Aggregate the gene count across all Malignant/Non-malignant spots for each cancer sample.
#' 
#' @param SpatialNeighborhood The spatial neighborhood data got with importFiles_SMTdb().
#' @param STexpr The spatial transcriptomcis expression data got with importFiles_SMTdb().
#' 
#' @return A list of pseudobulk RNA counts, with each component represent a cancer type.
#' 
#' @examples 
#' # getPseudoBulk(SpatialNeighborhood, STexpr)
#' 
getPseudoBulk_malignant = function(SpatialNeighborhood, STexpr) {
    lapply(names(SpatialNeighborhood), function(c) {
        print(c)
        sp_nei = SpatialNeighborhood[[c]]
        st_expr = STexpr[[c]]
        lapply(names(sp_nei), function(sample) {
            sp_nei_sample = sp_nei[[sample]] %>% na.omit() 
            st_expr_sample = st_expr[[sample]] %>% na.omit() %>%
                tibble::column_to_rownames(var="row.names")
            st_expr_sample = st_expr_sample[rowSums(st_expr_sample, na.rm=T)>1,]
            st_expr_sample = st_expr_sample[, colSums(st_expr_sample, na.rm=T)>0]
            Malignant_spots = sp_nei_sample %>% 
                filter(Location%in%c("Malignant", "Mal")) %>% pull(row.names) %>%
                gsub("-", ".", .) %>%
                intersect(., colnames(st_expr_sample))
            Non_malignant_spots = sp_nei_sample %>% 
                filter(Location%in%c("Non-malignant", "nMal")) %>% pull(row.names) %>%
                gsub("-", ".", .) %>%
                intersect(., colnames(st_expr_sample))
                # we only keep cancers with at least 3 malignant samples
            if (length(Malignant_spots)>=3) {
                Malignant =  rowSums(st_expr_sample[, Malignant_spots], na.rm=T)} else {
                    Malignant = NA
                }
                # we only keep cancers with at least 3 stromal samples
            if (length(Non_malignant_spots)>=3) {
                Non_malignant =  rowSums(st_expr_sample[, Non_malignant_spots], na.rm=T)} else {
                    Non_malignant = NA
                }
            data.frame(Feature=rownames(st_expr_sample),
                Malignant =  Malignant,
                Non_malignant = Non_malignant
            ) %>% setNames(c("Feature", paste0(c("Malignant", "nonMalignant"), "_", sample)))
        }) %>% Reduce(full_join, .)
    }) %>% setNames(names(SpatialNeighborhood))
}

pseudoBulk_cancer = getPseudoBulk_malignant(SpatialNeighborhood, STexpr_cancer)
pseudoBulk_cancer = pseudoBulk_cancer[!is.na(pseudoBulk_cancer)]
# Remove empty samples:
pseudoBulk_cancer = lapply(pseudoBulk_cancer, function(p) {
    samples2keep = 
        names(which(colSums(as.data.frame(p) %>% 
        tibble::column_to_rownames(var="Feature"), na.rm=T)>0))
    p[,c("Feature", samples2keep)]
})
saveRDS(pseudoBulk_cancer, paste0(OUTDIR, "/pseudoBulk_cancer.RDS"))
# pseudoBulk = readRDS(paste0(OUTDIR, "/pseudoBulk_cancer.RDS"))
openxlsx::write.xlsx(pseudoBulk_cancer, 
    file=paste0(OUTDIR, "/PseudoCount_filtered_cancer.xlsx"))

head(pseudoBulk_cancer[[1]])

```


# Get GSVA metabolic activity scores

## Get KEGG metabolic pathways

```{r}
# Get KEGG metabolic pathways
paste0(OUTDIR, "/KEGG_metabolism/") %>% 
    dir.create(., recursive=TRUE, showWarnings=FALSE)

kegg_metab_dbs = 
    getdb_metabolism(databaseDIR=paste0(OUTDIR, "/KEGG_metabolism/"))

kegg_metab_db = kegg_metab_dbs$kegg_metab_db

# databaseDIR=paste0(OUTDIR, "/KEGG_metabolism/")
# kegg_metab_dbs = 
# list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
#     kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))

# Make a circular packing plot to show the number of pathways in each class
getCirPackingPlot_GSVA(
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table,
    outdir = OUTDIR)

```

# Get sample GSVA scores

```{r}
OUTDIR = "../results/SMTdb/"
getGSVAscores = function(
    pseudoBulk_cancer, Feature.filtering=T, 
    min.count = 5, samplProp2rm = 0.1, kegg_metab_db, OUTDIR) {
    lapply(names(pseudoBulk_cancer), function(c) {
        dat = pseudoBulk_cancer[[c]] 
        rownames(dat) = NULL
        dat = dat %>% as.data.frame() %>%
            tibble::column_to_rownames(var="Feature") #%>% na.omit()
        meta0 = data.frame(Slice = colnames(dat) %>% 
                    gsub("nonMalignant_", "", .) %>%
                    gsub("Malignant_", "", .),
                Group = gsub("_.*", "", colnames(dat)),
                Sample = colnames(dat))
        # We need to make sure there are more than 2 samples in each condition:
        if (sum(table(meta0$Group)>2)==2) {
            meta0$Group = factor(meta0$Group, levels=c("nonMalignant", "Malignant")) # Reference first!
            paste0(OUTDIR, "/GSVAmetabolicScores/", c, "/") %>%
                dir.create(., recursive=TRUE, showWarnings=FALSE)
                meta0 = filter(meta0, Sample%in%colnames(dat))

            # Make sure the sample are in the same order in both meta and dat:
            dat = dat[, meta0$Sample]
            meta0 = meta0 %>% mutate(across(where(is.factor), droplevels))
            mapping = setNames(meta0$Group, meta0$Sample)
            print("Only keep features that have at least min.nr number of values that are >5 in each group...")
            min.nr = round(max((ncol(dat)/2)*samplProp2rm, 2), 0)
            if (Feature.filtering) {
                B = mapping[colnames(dat)] %>% as.character()
                featureskeep = apply(dat>min.count, 1, function(row) {
                    A = row
                    sum(aggregate(A~B, data=data.frame(A=A,B=B), sum, na.rm=T)[,"A"]>=min.nr)>1
                })
                dat = dat[featureskeep, ]
            }
            temp = replace(dat, is.na(dat), 0)
            dge = edgeR::DGEList(counts=temp)
            dge = edgeR::calcNormFactors(dge, method = "TMMwsp")
            effect.lib.sizes = (dge$samples$lib.size) * (dge$samples$norm.factors)
            normalizedCount = (t(t(dat)/effect.lib.sizes))*median(effect.lib.sizes)
            write.table(
                normalizedCount %>% as.data.frame() %>%
                        tibble::rownames_to_column(var="Features"),
                paste0(OUTDIR, "/GSVAmetabolicScores/", c, "/NormalizedCount.csv"),
                quote = F, row.names = F, col.names = T, sep="\t")

            # Calcualte GSVA per sample
            # BiocManager::install("GSVA")
            normalizedCount = (t(t(temp)/effect.lib.sizes))*median(effect.lib.sizes)
            normalizedCount = as.matrix(normalizedCount)
            normalizedCount = matrix(as.numeric(normalizedCount),
                nrow = nrow(normalizedCount),
                ncol = ncol(normalizedCount), dimnames = dimnames(normalizedCount))

            param = GSVA::gsvaParam(normalizedCount, kegg_metab_db, minSize=5, maxSize=500)
            gsva_scores = 
                tryCatch(GSVA::gsva(param, verbose=TRUE) %>% 
                as.data.frame(), error = function(e) NA)
            write.table(
                gsva_scores %>% as.data.frame() %>%
                        tibble::rownames_to_column(var="Features"),
                paste0(OUTDIR, "/GSVAmetabolicScores/", c, "/GSVAmetabolicScores.csv"),
                quote = F, row.names = F, col.names = T, sep="\t")
            return(as.data.frame(gsva_scores))
        } else {return(NA)}
    }) %>% setNames(names(pseudoBulk_cancer)) %>% .[!is.na(.)]
} 

GSVAscores = getGSVAscores(
    pseudoBulk_cancer, Feature.filtering=T, 
    min.count = 5, samplProp2rm = 0.1, 
    kegg_metab_db=kegg_metab_dbs$kegg_metab_db, 
    OUTDIR)

saveRDS(GSVAscores, 
    paste0(OUTDIR, "/GSVAmetabolicScores/GSVAscores.RDS"))
# OUTDIR = "../results/SMTdb/"
# GSVAscores = readRDS(paste0(OUTDIR, "/GSVAmetabolicScores/GSVAscores.RDS"))
# names(GSVAscores)
colnames(GSVAscores[[1]])
Sample_info = lapply(names(GSVAscores), function(cancer_type) {
    GSVAdat = GSVAscores[[cancer_type]]
    dat = as.data.frame(list(Tumor = grep("^Malignant_",colnames(GSVAdat)) %>% length(),
                        Normal = grep("^nonMalignant_",colnames(GSVAdat)) %>% length()))
    dat$Cancer_type = cancer_type
    dat$Data.type = "SpatialTranscriptomcis"
    dat$Source = "SMTdb"
    dat
}) %>% Reduce(rbind, .)
Sample_info$Description = 
    c("Colorectal cancer", "Brain glioblastoma", "Oral squamous cell carcinoma", "Ovarian carcinoma")
Sample_info = Sample_info %>%
    select(Cancer_type, Description, Data.type, Normal, Tumor, Source)
openxlsx::write.xlsx(Sample_info, paste0(OUTDIR, "Sample_info.xlsx"))

```

## Visualize sample GSVA scores_heatmap

```{r}
w = 9.5
height = 15.5
lapply(names(GSVAscores), function(c) {
            sub = GSVAscores[[c]]
            colAnn = data.frame(Group = gsub("_.*", "", colnames(sub)))
            # colnames(sub) = gsub("nonMalignant|Malignant", "", colnames(sub))
            rownames(colAnn) = colnames(sub)
            colAnn = arrange(colAnn, Group)
            sub = sub[rownames(colAnn)]
            L = ncol(sub)
            rowAnn = kegg_metab_dbs$kegg_metab_db_table %>% 
                dplyr::select(gs_name, class) %>% unique() %>% 
                as.data.frame() %>% mutate(gs_name2 = gs_name) %>%
                tibble::column_to_rownames(var="gs_name2")
            rowAnn = rowAnn[rownames(sub),]
            rowAnn = arrange(rowAnn, class)
            sub = sub[rownames(rowAnn), ]
            rowAnn = rowAnn[,"class", drop=FALSE]
            pheatmap::pheatmap(sub,
                            show_colnames=TRUE,
                            color=viridisLite::magma(50),
                            breaks=seq(-2,2,len=50),
                            border_color="black",
                            scale="row",
                            annotation_col=colAnn,
                            annotation_row=rowAnn,
                            main="Task Scores",
                            cluster_cols = FALSE,
                            cluster_rows = FALSE,
                            filename=paste0(OUTDIR, "/GSVAmetabolicScores/", 
                                c, "/GSVAmetabolicScoreHeatmap_wSampleNames.pdf"), 
                            width=w+0.14*L, height=height)
            pheatmap::pheatmap(sub,
                            labels_col=rep("", ncol(sub)),
                            color=viridisLite::magma(50),
                            breaks=seq(-2,2,len=50),
                            border_color="black",
                            scale="row",
                            annotation_col=colAnn,
                            annotation_row=rowAnn,
                            main="Task Scores",
                            cluster_cols = FALSE,
                            cluster_rows = FALSE,
                            filename=paste0(OUTDIR, "/GSVAmetabolicScores/", 
                                c, "/GSVAmetabolicScoreHeatmap.pdf"), 
                            width=w+0.14*L, height=height)
})

```

## Summarize metabolic pathway size

```{r}
visualizeGSsizeClass(
    GSVA_limma_rslt_gsva = GSVAscores, 
    genes4GSVA = lapply(names(GSVAscores), function(c) {
        read.table(
            paste0(OUTDIR, "/GSVAmetabolicScores/", c, "/NormalizedCount.csv"),
            header=TRUE) %>% pull(Features)
        }) %>% unlist() %>% as.character() %>% unique(),
    kegg_metab_db = kegg_metab_dbs$kegg_metab_db, 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table,
    OUTDIR=paste0(OUTDIR, "/GSVAmetabolicScores/"))

getCirPackingPlot_GSVA(
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table %>%
        filter(gs_name %in% 
            (sapply(GSVAscores, function(g) 
                rownames(g)) %>% unlist() %>% as.character() %>% unique())), 
    outdir = OUTDIR)

```

## Visualize sample GSVA score Malignant/nonMalignant ratios

```{r}
w = 9
height = 14.5
lapply(names(GSVAscores), function(c) {
    sub = GSVAscores[[c]]
    paired_samples = which((gsub("nonMalignant_|Malignant_", "", colnames(sub)) %>% 
        table())==2) %>% names()
    MalvsnonMasRatio = 
        sub[,paste0("Malignant_", paired_samples)]-sub[,paste0("nonMalignant_", paired_samples)]
    colnames(MalvsnonMasRatio) = gsub("Malignant_", "", colnames(MalvsnonMasRatio))
    L = ncol(MalvsnonMasRatio)
    rowAnn = kegg_metab_dbs$kegg_metab_db_table %>% 
        dplyr::select(gs_name, class) %>% unique() %>% 
        as.data.frame() %>% mutate(gs_name2 = gs_name) %>%
        tibble::column_to_rownames(var="gs_name2")
    rowAnn = rowAnn[rownames(MalvsnonMasRatio),]
    rowAnn = arrange(rowAnn, class)
    MalvsnonMasRatio = MalvsnonMasRatio[rownames(rowAnn), ]
    rowAnn = rowAnn[,"class", drop=FALSE]
    pheatmap::pheatmap(MalvsnonMasRatio,
                    show_colnames=TRUE,
                    color=viridisLite::magma(50),
                    breaks=seq(-2,2,len=50),
                    border_color="black",
                    scale="row",
                    # annotation_col=colAnn,
                    annotation_row=rowAnn,
                    main="Task Scores",
                    cluster_cols = FALSE,
                    cluster_rows = FALSE,
                    filename=paste0(OUTDIR, "/GSVAmetabolicScores/", c, 
                        "/GSVAscore_MalvsnonMalRatio_Heatmap_wSampleNames.pdf"), 
                    width=w+0.14*L, height=height)
    pheatmap::pheatmap(MalvsnonMasRatio,
                    labels_col=rep("", ncol(MalvsnonMasRatio)),
                    color=viridisLite::magma(50),
                    breaks=seq(-2,2,len=50),
                    border_color="black",
                    scale="row",
                    # annotation_col=colAnn,
                    annotation_row=rowAnn,
                    main="Task Scores",
                    cluster_cols = FALSE,
                    cluster_rows = FALSE,
                    filename=paste0(OUTDIR, "/GSVAmetabolicScores/", 
                        c, "/GSVAscore_MalvsnonMalRatio_Heatmap.pdf"), 
                    width=w+0.14*L, height=height)
})

```

## Summarize sample GSVA scores

```{r}
visualizeGSVSscoresGroup(
    GSVA_limma_rslt_gsva = GSVAscores,
    paired_data = lapply(names(GSVAscores), function(c) {
            g = GSVAscores[[c]]
            meta0 = data.frame(Sample = colnames(g),
                Group = gsub("_.*", "", colnames(g)),
                Sample = gsub("nonMalignant_|Malignant_", "", colnames(g)))
            meta0
        })  %>% setNames(names(GSVAscores)),
    OUTDIR = paste0(OUTDIR, "/GSVAmetabolicScores/"),
    Condition_column = "Group",
    Condition.control = "nonMalignant",
    sample_column = "Sample",
    w=3, h=3.5, title="SMTdb"
)

visualizeGSVSscoresClass(
    GSVA_limma_rslt = 
        lapply(GSVAscores, function(g) list(gsva=g)), 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table, 
    paired_data=lapply(names(GSVAscores), function(c) {
        g = GSVAscores[[c]]
        meta0 = data.frame(Sample = colnames(g),
            Group = gsub("_.*", "", colnames(g)),
            Sample = gsub("nonMalignant_|Malignant_", "", colnames(g)))
        meta0
    })  %>% setNames(names(GSVAscores)), 
    OUTDIR = paste0(OUTDIR, "/GSVAmetabolicScores/"),
    Condition_column = "Group",
    Condition.control = "nonMalignant",
    sample_column = "Sample",
    w=5, h=3.5)

```



# Paired differential analysis on sample GSVA scores

```{r}
OUTDIR = "../results/SMTdb/"
GSVAlimmaTest_inner = function(
    GSVAscores, currentCovariate=NULL, paired=T, OUTDIR) {
    lapply(names(GSVAscores), function(c) {
        g = GSVAscores[[c]]
        meta0 = data.frame(Sample = colnames(g),
                        Group = gsub("_.*", "", colnames(g)),
                        Sample = gsub("nonMalignant_|Malignant_", "", colnames(g)))
        paired_variable = "Sample"
        baseFormula = ~ Group
        if (is.null(currentCovariate)) {
            currentFormula = baseFormula
        } else {
            currentFormula = as.formula(paste0("~ ", currentCovariate, " + Group"))}

        design = model.matrix(currentFormula, data=meta0) #
        colnames(design) = gsub("Group", "", colnames(design))
        if (paired) { #
            #Fit with correlated arrays
            dupcor = limma::duplicateCorrelation(g, design, #
                block=meta0[, paired_variable, drop=T]) #
            fit = limma::lmFit(g, design, block=meta0[,paired_variable,drop=T], #
                correlation=dupcor$consensus)
        } else { fit = limma::lmFit(g, design) }

        # # Limma trend to refine gene-level variance estimates
        # gssizes = sapply(kegg_metab_db[rownames(fit$coefficients)], function(gs) {
        #     length(gs)
        # })
        # fit.con = limma::eBayes(fit, robust = TRUE, trend=gssizes)
        fit.con = limma::eBayes(fit, robust = TRUE, trend=TRUE)
        rlst_interest =
            limma::topTable(fit.con, n=Inf, coef=levels(meta0$Group)[2]) %>%
            arrange(-t)
        paste0(OUTDIR, "/Limma_GSVA/", c, "/") %>% dir.create(., recursive=T, showWarnings=F)
        write.table(
            rlst_interest %>% as.data.frame() %>%
                    tibble::rownames_to_column(var="Features"),
            paste0(OUTDIR, "/Limma_GSVA/", c, "/limmaResult.csv"), #
            quote = F, row.names = F, col.names = T, sep="\t")
        return(rlst_interest)
    }) %>% setNames(names(GSVAscores))
}

GSVA_limma_rslt = 
    GSVAlimmaTest_inner(GSVAscores, currentCovariate=NULL, paired=T, OUTDIR)

saveRDS(GSVA_limma_rslt, 
    paste0(OUTDIR, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))

# GSVA_limma_rslt = readRDS(paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
```

## Visualize differential results

```{r}
makeLimmaDotplot_TCGA(rlst_list=GSVA_limma_rslt, sig_statistic="P.Value", 
        sig.cutoff=1, w=8, h=16, 
        OUTDIRV=paste0(OUTDIR, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title="Dys.Path_SMTdb", 
        suffix="SMTdb", nested=T,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list=GSVA_limma_rslt, sig_statistic="P.Value", 
        sig.cutoff=0.5, w=8, h=15, 
        OUTDIRV=paste0(OUTDIR, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title="Dys.Path_SMTdb", 
        suffix="SMTdb", nested=T,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list=GSVA_limma_rslt, sig_statistic="P.Value", 
        sig.cutoff=0.1, w=8, h=8, 
        OUTDIRV=paste0(OUTDIR, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title="Dys.Path_SMTdb", 
        suffix="SMTdb", nested=T,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list=GSVA_limma_rslt, sig_statistic="P.Value", 
        sig.cutoff=0.05, w=8, h=6.5, 
        OUTDIRV=paste0(OUTDIR, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title="Dys.Path_SMTdb", 
        suffix="SMTdb", nested=T,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())
```

## Visualize the number of significant pathways

```{r}
visualizeSigNr(OUTDIR=paste0(OUTDIR, "/Limma_GSVA"), 
    GSVA_limma_rslt =
        lapply(GSVA_limma_rslt, function(g) list(rslt_of_interest=g)),
    significance_statistic="P.Value", sig.cutoff=0.1,
    w=2, h=3.5)

```

###Visualize differential effect

### Differential effect boxplot by class

```{r}
OUTDIR = "../results/SMTdb/"
GSVA_limma_rslt = 
    readRDS(paste0(OUTDIR, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
databaseDIR=paste0(OUTDIR, "/KEGG_metabolism/")
kegg_metab_dbs = 
    list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
    kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))

diffEffectBoxplot_bySystem_GSVA(
    rlst_list = lapply(GSVA_limma_rslt, function(g) list(rslt_of_interest=g)), 
    taskHierarchy=
        kegg_metab_dbs$kegg_metab_db_table %>% select(gs_name, class) %>% unique(), 
    weight.effect.by.gene = TRUE,
    OUT_DIR=paste0(OUTDIR, "/Limma_GSVA"),
    w=2, h=3.5, title="SMTdb",
    winsorize=TRUE)

```

### Differential effect boxplot by cancer

```{r}
diffEffectBoxplot_byCancer_GSVA(
  rlst_list = lapply(GSVA_limma_rslt, function(g) list(rslt_of_interest=g)), 
  OUT_DIR=paste0(OUTDIR, "/Limma_GSVA"), w=1.5, h=3.0, title="SMTdb")
```




