---
title: "TCGA"
author: Yuan Li (yuan.li@nbis.se)
format: html
---



```{r}
# BiocManager::install("TCGAbiolinks")
library(TCGAbiolinks)
library(SummarizedExperiment)
library(dplyr)
library(dplyr)
library(ggplot2)
library(parseMetab)
# remove.packages("parseMetab")
# detach("package:KinaseDownstream", unload = TRUE)
# remotes::install_github("Maj18/parseMetab", dependencies = TRUE)
# library(parseMetab)

```

# Download SMTdb cancer sample data

```{bash}
# #!/bin/bash

# (
#   echo -e "cancer\ttissue\tpaper\tpaper_html\tdatasetID\tspot_num\tcell_num\tbac_num\tslice_type\tsample"
#   jq -r '
#     .data[] | [
#       .cancer,
#       .tissue,
#       .paper,
#       .paper_html,
#       .datasetID,
#       .spot_num,
#       .cell_num,
#       .bac_num,
#       .slice_type,
#       .sample
#     ] | @tsv
#   ' ../data/SMTdb/download_file.json
# ) > ../data/SMTdb/output.tsv

# # First get the header to know column numbers
# header=$(head -1 ../data/SMTdb/output.tsv)

# # Example columns based on your JSON (adjust if different):
# # cancer = column 1
# # datasetID = column 5
# # slice_type = column 9

# # List of files to download
# # "sc_celltype_file.txt", "sc_gene_exp_count.txt"
# files=("st_cluster.txt" "st_neb.txt" "st_dev_rctd.txt" "st_gene_exp_count.txt")

# tail -n +2 ../data/SMTdb/output.tsv | awk -F '\t' '{ print $1 "\t" $5 }' | while IFS=$'\t' read -r cancer dataset_id; do
#   mkdir -p "../data/SMTdb/$cancer"/${dataset_id}

#   for file in "../data/SMTdb/${files[@]}"; do
#     output_file="../data/SMTdb/${cancer}/${dataset_id}/${file}"
#     if [ -f "$output_file" ]; then
#       echo "File $output_file already exists, skipping download."
#     else
#       echo "Downloading $file for $dataset_id into folder $cancer"
#       curl -L -o "$output_file" "http://bio-bigdata.hrbmu.edu.cn/SMTdb/downloadCopy.jsp?filename=${file}&loc=${dataset_id}"
#     fi
#   done

# done

```

# Process the data

```{r}
INDIR = "../data/SMTdb/"
meta = read.table(
  paste0(INDIR, "/output.tsv"),
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE,
  fill = TRUE,
  quote = "",
  comment.char = ""
)

# We will only keep cancers that have >5 samples:
print("We will only keep cancer samples that have >5 cancer samples...")
cancers2keep = meta %>% filter(slice_type=="cancer") %>%
    group_by(cancer) %>% 
    summarise(count = n()) %>%
    filter(count > 5) %>% pull(cancer)
meta = meta %>% 
    filter(cancer %in% cancers2keep) %>% filter(slice_type=="cancer")
OUTDIR = "../results/SMTdb/"
OUTDIR %>% dir.create(recursive=T, showWarnings=FALSE)
write.table(meta, paste0(OUTDIR, "/meta.csv"),
                    quote=F, sep="\t", row.names=F)
saveRDS(meta, paste0(OUTDIR, "/meta.RDS"))
# meta = readRDS(paste0(OUTDIR, "/meta.RDS"))
head(meta)
dim(meta)
```

## Import data

```{r}
INDIR = "../data/SMTdb/"
SpatialNeighborhood = importFiles_SMTdb(INDIR, pattern="st_neb.txt", meta=meta)
STexpr_all = importFiles_SMTdb(INDIR, pattern="st_gene_exp_count.txt", meta=meta)

# Save the objects
OUTDIR = "../results/SMTdb/"
OUTDIR %>% dir.create(., recursive=TRUE, showWarnings=FALSE)
saveRDS(SpatialNeighborhood, paste0(OUTDIR, "/SpatialNeighborhood.RDS"))
# SpatialNeighborhood = readRDS(paste0(OUTDIR, "/SpatialNeighborhood.RDS"))
saveRDS(STexpr_all, paste0(OUTDIR, "/STexpr_all.RDS"))
# STexpr_all = readRDS(paste0(OUTDIR, "/STexpr_all.RDS"))
lapply(STexpr_all, function(STexpr) names(STexpr))
lapply(SpatialNeighborhood, function(STexpr) names(STexpr))
lapply(SpatialNeighborhood, function(x) length(x))
lapply(STexpr_all, function(x) length(x))

# Get ST data for only cancer
STexpr_cancer = lapply(names(STexpr_all), function(c) {
    cancer_samples = names(SpatialNeighborhood[[c]])
    STexpr_all[[c]][cancer_samples]
}) %>% setNames(names(STexpr_all))
saveRDS(STexpr_cancer, paste0(OUTDIR, "/STexpr_cancer.RDS"))
lapply(STexpr_cancer, function(STexpr) names(STexpr))

rm(STexpr_all)
gc()

```

## Make pseudo-bulk per sample

### Malignant spots on cacner samples

```{r}
pseudoBulk_cancer = getPseudoBulk_malignant(SpatialNeighborhood, STexpr_cancer)
pseudoBulk_cancer = pseudoBulk_cancer[!is.na(pseudoBulk_cancer)]
# Remove empty samples:
pseudoBulk_cancer = lapply(pseudoBulk_cancer, function(p) {
    samples2keep = 
        names(which(colSums(as.data.frame(p) %>% 
        tibble::column_to_rownames(var="Feature"), na.rm=T)>0))
    p[,c("Feature", samples2keep)]
})
saveRDS(pseudoBulk_cancer, paste0(OUTDIR, "/pseudoBulk_cancer.RDS"))
# pseudoBulk = readRDS(paste0(OUTDIR, "/pseudoBulk_cancer.RDS"))
openxlsx::write.xlsx(pseudoBulk_cancer, 
    file=paste0(OUTDIR, "/PseudoCount_filtered_cancer.xlsx"))

head(pseudoBulk_cancer[[1]])
names(pseudoBulk_cancer)
```


# Get GSVA metabolic activity scores

## Get KEGG metabolic pathways

```{r}
# Get KEGG metabolic pathways
paste0(OUTDIR, "/KEGG_metabolism/") %>% 
    dir.create(., recursive=TRUE, showWarnings=FALSE)

kegg_metab_dbs = 
    getdb_metabolism(databaseDIR=paste0(OUTDIR, "/KEGG_metabolism/"))

kegg_metab_db = kegg_metab_dbs$kegg_metab_db

# databaseDIR=paste0(OUTDIR, "/KEGG_metabolism/")
# kegg_metab_dbs = 
# list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
#     kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))

# Make a circular packing plot to show the number of pathways in each class
getCirPackingPlot_GSVA(
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table,
    outdir = OUTDIR)

```

# Get sample GSVA scores

```{r}
OUTDIR = "../results/SMTdb/"
GSVAscores = getGSVAscores(
    pseudoBulk_cancer, Feature.filtering=T, 
    min.count = 5, samplProp2rm = 0.1, 
    kegg_metab_db=kegg_metab_dbs$kegg_metab_db, 
    OUTDIR)

saveRDS(GSVAscores, 
    paste0(OUTDIR, "/GSVAmetabolicScores/GSVAscores.RDS"))
# OUTDIR = "../results/SMTdb/"
# GSVAscores = readRDS(paste0(OUTDIR, "/GSVAmetabolicScores/GSVAscores.RDS"))
# names(GSVAscores)
colnames(GSVAscores[[1]])
Sample_info = lapply(names(GSVAscores), function(cancer_type) {
    GSVAdat = GSVAscores[[cancer_type]]
    dat = as.data.frame(list(Tumor = grep("^Malignant_",colnames(GSVAdat)) %>% length(),
                        Normal = grep("^nonMalignant_",colnames(GSVAdat)) %>% length()))
    dat$Cancer_type = cancer_type
    dat$Data.type = "SpatialTranscriptomcis"
    dat$Source = "SMTdb"
    dat
}) %>% Reduce(rbind, .)
Sample_info$Description = 
    c("Colorectal cancer", "Brain glioblastoma", "Oral squamous cell carcinoma", "Ovarian carcinoma")
Sample_info = Sample_info %>%
    select(Cancer_type, Description, Data.type, Normal, Tumor, Source)
openxlsx::write.xlsx(Sample_info, paste0(OUTDIR, "Sample_info.xlsx"))

```

## Visualize sample GSVA scores_heatmap

```{r}
w = 9.5
height = 15.5
lapply(names(GSVAscores), function(c) {
            sub = GSVAscores[[c]]
            colAnn = data.frame(Group = gsub("_.*", "", colnames(sub)))
            # colnames(sub) = gsub("nonMalignant|Malignant", "", colnames(sub))
            rownames(colAnn) = colnames(sub)
            colAnn = arrange(colAnn, Group)
            sub = sub[rownames(colAnn)]
            L = ncol(sub)
            rowAnn = kegg_metab_dbs$kegg_metab_db_table %>% 
                dplyr::select(gs_name, class) %>% unique() %>% 
                as.data.frame() %>% mutate(gs_name2 = gs_name) %>%
                tibble::column_to_rownames(var="gs_name2")
            rowAnn = rowAnn[rownames(sub),]
            rowAnn = arrange(rowAnn, class)
            sub = sub[rownames(rowAnn), ]
            rowAnn = rowAnn[,"class", drop=FALSE]
            pheatmap::pheatmap(sub,
                            show_colnames=TRUE,
                            color=viridisLite::magma(50),
                            breaks=seq(-2,2,len=50),
                            border_color="black",
                            scale="row",
                            annotation_col=colAnn,
                            annotation_row=rowAnn,
                            main="Task Scores",
                            cluster_cols = FALSE,
                            cluster_rows = FALSE,
                            filename=paste0(OUTDIR, "/GSVAmetabolicScores/", 
                                c, "/GSVAmetabolicScoreHeatmap_wSampleNames.pdf"), 
                            width=w+0.14*L, height=height)
            pheatmap::pheatmap(sub,
                            labels_col=rep("", ncol(sub)),
                            color=viridisLite::magma(50),
                            breaks=seq(-2,2,len=50),
                            border_color="black",
                            scale="row",
                            annotation_col=colAnn,
                            annotation_row=rowAnn,
                            main="Task Scores",
                            cluster_cols = FALSE,
                            cluster_rows = FALSE,
                            filename=paste0(OUTDIR, "/GSVAmetabolicScores/", 
                                c, "/GSVAmetabolicScoreHeatmap.pdf"), 
                            width=w+0.14*L, height=height)
})

```

## Summarize metabolic pathway size

```{r}
visualizeGSsizeClass(
    GSVA_limma_rslt_gsva = GSVAscores, 
    genes4GSVA = lapply(names(GSVAscores), function(c) {
        read.table(
            paste0(OUTDIR, "/GSVAmetabolicScores/", c, "/NormalizedCount.csv"),
            header=TRUE) %>% pull(Features)
        }) %>% unlist() %>% as.character() %>% unique(),
    kegg_metab_db = kegg_metab_dbs$kegg_metab_db, 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table,
    OUTDIR=paste0(OUTDIR, "/GSVAmetabolicScores/"))

getCirPackingPlot_GSVA(
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table %>%
        filter(gs_name %in% 
            (sapply(GSVAscores, function(g) 
                rownames(g)) %>% unlist() %>% as.character() %>% unique())), 
    outdir = OUTDIR)

```

## Visualize sample GSVA score Malignant/nonMalignant ratios

```{r}
w = 9
height = 14.5
lapply(names(GSVAscores), function(c) {
    sub = GSVAscores[[c]]
    paired_samples = which((gsub("nonMalignant_|Malignant_", "", colnames(sub)) %>% 
        table())==2) %>% names()
    MalvsnonMasRatio = 
        sub[,paste0("Malignant_", paired_samples)]-sub[,paste0("nonMalignant_", paired_samples)]
    colnames(MalvsnonMasRatio) = gsub("Malignant_", "", colnames(MalvsnonMasRatio))
    L = ncol(MalvsnonMasRatio)
    rowAnn = kegg_metab_dbs$kegg_metab_db_table %>% 
        dplyr::select(gs_name, class) %>% unique() %>% 
        as.data.frame() %>% mutate(gs_name2 = gs_name) %>%
        tibble::column_to_rownames(var="gs_name2")
    rowAnn = rowAnn[rownames(MalvsnonMasRatio),]
    rowAnn = arrange(rowAnn, class)
    MalvsnonMasRatio = MalvsnonMasRatio[rownames(rowAnn), ]
    rowAnn = rowAnn[,"class", drop=FALSE]
    pheatmap::pheatmap(MalvsnonMasRatio,
                    show_colnames=TRUE,
                    color=viridisLite::magma(50),
                    breaks=seq(-2,2,len=50),
                    border_color="black",
                    scale="row",
                    # annotation_col=colAnn,
                    annotation_row=rowAnn,
                    main="Task Scores",
                    cluster_cols = FALSE,
                    cluster_rows = FALSE,
                    filename=paste0(OUTDIR, "/GSVAmetabolicScores/", c, 
                        "/GSVAscore_MalvsnonMalRatio_Heatmap_wSampleNames.pdf"), 
                    width=w+0.14*L, height=height)
    pheatmap::pheatmap(MalvsnonMasRatio,
                    labels_col=rep("", ncol(MalvsnonMasRatio)),
                    color=viridisLite::magma(50),
                    breaks=seq(-2,2,len=50),
                    border_color="black",
                    scale="row",
                    # annotation_col=colAnn,
                    annotation_row=rowAnn,
                    main="Task Scores",
                    cluster_cols = FALSE,
                    cluster_rows = FALSE,
                    filename=paste0(OUTDIR, "/GSVAmetabolicScores/", 
                        c, "/GSVAscore_MalvsnonMalRatio_Heatmap.pdf"), 
                    width=w+0.14*L, height=height)
})

```

## Summarize sample GSVA scores

```{r}
visualizeGSVSscoresGroup(
    GSVA_limma_rslt_gsva = GSVAscores,
    paired_data = lapply(names(GSVAscores), function(c) {
            g = GSVAscores[[c]]
            meta0 = data.frame(# Sample = colnames(g),
                Group = gsub("_.*", "", colnames(g)),
                Sample = gsub("nonMalignant_|Malignant_", "", colnames(g)))
            meta0
        })  %>% setNames(names(GSVAscores)),
    OUTDIR = paste0(OUTDIR, "/GSVAmetabolicScores/"),
    Condition_column = "Group",
    Condition.control = "nonMalignant",
    sample_column = "Sample",
    w=3, h=3.5, title="SMTdb"
)

visualizeGSVSscoresClass(
    GSVA_limma_rslt = 
        lapply(GSVAscores, function(g) list(gsva=g)), 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table, 
    paired_data=lapply(names(GSVAscores), function(c) {
        g = GSVAscores[[c]]
        meta0 = data.frame(Sample = colnames(g),
            Group = gsub("_.*", "", colnames(g)),
            Sample = gsub("nonMalignant_|Malignant_", "", colnames(g)))
        meta0
    })  %>% setNames(names(GSVAscores)), 
    OUTDIR = paste0(OUTDIR, "/GSVAmetabolicScores/"),
    Condition_column = "Group",
    Condition.control = "nonMalignant",
    sample_column = "Sample",
    w=5, h=3.5)

```



# Paired differential analysis on sample GSVA scores

```{r}
library(parseMetab)
OUTDIR = "../results/SMTdb/"
GSVA_limma_rslt = 
    GSVAlimmaTest_inner(GSVAscores, currentCovariate=NULL, paired=T, OUTDIR)

saveRDS(GSVA_limma_rslt, 
    paste0(OUTDIR, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))

# GSVA_limma_rslt = readRDS(paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
```

## Visualize differential results

```{r}
makeLimmaDotplot_TCGA(rlst_list=GSVA_limma_rslt, sig_statistic="P.Value", 
        sig.cutoff=1, w=8, h=16, 
        OUTDIRV=paste0(OUTDIR, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title="Dys.Path_SMTdb", 
        suffix="SMTdb", nested=T,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list=GSVA_limma_rslt, sig_statistic="P.Value", 
        sig.cutoff=0.5, w=8, h=15, 
        OUTDIRV=paste0(OUTDIR, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title="Dys.Path_SMTdb", 
        suffix="SMTdb", nested=T,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list=GSVA_limma_rslt, sig_statistic="P.Value", 
        sig.cutoff=0.1, w=8, h=8, 
        OUTDIRV=paste0(OUTDIR, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title="Dys.Path_SMTdb", 
        suffix="SMTdb", nested=T,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list=GSVA_limma_rslt, sig_statistic="P.Value", 
        sig.cutoff=0.05, w=8, h=6.5, 
        OUTDIRV=paste0(OUTDIR, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title="Dys.Path_SMTdb", 
        suffix="SMTdb", nested=T,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())
```

## Visualize the number of significant pathways

```{r}
visualizeSigNr(OUTDIR=paste0(OUTDIR, "/Limma_GSVA"), 
    GSVA_limma_rslt =
        lapply(GSVA_limma_rslt, function(g) list(rslt_of_interest=g)),
    significance_statistic="P.Value", sig.cutoff=0.1,
    w=2, h=3.5)

```

###Visualize differential effect

### Differential effect boxplot by class

```{r}
OUTDIR = "../results/SMTdb/"
GSVA_limma_rslt = 
    readRDS(paste0(OUTDIR, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
databaseDIR=paste0(OUTDIR, "/KEGG_metabolism/")
kegg_metab_dbs = 
    list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
    kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))

diffEffectBoxplot_bySystem_GSVA(
    rlst_list = lapply(GSVA_limma_rslt, function(g) list(rslt_of_interest=g)), 
    taskHierarchy=
        kegg_metab_dbs$kegg_metab_db_table %>% select(gs_name, class) %>% unique(), 
    weight.effect.by.gene = TRUE, effect.statistic="logFC",
    OUT_DIR=paste0(OUTDIR, "/Limma_GSVA"),
    w=4, h=3.5, title="SMTdb",
    winsorize=TRUE, sig.statistic="P.Value", sig.cutoff=0.1
    )
     
```

### Differential effect boxplot by cancer

```{r}
diffEffectBoxplot_byCancer_GSVA(
  rlst_list = lapply(GSVA_limma_rslt, function(g) list(rslt_of_interest=g)), 
  OUT_DIR=paste0(OUTDIR, "/Limma_GSVA"), w=1.5, h=3.0, title="SMTdb")
```

```{r}

```


