---
title: "Prepare for publication"
format: html
---



```{r}
# install.packages("ggh4x")
library(ggh4x)
library(dplyr)
library(ggplot2)
library(tidyr)
library(parseMetab)
```

# Sample size circular barplot (combined)

```{r}
# Cancer, source, condition, count, id)
meta_Zhou = 
    readRDS("../results/CellFie_Zhou2020/data/Zhou2020/meta.RDS")
sample_sizes_Zhou = 
    meta_Zhou %>% group_by(Cancer_type, Condition) %>%
    summarize(Count=n()) %>% ungroup() %>% 
    mutate(Source = "Zhou2020") 

meta_Hu = 
    readRDS("../results/CellFie_Hu2025/data/Hu2025/Meta_hu2025.RDS")
sample_sizes_Hu = 
    meta_Hu %>% group_by(Cancer_type, Condition) %>%
    summarize(Count=n()) %>% ungroup() %>% 
    mutate(Source = "Hu2025")
   
sample_sizes_TCGA = 
    readRDS("../results/TCGA/Summary/sample_sizes.RDS") %>% 
    mutate(Cancer_type = gsub("\n.*", "", Cancer_type)) %>%
    mutate(Condition = sample_type) %>%
    dplyr::select(Cancer_type, Condition, Count) %>%
    mutate(Source = "TCGA") 

sample_sizes_SMTdb =
    openxlsx::read.xlsx("../results/SMTdb/Sample_info.xlsx") %>% 
    gather(key="Condition", value="Count", -Cancer_type, -Description, -Data.type, -Source) %>%
    dplyr::select(-Description, -Data.type, -Source) %>% 
    mutate(Source = "SMTdb")

sample_sizes = 
    list(sample_sizes_Zhou,
        sample_sizes_Hu,
        sample_sizes_TCGA,
        sample_sizes_SMTdb
    ) %>% Reduce(rbind, .)
    
getCircularBarplot(
    data = setNames(sample_sizes, 
        c("individual", "observation", "value", "group")), 
    outdir = "../results/Summary/", h=19, w=21
)

openxlsx::write.xlsx(list(sample_sizes), "../results/Summary/SampleSizes.xlsx")

```

# Study summary table

```{r}
OUTDIR_proteomics = "../results/CellFie_combined/"
OUTDIR_proteomics %>% dir.create(., recursive=T, showWarnings=FALSE)

# Combine the sample summary from both Hu2025 and Zhou2020 together:
meta_Hu2025 = openxlsx::read.xlsx(
  "../results/CellFie_Hu2025/data/Hu2025/Meta_hu2025.xlsx", 
  sheet="SampleSummary") %>% spread(Condition, n) # Conditon: Tumor/Normal
meta_Zhou2020 = openxlsx::read.xlsx(
  "../results/CellFie_Zhou2020/data/Zhou2020/meta.xlsx",
  sheet="SampleSummary"
) %>% spread(Condition, n)
Hu2025_mmc1 = openxlsx::read.xlsx(
    "../data/Hu2025/mmc1.xlsx",
    sheet = "Datasets"
    ) %>% filter(Data.type=="Proteomic") %>%
    dplyr::mutate(Cancer_type=Cancer.type) %>% 
    dplyr::mutate(Source="Hu2025") %>%
    dplyr::select(Cancer_type, Description, Data.type, PMID, Publication.title, Source,
        -Tumor,-Cancer.type,-Source.ID, -Note,-`Non-tumor`) %>%
    dplyr::right_join(meta_Hu2025) %>% 
    dplyr::select(Cancer_type, Description, Normal, Tumor, Data.type, PMID, Publication.title, Source) %>%
    unique()
Zhou2020_MOESM3 = openxlsx::read.xlsx(
  "../data/Zhou2020/13045_2020_1013_MOESM2_ESM*.xlsx",
  sheet = "Patient_clinical charateristics"
  ) %>% mutate(Cancer_type=Cancer.type) %>%
  mutate(Cancer_type=gsub("ESCA", "ESCC", Cancer_type)) %>%
  mutate(Cancer_type=gsub("LUSC", "SCLC", Cancer_type)) %>%
  mutate(Cancer_type=gsub("LUSQ", "LUSC", Cancer_type)) %>%
  mutate(Description = gsub("[(]|[)]", "", Histologic.type)) %>% ###
  dplyr::mutate(Source="Zhou2020") %>%
  dplyr::select(Cancer_type, Description, Source) %>% ###
  dplyr::right_join(meta_Zhou2020) %>%
  mutate(Data.type="Proteomic") %>% ###
  mutate(PMID="33287876") %>% ###
  mutate(Publication.title="Proteomic signatures of 16 major types of human cancer reveal universal and cancer-type‑specific proteins for the identification of potential therapeutic targets") %>%
  dplyr::select(Cancer_type, Description, Normal, Tumor, Data.type, PMID, Publication.title, Source) %>%
  unique()

meta_combined = rbind(Hu2025_mmc1, Zhou2020_MOESM3) %>% arrange(Cancer_type) %>%
  mutate(Description = 
    gsub("Glioblastoma multiforme", "Brain glioblastoma", Description) %>%
    gsub("Esophagus squamous cell carcinoma","Esophageal squamous cell Carcinoma",.) %>%
    gsub("Head&neck pharynx squamous cell carcinoma|Head&neck tongue squamous cell carcinoma|Head&neck larynx squamous cell carcinoma", "Head and Neck squamous cell carcinoma", .) %>%
    gsub("Kidney clear cell renal cell carcinoma","Kidney renal clear cell carcinoma",.) %>%
    gsub("Ovary high grade serous carcinoma|Normal ovary","Ovarian serous cystadenocarcinoma",.) %>%
    gsub("Pancreas ductal adenocarcinoma", "Pancreatic adenocarcinoma",.) %>%
    gsub("Prostate non-aggressive adenocarcinoma|Prostate aggressive adenocarcinoma|Normal prostate","Prostate adenocarcinoma",.) %>%
    gsub("Uterus endometriod carcinoma","Uterine Corpus Endometrial Carcinoma",.)) %>%
  unique()

openxlsx::write.xlsx(list(Meta_combined = meta_combined), 
    file=paste0(OUTDIR_proteomics, "/Datasets.xlsx"))

```

# Sample summary
```{r}
OUTDIR = "../results/Summary/"
OUTDIR %>% dir.create(., recursive=TRUE, showWarnings=FALSE)
sample_sizes$Cancer_type = gsub("\n.*", "", sample_sizes$Cancer_type) 
library(tidyr)
sample_sizes = sample_sizes %>% filter(Source=="TCGA") %>% spread(., key=Condition, value=Count)
TCGA_info = openxlsx::read.xlsx("../data/TCGA/TCGA_projects.xlsx")
samples_info_TCGA = sample_sizes %>% 
    dplyr::left_join(TCGA_info %>% 
    mutate(Cancer_type=tumor)%>%select(Cancer_type, name)) %>%
    mutate(Description=name) %>%
    mutate(Data.type="Transcriptomic") %>%
    mutate(Source="TCGA") %>%
    select(Cancer_type, Description, Data.type, Normal, Tumor, Source)
proteomics_sample_info = 
    openxlsx::read.xlsx(paste0(OUTDIR_proteomics, "/Datasets.xlsx")) %>%
    mutate(Source=ifelse(PMID=="33287876", "Zhou2020", "Hu2025")) %>%
    select(Cancer_type, Description, Data.type, Normal, Tumor, Source)

proteomics_TCGA = rbind(samples_info_TCGA,proteomics_sample_info )

OUTDIR_SMTdb = "../results/SMTdb/"
SMTdb = openxlsx::read.xlsx(paste0(OUTDIR_SMTdb, "/Sample_info.xlsx"))
all = list(proteomics_TCGA, SMTdb) %>% Reduce(rbind, .)
openxlsx::write.xlsx(all, 
    paste0(OUTDIR, "/Sample_summary.xlsx"))
# SampleSummary = openxlsx::read.xlsx("../results/Summary/Sample_summary.xlsx")

print(paste0("We have in total ", length(unique(all$Cancer_type)), " Cancer types studied!"))
print(paste0("We have in total ", 
    sum(all$Tumor)+sum(all$Normal,na.rm=T), 
    " samples studies, which include ", sum(all$Tumor), " tumor samples and ",
    sum(all$Normal,na.rm=T), " normal samples!"))
# print(all, n=Inf)

# Cancer_type, Description, Data.type, Normal, Tumor, Scource, PMID?,  
```

[1] "We have in total 24 Cancer types studied!"
[1] "We have in total 3676 samples studies, which include 2041 tumor samples and 1635 normal samples!"



# Feature(Metabolic/gene) summary

```{r}
# Read metabolic scores from all datasets
library(readxl)
# Path to your Excel file
file_path = "../results/Summary/MetabolicScores.xlsx"
# Get all sheet names
sheet_names = excel_sheets(file_path)
# Read all sheets into a named list
all_sheets = lapply(sheet_names, read_excel, path = file_path)
# Name the list elements by sheet names
names(all_sheets) = sheet_names

# Zhou2020 & Hu2025
protein = all_sheets[grep("Zhou2020|Hu2025", names(all_sheets), value=T)]
protein = lapply(names(protein), function(c) {
    dat = protein[[c]]
    data.frame(Metabolic_units = nrow(dat),
        Genes = length(unique(unlist(strsplit(dat$Genes, split = ",")))),
        Source= gsub("_.*", "", c),
        Cancer_type = gsub(".*_", "", c)
    ) 
}) %>% Reduce(rbind,.) %>%
    gather(key="Unit", value="Count", -Cancer_type, -Source)


# TCGA & SMTdb
RNA = all_sheets[grep("TCGA|SMTdb", names(all_sheets), value=T)]
RNA = lapply(names(RNA), function(c) {
    dat = RNA[[c]]
    data.frame(Metabolic_units = nrow(dat),
        Genes = length(unique(unlist(strsplit(dat$gene_symbol, split = ",")))),
        Source= gsub("_.*", "", c),
        Cancer_type = gsub(".*_", "", c)
    ) 
}) %>% Reduce(rbind,.) %>%
    gather(key="Unit", value="Count", -Cancer_type, -Source)

genem = rbind(RNA, protein)
genem = 
    setNames(genem, 
    c("group", "individual", "observation", "value"))

openxlsx::write.xlsx(list(genem), "../results/Summary/FeatureSummary.xlsx")
genem$group = factor(genem$group, levels=c("TCGA", "SMTdb", "Zhou2020", "Hu2025"))
getCircularBarplot(
    data=genem, 
    outdir="../results/Summary/", name="FeatureSummary", h=25, w=29)

genem = rbind(RNA, protein)
sample_summary = 
    openxlsx::read.xlsx("../results/Summary/Sample_summary.xlsx") %>%
    mutate(id = paste0(Source, Cancer_type)) %>%
    dplyr::full_join(genem%>%mutate(id = paste0(Source, Cancer_type))%>%
    spread(key=Unit, value=Count) %>% mutate(`Task/Pathway`= Metabolic_units)) %>%
    dplyr::select(-id, -Metabolic_units)
openxlsx::write.xlsx(sample_summary, "../results/Summary/Sample_summary.xlsx")

```


# Prepare metabolic activitiy score table for publication

```{r}
# Zhou2020
INFILE = "../results/CellFie_Zhou2020/CellFieOut/all/detailScoring_new.csv"
dat = read.csv2(INFILE,sep="\t") %>%
    filter( ExpressionScoreEssentialRxnsTask != -1) %>%
    mutate(Task=Depth3, Class=Depth1, EntrezID=GeneAssociatedToEssentialRxnsTask) %>%
    dplyr::select(Task, Class, EntrezID, Sample, TaskScore, Cancer_type)
# mapping = mygene::queryMany(dat$EntrezID, 
#             scopes = "entrezgene", 
#             fields = "symbol", species = "human") 
# saveRDS(mapping, "../results/Summary/mapping.RDS")
mapping = readRDS("../results/Summary/mapping.RDS")
mappings2 = setNames(mapping$symbol, mapping$query) 
ZhouActivity = lapply(unique(dat$Cancer_type), function(c) {
    filter(dat, Cancer_type==c) %>%
    mutate(Genes = mappings2[EntrezID])  %>% 
    group_by(Task, Class, Sample,TaskScore) %>%
    summarize(
        Genes = paste(unique(Genes), collapse = ",")
    ) %>% ungroup() %>%
    spread(key=Sample, value=TaskScore)
}) %>% setNames(paste0("Zhou2020_", unique(dat$Cancer_type)))
# Only keep those samples used in differential analysis
meta_Zhou = 
    readRDS("../results/CellFie_Zhou2020/data/Zhou2020/meta.RDS")
normSamples = meta_Zhou$Sample
ZhouActivity = lapply(ZhouActivity, function(c) {
    c[, c("Task", "Class", "Genes", intersect(colnames(c), normSamples))]
})
# Only cancer types that end up in differential results
ZhouActivity = ZhouActivity[paste0("Zhou2020_", unique(meta_Zhou$Cancer_type))]
saveRDS(ZhouActivity, "../results/Summary/MetabolicActivity_Zhou2020.RDS")
#head(ZhouActivity)


# Hu2025
INDIR = "../results/CellFie_Hu2025/CellFieOut/"
cancers = list.dirs(INDIR, full.names=FALSE) %>% setdiff(.,"")
library(tidyr)
HuActivity = lapply(cancers, function(c) {
    dat = read.csv2(paste0(INDIR,"/",c,"/detailScoring_new.csv"),sep="\t") %>%
        filter( ExpressionScoreEssentialRxnsTask != -1) %>%
        mutate(Task=Depth3, Class=Depth1, EntrezID=GeneAssociatedToEssentialRxnsTask) %>%
        dplyr::select(Task, Class, EntrezID, Sample, TaskScore) %>%
        mutate(Genes = mappings2[EntrezID])  %>% 
        group_by(Task, Class, Sample,TaskScore) %>%
        summarize(
            Genes = paste(unique(Genes), collapse = ",")
        ) %>% ungroup() %>%
        spread(key=Sample, value=TaskScore)
}) %>% setNames(paste0("Hu2025_",cancers))
# Only keep those samples used in differential analysis
meta_Hu = readRDS("../results/CellFie_Hu2025/data/Hu2025/Meta_hu2025.RDS")
normSamples2 = meta_Hu$Sample
HuActivity = lapply(HuActivity, function(c) {
    c[, c("Task", "Class", "Genes", intersect(colnames(c),normSamples2))]
})
# Only cancer types that end up in differential results
HuActivity = HuActivity[paste0("Hu2025_", unique(meta_Hu$Cancer_type))]
saveRDS(HuActivity, "../results/Summary/MetabolicActivity_Hu2025.RDS")
# head(HuActivity)


# TCGA
INDIR2 = "../results/TCGA/Limma_GSVA/"
cancers2 = list.dirs(INDIR2, full.names=FALSE) %>% setdiff(.,"")
databaseDIR = "../results/TCGA/KEGG_metabolism/"
kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS"))
TCGAactivity = lapply(cancers2, function(c) {
    dat = kegg_metab_db_table %>%
        right_join(., 
        read.csv2(paste0(INDIR2,"/",c,"/GSVAscores.csv"), sep="\t") %>%
        mutate(gs_name = Features)) %>%
        mutate(pathway_name = gs_name) %>%
        dplyr::select(-Features, -gs_name) %>%
        dplyr::select(pathway_name, gene_symbol, class, everything()) %>%
        group_by(pathway_name, class) %>%
        summarise(
            gene_symbol = paste(unique(gene_symbol), collapse = ","),
            across(-gene_symbol, dplyr::first),
            .groups = "drop"
        )
}) %>% setNames(paste0("TCGA_", cancers2))
# Only cancer types that end up in differential results
outdir3 = "../results/TCGA"
GSVA_limma_rslt_TCGA = 
    readRDS(paste0(outdir3, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
normSamples3 = sapply(GSVA_limma_rslt_TCGA, function(c) {
    names(c$gsva)
}) %>% unlist() %>% as.character()
TCGAactivity = TCGAactivity[paste0("TCGA_",names(GSVA_limma_rslt_TCGA))]
saveRDS(TCGAactivity, "../results/Summary/MetabolicActivity_TCGA.RDS")


# SMTdb
INDIR3 = "../results/SMTdb/GSVAmetabolicScores/"
cancers3 = list.dirs(INDIR3, full.names=FALSE) %>% setdiff(.,"")
databaseDIR = "../results/TCGA/KEGG_metabolism/"
kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS"))
STMdbActivity = lapply(cancers3, function(c) {
    dat = kegg_metab_db_table %>%
        right_join(., 
        read.csv2(paste0(INDIR3,"/",c,"/GSVAmetabolicScores.csv"), sep="\t") %>%
        mutate(gs_name = Features)) %>%
        mutate(pathway_name = gs_name) %>%
        dplyr::select(-Features, -gs_name) %>%
        dplyr::select(pathway_name, gene_symbol, class, everything()) %>%
        group_by(pathway_name, class) %>%
        summarise(
            gene_symbol = paste(unique(gene_symbol), collapse = ","),
            across(-gene_symbol, dplyr::first),
            .groups = "drop"
        )
}) %>% setNames(paste0("SMTdb_", cancers3))
saveRDS(STMdbActivity, "../results/Summary/MetabolicActivity_STMdb.RDS")

openxlsx::write.xlsx(
    c(HuActivity, ZhouActivity, TCGAactivity, STMdbActivity),
    file="../results/Summary/MetabolicScores.xlsx"
)

sapply(c(HuActivity, ZhouActivity, TCGAactivity, STMdbActivity), function(data) {
    dim(data)
})

```

# Prepare TCGA meta data table used for survival analysis

```{r}
cancers2analyze =
    readRDS("../results/SurvivalAnalysis_TCGA/cancers.RDS")
TCGA_meta4survival = lapply(cancers2analyze, function(c) {
    clinic = openxlsx::read.xlsx(paste0("../data/TCGA/",c,"/Clinical_",c,".xlsx")) %>%
        dplyr::select(submitter_id, days_to_death, 
        days_to_last_follow_up, 
        age_at_diagnosis, gender, ajcc_pathologic_stage) %>% 
        mutate(Cancer_type = c) %>%
        mutate(Source = "TCGA")
    meta = openxlsx::read.xlsx(paste0("../data/TCGA/",c,"/Meta_",c,".xlsx")) %>%
        mutate(submitter_id = cases.submitter_id) %>%
        mutate(Sample=sample.submitter_id) %>%
        dplyr::select(Sample, everything())
    clinic %>% left_join(meta) %>% dplyr::select(-submitter_id) %>%
        dplyr::select(Sample, everything())
}) %>% Reduce(rbind, .)
openxlsx::write.xlsx(TCGA_meta4survival, "../results/Summary/TCGA_meta4survival.xlsx")

```


# Prepare full metadata table

```{r}
columns2rm = c("pathway_name", "class", "gene_symbol", "Task", "Class Genes")
ZhouActivity = readRDS("../results/Summary/MetabolicActivity_Zhou2020.RDS")
ZhouSamples = lapply(ZhouActivity, function(c) {colnames(c)}) %>% 
    unlist() %>% as.character() %>% setdiff(., columns2rm) 
ZhouMeta = readRDS("../results/CellFie_Zhou2020/data/Zhou2020/meta.RDS") %>% 
    filter(Sample %in% ZhouSamples) %>%
    mutate(Stage = Tumor_Stage_Pathological) %>%
    dplyr::select(Sample, Condition, Patient, Age, Gender, Stage, Cancer_type)
table(ZhouMeta$Batch)


HuActivity = readRDS("../results/Summary/MetabolicActivity_Hu2025.RDS")
HuSamples = lapply(HuActivity, function(c) {colnames(c)}) %>% 
    unlist() %>% as.character() %>% setdiff(., columns2rm)  
HuMeta = readRDS("../results/CellFie_Hu2025/data/Hu2025/Meta_hu2025.RDS") %>% 
    filter(Sample %in% HuSamples) %>%
    mutate(Sample = Sample_ID) %>%
    mutate(Condition = gsub("NAT", "Normal", Sample_type)) %>%
    mutate(Patient = Patient_ID) %>%
    mutate(Cancer_type = Tumor_code) %>%
    mutate(Age = NA) %>%
    mutate(Gender = NA) %>%
    mutate(Stage = NA) %>%
    mutate(days_to_death = NA) %>%
    mutate(days_to_last_follow_up = NA) %>%
    dplyr::select(Sample, Condition, Patient, Cancer_type, Batch)
head(HuMeta)


TCGAactivity = readRDS("../results/Summary/MetabolicActivity_TCGA.RDS")
TCGAsamples = lapply(TCGAactivity, function(c) {colnames(c)}) %>% 
    unlist() %>% as.character() %>% setdiff(., columns2rm)
TCGA_meta = lapply(gsub("TCGA_", "", names(TCGAactivity)), function(c) {
        clinic = openxlsx::read.xlsx(paste0("../data/TCGA/",c,"/Clinical_",c,".xlsx")) %>%
            dplyr::select(submitter_id, days_to_death, 
            days_to_last_follow_up, 
            age_at_diagnosis, gender, ajcc_pathologic_stage) %>% 
            mutate(Cancer_type = c) %>%
            mutate(Source = "TCGA")
        meta = openxlsx::read.xlsx(paste0("../data/TCGA/",c,"/Meta_",c,".xlsx")) %>%
            mutate(submitter_id = cases.submitter_id) %>%
            mutate(Sample=sample.submitter_id) %>%
            dplyr::select(Sample, everything())
        clinic %>% left_join(meta) %>% dplyr::select(-submitter_id) %>%
            dplyr::select(Sample, everything())
    }) %>% Reduce(rbind, .) %>%
    mutate(Sample = gsub("-", ".", Sample)) %>%
    filter(Sample %in% TCGAsamples) %>% 
    mutate(Patient = cases.submitter_id)
# Only keep the first samples, if one patient has more than one row#######
paired_data = readRDS("../results/TCGA/Summary/paired_data.RDS")
file_ids = sapply(paired_data, function(c) c%>%pull(file_id)) %>% unlist() %>% as.character()
TCGA_meta = TCGA_meta %>% filter(file_id %in% file_ids)
TCGA_meta2 = TCGA_meta %>% mutate(Condition = sample_type) %>% 
    dplyr::select(Sample, Condition, Patient, age_at_diagnosis, gender, ajcc_pathologic_stage, Cancer_type, days_to_death, days_to_last_follow_up)
head(TCGA_meta2)


# SMTdb
SMTdbActivity = readRDS("../results/Summary/MetabolicActivity_STMdb.RDS")
SMTdbSamples = lapply(SMTdbActivity, function(c) {colnames(c)}) %>% 
    unlist() %>% as.character() %>% setdiff(., columns2rm) %>%
    gsub("^Malignant_|^nonMalignant_", "", .)
SMTdbMeta = readRDS("../results/SMTdb/meta.RDS") %>%
    filter(datasetID %in% SMTdbSamples) %>%
    mutate(Cancer_type = cancer) %>%
    dplyr::select(datasetID, slice_type, spot_num, Cancer_type)
head(SMTdbMeta)

SampleSummary = openxlsx::read.xlsx("../results/Summary/Sample_summary.xlsx")
openxlsx::write.xlsx(list(SampleSummary=SampleSummary,
    Hu2025 = HuMeta,
    Zhou2020 = ZhouMeta,
    TCGA = TCGA_meta2,
    SMTdb = SMTdbMeta),
    "../results/Summary/Sample_summary_all.xlsx")
# Select Samples, pair, batch to keep: more clinical data for TCGA to keep
TCGA_meta2 %>% group_by(Cancer_type, Condition) %>% summarise(count = n())

# length(unique(SampleSummary$Cancer_type))
# sum(SampleSummary$Normal, na.rm=T)
# sum(SampleSummary$Tumor, na.rm=T)

```


# Prepare differential analysis result table for publication

```{r}
# Zhou2020_meta
OUTDIR1 = "../results/CellFie_Zhou2020/"
OUTDIRV1 = paste0(OUTDIR1, "/Limma/TaskDepth3/")
result3_Zhou = readRDS(paste0(OUTDIRV1, "/result3.RDS"))
result3_Zhou = result3_Zhou$rlst_list
result3_Zhou = lapply(result3_Zhou, function(c) {
    c$rslt_of_interest %>% as.data.frame() %>%
    tibble::rownames_to_column(var="Task")
})
names(result3_Zhou) = paste0("MA_Zhou2020_", names(result3_Zhou))
result3_Zhou = lapply(names(result3_Zhou), function(c) {
    ZhouActivity[[gsub("MA_","",c)]][,c("Task","Class","Genes")] %>% 
        dplyr::left_join(.,result3_Zhou[[c]])
}) %>% setNames(names(result3_Zhou))

# Hu2025_meta
OUTDIR2 = "../results/CellFie_Hu2025/"
OUTDIRV2 = paste0(OUTDIR2, "/Limma/TaskDepth3/")
result3_Hu = readRDS(paste0(OUTDIRV2, "/result3.RDS"))
result3_Hu = result3_Hu$rlst_list
result3_Hu = lapply(result3_Hu, function(c) {
    c$rslt_of_interest %>%
    tibble::rownames_to_column(var="Task")
})
names(result3_Hu) = paste0("MA_Hu2025_", names(result3_Hu))
result3_Hu = lapply(names(result3_Hu), function(c) {
    HuActivity[[gsub("MA_","",c)]][,c("Task","Class","Genes")] %>% 
        dplyr::left_join(.,result3_Hu[[c]])
}) %>% setNames(names(result3_Hu))

# Hu2025/Zhou2020_Genes
rslt_lists = readRDS("../results/FeatureDifferential_CellFie/rslt_list.RDS")
diffGenes  = lapply(c("Zhou", "Hu"), function(author) {
    study = paste0("limma_rslt_list", author)
    studyYear = paste0(author, ifelse(author=="Hu", "2025", "2020"))
    lapply(names(rslt_lists), function(system) {
        lapply(names(rslt_lists[[system]][[study]]),function(c) {
            temp = rslt_lists[[system]][[study]][[c]]
            temp$Cancer_type = c
            temp
        }) %>% Reduce(rbind, .)
    }) %>% 
    setNames(paste0("Protein_",studyYear,"_", sapply(strsplit(names(rslt_lists)," "),function(n) n[1])))
}) 
diffGenes = c(diffGenes[[1]], diffGenes[[2]])

# TCGA
outdir3 = "../results/TCGA"
GSVA_limma_rslt_TCGA = 
    readRDS(paste0(outdir3, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
GSVA_limma_rslt_TCGA = lapply(GSVA_limma_rslt_TCGA, function(c) {
    c$rslt_of_interest %>% as.data.frame() %>%
        tibble::rownames_to_column(var="pathway_name")
})
names(GSVA_limma_rslt_TCGA) = paste0("MA_TCGA_", names(GSVA_limma_rslt_TCGA))
GSVA_limma_rslt_TCGA = lapply(names(GSVA_limma_rslt_TCGA), function(c) {
    TCGAactivity[[gsub("MA_","",c)]][,c("pathway_name","class","gene_symbol")] %>% 
        dplyr::left_join(.,GSVA_limma_rslt_TCGA[[c]])
}) %>% setNames(names(GSVA_limma_rslt_TCGA))


# SMTdb
OUTDIR4 = "../results/SMTdb/"
GSVA_limma_rslt_SMTdb = 
    readRDS(paste0(OUTDIR4, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
GSVA_limma_rslt_SMTdb = lapply(GSVA_limma_rslt_SMTdb, function(c) {
    c %>% as.data.frame() %>%
        tibble::rownames_to_column(var="pathway_name")
})
names(GSVA_limma_rslt_SMTdb) = paste0("MA_SMTdb_", names(GSVA_limma_rslt_SMTdb))
GSVA_limma_rslt_SMTdb = lapply(names(GSVA_limma_rslt_SMTdb), function(c) {
    STMdbActivity[[gsub("MA_","",c)]][,c("pathway_name","class","gene_symbol")] %>% 
        dplyr::left_join(.,GSVA_limma_rslt_SMTdb[[c]])
}) %>% setNames(names(GSVA_limma_rslt_SMTdb))


openxlsx::write.xlsx(
    c(result3_Zhou, result3_Hu, GSVA_limma_rslt_TCGA, GSVA_limma_rslt_SMTdb, diffGenes),
    file = "../results/Summary/limmadifferentialtable.xlsx"
)

```


# Differential effect (t) plots

Be aware that strictly speaking, t is not effect size! logFC is! But the direction is the same between the two.

## Proteomics (task-level, protein-levels)
```{r}
# Differential task scores
OUTDIR = "../results/CellFie_Zhou2020/"
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")
result3Zhou = readRDS(paste0(OUTDIRV, "/result3.RDS"))
cfsZhou = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))

OUTDIR = "../results/CellFie_Hu2025/"
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")
result3Hu = readRDS(paste0(OUTDIRV, "/result3.RDS"))
cfsHu = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))

# Differential gene 
rslt_lists = readRDS("../results/FeatureDifferential_CellFie/rslt_list.RDS")
# system-study-cancer-rsltofinterest
result3_proteins = lapply(names(rslt_lists[[1]])[1:2], function(study) {
    lapply(names(rslt_lists[[1]][[1]]), function(c) {
        list(rslt_of_interest = lapply(names(rslt_lists), function(system) {
            temp = rslt_lists[[system]][[study]][[c]] %>%
                as.data.frame() 
            if (nrow(temp)>0) {
                temp = temp %>% 
                    mutate(Depth1 = gsub("METABOLISM","M.",system), Depth3=Feature)
            }
            temp
        })%>% Reduce(rbind,.) %>% as.data.frame())
    }) %>% setNames(names(rslt_lists[[1]][[1]]))
}) %>% setNames(c("Zhou2020_protein", "Hu2025_protein")) 
result3_proteins = result3_proteins %>% lapply(., function(study) {
    study[sapply(study, function(c) {
        nrow(c$rslt_of_interest)>0
    })]
    }) %>% lapply(., function(study) {
        study[sapply(study, function(c) {
            !is.null(c$rslt_of_interest)
        })]
    }) 
result3_proteins = result3_proteins %>% lapply(., function(study) {
        lapply(study, function(c) {
            lapply(c, function(r) {
                r %>% dplyr::select(-EntrezID, -Feature)
            })
        })
    })

OUTDIR = "../results/Summary/"
diffEffectBoxplot_bySystem(
  rlst_list = c(list(
    Zhou2020_Mtask=result3Zhou$rlst_list,
    Hu2025_Mtask=result3Hu$rlst_list), 
    result3_proteins), 
  cf=list(Zhou2020_Mtask=cfsZhou%>%mutate(Depth1=gsub("METABOLISM","M.", Depth1)),
          Hu2025_Mtask=cfsHu%>%mutate(Depth1=gsub("METABOLISM","M.", Depth1)),
          Zhou2020_protein=cfsZhou%>%mutate(Depth1=gsub("METABOLISM","M.", Depth1)),
          Hu2025_protein=cfsHu%>%mutate(Depth1=gsub("METABOLISM","M.", Depth1))), 
  OUT_DIR=paste0(OUTDIR, "/"), sig.statistic="adj.P.Val", sig.cutoff=0.2,
  w=6, h=3, title="Proteomics", prob=0.15)

```

## Transcriptomics (bulk, spatial)

```{r}
outdir = "../results/TCGA"
GSVA_limma_rslt_TCGA = 
    readRDS(paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
databaseDIR=paste0(outdir, "/KEGG_metabolism/")
kegg_metab_dbs = 
    list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
    kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))

OUTDIR = "../results/SMTdb/"
GSVA_limma_rslt_SMTdb = 
    readRDS(paste0(OUTDIR, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
databaseDIR=paste0(OUTDIR, "/KEGG_metabolism/")

OUTDIR = "../results/Summary/"
diffEffectBoxplot_bySystem_GSVA(
    rlst_list = list(
        TCGA=GSVA_limma_rslt_TCGA,
        SMTdb=lapply(GSVA_limma_rslt_SMTdb, function(g) list(rslt_of_interest=g))),
    taskHierarchy =
        kegg_metab_dbs$kegg_metab_db_table %>% select(gs_name, class) %>% unique(), 
    OUT_DIR = OUTDIR, winsorize=TRUE, probs=0.2,
    weight.effect.by.gene=FALSE, sig.statistic="P.Value", sig.cutoff=0.2,
    w=3.5, h=3, title="(spatial) Transcriptomics", suffix="Transcritpomics")

```


# Differential number plot

```{r}
# Zhou
## adj.P.Val < 0.1
OUTDIR1 = "../results/CellFie_Zhou2020/"
OUTDIRV1 = paste0(OUTDIR1, "/Limma/TaskDepth3/")
result3 = readRDS(paste0(OUTDIRV1, "/result3.RDS"))
diff_Zhou = result3$rlst_list  
diff_Zhou = lapply(names(diff_Zhou), function(c) {
        dat = diff_Zhou[[c]]$rslt_of_interest
        data.frame(up=sum(dat$adj.P.Val<0.1&dat$logFC>0),
            down = sum(dat$adj.P.Val<0.1&dat$logFC<0),
            Cancer_type = c,
            Source="Zhou2020")
    }) %>% Reduce(rbind, .) %>%
    mutate(down = -down) %>%
    gather(key="Regulation", value="Count", -Cancer_type, -Source)


# Hu
## adj.P.Val < 0.05
OUTDIR2 = "../results/CellFie_Hu2025/"
OUTDIRV2 = paste0(OUTDIR2, "/Limma/TaskDepth3/")
result3_Hu = readRDS(paste0(OUTDIRV2, "/result3.RDS")) 
diff_Hu = result3_Hu$rlst_list  
diff_Hu = lapply(names(diff_Hu), function(c) {
        dat = diff_Hu[[c]]$rslt_of_interest
        data.frame(up=sum(dat$adj.P.Val<0.05&dat$logFC>0),
            down = sum(dat$adj.P.Val<0.05&dat$logFC<0),
            Cancer_type = c,
            Source = "Hu2025")
    }) %>% Reduce(rbind, .) %>%
    mutate(down = -down) %>%
    gather(key="Regulation", value="Count", -Cancer_type, -Source)


# TCGA
## adj.P.Val < 0.05
outdir3 = "../results/TCGA"
GSVA_limma_rslt_TCGA = 
    readRDS(paste0(outdir3, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
diff_TCGA = lapply(names(GSVA_limma_rslt_TCGA), function(c) {
        dat = GSVA_limma_rslt_TCGA[[c]]$rslt_of_interest
        data.frame(up=sum(dat$adj.P.Val<0.05&dat$logFC>0),
            down = sum(dat$adj.P.Val<0.05&dat$logFC<0),
            Cancer_type = c,
            Source = "TCGA")
    }) %>% Reduce(rbind, .) %>%
    mutate(down = -down) %>%
    gather(key="Regulation", value="Count", -Cancer_type, -Source)


# SMTdb
## P.Value < 0.1
OUTDIR4 = "../results/SMTdb/"
GSVA_limma_rslt_SMTdb = 
    readRDS(paste0(OUTDIR4, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
diff_SMTdb = lapply(names(GSVA_limma_rslt_SMTdb), function(c) {
        dat = GSVA_limma_rslt_SMTdb[[c]]
        data.frame(up=sum(dat$P.Value<0.1&dat$logFC>0),
            down = sum(dat$P.Value<0.1&dat$logFC<0),
            Cancer_type = c,
            Source="SMTdb")
    }) %>% Reduce(rbind, .) %>%
    mutate(down = -down) %>%
    gather(key="Regulation", value="Count", -Cancer_type, -Source)

diffs = list(diff_Hu, diff_Zhou, diff_TCGA, diff_SMTdb) %>%
    Reduce(rbind, .)
cancer_order = diffs %>% group_by(Cancer_type, Regulation) %>% 
    summarize(mean = mean(Count, na.rm=T)) %>% 
    arrange(-mean) %>% pull(Cancer_type)
diffs$Cancer_type = factor(diffs$Cancer_type, levels=unique(cancer_order))

pdf("../results/Summary/diffnr.pdf", w=6, h=3)
ggplot(diffs) +
    geom_col(
        aes(x = Cancer_type, y = Count, fill=Source), 
        stat = "identity" ,
        alpha = 0.5
    ) +
    ggplot2::scale_fill_manual(
        values=c("gray70", "green4", "#6A3D9A","#FF7F00")) +
    ggplot2::geom_hline(yintercept=0, color="black") +
    ggplot2::labs(title = "",
        x="Cancer_type", y="# Significant")+
    ggplot2::theme_minimal() +
    ggplot2::theme(legend.position="right",
        axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0))
dev.off()

```




```{r}
STOP!
```

