---
title: "Prepare for publication"
format: html
---



# Sample summary_circular barplot

```{r}
# install.packages("ggh4x")
library(ggh4x)
library(dplyr)
library(ggplot2)
# Combine the sample summary from both Hu2025 and Zhou2020 together:
library(tidyr)
meta_Hu2025 = openxlsx::read.xlsx(
  "../results/CellFie_Hu2025/data/Hu2025/Meta_hu2025.xlsx", 
  sheet="SampleSummary") %>% mutate(Condition=Group) %>%
  dplyr::select(-Group) %>%
  spread(Condition, n)
# meta_Hu2025$Normal+meta_Hu2025$Tumor
#  [1] 197 248 109 171 176 330 211 208 103 207 160  95
meta_Zhou2020 = openxlsx::read.xlsx(
  "../results/CellFie_Zhou2020/data/Zhou2020/meta.xlsx",
  sheet="SampleSummary"
) %>% spread(Condition, n)
# meta_Zhou2020$Normal+meta_Zhou2020$Tumor
#  [1] 22 10 22 13 NA 11 19 16 17 15 13 13 14 12 15  8
Hu2025_mmc1 = openxlsx::read.xlsx(
  "../data/Hu2025/mmc1.xlsx",
  sheet = "Datasets"
  ) %>% filter(Data.type=="Proteomic") %>%
  mutate(Cancer_type=Cancer.type) %>% 
  mutate(Normal=`Non-tumor`) %>%
  dplyr::select(Cancer_type, Description, Data.type,PMID,Publication.title,
      -Normal,-Tumor,-Cancer.type,-Source,-Source.ID, -Note,-`Non-tumor`) %>%
      dplyr::right_join(meta_Hu2025) %>% 
      dplyr::select(Cancer_type, Description, Normal, Tumor, Data.type, PMID, Publication.title) %>%
      unique()
Zhou2020_MOESM3 = openxlsx::read.xlsx(
  "../data/Zhou2020/13045_2020_1013_MOESM2_ESM*.xlsx",
  sheet = "Patient_clinical charateristics"
  ) %>% mutate(Cancer_type=Cancer.type) %>%
  mutate(Cancer_type=gsub("ESCA", "ESCC", Cancer_type)) %>%
  mutate(Cancer_type=gsub("LUSC", "SCLC", Cancer_type)) %>%
  mutate(Cancer_type=gsub("LUSQ", "LUSC", Cancer_type)) %>%
  mutate(Description = gsub("[(]|[)]", "", Histologic.type)) %>%
  dplyr::select(Cancer_type, Description) %>% 
  dplyr::right_join(meta_Zhou2020) %>%
  mutate(Data.type="Proteomic") %>%
  mutate(PMID="33287876") %>%
  mutate(Publication.title="Proteomic signatures of 16 major types of human cancer reveal universal and cancer-type‑specific proteins for the identification of potential therapeutic targets") %>%
  dplyr::select(Cancer_type, Description, Normal, Tumor, Data.type, PMID, Publication.title) %>%
  unique()

meta_combined = rbind(Hu2025_mmc1, Zhou2020_MOESM3) %>% arrange(Cancer_type) %>%
  mutate(Description = 
    gsub("Glioblastoma multiforme", "Brain glioblastoma", Description) %>%
    gsub("Esophagus squamous cell carcinoma","Esophageal squamous cell Carcinoma",.) %>%
    gsub("Head&neck pharynx squamous cell carcinoma|Head&neck tongue squamous cell carcinoma|Head&neck larynx squamous cell carcinoma", "Head and Neck squamous cell carcinoma", .) %>%
    gsub("Kidney clear cell renal cell carcinoma","Kidney renal clear cell carcinoma",.) %>%
    gsub("Ovary high grade serous carcinoma|Normal ovary","Ovarian serous cystadenocarcinoma",.) %>%
    gsub("Pancreas ductal adenocarcinoma", "Pancreatic adenocarcinoma",.) %>%
    gsub("Prostate non-aggressive adenocarcinoma|Prostate aggressive adenocarcinoma|Normal prostate","Prostate adenocarcinoma",.) %>%
    gsub("Uterus endometriod carcinoma","Uterine Corpus Endometrial Carcinoma",.)) %>%
  unique()
# meta_combined$Cancer_type[18:19] = c("SCLC", "LUSC")
# meta_combined$Cancer_type = gsub("ESCA", "ESCC", meta_combined$Cancer_type)
# sum(meta_combined$Normal,na.rm=T)+sum(meta_combined$Tumor,na.rm=T)
# [1] 2447
TableNames = list(Meta_combined = meta_combined)
openxlsx::write.xlsx(TableNames, 
  file=paste0("../doc/FiguresTables/Datasets.xlsx"))

temp = meta_combined %>% 
  filter(PMID=="33287876") %>% 
  .[, c("Cancer_type", "Normal", "Tumor")] %>%
  mutate(Sum = ifelse(is.na(Normal), 0, Normal)+ifelse(is.na(Tumor), 0, Tumor)) %>% 
  mutate(Cancer_type = paste0(Cancer_type, "\n(",Sum,")")) %>%
  dplyr::select(-Sum) %>%
  gather(key="Sample_type", value="Count", -Cancer_type)
temp$Sample_type = factor(temp$Sample_type, levels=c("Tumor","Normal"))
Zhou2020_sample = temp %>%
  ggplot(., aes(x=Cancer_type, y=Count, fill=Sample_type)) +
  geom_bar(stat="identity", show.legend=FALSE) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")) + 
  coord_polar() +
  ggtitle("Zhou2020") 
temp = meta_combined %>% 
  filter(PMID!="33287876") %>% 
  .[, c("Cancer_type", "Normal", "Tumor")] %>%
  mutate(Sum = ifelse(is.na(Normal), 0, Normal)+ifelse(is.na(Tumor), 0, Tumor)) %>% 
  mutate(Cancer_type = paste0(Cancer_type, "\n(",Sum,")")) %>%
  dplyr::select(-Sum) %>%
  gather(key="Sample_type", value="Count", -Cancer_type)
temp$Sample_type = factor(temp$Sample_type, levels=c("Tumor","Normal"))
Hu2025_sample =  temp %>%
  ggplot(., aes(x=Cancer_type, y=Count, fill=Sample_type)) +
  geom_bar(stat="identity") +
  theme_minimal() +
  theme(legend.position = "left",
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        plot.margin = unit(c(-0,-0,-0,-0), "cm")) +
  coord_polar() +
  ggtitle("Others") 
pdf(paste0("../doc/FiguresTables/SampleSummary.pdf"), h=4,w=9)
cowplot::plot_grid(Zhou2020_sample, Hu2025_sample, ncol=2)
dev.off()
rm(temp)
gc()

# Combine the sample information from both Hu2025 and Zhou2020
Zhou2020_meta = openxlsx::read.xlsx(
  "../results/CellFie_Zhou2020/data/Zhou2020/meta.xlsx",
  sheet="MetaData"
) %>% mutate(Sample_ID=Sample) %>%
  mutate(Patient_ID=Pair_ID) %>%
  mutate(Sample_type=Condition) %>%
  mutate(Paired="Single") %>%
  mutate(Batch="Batch1") %>%
  mutate(TMT_plex="") %>% 
  dplyr::select(Sample_ID,Cancer_type,Patient_ID,Sample_type,Paired,Batch,TMT_plex) 
Zhou2020_meta[!is.na(Zhou2020_meta$Patient_ID),"Paired"] = "Paired"
# Zhou2020_meta$Cancer_type = gsub("LUSC","SCLC", Zhou2020_meta$Cancer_type) %>%
#   gsub("LUSQ","LUSC",.) %>% gsub("ESCA","ESCC",.)
Zhou2020_meta = Zhou2020_meta %>% left_join(meta_combined%>%filter(PMID=="33287876"))
Zhou2020_meta$Sample_type = gsub("NAT", "Normal", Zhou2020_meta$Sample_type)

Hu2025_meta = openxlsx::read.xlsx(
  "../results/CellFie_Hu2025/data/Hu2025/Meta_hu2025.xlsx",
  sheet="meta") %>% mutate(Cancer_type=Tumor_code) %>%
  mutate(Batch = Run_ID) %>%
  dplyr::select(Sample_ID,Cancer_type,Patient_ID,Sample_type,Paired,Batch,TMT_plex) %>%
  left_join(meta_combined%>%filter(PMID!="33287876"))
Hu2025_meta$Sample_type = gsub("NAT", "Normal", Hu2025_meta$Sample_type)

Sample_info = rbind(Zhou2020_meta%>%dplyr::select(-Normal,-Tumor), 
  Hu2025_meta%>%dplyr::select(-Normal,-Tumor))
# table(Sample_info$Paired)
# Paired Single 
#   2014    433 
# > 2014/2 + 433
# [1] 1440
TableNames = list(
  Sample_info = Sample_info,
  Cancer_type_summary = Sample_info %>% dplyr::count(Cancer_type, Sample_type)
  )
openxlsx::write.xlsx(TableNames, 
  file=paste0("../doc/FiguresTables/SampleInfo.xlsx"))
rm(Sample_info, TableNames)
# sum(is.na(Sample_info$Patient_ID))
# unique(Sample_info$Patient_ID)
# Sample_info %>% dplyr::count(Cancer_type, Sample_type)


# Collect all normalized datasets to one single excel file
# zhou2020_data = 
#     read.csv("../results/CellFie_Zhou2020/data/Zhou2020/CancerProteomics.csv", 
#     header = TRUE, stringsAsFactors = FALSE)
# # 1. List all CSV files in the folder
# file_list = list.files(path = "../results/CellFie_Hu2025/data/Hu2025/", 
#     pattern = "\\.csv$", full.names = TRUE)
# file_names = list.files(path = "../results/CellFie_Hu2025/data/Hu2025/", 
#     pattern = "\\.csv$", full.names = F) %>% 
#     gsub("CancerProteomicsHu2025_", "", .) %>%
#     gsub(".csv", "", .)
# # 2. Read each CSV file into a list
# Hu2025_data = lapply(file_list, read.csv, header = TRUE, stringsAsFactors = FALSE) %>%
#     setNames(file_names)
# openxlsx::write.xlsx(c(list(Zhou2020 = zhou2020_data), 
#     Hu2025_data[setdiff(file_names, "CancerProteomicsHu2025")]), 
#     file=paste0("../doc/FiguresTables/ProcessedProteinAbundances.xlsx"))
# rm(Sample_info, TableNames, zhou2020_data, file_list, file_names, Hu2025_data)

```

# Collect all task score results

```{r}
taskScores_files = "../results/CellFie_Hu2025/Summary/TaskScores_differentDepths.xlsx"
taskScores_Hu2025 = lapply(openxlsx::getSheetNames(taskScores_files), function(sheet) {
  openxlsx::read.xlsx(taskScores_files, sheet = sheet) 
}) %>% setNames(paste0(openxlsx::getSheetNames(taskScores_files), "_Others"))

taskScores_files2 = "../results/CellFie_Zhou2020/Summary/TaskScores_differentDepths.xlsx"
taskScores_Zhou2020 = lapply(openxlsx::getSheetNames(taskScores_files2), function(sheet) {
  openxlsx::read.xlsx(taskScores_files2, sheet = sheet) 
}) %>% setNames(paste0(openxlsx::getSheetNames(taskScores_files2), "_Zhou2020"))

openxlsx::write.xlsx(c(taskScores_Hu2025, taskScores_Zhou2020), 
  file=paste0("../doc/FiguresTables/TaskScores.xlsx"))
rm(taskScores_Hu2025, taskScores_Zhou2020, taskScores_files, taskScores_files2)

```

# Task hierachy_cicular packing plot
```{r}
# Summarize the CellFie output
## Hu et al. 2025
cfs = readRDS("../results/CellFie_Hu2025/Summary/cfs.RDS")
head(cfs)
min(cfs$TaskScore)
unique(cfs$Depth3) %>% length()
# 138 metabolic tasks
unique(cfs$Depth2) %>% length()
# 58 metabolic subsystems
unique(cfs$Depth1) %>% length()
# 7 metabolic systems

# Libraries
library(ggraph)
library(igraph)
library(tidyverse)
library(viridis)
edges = cfs%>%mutate(Depth0="Metabolism")%>%.[,c("Depth0","Depth1")]%>%
    unique()%>%setNames(c("from", "to")) %>%
    rbind(cfs[, c("Depth1","Depth2")]%>%unique()%>%setNames(c("from", "to"))) %>%
    rbind(cfs[, c("Depth2","Depth3")]%>%unique()%>%setNames(c("from","to")))
vertices = cfs%>%mutate(Depth0="Metabolism")%>%.[,c("Depth0","Depth3")]%>%unique()%>%
    pull(Depth0)%>%table()%>%as.data.frame()%>%setNames(c("name", "size")) %>%
    rbind(cfs[,c("Depth1","Depth3")]%>%unique()%>%pull(Depth1)%>%table()%>%
    as.data.frame()%>%setNames(c("name", "size"))) %>%
    rbind(cfs[,c("Depth2","Depth3")]%>%unique()%>%pull(Depth2)%>%table()%>%
    as.data.frame()%>%setNames(c("name", "size"))) %>%
    rbind(cfs%>%pull(Depth3)%>%unique()%>%table()%>%as.data.frame()%>%
    setNames(c("name","size"))) %>%
    mutate(size = size) %>%
    mutate(shortName=gsub("METABOLISM","M.",paste0(name,"(",size,")")))
vertices$shortName[!vertices$name%in%unique(cfs$Depth1)] = NA
mygraph = graph_from_data_frame(edges, vertices=vertices)       
p1 = ggraph(mygraph, layout = 'circlepack', weight=size) +
  geom_node_circle(aes(col=depth)) +
  geom_node_label(aes(label=shortName),repel=T,size=4,color="darkgreen") +
  theme_void() +
  scale_color_viridis() +
  theme(legend.position="FALSE") +
  ggtitle("Others")
rm(edges, vertices, mygraph)

## Zhou et al. 2020
cf5 = readRDS("../results/CellFie_Zhou2020/Summary/cf5.RDS")
head(cf5)
min(cf5$TaskScore)
unique(cf5$Depth3) %>% length()
# 176 metabolic tasks
unique(cf5$Depth2) %>% length()
# 68 metabolic subsystems
unique(cf5$Depth1) %>% length()
# 7 metabolic systems
edges = cf5%>%mutate(Depth0="Metabolism")%>%.[,c("Depth0","Depth1")]%>%
    unique()%>%setNames(c("from", "to")) %>%
    rbind(cf5[, c("Depth1","Depth2")]%>%unique()%>%setNames(c("from", "to"))) %>%
    rbind(cf5[, c("Depth2","Depth3")]%>%unique()%>%setNames(c("from","to")))
vertices = cf5%>%mutate(Depth0="Metabolism")%>%.[,c("Depth0","Depth3")]%>%
    unique()%>%pull(Depth0)%>%table()%>%as.data.frame()%>%setNames(c("name", "size")) %>%
    rbind(cf5[,c("Depth1","Depth3")]%>%unique()%>%pull(Depth1)%>%table()%>%
    as.data.frame()%>%setNames(c("name", "size"))) %>%
    rbind(cf5[,c("Depth2","Depth3")]%>%unique()%>%pull(Depth2)%>%table()%>%
    as.data.frame()%>%setNames(c("name", "size"))) %>%
    rbind(cf5%>%pull(Depth3)%>%unique()%>%table()%>%
    as.data.frame()%>%setNames(c("name","size"))) %>%
    mutate(size = size) %>%
    mutate(shortName=gsub("METABOLISM","M.",paste0(name,"(",size,")")))
vertices$shortName[!vertices$name%in%unique(cf5$Depth1)] = NA
mygraph = graph_from_data_frame(edges, vertices=vertices)       
p2 = ggraph(mygraph, layout = 'circlepack', weight=size) +
  geom_node_circle(aes(col=depth)) +
  geom_node_label(aes(label=shortName),repel=T,size=4,color="darkgreen") +
  theme_void() +
  scale_color_viridis() +
  theme(legend.position="FALSE") +
  ggtitle("Zhou2020")

pdf("../doc/FiguresTables/TaskSummary.pdf", h=5, w=9.7)
cowplot::plot_grid(p2,p1,ncol=2)
dev.off()
rm(edges, vertices, mygraph)
gc()



# Zhou total active task summary among tissues

```



# Active task percentage by cancers_circular barplot

## Depth 1
```{r}
plotActiveTaskPerc = function(dat=cf5, meta=Zhou2020_meta, system, title="Zhou2020_AllTasks") {
    temp = dat %>% filter(Depth1%in%system) %>% 
        left_join(meta%>%mutate(Sample=Sample_ID)) %>%
        group_by(Cancer_type) %>%
        summarize(ActiveTaskPerc = mean(BinaryTaskScore)) %>%
        arrange(ActiveTaskPerc)
    temp$Cancer_type = factor(temp$Cancer_type, levels=unique(temp$Cancer_type))
    P = ggplot(temp, aes(x=Cancer_type, y=ActiveTaskPerc, fill=Cancer_type)) +
    geom_bar(stat="identity") +
    theme_minimal() +
    ylim(0, 1) +
    theme(legend.position = "none",
            axis.text.y = element_blank(),
            axis.ticks = element_blank(),
            axis.title = element_blank(),
            # panel.grid = element_blank(),
            plot.margin = unit(c(-0,-0,-0,-0), "cm")) +
    coord_polar() +
    ggtitle(title) 
    system = ifelse(length(system)>1, "All_tasks", system)
    return(list(Perc=temp%>%setNames(c("Cancer_type",system)), plot=P))   
}

dir.create("../doc/FiguresTables/ActiveTaskPerc/Depth1/", 
    recursive=T, showWarnings=F)
# All samples:
## Zhou2020
P1 = plotActiveTaskPerc(dat=cf5, meta=Zhou2020_meta, 
    system=unique(cf5$Depth1), 
    title="    Zhou2020_AllTasks")
Ps = lapply(unique(cf5$Depth1), function(system) {
    plotActiveTaskPerc(dat=cf5, meta=Zhou2020_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cf5$Depth1))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth1/ActiveTaskPerc_Cancer_Zhou2020.pdf", h=10, w=12)
    cowplot::plot_grid(P1[[2]], Ps[[1]][[2]], Ps[[2]][[2]], Ps[[3]][[2]], 
        Ps[[4]][[2]], Ps[[5]][[2]], Ps[[6]][[2]], Ps[[7]][[2]], ncol=4)
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=list(P1[[1]],Ps[[1]][[1]],
    Ps[[2]][[1]],Ps[[3]][[1]],Ps[[4]][[1]],
    Ps[[5]][[1]],Ps[[6]][[1]],Ps[[7]][[1]])%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth1/ActiveTaskPerc_Zhou2020.xlsx"))
rm(P1, Ps)
gc()

## Hu2025
P1 = plotActiveTaskPerc(dat=cfs, meta=Hu2025_meta, 
    system=unique(cfs$Depth1), 
    title="    Others_AllTasks")
Ps = lapply(unique(cfs$Depth1), function(system) {
    plotActiveTaskPerc(dat=cfs, meta=Hu2025_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cfs$Depth1))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth1/ActiveTaskPerc_Cancer_Hu2025.pdf", h=10, w=12)
    cowplot::plot_grid(P1[[2]], Ps[[1]][[2]], Ps[[2]][[2]], Ps[[3]][[2]], 
        Ps[[4]][[2]], Ps[[5]][[2]], Ps[[6]][[2]], Ps[[7]][[2]], ncol=4)
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=list(P1[[1]],Ps[[1]][[1]],
    Ps[[2]][[1]],Ps[[3]][[1]],Ps[[4]][[1]],
    Ps[[5]][[1]],Ps[[6]][[1]],Ps[[7]][[1]])%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth1/ActiveTaskPerc_Hu2025.xlsx"))
rm(P1, Ps)
gc()

# Normal samples
## Zhou2020
P1 = plotActiveTaskPerc(dat=cf5[grep("_N",cf5$Group),], meta=Zhou2020_meta, 
    system=unique(cf5$Depth1), 
    title="    Zhou2020_AllTasks")
Ps = lapply(unique(cf5$Depth1), function(system) {
    plotActiveTaskPerc(dat=cf5[grep("_N",cf5$Group),], meta=Zhou2020_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cf5$Depth1))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth1/ActiveTaskPerc_Cancer_Zhou2020_NormalSamples.pdf", h=10, w=12)
    cowplot::plot_grid(P1[[2]], Ps[[1]][[2]], Ps[[2]][[2]], Ps[[3]][[2]], 
        Ps[[4]][[2]], Ps[[5]][[2]], Ps[[6]][[2]], Ps[[7]][[2]], ncol=4)
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=list(P1[[1]],Ps[[1]][[1]],
    Ps[[2]][[1]],Ps[[3]][[1]],Ps[[4]][[1]],
    Ps[[5]][[1]],Ps[[6]][[1]],Ps[[7]][[1]])%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth1/ActiveTaskPerc_Zhou2020_NormalSamples.xlsx"))
ActiveTaskPercentage_ZhouNormal=list(P1[[1]],Ps[[1]][[1]],
    Ps[[2]][[1]],Ps[[3]][[1]],Ps[[4]][[1]],
    Ps[[5]][[1]],Ps[[6]][[1]],Ps[[7]][[1]])%>%
    Reduce(dplyr::left_join,.)
rm(P1, Ps)
gc()


## Hu2025
P1 = plotActiveTaskPerc(dat=cfs[grep("_NAT|_Normal",cfs$Group),], meta=Hu2025_meta, 
    system=unique(cfs$Depth1), 
    title="    Others_AllTasks")
Ps = lapply(unique(cfs$Depth1), function(system) {
    plotActiveTaskPerc(dat=cfs[grep("_NAT|_Normal",cfs$Group),], meta=Hu2025_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cfs$Depth1))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth1/ActiveTaskPerc_Cancer_Hu2025_NormalSamples.pdf", h=10, w=12)
    cowplot::plot_grid(P1[[2]], Ps[[1]][[2]], Ps[[2]][[2]], Ps[[3]][[2]], 
        Ps[[4]][[2]], Ps[[5]][[2]], Ps[[6]][[2]], Ps[[7]][[2]], ncol=4)
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=list(P1[[1]],Ps[[1]][[1]],
    Ps[[2]][[1]],Ps[[3]][[1]],Ps[[4]][[1]],
    Ps[[5]][[1]],Ps[[6]][[1]],Ps[[7]][[1]])%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth1/ActiveTaskPerc_Hu2025_NormalSamples.xlsx"))
ActiveTaskPercentage_HuNormal = list(P1[[1]],Ps[[1]][[1]],
    Ps[[2]][[1]],Ps[[3]][[1]],Ps[[4]][[1]],
    Ps[[5]][[1]],Ps[[6]][[1]],Ps[[7]][[1]])%>%
    Reduce(dplyr::left_join,.)
rm(P1, Ps)
gc()


# Cancer samples
## Zhou2020
P1 = plotActiveTaskPerc(dat=cf5[grep("_T",cf5$Group),], meta=Zhou2020_meta, 
    system=unique(cf5$Depth1), 
    title="    Zhou2020_AllTasks")
Ps = lapply(unique(cf5$Depth1), function(system) {
    plotActiveTaskPerc(dat=cf5[grep("_T",cf5$Group),], meta=Zhou2020_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cf5$Depth1))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth1/ActiveTaskPerc_Cancer_Zhou2020_TumorSamples.pdf", h=10, w=12)
    cowplot::plot_grid(P1[[2]], Ps[[1]][[2]], Ps[[2]][[2]], Ps[[3]][[2]], 
        Ps[[4]][[2]], Ps[[5]][[2]], Ps[[6]][[2]], Ps[[7]][[2]], ncol=4)
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=list(P1[[1]],Ps[[1]][[1]],
    Ps[[2]][[1]],Ps[[3]][[1]],Ps[[4]][[1]],
    Ps[[5]][[1]],Ps[[6]][[1]],Ps[[7]][[1]])%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth1/ActiveTaskPerc_Zhou2020_TumorSamples.xlsx"))
ActiveTaskPercentage_ZhouTumor=list(P1[[1]],Ps[[1]][[1]],
    Ps[[2]][[1]],Ps[[3]][[1]],Ps[[4]][[1]],
    Ps[[5]][[1]],Ps[[6]][[1]],Ps[[7]][[1]])%>%
    Reduce(dplyr::left_join,.)
rm(P1, Ps)
gc()

## Hu2025
P1 = plotActiveTaskPerc(dat=cfs[grep("_Tumor",cfs$Group),], meta=Hu2025_meta, 
    system=unique(cfs$Depth1), 
    title="    Others_AllTasks")
Ps = lapply(unique(cfs$Depth1), function(system) {
    plotActiveTaskPerc(dat=cfs[grep("_Tumor",cfs$Group),], meta=Hu2025_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cfs$Depth1))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth1/ActiveTaskPerc_Cancer_Hu2025_TumorSamples.pdf", h=10, w=12)
    cowplot::plot_grid(P1[[2]], Ps[[1]][[2]], Ps[[2]][[2]], Ps[[3]][[2]], 
        Ps[[4]][[2]], Ps[[5]][[2]], Ps[[6]][[2]], Ps[[7]][[2]], ncol=4)
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=list(P1[[1]],Ps[[1]][[1]],
    Ps[[2]][[1]],Ps[[3]][[1]],Ps[[4]][[1]],
    Ps[[5]][[1]],Ps[[6]][[1]],Ps[[7]][[1]])%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth1/ActiveTaskPerc_Hu2025_TumorSamples.xlsx"))
ActiveTaskPercentage_HuTumor=list(P1[[1]],Ps[[1]][[1]],
    Ps[[2]][[1]],Ps[[3]][[1]],Ps[[4]][[1]],
    Ps[[5]][[1]],Ps[[6]][[1]],Ps[[7]][[1]])%>%
    Reduce(dplyr::left_join,.)
rm(P1, Ps)
gc()

openxlsx::write.xlsx(list(
    Zhou2020_Normal=ActiveTaskPercentage_ZhouNormal,
    Zhou2020_Tumor=ActiveTaskPercentage_ZhouTumor,
    Others_Normal=ActiveTaskPercentage_HuNormal,
    Others_Tumor=ActiveTaskPercentage_HuTumor), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth1/ActiveTaskPerc_all.xlsx"))
```

## Depth 2
```{r}
plotActiveTaskPerc2 = function(dat=cf5, meta=Zhou2020_meta, system, title="Zhou2020_AllTasks") {
    temp = dat %>% filter(Depth2%in%system) %>% 
        left_join(meta%>%mutate(Sample=Sample_ID)) %>%
        group_by(Cancer_type) %>%
        summarize(ActiveTaskPerc = mean(BinaryTaskScore)) %>%
        arrange(ActiveTaskPerc)
    temp$Cancer_type = factor(temp$Cancer_type, levels=unique(temp$Cancer_type))
    P = ggplot(temp, aes(x=Cancer_type, y=ActiveTaskPerc, fill=Cancer_type)) +
    geom_bar(stat="identity") +
    theme_minimal() +
    ylim(0, 1) +
    theme(legend.position = "none",
            axis.text.y = element_blank(),
            axis.ticks = element_blank(),
            axis.title = element_blank(),
            # panel.grid = element_blank(),
            plot.margin = unit(c(-0,-0,-0,-0), "cm")) +
    coord_polar() +
    ggtitle(title) 
    system = ifelse(length(system)>1, "All_tasks", system)
    return(list(Perc=temp%>%setNames(c("Cancer_type",system)), plot=P))   
}

dir.create("../doc/FiguresTables/ActiveTaskPerc/Depth2/", 
    recursive=T, showWarnings=F)
# All samples:
## Zhou2020
Ps = lapply(unique(cf5$Depth2), function(system) {
    plotActiveTaskPerc2(dat=cf5, meta=Zhou2020_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cf5$Depth2))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth2/ActiveTaskPerc_Cancer_Zhou2020_depth2.pdf", h=4, w=4)
    lapply(Ps, function(P) {
        show(P$plot)
    })
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=lapply(Ps, function(p) {
    p$Perc})%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth2/ActiveTaskPerc_Zhou2020_depth2.xlsx"))
rm(Ps)
gc()

## Hu2025
Ps = lapply(unique(cfs$Depth2), function(system) {
    plotActiveTaskPerc2(dat=cfs, meta=Hu2025_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cfs$Depth2))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth2/ActiveTaskPerc_Cancer_Hu2025_depth2.pdf", h=4, w=4)
    lapply(Ps, function(P) {
        show(P$plot)
    })
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=lapply(Ps, function(p) {
    p$Perc})%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth2/ActiveTaskPerc_Hu2025_depth2.xlsx"))
rm(Ps)
gc()

# Normal samples
## Zhou2020
Ps = lapply(unique(cf5$Depth2), function(system) {
    plotActiveTaskPerc2(dat=cf5[grep("_N",cf5$Group),], meta=Zhou2020_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cf5$Depth2))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth2/ActiveTaskPerc_Cancer_Zhou2020_NormalSamples_depth2.pdf", h=4, w=4)
    lapply(Ps, function(P) {
        show(P$plot)
    })
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=lapply(Ps, function(p) {
    p$Perc})%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth2/ActiveTaskPerc_Zhou2020_NormalSamples_depth2.xlsx"))
rm(Ps)
gc()

## Hu2025
Ps = lapply(unique(cfs$Depth2), function(system) {
    plotActiveTaskPerc2(dat=cfs[grep("_NAT|_Normal",cfs$Group),], meta=Hu2025_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cfs$Depth2))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth2/ActiveTaskPerc_Cancer_Hu2025_NormalSamples_depth2.pdf", h=4, w=4)
    lapply(Ps, function(P) {
        show(P$plot)
    })
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=lapply(Ps, function(p) {
    p$Perc})%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth2/ActiveTaskPerc_Hu2025_NormalSamples_depth2.xlsx"))
rm(Ps)
gc()


# Cancer samples
## Zhou2020
Ps = lapply(unique(cf5$Depth2), function(system) {
    plotActiveTaskPerc2(dat=cf5[grep("_T",cf5$Group),], meta=Zhou2020_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cf5$Depth2))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth2/ActiveTaskPerc_Cancer_Zhou2020_TumorSamples_depth2.pdf", h=4, w=4)
    lapply(Ps, function(P) {
        show(P$plot)
    })
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=lapply(Ps, function(p) {
    p$Perc})%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth2/ActiveTaskPerc_Zhou2020_TumorSamples_depth2.xlsx"))
rm(Ps)
gc()

## Hu2025
Ps = lapply(unique(cfs$Depth2), function(system) {
    plotActiveTaskPerc2(dat=cfs[grep("_Tumor",cfs$Group),], meta=Hu2025_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cfs$Depth2))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth2/ActiveTaskPerc_Cancer_Hu2025_TumorSamples_depth2.pdf", h=4, w=4)
    lapply(Ps, function(P) {
        show(P$plot)
    })
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=lapply(Ps, function(p) {
    p$Perc})%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth2/ActiveTaskPerc_Hu2025_TumorSamples_depth2.xlsx"))
rm(Ps)
gc()
```

## Depth 3
```{r}
plotActiveTaskPerc3 = function(dat=cf5, meta=Zhou2020_meta, system, title="Zhou2020_AllTasks") {
    temp = dat %>% filter(Depth3%in%system) %>% 
        left_join(meta%>%mutate(Sample=Sample_ID)) %>%
        group_by(Cancer_type) %>%
        summarize(ActiveTaskPerc = mean(BinaryTaskScore)) %>%
        arrange(ActiveTaskPerc)
    temp$Cancer_type = factor(temp$Cancer_type, levels=unique(temp$Cancer_type))
    P = ggplot(temp, aes(x=Cancer_type, y=ActiveTaskPerc, fill=Cancer_type)) +
    geom_bar(stat="identity") +
    theme_minimal() +
    ylim(0, 1) +
    theme(legend.position = "none",
            axis.text.y = element_blank(),
            axis.ticks = element_blank(),
            axis.title = element_blank(),
            # panel.grid = element_blank(),
            plot.margin = unit(c(-0,-0,-0,-0), "cm")) +
    coord_polar() +
    ggtitle(title) 
    system = ifelse(length(system)>1, "All_tasks", system)
    return(list(Perc=temp%>%setNames(c("Cancer_type",system)), plot=P))   
}

dir.create("../doc/FiguresTables/ActiveTaskPerc/Depth3/", 
    recursive=T, showWarnings=F)
# All samples:
## Zhou2020
Ps = lapply(unique(cf5$Depth3), function(system) {
    plotActiveTaskPerc3(dat=cf5, meta=Zhou2020_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cf5$Depth3))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth3/ActiveTaskPerc_Cancer_Zhou2020_depth3.pdf", h=4, w=4)
    lapply(Ps, function(P) {
        show(P$plot)
    })
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=lapply(Ps, function(p) {
    p$Perc})%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth3/ActiveTaskPerc_Zhou2020_depth3.xlsx"))
rm(Ps)
gc()

## Hu2025
Ps = lapply(unique(cfs$Depth3), function(system) {
    plotActiveTaskPerc3(dat=cfs, meta=Hu2025_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cfs$Depth3))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth3/ActiveTaskPerc_Cancer_Hu2025_depth3.pdf", h=4, w=4)
    lapply(Ps, function(P) {
        show(P$plot)
    })
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=lapply(Ps, function(p) {
    p$Perc})%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth3/ActiveTaskPerc_Hu2025_depth3.xlsx"))
rm(Ps)
gc()

# Normal samples
## Zhou2020
Ps = lapply(unique(cf5$Depth3), function(system) {
    plotActiveTaskPerc3(dat=cf5[grep("_N",cf5$Group),], meta=Zhou2020_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cf5$Depth3))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth3/ActiveTaskPerc_Cancer_Zhou2020_NormalSamples_depth3.pdf", h=4, w=4)
    lapply(Ps, function(P) {
        show(P$plot)
    })
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=lapply(Ps, function(p) {
    p$Perc})%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth3/ActiveTaskPerc_Zhou2020_NormalSamples_depth3.xlsx"))
rm(Ps)
gc()

## Hu2025
Ps = lapply(unique(cfs$Depth3), function(system) {
    plotActiveTaskPerc3(dat=cfs[grep("_NAT|_Normal",cfs$Group),], meta=Hu2025_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cfs$Depth3))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth3/ActiveTaskPerc_Cancer_Hu2025_NormalSamples_depth3.pdf", h=4, w=4)
    lapply(Ps, function(P) {
        show(P$plot)
    })
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=lapply(Ps, function(p) {
    p$Perc})%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth3/ActiveTaskPerc_Hu2025_NormalSamples_depth3.xlsx"))
rm(Ps)
gc()


# Cancer samples
## Zhou2020
Ps = lapply(unique(cf5$Depth3), function(system) {
    plotActiveTaskPerc3(dat=cf5[grep("_T",cf5$Group),], meta=Zhou2020_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cf5$Depth3))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth3/ActiveTaskPerc_Cancer_Zhou2020_TumorSamples_depth3.pdf", h=4, w=4)
    lapply(Ps, function(P) {
        show(P$plot)
    })
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=lapply(Ps, function(p) {
    p$Perc})%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth3/ActiveTaskPerc_Zhou2020_TumorSamples_depth3.xlsx"))
rm(Ps)
gc()

## Hu2025
Ps = lapply(unique(cfs$Depth3), function(system) {
    plotActiveTaskPerc3(dat=cfs[grep("_Tumor",cfs$Group),], meta=Hu2025_meta, 
        system=system, 
        title=paste0("    ",gsub("METABOLISM","M.",system)))
}) %>% setNames(unique(cfs$Depth3))
pdf("../doc/FiguresTables/ActiveTaskPerc/Depth3/ActiveTaskPerc_Cancer_Hu2025_TumorSamples_depth3.pdf", h=4, w=4)
    lapply(Ps, function(P) {
        show(P$plot)
    })
dev.off()
openxlsx::write.xlsx(list(ActiveTaskPercentage=lapply(Ps, function(p) {
    p$Perc})%>%
    Reduce(dplyr::left_join,.)), 
  file=paste0("../doc/FiguresTables/ActiveTaskPerc/Depth3/ActiveTaskPerc_Hu2025_TumorSamples_depth3.xlsx"))
rm(Ps)
gc()
```



# Upset plots

## Depth3
```{r}
deg_d3_Z = readRDS("../results/CellFie_Zhou2020/Limma/TaskDepth3/Upset_overlapping/Sig.Diff.features_list.RDS")
# Stop! Are you sure the cancer type names of Zhou2020 has been corrected or not? The original LUSC—>SCLC, LUSQ—>LUSC, ESCA—>ESCC!, if it has already been changed, the command line below should NOT be run:
# names(deg_d3_Z) = gsub("LUSC","SCLC",names(deg_d3_Z)) %>%
#     gsub("LUSQ","LUSC",.) %>% gsub("ESCA","ESCC",.)

deg_d3_H = readRDS("../results/CellFie_Hu2025/Limma/TaskDepth3/Upset_overlapping/Sig.Diff.features_list.RDS")

# The number of upregulated tasks shared by more than 5 cancer types in Zhou2020
sum((deg_d3_Z[grep("up", names(deg_d3_Z), value=TRUE)] %>%
    unlist() %>% table()) > 5)
# 8
# The number of upregulated tasks shared by more than 5 cancer types in Hou2025
sum((deg_d3_H[grep("up", names(deg_d3_H), value=TRUE)] %>%
    unlist() %>% table()) > 5)
# 27

# The number of downregulated tasks shared by more than 5 cancer types in Zhou2020
sum((deg_d3_Z[grep("down", names(deg_d3_Z), value=TRUE)] %>%
    unlist() %>% table()) > 5)
# 0
# The number of downregulated tasks shared by more than 5 cancer types in Hou2025
sum((deg_d3_H[grep("down", names(deg_d3_H), value=TRUE)] %>%
    unlist() %>% table()) > 5)
# 66

orders_Z = lapply(deg_d3_Z, function(d){
    length(d)
}) %>% unlist() %>% .[order(.)]
orders_H = lapply(deg_d3_H, function(d){
    length(d)
}) %>% unlist() %>% .[order(.)]

"../doc/FiguresTables/upset/depth3/" %>% dir.create(., recursive=T, showWarnings=F)
library(ComplexHeatmap)
# Zhou 2020
sig.feature.list_Z = deg_d3_Z[c(grep("up",names(orders_Z),value=T),grep("down",names(orders_Z),value=T))]
m = sig.feature.list_Z %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth3/Sig.Diff.features_upset_Zhou2020_depth3.pdf", h=6.3, w=20)
      print(UpSet(m,
              top_annotation = upset_top_annotation(m, add_numbers = TRUE),
              right_annotation = upset_right_annotation(m, add_numbers = TRUE),
              comb_order = rev(order(comb_size(m))),
              set_order = Orders,
              column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()

# Hu 2025
sig.feature.list_H = deg_d3_H[c(grep("up",names(orders_H),value=T),grep("down",names(orders_H),value=T))]
m = sig.feature.list_H %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth3/Sig.Diff.features_upset_Hu2025_depth3.pdf", h=5.7, w=18)
      print(UpSet(m,
              top_annotation = upset_top_annotation(m, add_numbers = TRUE),
              right_annotation = upset_right_annotation(m, add_numbers = TRUE),
              comb_order = rev(order(comb_size(m))),
              set_order = Orders,
              column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()

# up: Zhou2020 + Hu2025
sig.feature.list_up = c(
    deg_d3_H[grep("up",names(orders_H),value=T)]%>%setNames(paste0(names(.),"_Others")),
    deg_d3_Z[rev(grep("up",names(orders_Z),value=T))]%>%setNames(paste0(names(.),"_Zhou2020")))
m = sig.feature.list_up %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth3/Sig.Diff.features_upset_up_depth3.pdf", h=5.7, w=18)
      print(UpSet(m,
              top_annotation = upset_top_annotation(m, add_numbers = TRUE),
              right_annotation = upset_right_annotation(m, add_numbers = TRUE),
              comb_order = rev(order(comb_size(m))),
              set_order = Orders,
              column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()

# down: Zhou2020 + Hu2025
sig.feature.list_down = c(
    deg_d3_H[grep("down",names(orders_H),value=T)]%>%setNames(paste0(names(.),"_Others")),
    deg_d3_Z[rev(grep("down",names(orders_Z),value=T))]%>%setNames(paste0(names(.),"_Zhou2020")))
m = sig.feature.list_down %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth3/Sig.Diff.features_upset_down_depth3.pdf", h=5.7, w=18)
      print(UpSet(m,
              top_annotation = upset_top_annotation(m, add_numbers = TRUE),
              right_annotation = upset_right_annotation(m, add_numbers = TRUE),
              comb_order = rev(order(comb_size(m))),
              set_order = Orders,
              column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()


# up: Hu2025 
sig.feature.list_upH = c(
    deg_d3_H[grep("up",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")))
m = sig.feature.list_upH %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth3/Sig.Diff.features_upset_Hu2025_up_depth3.pdf", h=4, w=10)
      print(UpSet(m,
              top_annotation = upset_top_annotation(m, add_numbers = TRUE),
              right_annotation = upset_right_annotation(m, add_numbers = TRUE),
              comb_order = rev(order(comb_size(m))),
              set_order = Orders,
              column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()

# down: Zhou2020
sig.feature.list_downH = c(
    deg_d3_H[grep("down",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")))
m = sig.feature.list_downH %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth3/Sig.Diff.features_upset_Hu2025_down_depth3.pdf", h=4, w=10)
      print(UpSet(m,
              top_annotation = upset_top_annotation(m, add_numbers = TRUE),
              right_annotation = upset_right_annotation(m, add_numbers = TRUE),
              comb_order = rev(order(comb_size(m))),
              set_order = Orders,
              column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()


# Pan-cancer upregulated tasks
pancancerUP = capture.output(
    cat("Pan cancer task identified by Hu2025: "),
    deg_d3_H[grep("up",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
        unlist() %>% table() %>% .[.>0] %>% names(),
    cat(""),
    cat(""),
    cat("Pan cancer tasks identified by Zhou2020: "),
    deg_d3_Z[grep("up",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
        unlist() %>% table() %>% .[.>0] %>% names()
)
writeLines(pancancerUP, "../doc/FiguresTables/upset/depth3/Pancancer_upregulatedTasks_depth3.txt")

# Pan-cancer upregulated tasks
pancancerDN = capture.output(
    cat("Pan cancer task identified by Hu2025: "),
    deg_d3_H[grep("down",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
        unlist() %>% table() %>% .[.>0] %>% names(),
    cat(""),
    cat(""),
    cat("Pan cancer tasks identified by Zhou2020: "),
    deg_d3_Z[grep("down",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
        unlist() %>% table() %>% .[.>0] %>% names()
)
writeLines(pancancerDN, "../doc/FiguresTables/upset/depth3/Pancancer_downregulatedTasks_depth3.txt")

```

## Depth2
```{r}
deg_d2_Z = readRDS("../results/CellFie_Zhou2020/Limma/TaskDepth2/Upset_overlapping/Sig.Diff.features_list.RDS")
# Stop! Are you sure the cancer type names of Zhou2020 has been corrected or not? The original LUSC—>SCLC, LUSQ—>LUSC, ESCA—>ESCC!, if it has already been changed, the command line below should NOT be run:
# names(deg_d2_Z) = gsub("LUSC","SCLC",names(deg_d2_Z)) %>%
#     gsub("LUSQ","LUSC",.) %>% gsub("ESCA","ESCC",.)

deg_d2_H = readRDS("../results/CellFie_Hu2025/Limma/TaskDepth2/Upset_overlapping/Sig.Diff.features_list.RDS")
deg_d2_H

orders_Z = lapply(deg_d2_Z, function(d){
    length(d)
}) %>% unlist() %>% .[order(.)]
orders_H = lapply(deg_d2_H, function(d){
    length(d)
}) %>% unlist() %>% .[order(.)]

"../doc/FiguresTables/upset/depth2/" %>% dir.create(., recursive=T, showWarnings=F)
library(ComplexHeatmap)
# Zhou 2020
sig.feature.list_Z = deg_d2_Z[c(grep("up",names(orders_Z),value=T),grep("down",names(orders_Z),value=T))]
m = sig.feature.list_Z %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth2/Sig.Diff.features_upset_Zhou2020_depth2.pdf", h=6.3, w=15)
    print(UpSet(m,
        top_annotation = upset_top_annotation(m, add_numbers = TRUE),
        right_annotation = upset_right_annotation(m, add_numbers = TRUE),
        comb_order = rev(order(comb_size(m))),
        set_order = Orders,
        column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()

# Hu 2025
sig.feature.list_H = deg_d2_H[c(grep("up",names(orders_H),value=T),grep("down",names(orders_H),value=T))]
m = sig.feature.list_H %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth2/Sig.Diff.features_upset_Hu2025_depth2.pdf", h=5.7, w=14)
    print(UpSet(m,
        top_annotation = upset_top_annotation(m, add_numbers = TRUE),
        right_annotation = upset_right_annotation(m, add_numbers = TRUE),
        comb_order = rev(order(comb_size(m))),
        set_order = Orders,
        column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()

# up: Zhou2020 + Hu2025
sig.feature.list_up = c(
    deg_d2_H[grep("up",names(orders_H),value=T)]%>%setNames(paste0(names(.),"_Others")),
    deg_d2_Z[rev(grep("up",names(orders_Z),value=T))]%>%setNames(paste0(names(.),"_Zhou2020")))
m = sig.feature.list_up %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth2/Sig.Diff.features_upset_up_depth2.pdf", h=5.7, w=14)
    print(UpSet(m,
        top_annotation = upset_top_annotation(m, add_numbers = TRUE),
        right_annotation = upset_right_annotation(m, add_numbers = TRUE),
        comb_order = rev(order(comb_size(m))),
        set_order = Orders,
        column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()

# down: Zhou2020 + Hu2025
sig.feature.list_down = c(
    deg_d2_H[grep("down",names(orders_H),value=T)]%>%setNames(paste0(names(.),"_Others")),
    deg_d2_Z[rev(grep("down",names(orders_Z),value=T))]%>%setNames(paste0(names(.),"_Zhou2020")))
m = sig.feature.list_down %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth2/Sig.Diff.features_upset_down_depth2.pdf", h=5.7, w=14)
    print(UpSet(m,
        top_annotation = upset_top_annotation(m, add_numbers = TRUE),
        right_annotation = upset_right_annotation(m, add_numbers = TRUE),
        comb_order = rev(order(comb_size(m))),
        set_order = Orders,
        column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()


# up: Hu2025 
sig.feature.list_upH = c(
    deg_d2_H[grep("up",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")))
m = sig.feature.list_upH %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth2/Sig.Diff.features_upset_Hu2025_up_depth2.pdf", h=4, w=8)
    print(UpSet(m,
        top_annotation = upset_top_annotation(m, add_numbers = TRUE),
        right_annotation = upset_right_annotation(m, add_numbers = TRUE),
        comb_order = rev(order(comb_size(m))),
        set_order = Orders,
        column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()

# down: Zhou2020
sig.feature.list_downH = c(
    deg_d2_H[grep("down",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")))
m = sig.feature.list_downH %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth2/Sig.Diff.features_upset_Hu2025_down_depth2.pdf", h=4, w=8)
    print(UpSet(m,
        top_annotation = upset_top_annotation(m, add_numbers = TRUE),
        right_annotation = upset_right_annotation(m, add_numbers = TRUE),
        comb_order = rev(order(comb_size(m))),
        set_order = Orders,
        column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()


# Pan-cancer upregulated tasks
pancancerUP = capture.output(
    cat("Pan cancer task identified by Hu2025: "),
    deg_d2_H[grep("up",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
        unlist() %>% table() %>% .[.>0] %>% names(),
    cat(""),
    cat(""),
    cat("Pan cancer tasks identified by Zhou2020: "),
    deg_d2_Z[grep("up",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
        unlist() %>% table() %>% .[.>0] %>% names()
)
writeLines(pancancerUP, "../doc/FiguresTables/upset/depth2/Pancancer_upregulatedTasks_depth2.txt")

# Pan-cancer upregulated tasks
pancancerDN = capture.output(
    cat("Pan cancer task identified by Hu2025: "),
    deg_d2_H[grep("down",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
        unlist() %>% table() %>% .[.>0] %>% names(),
    cat(""),
    cat(""),
    cat("Pan cancer tasks identified by Zhou2020: "),
    deg_d2_Z[grep("down",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
        unlist() %>% table() %>% .[.>0] %>% names()
)
writeLines(pancancerDN, "../doc/FiguresTables/upset/depth2/Pancancer_downregulatedTasks_depth2.txt")
```

## Depth1
```{r}
deg_d1_Z = readRDS("../results/CellFie_Zhou2020/Limma/TaskDepth1/Upset_overlapping/Sig.Diff.features_list.RDS")
# Stop! Are you sure the cancer type names of Zhou2020 has been corrected or not? The original LUSC—>SCLC, LUSQ—>LUSC, ESCA—>ESCC!, if it has already been changed, the command line below should NOT be run:
# names(deg_d1_Z) = gsub("LUSC","SCLC",names(deg_d1_Z)) %>%
#     gsub("LUSQ","LUSC",.) %>% gsub("ESCA","ESCC",.)

deg_d1_H = readRDS("../results/CellFie_Hu2025/Limma/TaskDepth1/Upset_overlapping/Sig.Diff.features_list.RDS")
deg_d1_H

orders_Z = lapply(deg_d1_Z, function(d){
    length(d)
}) %>% unlist() %>% .[order(.)]
orders_H = lapply(deg_d1_H, function(d){
    length(d)
}) %>% unlist() %>% .[order(.)]

"../doc/FiguresTables/upset/depth1/" %>% dir.create(., recursive=T, showWarnings=F)
library(ComplexHeatmap)
# Zhou 2020
sig.feature.list_Z = deg_d1_Z[c(grep("up",names(orders_Z),value=T),grep("down",names(orders_Z),value=T))]
m = sig.feature.list_Z %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth1/Sig.Diff.features_upset_Zhou2020_depth1.pdf", h=6.3, w=4)
    print(UpSet(m,
        top_annotation = upset_top_annotation(m, add_numbers = TRUE),
        right_annotation = upset_right_annotation(m, add_numbers = TRUE),
        comb_order = rev(order(comb_size(m))),
        set_order = Orders,
        column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()

# Hu 2025
sig.feature.list_H = deg_d1_H[c(grep("up",names(orders_H),value=T),grep("down",names(orders_H),value=T))]
m = sig.feature.list_H %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth1/Sig.Diff.features_upset_Hu2025_depth1.pdf", h=5.7, w=4)
      print(UpSet(m,
              top_annotation = upset_top_annotation(m, add_numbers = TRUE),
              right_annotation = upset_right_annotation(m, add_numbers = TRUE),
              comb_order = rev(order(comb_size(m))),
              set_order = Orders,
              column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()

# up: Zhou2020 + Hu2025
sig.feature.list_up = c(
    deg_d1_H[grep("up",names(orders_H),value=T)]%>%setNames(paste0(names(.),"_Others")),
    deg_d1_Z[rev(grep("up",names(orders_Z),value=T))]%>%setNames(paste0(names(.),"_Zhou2020")))
m = sig.feature.list_up %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth1/Sig.Diff.features_upset_up_depth1.pdf", h=5.7, w=4)
      print(UpSet(m,
              top_annotation = upset_top_annotation(m, add_numbers = TRUE),
              right_annotation = upset_right_annotation(m, add_numbers = TRUE),
              comb_order = rev(order(comb_size(m))),
              set_order = Orders,
              column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()

# down: Zhou2020 + Hu2025
sig.feature.list_down = c(
    deg_d1_H[grep("down",names(orders_H),value=T)]%>%setNames(paste0(names(.),"_Others")),
    deg_d1_Z[rev(grep("down",names(orders_Z),value=T))]%>%setNames(paste0(names(.),"_Zhou2020")))
m = sig.feature.list_down %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth1/Sig.Diff.features_upset_down_depth1.pdf", h=5.7, w=4)
      print(UpSet(m,
              top_annotation = upset_top_annotation(m, add_numbers = TRUE),
              right_annotation = upset_right_annotation(m, add_numbers = TRUE),
              comb_order = rev(order(comb_size(m))),
              set_order = Orders,
              column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()


# up: Hu2025 
sig.feature.list_upH = c(
    deg_d1_H[grep("up",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")))
m = sig.feature.list_upH %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth1/Sig.Diff.features_upset_Hu2025_up_depth1.pdf", h=4, w=3)
      print(UpSet(m,
              top_annotation = upset_top_annotation(m, add_numbers = TRUE),
              right_annotation = upset_right_annotation(m, add_numbers = TRUE),
              comb_order = rev(order(comb_size(m))),
              set_order = Orders,
              column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()

# down: Zhou2020
sig.feature.list_downH = c(
    deg_d1_H[grep("down",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")))
m = sig.feature.list_downH %>% make_comb_mat()
Orders = names(m)
pdf("../doc/FiguresTables/upset/depth1/Sig.Diff.features_upset_Hu2025_down_depth1.pdf", h=4, w=3)
      print(UpSet(m,
              top_annotation = upset_top_annotation(m, add_numbers = TRUE),
              right_annotation = upset_right_annotation(m, add_numbers = TRUE),
              comb_order = rev(order(comb_size(m))),
              set_order = Orders,
              column_title = paste0("Significant features (adj.P.Val<0.05)")))
dev.off()


# Pan-cancer upregulated tasks
pancancerUP = capture.output(
    cat("Pan cancer task identified by Hu2025: "),
    deg_d1_H[grep("up",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
        unlist() %>% table() %>% .[.>0] %>% names(),
    cat(""),
    cat(""),
    cat("Pan cancer tasks identified by Zhou2020: "),
    deg_d1_Z[grep("up",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
        unlist() %>% table() %>% .[.>0] %>% names()
)
writeLines(pancancerUP, "../doc/FiguresTables/upset/depth1/Pancancer_upregulatedTasks_depth1.txt")

# Pan-cancer downregulated tasks
pancancerDN = capture.output(
    cat("Pan cancer task identified by Hu2025: "),
    deg_d1_H[grep("down",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
        unlist() %>% table() %>% .[.>0] %>% names(),
    cat(""),
    cat(""),
    cat("Pan cancer tasks identified by Zhou2020: "),
    deg_d1_Z[grep("down",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
        unlist() %>% table() %>% .[.>0] %>% names()
)
writeLines(pancancerDN, "../doc/FiguresTables/upset/depth1/Pancancer_downregulatedTasks_depth1.txt")

```


# Barplot
```{r}
# Hu2025
temp = sapply(deg_d3_H, function(H) {
    length(H)
})
B3_H = data.frame(Cancer_type = names(temp), 
    Regulation = names(temp),
    Count=as.numeric(temp)) %>%
    mutate(Cancer_type=gsub("_.*","",Cancer_type)) %>%
    mutate(Regulation=gsub(".*_","",Regulation)) %>%
    mutate(Count = ifelse(Regulation=="up",Count,-Count)) %>%
    arrange(.,Count) %>%
    mutate(Cancer_type=factor(Cancer_type,levels=unique(Cancer_type))) %>%
    ggplot(., aes(x=Cancer_type, y=Count, fill=Regulation)) +
    geom_col() +
    scale_fill_manual(values=c("up"="firebrick","down"="steelblue")) +
    # scale_fill_manual(values=c("up"="magenta","down"="lightblue")) +
    geom_hline(yintercept=0, color="black") +
    labs(title = "Sig. diff. tasks_others",
         x="Cancer_type", y="Count")+
    theme_minimal() +
    theme(legend.position="none",
          axis.text.x = element_text(angle = 90, hjust=1, vjust=0))
rm(temp)

temp = sapply(deg_d2_H, function(H) {
    length(H)
})
B2_H = data.frame(Cancer_type = names(temp), 
    Regulation = names(temp),
    Count=as.numeric(temp)) %>%
    mutate(Cancer_type=gsub("_.*","",Cancer_type)) %>%
    mutate(Regulation=gsub(".*_","",Regulation)) %>%
    mutate(Count = ifelse(Regulation=="up",Count,-Count)) %>%
    arrange(.,Count) %>%
    mutate(Cancer_type=factor(Cancer_type,levels=unique(Cancer_type))) %>%
    ggplot(., aes(x=Cancer_type, y=Count, fill=Regulation)) +
    geom_col() +
    scale_fill_manual(values=c("up"="firebrick","down"="steelblue")) +
    # scale_fill_manual(values=c("up"="magenta","down"="lightblue")) +
    geom_hline(yintercept=0, color="black") +
    labs(title = "Sig. diff. subsystems",
         x="Cancer_type", y="Count")+
    theme_minimal() +
    theme(legend.position="none",
          axis.text.x = element_text(angle = 90, hjust=1, vjust=0))
rm(temp)

temp = sapply(deg_d1_H, function(H) {
    length(H)
})
B1_H = data.frame(Cancer_type = names(temp), 
    Regulation = names(temp),
    Count=as.numeric(temp)) %>%
    mutate(Cancer_type=gsub("_.*","",Cancer_type)) %>%
    mutate(Regulation=gsub(".*_","",Regulation)) %>%
    mutate(Count = ifelse(Regulation=="up",Count,-Count)) %>%
    arrange(.,Count) %>%
    mutate(Cancer_type=factor(Cancer_type,levels=unique(Cancer_type))) %>%
    ggplot(., aes(x=Cancer_type, y=Count, fill=Regulation)) +
    geom_col() +
    scale_fill_manual(values=c("up"="firebrick","down"="steelblue")) +
    # scale_fill_manual(values=c("up"="magenta","down"="lightblue")) +
    geom_hline(yintercept=0, color="black") +
    labs(title = "Sig. diff. systems",
         x="Cancer_type", y="Count")+
    theme_minimal() +
    theme(legend.position="none",
          axis.text.x = element_text(angle = 90, hjust=1, vjust=0))
rm(temp)


# Zhou2020
temp = sapply(deg_d3_Z, function(H) {
    length(H)
})
B3_Z = data.frame(Cancer_type = names(temp), 
    Regulation = names(temp),
    Count=as.numeric(temp)) %>%
    mutate(Cancer_type=gsub("_.*","",Cancer_type)) %>%
    mutate(Regulation=gsub(".*_","",Regulation)) %>%
    mutate(Count = ifelse(Regulation=="up",Count,-Count)) %>%
    arrange(.,Count) %>%
    mutate(Cancer_type=factor(Cancer_type,levels=unique(Cancer_type))) %>%
    ggplot(., aes(x=Cancer_type, y=Count, fill=Regulation)) +
    geom_col() +
    scale_fill_manual(values=c("up"="firebrick","down"="steelblue")) +
    # scale_fill_manual(values=c("up"="magenta","down"="lightblue")) +
    geom_hline(yintercept=0, color="black") +
    labs(title = "Sig. diff. tasks_Zhou2020",
         x="Cancer_type", y="Count")+
    theme_minimal() +
    theme(legend.position="none",
          axis.text.x = element_text(angle = 90, hjust=1, vjust=0))
rm(temp)

temp = sapply(deg_d2_Z, function(H) {
    length(H)
})
B2_Z = data.frame(Cancer_type = names(temp), 
    Regulation = names(temp),
    Count=as.numeric(temp)) %>%
    mutate(Cancer_type=gsub("_.*","",Cancer_type)) %>%
    mutate(Regulation=gsub(".*_","",Regulation)) %>%
    mutate(Count = ifelse(Regulation=="up",Count,-Count)) %>%
    arrange(.,Count) %>%
    mutate(Cancer_type=factor(Cancer_type,levels=unique(Cancer_type))) %>%
    ggplot(., aes(x=Cancer_type, y=Count, fill=Regulation)) +
    geom_col() +
    scale_fill_manual(values=c("up"="firebrick","down"="steelblue")) +
    # scale_fill_manual(values=c("up"="magenta","down"="lightblue")) +
    geom_hline(yintercept=0, color="black") +
    labs(title = "Sig. diff. subsystems",
         x="Cancer_type", y="Count")+
    theme_minimal() +
    theme(legend.position="none",
          axis.text.x = element_text(angle = 90, hjust=1, vjust=0))
rm(temp)

temp = sapply(deg_d1_Z, function(H) {
    length(H)
})
B1_Z = data.frame(Cancer_type = names(temp), 
    Regulation = names(temp),
    Count=as.numeric(temp)) %>%
    mutate(Cancer_type=gsub("_.*","",Cancer_type)) %>%
    mutate(Regulation=gsub(".*_","",Regulation)) %>%
    mutate(Count = ifelse(Regulation=="up",Count,-Count)) %>%
    arrange(.,Count) %>%
    mutate(Cancer_type=factor(Cancer_type,levels=unique(Cancer_type))) %>%
    ggplot(., aes(x=Cancer_type, y=Count, fill=Regulation)) +
    geom_col() +
    scale_fill_manual(values=c("up"="firebrick","down"="steelblue")) +
    # scale_fill_manual(values=c("up"="magenta","down"="lightblue")) +
    geom_hline(yintercept=0, color="black") +
    labs(title = "Sig. diff. systems",
         x="Cancer_type", y="Count")+
    theme_minimal() +
    theme(legend.position="none",
          axis.text.x = element_text(angle = 90, hjust=1, vjust=0))
rm(temp)


library(patchwork)
dir.create("../doc/FiguresTables/Barplot/", 
    recursive=T, showWarnings=F)
pdf("../doc/FiguresTables/Barplot/SigDiffMetaProcesses_Hu2025.pdf", w=7.5, h=3.0)
# pdf("../doc/FiguresTables/Barplot/SigDiffMetaProcesses.pdf", w=7, h=6)
    # left_column = B3_H/B2_H/B1_H
    # right_column = B3_Z/B2_Z/B1_Z
    # (left_column|right_column) +
    #     plot_annotation(tag_levels = 'A')
    top_rows = B3_H|B2_H|B1_H
    (top_rows) 
dev.off()

pdf("../doc/FiguresTables/Barplot/SigDiffMetaProcesses_Zhou2020.pdf", w=7.5, h=3.0)
# pdf("../doc/FiguresTables/Barplot/SigDiffMetaProcesses.pdf", w=7, h=6)
    # left_column = B3_H/B2_H/B1_H
    # right_column = B3_Z/B2_Z/B1_Z
    # (left_column|right_column) +
    #     plot_annotation(tag_levels = 'A')
    top_rows = B3_Z|B2_Z|B1_Z
    (top_rows) 
dev.off()

pdf("../doc/FiguresTables/Barplot/SigDiffMetaProcesses2.pdf", w=9, h=4)
    top_row = B3_H|B2_H|B1_H
    lower_row = B3_Z|B2_Z|B1_Z
    (top_row/lower_row) 
dev.off()

pdf("../doc/FiguresTables/Barplot/SigDiffMetaProcesses_all.pdf", w=7, h=6)
    left_column = B3_H/B2_H/B1_H
    right_column = B3_Z/B2_Z/B1_Z
    (left_column|right_column) +
        plot_annotation(tag_levels = 'A')
dev.off()
rm(B3_H, B2_H, B1_H, B3_Z, B2_Z, B1_Z)
gc()

```


# Dotplot of pan-cancer metabolic processes

## Depth3
```{r}
library(dplyr)
makeLimmaDotplot = function(rlst_list, ups, downs, 
    adj.P.Val.cutoff, w=15, h=30, OUTDIRV, 
    Depth="Depth3", lowfigure=F,
    title = "Pan-cancer tasks", suffix="Hu2025",
    taskHierarchy=cfs, nested=T){
    rlst_list$OV$rslt_of_interest %>% colnames()
    taskHierarchy$Task = taskHierarchy[, Depth]
    rlst_long = lapply(seq_along(rlst_list), function(i) {
      rlst = rlst_list[[i]]$rslt_of_interest
      rlst$Cancer = names(rlst_list)[i]
      rlst$Regulation = ifelse(rlst$logFC>0&rlst$adj.P.Val<adj.P.Val.cutoff, "Up", 
        ifelse (rlst$logFC<0&rlst$adj.P.Val<adj.P.Val.cutoff, "Down", "NonSig"))
      rlst %>% tibble::rownames_to_column(var="Task") %>% 
        filter(Task %in% c(ups, downs)) %>%
        filter(adj.P.Val<adj.P.Val.cutoff)
    }) %>% Reduce(rbind, .) %>%
    left_join(taskHierarchy) %>%
    mutate(logFC=round(logFC, 2)) %>%
    mutate(Depth2 = stringr::str_sub(Depth2, 1, 3)) %>%
    mutate(Depth1 = stringr::str_sub(Depth1, 1, 3))

    rlst_long$Task = factor(rlst_long$Task,
      levels=rev(c(ups, downs) %>% unique()))
    rlst_long$t[rlst_long$t>5] = 5
    rlst_long$t[rlst_long$t< -5] = -5
    # table(rlst_long$Cancer)
    library(ggplot2)
    library(ggh4x) 
    p = ggplot(data=rlst_long, mapping=aes_string(x="Cancer", y="Task")) +
        geom_point(aes(size=adj.P.Val, shape=Regulation, color=logFC))
    if (nested) p = p+facet_nested(Depth1~., scales = "free_y", space = "free")
    p = p + scale_shape_manual(values = c("Up" = 18, "Down" = 20)) +
        ggtitle(title) +
        ylab("") + # labs(fill="adj.P.Val") +
        scale_color_gradient2(low="blue", mid="snow2", high="red", midpoint=0) +
        theme_bw(base_size = 15) + 
        scale_y_discrete(labels = function(x) str_wrap(x, width = 1000)) +
        scale_size(trans="reverse") +
        guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
        theme(
            axis.title = element_blank(),
            strip.text.y = element_text(size = 10, angle = 90, hjust=0.5),
            axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) +
        labs(caption=paste0("adj.P.ValCutoff = ",adj.P.Val.cutoff,",pAdjustMethod = BH"))

    if (lowfigure) {
      p = p + theme(
              axis.title = element_blank(),
              legend.position = "bottom",
              legend.box = "vertical",           # Stack legends vertically
              legend.box.just = "left",        # Center the whole legend box
              legend.text = element_text(size = 7),
              legend.title = element_text(size = 11),
              legend.spacing.y = unit(0, "cm") # Reduce spacing between legend rows
            ) +
            guides(
              fill  = guide_legend(ncol=1, title.position = "top", byrow = F),
              shape = guide_legend(title.position = "left"),
              size  = guide_legend(title.position = "left")
            )
    }

    pdf(paste0(OUTDIRV,"/Dotplot_all_adj.P.ValCutoff",adj.P.Val.cutoff,"_",suffix,".pdf"),w=w,h=h)
    show(p)
    dev.off()
    return(p)
}

# Hu 2025
rlst_list3 = readRDS("../results/CellFie_Hu2025/Limma/TaskDepth3/result3.RDS")$rlst_list
ups = deg_d3_H[grep("up",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>0] %>% names()
downs = deg_d3_H[grep("down",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>0] %>% names()
OUTDIRV =  "../doc/FiguresTables/Dotplot/depth3/"
dir.create(OUTDIRV, recursive=T, showWarnings=F)
adj.P.Val.cutoff = 0.05
D3_H = makeLimmaDotplot(rlst_list=rlst_list3, ups, downs, 
  adj.P.Val.cutoff, w=12.5, h=23, OUTDIRV, 
  title="Tasks_others", suffix="Hu2025",
  taskHierarchy=cfs, nested=T)
rm(ups, downs, OUTDIRV)
gc()

# Zhou 2020
rlst_list3 = readRDS("../results/CellFie_Zhou2020/Limma/TaskDepth3/result.RDS")
ups = deg_d3_Z[grep("up",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>0] %>% names()
downs = deg_d3_Z[grep("down",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>0] %>% names()
OUTDIRV =  "../doc/FiguresTables/Dotplot/depth3/"
adj.P.Val.cutoff = 0.05
D3_Z = makeLimmaDotplot(rlst_list=rlst_list3, ups, downs, 
  adj.P.Val.cutoff, w=12.5, h=29, OUTDIRV, 
  title="Tasks_Zhou2020", suffix="Zhou2020",
  lowfigure=T, taskHierarchy=cf5, nested=T)
rm(ups, downs, OUTDIRV)
gc()


# Hu 2025
rlst_list3 = readRDS("../results/CellFie_Hu2025/Limma/TaskDepth3/result3.RDS")$rlst_list
ups = deg_d3_H[grep("up",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>7] %>% names()
downs = deg_d3_H[grep("down",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>7] %>% names()
OUTDIRV =  "../doc/FiguresTables/Dotplot/depth3/"
dir.create(OUTDIRV, recursive=T, showWarnings=F)
adj.P.Val.cutoff = 0.05
D3_H = makeLimmaDotplot(rlst_list=rlst_list3, ups, downs, 
  adj.P.Val.cutoff, w=9, h=6, OUTDIRV, 
  title="Tasks_others", suffix="Hu2025_tops",
  taskHierarchy=cfs, nested=T)
rm(ups, downs, OUTDIRV)
gc()

# Zhou 2020
rlst_list3 = readRDS("../results/CellFie_Zhou2020/Limma/TaskDepth3/result.RDS")
ups = deg_d3_Z[grep("up",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>4] %>% names()
downs = deg_d3_Z[grep("down",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>4] %>% names()
OUTDIRV =  "../doc/FiguresTables/Dotplot/depth3/"
adj.P.Val.cutoff = 0.05
D3_Z = makeLimmaDotplot(rlst_list=rlst_list3, ups, downs, 
  adj.P.Val.cutoff, w=10.5, h=7.5, OUTDIRV, 
  title="Tasks_Zhou2020", suffix="Zhou2020_tops",
  lowfigure=T, taskHierarchy=cf5, nested=T)
# rm(ups, downs, OUTDIRV)
gc()
```

## Depth2
```{r}
# Hu 2025
rlst_list2 = readRDS("../results/CellFie_Hu2025/Limma/TaskDepth2/result2.RDS")$rlst_list
ups = deg_d2_H[grep("up",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>0] %>% names()
downs = deg_d2_H[grep("down",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>0] %>% names()
OUTDIRV =  "../doc/FiguresTables/Dotplot/depth2/"
dir.create(OUTDIRV, recursive=T, showWarnings=F)
adj.P.Val.cutoff = 0.05
D2_H = makeLimmaDotplot(rlst_list=rlst_list2, ups, downs, 
  adj.P.Val.cutoff, w=8.5, h=12, OUTDIRV, Depth="Depth2",
  title="Subsystems_Others", suffix="Hu2025",
  taskHierarchy=cfs, nested=T)
rm(ups, downs, OUTDIRV)
gc()

# Zhou 2020
rlst_list2 = readRDS("../results/CellFie_Zhou2020/Limma/TaskDepth2/result.RDS")
ups = deg_d2_Z[grep("up",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>0] %>% names()
downs = deg_d2_Z[grep("down",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>0] %>% names()
adj.P.Val.cutoff = 0.05
OUTDIRV =  "../doc/FiguresTables/Dotplot/depth2/"
D2_Z = makeLimmaDotplot(rlst_list=rlst_list2, ups, downs, 
  adj.P.Val.cutoff, w=7.5, h=15, OUTDIRV, Depth="Depth2",
  title="Subsystems_Zhou2020", suffix="Zhou2020",
  lowfigure=T, taskHierarchy=cf5, nested=T)
rm(ups, downs, OUTDIRV)
gc()

```

## Depth1
```{r}
# Hu 2025
rlst_list1 = readRDS("../results/CellFie_Hu2025/Limma/TaskDepth1/result1.RDS")$rlst_list
ups = deg_d1_H[grep("up",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>0] %>% names()
downs = deg_d1_H[grep("down",names(orders_H),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>0] %>% names()
OUTDIRV =  "../doc/FiguresTables/Dotplot/depth1/"
dir.create(OUTDIRV, recursive=T, showWarnings=F)
adj.P.Val.cutoff = 0.05
D1_H = makeLimmaDotplot(rlst_list=rlst_list1, ups, downs, 
  adj.P.Val.cutoff, w=6.3, h=4.7, OUTDIRV, Depth="Depth1", 
  title="Systems_others", suffix="Hu2025",
  lowfigure=TRUE, taskHierarchy=cfs, nested=F)
rm(ups, downs, OUTDIRV)
gc()

# Zhou2020
rlst_list1 = readRDS("../results/CellFie_Zhou2020/Limma/TaskDepth1/result.RDS")
ups = deg_d1_Z[grep("up",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>0] %>% names()
downs = deg_d1_Z[grep("down",names(orders_Z),value=T)]%>%setNames(paste0(names(.),"")) %>%
    unlist() %>% table() %>% .[.>0] %>% names()
OUTDIRV =  "../doc/FiguresTables/Dotplot/depth1/"
adj.P.Val.cutoff = 0.05
D1_Z = makeLimmaDotplot(rlst_list=rlst_list1, ups, downs, 
  adj.P.Val.cutoff, w=6.3, h=4.7, OUTDIRV, Depth="Depth1",
  title="Systems_Zhou2020", suffix="Zhou2020",
  lowfigure=TRUE, taskHierarchy=cf5, nested=F)
rm(ups, downs, OUTDIRV)
gc()

```

```{r}
library(patchwork)
pdf("../doc/FiguresTables/Dotplot/PanCancerMetabolicProcesses_combined.pdf",h=24,w=20)
left_column = D3_H/D3_Z/D1_Z/D1_H +
    plot_layout(heights=c(3,2,0.75,0.78))
right_column = D2_H/D2_Z + 
    plot_layout(heights=c(3.7,1.96))
(left_column | right_column +
    plot_layout(widths=c(1.5,1.5))) +
    plot_annotation(tag_levels = 'A')
dev.off()

```


# Normal vs differential 

## Depth3

### Task activity
```{r}
# Zhou2020
nested = T
p = cf5 %>% left_join(openxlsx::read.xlsx(
    "../results/CellFie_Zhou2020/data/Zhou2020/meta.xlsx",
    sheet="MetaData")) %>% 
    filter(Condition == "Normal") %>% 
    group_by(across(all_of(c("Depth3", "Cancer_type")))) %>%
    # group_by(across(all_of(c("Condition", "Depth3", "Cancer_type")))) %>% #
    summarize(
        BinaryTaskScoreSum = sum(BinaryTaskScore, na.rm=T),
        ActiveTaskPercentage = mean(BinaryTaskScore, na.rm=T)) %>%
    left_join(cf5) %>%
    mutate(Depth2 = stringr::str_sub(Depth2, 1, 3)) %>%
    mutate(Depth1 = stringr::str_sub(Depth1, 1, 3)) %>%
    # mutate(Group = paste0(Cancer_type, "_", Condition)) %>% #
    ggplot(., aes(x=Cancer_type, y=Depth3)) + #
    geom_point(aes(size=ActiveTaskPercentage, color=ActiveTaskPercentage))  #

if (nested) p = p+facet_nested(Depth1~., scales = "free_y", space = "free")

p = p +
    ylab("") + # labs(fill="adj.P.Val") +
    # scale_color_manual(values=c("Normal"="grey", "Tumor"="orange")) + #
    scale_color_gradient(low="blue", high="red") +
    theme_bw(base_size = 15) + 
    guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
    theme(
        axis.title = element_blank(),
        strip.text.y = element_text(size = 10, angle = 90, hjust=0.5),
        axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 

OUTDIRV =  "../doc/FiguresTables/Dotplot/depth3/"

pdf(paste0(OUTDIRV, "/NormalActiveTaskPercentage_Zhou2020.pdf"), w=16, h=32) #
p
dev.off()
rm(p)
gc()

# Hu2025
nested = T
p = cfs %>% dplyr::select(-Group) %>% left_join(openxlsx::read.xlsx(
    "../results/CellFie_Hu2025/data/Hu2025/Meta_hu2025.xlsx",
    sheet="meta")) %>% 
    filter(Group == "Normal") %>% 
    group_by(across(all_of(c("Depth3", "Cancer_type")))) %>%
    # group_by(across(all_of(c("Condition", "Depth3", "Cancer_type")))) %>% #
    summarize(
        BinaryTaskScoreSum = sum(BinaryTaskScore, na.rm=T),
        ActiveTaskPercentage = mean(BinaryTaskScore, na.rm=T)) %>%
    left_join(cfs) %>%
    mutate(Depth2 = stringr::str_sub(Depth2, 1, 3)) %>%
    mutate(Depth1 = stringr::str_sub(Depth1, 1, 3)) %>%
    # mutate(Group = paste0(Cancer_type, "_", Condition)) %>% # 
    ggplot(., aes(x=Cancer_type, y=Depth3)) + # 
    geom_point(aes(size=ActiveTaskPercentage, color=ActiveTaskPercentage))  # 

if (nested) p = p+facet_nested(Depth1~., scales = "free_y", space = "free")

p = p +
    ylab("") + # labs(fill="adj.P.Val") +
    # scale_color_manual(values=c("Normal"="grey", "Tumor"="orange")) + #
    scale_color_gradient(low="blue", high="red") +
    theme_bw(base_size = 15) + 
    guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
    theme(
        axis.title = element_blank(),
        strip.text.y = element_text(size = 10, angle = 90, hjust=0.5),
        axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 

OUTDIRV =  "../doc/FiguresTables/Dotplot/depth3/"

pdf(paste0(OUTDIRV, "/NormalActiveTaskPercentage_Hu2025.pdf"), w=14, h=32) #
p
dev.off()
rm(p)
gc()

```

## Depth2

### Subsystem activity
```{r}
# Zhou2020
nested = T
dep = "Depth2"
p = cf5 %>% left_join(openxlsx::read.xlsx(
    "../results/CellFie_Zhou2020/data/Zhou2020/meta.xlsx",
    sheet="MetaData")) %>% 
    filter(Condition == "Normal") %>% 
    group_by(across(all_of(c(dep, "Cancer_type")))) %>%
    # group_by(across(all_of(c("Condition", dep, "Cancer_type")))) %>% #
    summarize(
        BinaryTaskScoreSum = sum(BinaryTaskScore, na.rm=T),
        ActiveTaskPercentage = mean(BinaryTaskScore, na.rm=T)) %>%
    left_join(cf5) %>%
    mutate(Depth1 = stringr::str_sub(Depth1, 1, 6)) %>%
    # mutate(Group = paste0(Cancer_type, "_", Condition)) %>% #
    ggplot(., aes_string(x="Cancer_type", y=dep)) + # 
    geom_point(aes(size=ActiveTaskPercentage, color=ActiveTaskPercentage)) # 

if (nested) p = p+facet_nested(Depth1~., scales = "free_y", space = "free")

p = p +
    ylab("") + # labs(fill="adj.P.Val") +
    # scale_color_manual(values=c("Normal"="grey", "Tumor"="orange")) + #
    scale_color_gradient(low="blue", high="red") +
    theme_bw(base_size = 15) + 
    guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
    theme(
        axis.title = element_blank(),
        strip.text.y = element_text(size = 10, angle = 90, hjust=0.5),
        axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 

OUTDIRV =  paste0("../doc/FiguresTables/Dotplot/",dep,"/")
pdf(paste0(OUTDIRV, "/NormalActiveTaskPercentage_Zhou2020.pdf"), w=10, h=17) #
p
dev.off()
rm(p)
gc()

# Hu2025
nested = T
dep = "Depth2"
p = cfs %>% dplyr::select(-Group) %>% left_join(openxlsx::read.xlsx(
    "../results/CellFie_Hu2025/data/Hu2025/Meta_hu2025.xlsx",
    sheet="meta")) %>% 
    filter(Group == "Normal") %>% 
    group_by(across(all_of(c(dep, "Cancer_type")))) %>%
    # group_by(across(all_of(c("Condition", dep, "Cancer_type")))) %>%
    summarize(
        BinaryTaskScoreSum = sum(BinaryTaskScore, na.rm=T),
        ActiveTaskPercentage = mean(BinaryTaskScore, na.rm=T)) %>%
    left_join(cfs) %>%
    mutate(Depth1 = stringr::str_sub(Depth1, 1, 5)) %>%
    # mutate(Group = paste0(Cancer_type, "_", Condition)) %>%
    ggplot(., aes_string(x="Cancer_type", y=dep)) +
    geom_point(aes(size=ActiveTaskPercentage, color=ActiveTaskPercentage))

if (nested) p = p+facet_nested(Depth1~., scales = "free_y", space = "free")

p = p +
    ylab("") + # labs(fill="adj.P.Val") +
    # scale_color_manual(values=c("Normal"="grey", "Tumor"="orange")) +
    scale_color_gradient(low="blue", high="red") +
    theme_bw(base_size = 15) + 
    guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
    theme(
        axis.title = element_blank(),
        strip.text.y = element_text(size = 10, angle = 90, hjust=0.5),
        axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 

OUTDIRV =  paste0("../doc/FiguresTables/Dotplot/",dep,"/")

pdf(paste0(OUTDIRV, "/NormalActiveTaskPercentage_Hu2025.pdf"), w=10, h=15)
p
dev.off()
rm(p)
gc()

```

## Depth1

### Normal activity
```{r}
# Zhou2020
nested = F
dep = "Depth1"
p = cf5 %>% left_join(openxlsx::read.xlsx(
    "../results/CellFie_Zhou2020/data/Zhou2020/meta.xlsx",
    sheet="MetaData")) %>% 
    filter(Condition == "Normal") %>% 
    group_by(across(all_of(c(dep, "Cancer_type")))) %>%
    # group_by(across(all_of(c("Condition", dep, "Cancer_type")))) %>%
    summarize(
        BinaryTaskScoreSum = sum(BinaryTaskScore, na.rm=T),
        ActiveTaskPercentage = mean(BinaryTaskScore, na.rm=T)) %>%
    left_join(cf5) 
p$Depth1 = factor(p$Depth1, levels=p %>% group_by(Depth1) %>% 
    summarize(total = sum(ActiveTaskPercentage, na.rm=T)) %>% 
    arrange(total) %>%pull(Depth1) %>% rev())
p = p %>%
    # mutate(Group = paste0(Cancer_type, "_", Condition)) %>%
    ggplot(., aes_string(x="Cancer_type", y=dep)) +
    geom_point(aes(size=ActiveTaskPercentage, color=ActiveTaskPercentage))

if (nested) p = p+facet_nested(Depth1~., scales = "free_y", space = "free")

p = p +
    ylab("") + # labs(fill="adj.P.Val") +
    # scale_color_manual(values=c("Normal"="grey", "Tumor"="orange")) +
    scale_color_gradient(low="blue", high="red") +
    theme_bw(base_size = 15) + 
    guides(size="none") +
    theme(
        axis.title = element_blank(),
        strip.text.y = element_text(size = 10, angle = 90, hjust=0.5),
        axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 

OUTDIRV =  paste0("../doc/FiguresTables/Dotplot/",dep,"/")
pdf(paste0(OUTDIRV, "/NormalActiveTaskPercentage_Zhou2020.pdf"), w=9, h=3)
p
dev.off()
rm(p)
gc()

# Hu2020
nested = F
dep = "Depth1"
p = cfs %>% dplyr::select(-Group) %>% left_join(openxlsx::read.xlsx(
    "../results/CellFie_Hu2025/data/Hu2025/Meta_hu2025.xlsx",
    sheet="meta")) %>% 
    filter(Group == "Normal") %>% 
    group_by(across(all_of(c(dep, "Cancer_type")))) %>%
    # group_by(across(all_of(c("Condition", dep, "Cancer_type")))) %>%
    summarize(
        BinaryTaskScoreSum = sum(BinaryTaskScore, na.rm=T),
        ActiveTaskPercentage = mean(BinaryTaskScore, na.rm=T)) %>%
    left_join(cfs) 
p$Depth1 = factor(p$Depth1, levels=p %>% group_by(Depth1) %>% 
    summarize(total = sum(ActiveTaskPercentage, na.rm=T)) %>% 
    arrange(total) %>%pull(Depth1) %>% rev())
p = p %>%
    # mutate(Group = paste0(Cancer_type, "_", Condition)) %>%
    ggplot(., aes_string(x="Cancer_type", y=dep)) +
    geom_point(aes(size=ActiveTaskPercentage, color=ActiveTaskPercentage))

if (nested) p = p+facet_nested(Depth1~., scales = "free_y", space = "free")

p = p +
    #ggtitle(title) +
    ylab("") + # labs(fill="adj.P.Val") +
    # scale_color_manual(values=c("Normal"="grey", "Tumor"="orange")) +
    scale_color_gradient(low="blue", high="red") +
    theme_bw(base_size = 15) + 
    guides(size="none") +
    theme(
        axis.title = element_blank(),
        strip.text.y = element_text(size = 10, angle = 90, hjust=0.5),
        axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 

OUTDIRV =  paste0("../doc/FiguresTables/Dotplot/",dep,"/")
pdf(paste0(OUTDIRV, "/NormalActiveTaskPercentage_Hu2025.pdf"), w=9, h=3)
p
dev.off()
rm(p)
gc()

```




# Correlation between the nr of up/downregulated tasks/subsytems/systems and the Active task Nr/percentage of different tasks/subsystems/systems in normal tissues.

```{r}
# getCor = function(sum_statistic="ActiveTaskPerc") {
#     # Zhou 2020 
#     ## Depth3
#     tasks = unique(cf5$Depth3)
#     sigDiffCor_depth3 = lapply(tasks, function(t) {
#         ActiveTaskSum = cf5 %>% mutate(Sample_ID = Sample) %>% dplyr::select(-Sample) %>%
#             left_join(Zhou2020_meta) %>%
#             filter(Sample_type=="Normal") %>%  
#             filter(Depth3==t) %>%
#             group_by(Cancer_type) %>%
#             summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
#                     ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
#                     TaskScoreSum = sum(TaskScore, na.rm=T),
#                     TaskScoreMean = mean(TaskScore, na.rm=T))
#         temp = sapply(deg_d3_Z, function(H) {length(H)}) 
#         temp =data.frame(Cancer_type = names(temp), 
#             Regulation = names(temp),
#             Count=as.numeric(temp)) %>%
#             mutate(Cancer_type=gsub("_.*","",Cancer_type)) %>%
#             mutate(Regulation=gsub(".*_","",Regulation)) %>%
#             spread(., key="Regulation", value="Count") %>%
#             left_join(ActiveTaskSum)
#         temp1 = cor.test(log2(temp$down), temp[,sum_statistic,drop=T], method="spearman")
#         temp1 = data.frame(cor_dn=temp1$estimate, p.value_dn=temp1$p.value)
#         temp2 = cor.test(temp$up, temp$ActiveTaskPerc)
#         temp2 = data.frame(cor_up=temp2$estimate, p.value_up=temp2$p.value)
#         temp = cbind(temp1, temp2)
#         rownames(temp) = t
#         return(temp)
#     }) %>% Reduce(rbind, .) %>%
#         mutate(p.adj_up = p.adjust(p.value_up, method="BH")) %>%
#         mutate(p.adj_dn = p.adjust(p.value_dn, method="BH")) %>%
#         arrange(p.adj_dn)

#     ## Depth2
#     tasks = unique(cf5$Depth2)
#     sigDiffCor_depth2 = lapply(tasks, function(t) {
#         ActiveTaskSum = cf5 %>% mutate(Sample_ID = Sample) %>% dplyr::select(-Sample) %>%
#             left_join(Zhou2020_meta) %>%
#             filter(Sample_type=="Normal") %>%  
#             filter(Depth2==t) %>%
#             group_by(Cancer_type) %>%
#             summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
#                     ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
#                     TaskScoreSum = sum(TaskScore, na.rm=T),
#                     TaskScoreMean = mean(TaskScore, na.rm=T))
#         temp = sapply(deg_d2_Z, function(H) {length(H)}) 
#         temp =data.frame(Cancer_type = names(temp), 
#             Regulation = names(temp),
#             Count=as.numeric(temp)) %>%
#             mutate(Cancer_type=gsub("_.*","",Cancer_type)) %>%
#             mutate(Regulation=gsub(".*_","",Regulation)) %>%
#             spread(., key="Regulation", value="Count") %>%
#             left_join(ActiveTaskSum)
#         temp1 = cor.test(log2(temp$down), temp[,sum_statistic,drop=T], method="spearman")
#         temp1 = data.frame(cor_dn=temp1$estimate, p.value_dn=temp1$p.value)
#         temp2 = cor.test(temp$up, temp$ActiveTaskPerc)
#         temp2 = data.frame(cor_up=temp2$estimate, p.value_up=temp2$p.value)
#         temp = cbind(temp1, temp2)
#         rownames(temp) = t
#         return(temp)
#     }) %>% Reduce(rbind, .) %>%
#         mutate(p.adj_up = p.adjust(p.value_up, method="BH")) %>%
#         mutate(p.adj_dn = p.adjust(p.value_dn, method="BH")) %>%
#         arrange(p.adj_dn)

#     ## Depth1
#     tasks = unique(cf5$Depth1)
#     sigDiffCor_depth1 = lapply(tasks, function(t) {
#         ActiveTaskSum = cf5 %>% mutate(Sample_ID = Sample) %>% dplyr::select(-Sample) %>%
#             left_join(Zhou2020_meta) %>%
#             filter(Sample_type=="Normal") %>%  
#             filter(Depth1==t) %>%
#             group_by(Cancer_type) %>%
#             summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
#                     ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
#                     TaskScoreSum = sum(TaskScore, na.rm=T),
#                     TaskScoreMean = mean(TaskScore, na.rm=T))
#         temp = sapply(deg_d1_Z, function(H) {length(H)}) 
#         temp =data.frame(Cancer_type = names(temp), 
#             Regulation = names(temp),
#             Count=as.numeric(temp)) %>%
#             mutate(Cancer_type=gsub("_.*","",Cancer_type)) %>%
#             mutate(Regulation=gsub(".*_","",Regulation)) %>%
#             spread(., key="Regulation", value="Count") %>%
#             left_join(ActiveTaskSum) 
#         temp1 = cor.test(log2(temp$down+1), temp[,sum_statistic,drop=T], method="spearman")
#         t = gsub("METABOLISM", "M.", t)
#         show(temp %>% mutate(log2down = log2(down+1)) %>%
#             ggplot(., aes_string(x="log2down",y=sum_statistic)) +
#             geom_point() +
#             geom_smooth(method = "lm") +
#             theme_bw() +
#             labs(x = "log2(Nr of down-regualted systems)",
#                  y = "Active task percentage",
#                  title = paste0(t, "_Zhou2020","\n","cor=",round(temp1$estimate,2),
#                     "; p.adj=",round(temp1$p.value,2)))) 
#         temp1 = data.frame(cor_dn=temp1$estimate, p.value_dn=temp1$p.value)
#         temp2 = cor.test(temp$up, temp$ActiveTaskPerc)
#         temp2 = data.frame(cor_up=temp2$estimate, p.value_up=temp2$p.value)
#         temp = cbind(temp1, temp2)
#         rownames(temp) = t
#         return(temp)
#     }) %>% Reduce(rbind, .) %>%
#         mutate(p.adj_up = p.adjust(p.value_up, method="BH")) %>%
#         mutate(p.adj_dn = p.adjust(p.value_dn, method="BH")) %>%
#         arrange(p.adj_dn)


#     # Hu 2025
#     ## Depth3
#     tasks = unique(cfs$Depth3)
#     sigDiffCor_depth3H= lapply(tasks, function(t) {
#         ActiveTaskSum = cfs %>% mutate(Sample_ID = Sample) %>% dplyr::select(-Sample) %>%
#             left_join(Hu2025_meta) %>%
#             filter(Sample_type=="Normal") %>%  
#             filter(Depth3==t) %>% 
#             group_by(Cancer_type) %>%
#             summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
#                     ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
#                     TaskScoreSum = sum(TaskScore, na.rm=T),
#                     TaskScoreMean = mean(TaskScore, na.rm=T))
#         temp = sapply(deg_d3_H, function(H) {length(H)}) 
#         temp =data.frame(Cancer_type = names(temp), 
#             Regulation = names(temp),
#             Count=as.numeric(temp)) %>%
#             mutate(Cancer_type=gsub("_.*","",Cancer_type)) %>%
#             mutate(Regulation=gsub(".*_","",Regulation)) %>%
#             spread(., key="Regulation", value="Count") %>%
#             left_join(ActiveTaskSum)
#         temp1 = cor.test(log2(temp$down), temp[,sum_statistic,drop=T], method="spearman")
#         temp1 = data.frame(cor_dn=temp1$estimate, p.value_dn=temp1$p.value)
#         temp2 = cor.test(temp$up, temp$ActiveTaskPerc)
#         temp2 = data.frame(cor_up=temp2$estimate, p.value_up=temp2$p.value)
#         temp = cbind(temp1, temp2)
#         rownames(temp) = t
#         return(temp)
#     }) %>% Reduce(rbind, .) %>%
#         mutate(p.adj_up = p.adjust(p.value_up, method="BH")) %>%
#         mutate(p.adj_dn = p.adjust(p.value_dn, method="BH")) %>%
#         arrange(p.adj_dn)

#     ## Depth2
#     tasks = unique(cfs$Depth2)
#     sigDiffCor_depth2H = lapply(tasks, function(t) {
#         ActiveTaskSum = cfs %>% mutate(Sample_ID = Sample) %>% dplyr::select(-Sample) %>%
#             left_join(Hu2025_meta) %>%
#             filter(Sample_type=="Normal") %>%  
#             filter(Depth2==t) %>% 
#             group_by(Cancer_type) %>%
#             summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
#                     ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
#                     TaskScoreSum = sum(TaskScore, na.rm=T),
#                     TaskScoreMean = mean(TaskScore, na.rm=T))
#         temp = sapply(deg_d2_H, function(H) {length(H)}) 
#         temp =data.frame(Cancer_type = names(temp), 
#             Regulation = names(temp),
#             Count=as.numeric(temp)) %>%
#             mutate(Cancer_type=gsub("_.*","",Cancer_type)) %>%
#             mutate(Regulation=gsub(".*_","",Regulation)) %>%
#             spread(., key="Regulation", value="Count") %>%
#             left_join(ActiveTaskSum)
#         temp1 = cor.test(log2(temp$down), temp[,sum_statistic,drop=T])
#         temp1 = data.frame(cor_dn=temp1$estimate, p.value_dn=temp1$p.value)
#         temp2 = cor.test(temp$up, temp$ActiveTaskPerc)
#         temp2 = data.frame(cor_up=temp2$estimate, p.value_up=temp2$p.value)
#         temp = cbind(temp1, temp2)
#         rownames(temp) = t
#         return(temp)
#     }) %>% Reduce(rbind, .) %>%
#         mutate(p.adj_up = p.adjust(p.value_up, method="BH")) %>%
#         mutate(p.adj_dn = p.adjust(p.value_dn, method="BH")) %>%
#         arrange(p.adj_dn)

#     ## Depth1
#     tasks = unique(cfs$Depth1)
#     sigDiffCor_depth1H = lapply(tasks, function(t) {
#         ActiveTaskSum = cfs %>% mutate(Sample_ID = Sample) %>% dplyr::select(-Sample) %>%
#             left_join(Hu2025_meta) %>%
#             filter(Sample_type=="Normal") %>%  
#             filter(Depth1==t) %>% 
#             group_by(Cancer_type) %>%
#             summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
#                     ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
#                     TaskScoreSum = sum(TaskScore, na.rm=T),
#                     TaskScoreMean = mean(TaskScore, na.rm=T))
#         temp = sapply(deg_d1_H, function(H) {length(H)}) 
#         temp =data.frame(Cancer_type = names(temp), 
#             Regulation = names(temp),
#             Count=as.numeric(temp)) %>%
#             mutate(Cancer_type=gsub("_.*","",Cancer_type)) %>%
#             mutate(Regulation=gsub(".*_","",Regulation)) %>%
#             spread(., key="Regulation", value="Count") %>%
#             left_join(ActiveTaskSum)
#         temp1 = cor.test(log2(temp$down+1), temp[,sum_statistic,drop=T])
#         t = gsub("METABOLISM", "M.", t)
#         show(temp %>% mutate(log2down = log2(down+1)) %>%
#             ggplot(., aes_string(x="log2down",y=sum_statistic)) +
#             geom_point() +
#             geom_smooth(method = "lm") +
#             theme_bw() +
#             labs(x = "log2(Nr of down-regualted systems)",
#                  y = "Active task percentage",
#                  title = paste0(t, "_others","\n","cor=",
#                     round(temp1$estimate,2),"; p.adj=",round(temp1$p.value,2))))
#         temp1 = data.frame(cor_dn=temp1$estimate, p.value_dn=temp1$p.value)
#         temp2 = cor.test(temp$up, temp$ActiveTaskPerc)
#         temp2 = data.frame(cor_up=temp2$estimate, p.value_up=temp2$p.value)
#         temp = cbind(temp1, temp2)
#         rownames(temp) = t
#         return(temp)
#     }) %>% Reduce(rbind, .) %>%
#         mutate(p.adj_up = p.adjust(p.value_up, method="BH")) %>%
#         mutate(p.adj_dn = p.adjust(p.value_dn, method="BH")) %>%
#         arrange(p.adj_dn)

#     return(list(sigDiffCor_depth3Z = sigDiffCor_depth3,
#                 sigDiffCor_depth2Z = sigDiffCor_depth2,
#                 sigDiffCor_depth1Z = sigDiffCor_depth1,
#                 sigDiffCor_depth3H = sigDiffCor_depth3H,
#                 sigDiffCor_depth2H = sigDiffCor_depth2H,
#                 sigDiffCor_depth1H = sigDiffCor_depth1H))
# }

# # CORwActiveTaskNr = getCor(sum_statistic="ActiveTaskNr")
# dir.create("../doc/FiguresTables/correlation", recursive=T, showWarnings=F)
# pdf("../doc/FiguresTables/correlation/ScatterPlot.pdf", h=4.5, w=4.5)
# CORwActiveTaskPerc = getCor(sum_statistic="ActiveTaskPerc")
# dev.off()
# # CORwTaskScoreSum = getCor(sum_statistic="TaskScoreSum")
# # CORwTaskScoreMean = getCor(sum_statistic="TaskScoreMean")
# names(CORwActiveTaskPerc)
# CORwActiveTaskPerc[c(1,4)] = lapply(CORwActiveTaskPerc[c(1,4)], function(x) {
#     x$Task = rownames(x)
#     x})
# CORwActiveTaskPerc[c(2,5)] = lapply(CORwActiveTaskPerc[c(2,5)], function(x) {
#     x$Subsystem = rownames(x)
#     x})
# CORwActiveTaskPerc[c(3,6)] = lapply(CORwActiveTaskPerc[c(3,6)], function(x) {
#     x$System = rownames(x)
#     x})
# openxlsx::write.xlsx(CORwActiveTaskPerc, 
#   file=paste0("../doc/FiguresTables/correlation/CorrelationBetweenNormalTissueActiveTaskPercentageAndDifferentialTaskNr.xlsx"))

```

## Depth1
```{r}
# taskInfo = read.csv2("../results/CellFie_Hu2025/CellFieOut/ESCC/taskInfo.csv", 
#                      sep=",", stringsAsFactors = FALSE) %>%
#                      setNames(c("Depth3.ID", "Depth3", "Depth1", "Depth2"))

# cors = CORwActiveTaskPerc$sigDiffCor_depth1Z %>%
#     mutate(Study="Zhou2020") %>% 
#     filter(p.adj_dn<0.1) %>% 
#     dplyr::select(Study, cor_dn, p.value_dn, p.adj_dn) %>%
#     arrange(-cor_dn) %>% 
#     setNames(c("Study", "cor", "p.value", "p.adj")) %>%
#     mutate(Diff.Sys.="down") %>%
#     tibble::rownames_to_column(var="System") %>%
#     full_join(
#         CORwActiveTaskPerc$sigDiffCor_depth1Z %>%
#         mutate(Study="Zhou2020") %>% 
#         filter(p.adj_up<0.1) %>% 
#         dplyr::select(Study, cor_up, p.value_up, p.adj_up) %>%
#         arrange(-cor_up) %>% 
#         setNames(c("Study", "cor", "p.value", "p.adj")) %>%
#         mutate(Diff.Sys.="up") %>%
#         tibble::rownames_to_column(var="System")
#     ) %>%
# full_join(
#     CORwActiveTaskPerc$sigDiffCor_depth1H %>%
#     mutate(Study="Others") %>% 
#     filter(p.adj_dn<0.1) %>% 
#     dplyr::select(Study, cor_dn, p.value_dn, p.adj_dn) %>%
#     arrange(-cor_dn) %>% 
#     setNames(c("Study", "cor", "p.value", "p.adj")) %>%
#     mutate(Diff.Sys.="down") %>%
#     tibble::rownames_to_column(var="System") %>%
#     full_join(
#         CORwActiveTaskPerc$sigDiffCor_depth1H %>%
#         mutate(Study="Others") %>% 
#         filter(p.adj_up<0.1) %>% 
#         dplyr::select(Study, cor_up, p.value_up, p.adj_up) %>%
#         arrange(-cor_up) %>% 
#         setNames(c("Study", "cor", "p.value", "p.adj")) %>%
#         mutate(Diff.Sys.="up") %>%
#         tibble::rownames_to_column(var="System")
#     ) 
# ) %>% mutate(cor=round(cor, 2)) %>% arrange(-cor) 

# cors$System = factor(cors$System, 
#     levels=rev(unique(c(cors%>%filter(Diff.Sys.=="down")%>%pull(System),cors$System))))
# pdf("../doc/FiguresTables/correlation/CorrelationBetweenNormalTissueActiveTaskPercentageAndDifferentialTaskNr_depth1.pdf", h=3.5, w=7)
# ggplot(cors, aes(x=cor, y=System)) +
#     geom_point(aes(size=p.adj, shape=Diff.Sys., color=Diff.Sys.)) +
#     facet_wrap(~Study, ncol=2) + 
#     scale_shape_manual(values = c("up" = 20, "down" = 18)) +
#     ggtitle("Cor: normal meta. activity & diff. nr.") +
#     scale_color_manual(values=c("up"="blue", "down"="red")) +
#     theme_bw(base_size = 15) + 
#     scale_size(trans="reverse") +
#     guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
#     theme(
#         # axis.title = element_blank(),
#         strip.text.x = element_text(size = 14),
#         axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) +
#     labs(y="", x="Correlation",
#          caption=paste0("p.adj Cutoff = 0.1, p.adj Method = BH"))
# dev.off()

```

## Depth2
```{r}
# cors = CORwActiveTaskPerc$sigDiffCor_depth2Z %>%
#     mutate(Study="Zhou2020") %>% 
#     filter(p.value_dn<0.05) %>% 
#     dplyr::select(Study, cor_dn, p.value_dn, p.adj_dn) %>%
#     arrange(-cor_dn) %>% 
#     setNames(c("Study", "cor", "p.value", "p.adj")) %>%
#     mutate(Diff.Sys.="down") %>%
#     tibble::rownames_to_column(var="System") %>%
#     full_join(
#         CORwActiveTaskPerc$sigDiffCor_depth2Z %>%
#         mutate(Study="Zhou2020") %>% 
#         filter(p.adj_up<0.05) %>% 
#         dplyr::select(Study, cor_up, p.value_up, p.adj_up) %>%
#         arrange(-cor_up) %>% 
#         setNames(c("Study", "cor", "p.value", "p.adj")) %>%
#         mutate(Diff.Sys.="up") %>%
#         tibble::rownames_to_column(var="System")
#     ) %>%
# full_join(
#     CORwActiveTaskPerc$sigDiffCor_depth2H %>%
#     mutate(Study="Others") %>% 
#     filter(p.value_dn<0.1) %>% 
#     dplyr::select(Study, cor_dn, p.value_dn, p.adj_dn) %>%
#     arrange(-cor_dn) %>% 
#     setNames(c("Study", "cor", "p.value", "p.adj")) %>%
#     mutate(Diff.Sys.="down") %>%
#     tibble::rownames_to_column(var="System") %>%
#     full_join(
#         CORwActiveTaskPerc$sigDiffCor_depth2H %>%
#         mutate(Study="Others") %>% 
#         filter(p.adj_up<0.1) %>% 
#         dplyr::select(Study, cor_up, p.value_up, p.adj_up) %>%
#         arrange(-cor_up) %>% 
#         setNames(c("Study", "cor", "p.value", "p.adj")) %>%
#         mutate(Diff.Sys.="up") %>%
#         tibble::rownames_to_column(var="System")
#     ) 
# ) %>% mutate(cor=round(cor, 2)) %>% arrange(-cor) %>%
# left_join(taskInfo%>%mutate(System=Depth2)) %>%
# mutate(Depth1=stringr::str_sub(Depth1, 1, 3))

# cors$System = factor(cors$System, 
#     levels=rev(unique(c(cors%>%filter(Diff.Sys.=="down")%>%pull(System),cors$System))))
# pdf("../doc/FiguresTables/correlation/CorrelationBetweenNormalTissueActiveTaskPercentageAndDifferentialTaskNr_depth2.pdf", h=9.5, w=9.5)
# ggplot(cors, aes(x=cor, y=System)) +
#     geom_point(aes(size=p.adj, shape=Diff.Sys., color=Diff.Sys.)) +
#     facet_wrap(~Study, ncol=2) + 
#     # facet_nested(Depth1~Study, scales = "free_y", space = "free") +
#     scale_shape_manual(values = c("up" = 20, "down" = 18)) +
#     ggtitle("Cor: normal meta. activity & diff. nr.") +
#     scale_color_manual(values=c("up"="blue", "down"="red")) +
#     theme_bw(base_size = 15) + 
#     scale_size(trans="reverse") +
#     guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
#     theme(
#         # axis.title = element_blank(),
#         strip.text.x = element_text(size = 14),
#         axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) +
#     labs(y="", x="Correlation",
#          caption=paste0("p.value Cutoff = 0.05, p.adj Method = BH"))
# dev.off()

```

# Correlation between limma t and active task percentage

## Hu 2025
```{r}
taskInfo = read.csv2("../results/CellFie_Hu2025/CellFieOut/ESCC/taskInfo.csv", 
                     sep=",", stringsAsFactors = FALSE) %>%
                     setNames(c("Depth3.ID", "Depth3", "Depth1", "Depth2"))
DEG_files_H = "../results/CellFie_Hu2025/Limma/TaskDepth3/Limma_differentialAnalysis_result_Depth3.xlsx"
degs_H = lapply(openxlsx::getSheetNames(DEG_files_H), function(sheet) {
  openxlsx::read.xlsx(DEG_files_H, sheet = sheet) %>% 
    mutate(Cancer_type=sheet) %>%
    left_join(taskInfo %>% mutate(Task=Depth3))
}) %>% Reduce(rbind, .)

cancers = cfs %>% mutate(Cancer_type=gsub("_.*","",Group)) %>%
    pull(Cancer_type) %>% unique()
sigDiffCor_depth3_H = lapply(cancers, function(c) {
    TaskScore = cfs %>% mutate(Sample_ID = Sample) %>% dplyr::select(-Sample) %>%
        left_join(Hu2025_meta) %>% 
        filter(Sample_type=="Normal") %>%  
        filter(Cancer_type==c) %>% 
        mutate(Task = Depth3) %>% 
        group_by(Task) %>%
        summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
                ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
                TaskScoreSum = sum(TaskScore, na.rm=T),
                TaskScoreMean = mean(TaskScore, na.rm=T))
    temp = degs_H %>% filter(Cancer_type==c) %>% select(Task, t) %>%
        inner_join(TaskScore) %>% unique() %>% na.omit()
    n = nrow(temp)
    temp = cor.test(temp$t, temp$ActiveTaskPerc)
    temp = data.frame(cor=temp$estimate, p.value=temp$p.value)
    temp$Cancer_type = c
    temp$Study="Hu2025"
    temp$SampleSize = n
    temp
}) %>% Reduce(rbind, .) %>%
    mutate(p.adj = p.adjust(p.value, method="BH")) %>%
    arrange(p.adj) %>%
    mutate(Regulation = ifelse(p.adj<0.05, "Sig", "No"))
sigDiffCor_depth3_H$Cancer_type = factor(sigDiffCor_depth3_H$Cancer_type,
    levels=sigDiffCor_depth3_H$Cancer_type)

dir.create("../doc/FiguresTables/correlation_scoreVSt/", 
    recursive=T, showWarnings=F)
pdf("../doc/FiguresTables/correlation_scoreVSt/Cor_ScoreVSt_depth3_Hu2025.pdf", h=3.7, w=5)
ggplot(sigDiffCor_depth3_H, aes(x=Cancer_type, y=cor)) +
    geom_point(aes(size=p.adj, color=cor)) +
    # facet_wrap(~Study, ncol=2) + 
    # scale_shape_manual(values = c("Sig" = 20, "No" = 18)) +
    ggtitle("Normal.Task.Score vs t") +
    scale_size(trans="reverse") +
    scale_color_gradient2(low="slateblue1", mid="snow2", high="red", midpoint=0) +
    # scale_color_gradient(values=c("No"="grey", "Sig"="orange")) +
    theme_bw(base_size = 15) + 
    guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
    theme(
        # axis.title = element_blank(),
        # strip.text.x = element_text(size = 14),
        axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) +
    labs(X="Cancer_type", y="Pearson correlation",
         caption=paste0("n = 138, p.adj Method = BH"))
dev.off()

```

## Zhou2020
```{r}
taskInfo = read.csv2("../results/CellFie_Hu2025/CellFieOut/ESCC/taskInfo.csv", 
                     sep=",", stringsAsFactors = FALSE) %>%
                     setNames(c("Depth3.ID", "Depth3", "Depth1", "Depth2"))
DEG_files_Z = "../results/CellFie_Zhou2020/Limma/TaskDepth3/Limma_differentialAnalysis_result_Depth3.xlsx"
degs_Z = lapply(openxlsx::getSheetNames(DEG_files_Z), function(sheet) {
  openxlsx::read.xlsx(DEG_files_Z, sheet = sheet) %>% 
    mutate(Cancer_type=sheet) %>%
    left_join(taskInfo %>% mutate(Task=Depth3))
}) %>% Reduce(rbind, .)

cancers = Zhou2020_meta %>% pull(Cancer_type) %>% unique()
sigDiffCor_depth3_Z = lapply(cancers, function(c) {
    TaskScore = cf5 %>% mutate(Sample_ID = Sample) %>% dplyr::select(-Sample) %>%
        left_join(Zhou2020_meta) %>% 
        filter(Sample_type=="Normal") %>%  
        filter(Cancer_type==c) %>% 
        mutate(Task = Depth3) %>% 
        group_by(Task) %>%
        summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
                ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
                TaskScoreSum = sum(TaskScore, na.rm=T),
                TaskScoreMean = mean(TaskScore, na.rm=T))
    temp = degs_Z %>% filter(Cancer_type==c) %>% select(Task, t) %>%
        inner_join(TaskScore) %>% unique() %>% na.omit()
    n = nrow(temp)
    if (n>2) {
        temp = cor.test(temp$t, temp$ActiveTaskPerc)
        temp = data.frame(cor=temp$estimate, p.value=temp$p.value)
        temp$Cancer_type = c
        temp$Study="Zhou2020"
        temp$SampleSize = n
    } else {
        temp = data.frame(cor=NA, p.value=NA)
        temp$Cancer_type = c
        temp$Study="Zhou2020"
        temp$SampleSize = n
    }
    temp
}) %>% Reduce(rbind, .) %>%
    mutate(p.adj = p.adjust(p.value, method="BH")) %>%
    arrange(p.adj) %>%
    mutate(Regulation = ifelse(p.adj<0.05, "Sig", "No"))
sigDiffCor_depth3_Z$Cancer_type = factor(sigDiffCor_depth3_Z$Cancer_type,
    levels=sigDiffCor_depth3_Z$Cancer_type)

dir.create("../doc/FiguresTables/correlation_scoreVSt/", 
    recursive=T, showWarnings=F)
pdf("../doc/FiguresTables/correlation_scoreVSt/Cor_ScoreVSt_depth3_Zhou2020.pdf", h=3.85, w=5.5)
ggplot(sigDiffCor_depth3_Z, aes(x=Cancer_type, y=cor)) +
    geom_point(aes(size=p.adj, color=cor)) +
    # facet_wrap(~Study, ncol=2) + 
    # scale_shape_manual(values = c("Sig" = 20, "No" = 18)) +
    ggtitle("Normal.Task.Score vs t") +
    scale_size(trans="reverse") +
    scale_color_gradient2(low="slateblue1", mid="snow2", high="red", midpoint=0) +
    # scale_color_gradient(values=c("No"="grey", "Sig"="orange")) +
    theme_bw(base_size = 15) + 
    guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
    theme(
        # axis.title = element_blank(),
        # strip.text.x = element_text(size = 14),
        axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) +
    labs(X="Cancer_type", y="Pearson correlation",
         caption=paste0("n = 176, p.adj Method = BH"))
dev.off()


openxlsx::write.xlsx(list(Others=as.data.frame(sigDiffCor_depth3_H), 
                       Zhou2020=as.data.frame(sigDiffCor_depth3_Z)), 
  file=paste0("../doc/FiguresTables/correlation_scoreVSt/correlation_scoreVSt.xlsx"))

```


# Correlation between the nr of up/downregulated tasks and the Active task percentage in normal tissues for each cancer type.

```{r}
library(ggplot2)
library(dplyr)
# Hu2025 metabolic task score datasets
cfs = readRDS("../results/CellFie_Hu2025/Summary/cfs.RDS")
# Zhou2020 metabolic task score datasets
cf5 = readRDS("../results/CellFie_Zhou2020/Summary/cf5.RDS")
Sample_meta_Zhou2020 = readRDS(
    "../results/CellFie_Zhou2020/data/Zhou2020/meta.RDS")
Sample_meta_Hu2025 = readRDS(
    "../results/CellFie_Hu2025/data/Hu2025/Meta_hu2025.RDS")

taskInfo = read.csv2("../results/CellFie_Hu2025/CellFieOut/ESCC/taskInfo.csv", 
                     sep=",", stringsAsFactors = FALSE) %>%
                     setNames(c("Depth3.ID", "Depth3", "Depth1", "Depth2"))
DEG_files_H = "../results/CellFie_Hu2025/Limma/TaskDepth3/Limma_differentialAnalysis_result_Depth3.xlsx"
degs_H = lapply(openxlsx::getSheetNames(DEG_files_H), function(sheet) {
  openxlsx::read.xlsx(DEG_files_H, sheet = sheet) %>% 
    mutate(Cancer_type=sheet) %>%
    left_join(taskInfo %>% mutate(Task=Depth3))
}) %>% Reduce(rbind, .)

DEG_files_Z = "../results/CellFie_Zhou2020/Limma/TaskDepth3/Limma_differentialAnalysis_result_Depth3.xlsx"
degs_Z = lapply(openxlsx::getSheetNames(DEG_files_Z), function(sheet) {
  openxlsx::read.xlsx(DEG_files_Z, sheet = sheet) %>% 
    mutate(Cancer_type=sheet) %>%
    left_join(taskInfo %>% mutate(Task=Depth3))
}) %>% Reduce(rbind, .)


getCor = function(sum_statistic="ActiveTaskPerc") {
    # Zhou 2020 
    temp2 = cf5 %>% left_join(Sample_meta_Zhou2020, by="Sample")
    cancer_types = unique(temp2$Cancer_type)
    sigDiffCor_depth2 = lapply(cancer_types, function(t) {
        ActiveTaskSum = temp2 %>% mutate(Sample_ID = Sample) %>% 
            dplyr::select(-Sample) %>%
            filter(Condition=="Normal") %>%  
            filter(Cancer_type==t) %>%
            group_by(across(all_of(c("Depth2")))) %>%
            summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
                    ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
                    TaskScoreSum = sum(TaskScore, na.rm=T),
                    TaskScoreMean = mean(TaskScore, na.rm=T))
        temp = degs_Z[degs_Z$Cancer_type%in%t,] %>% 
            group_by(across(all_of(c("Depth2")))) %>%
            summarize(SigUpTaskPerc = mean(adj.P.Val<0.05&logFC>0),
                      SigDownTaskPerc = mean(adj.P.Val<0.05&logFC<0)) %>%
            left_join(ActiveTaskSum) %>% na.omit()
        if (nrow(temp)>2) {
            n = temp[, c("SigDownTaskPerc", sum_statistic)] %>%
                na.omit() %>% nrow()
            temp1 = cor.test(temp$SigDownTaskPerc, temp[,sum_statistic,drop=T])
            temp1 = data.frame(cor_dn=temp1$estimate, p.value_dn=temp1$p.value, N_dn=n)
            n = temp[, c("SigUpTaskPerc", sum_statistic)] %>%
                na.omit() %>% nrow()
            temp2 = cor.test(temp$SigUpTaskPerc, temp[,sum_statistic,drop=T])
            temp2 = data.frame(cor_up=temp2$estimate, p.value_up=temp2$p.value, N_up=n)
            temp = cbind(temp1, temp2)
            temp$Cancer_type = t
            temp$Study="Zhou2020"
        } else {
            temp = data.frame(
                cor_dn = numeric(),
                p.value_dn = numeric(),
                cor_up = numeric(),
                p.value_up = numeric(),
                Cancer_type = character(),
                Study = character(),
                N_dn = numeric(),
                N_up = numeric()
            )
        }
        return(temp)
    }) %>% Reduce(rbind, .) %>%
        mutate(p.adj_up = p.adjust(p.value_up, method="fdr")) %>%
        mutate(p.adj_dn = p.adjust(p.value_dn, method="fdr")) %>%
        arrange(p.adj_dn)

    # Hu 2025
    temp2 = cfs %>% left_join(Sample_meta_Hu2025%>%mutate(Condition=Group), by="Sample")
    cancer_types = unique(temp2$Cancer_type)
    sigDiffCor_depth2H = lapply(cancer_types, function(t) {
        ActiveTaskSum = temp2 %>% mutate(Sample_ID = Sample) %>% 
            dplyr::select(-Sample) %>%
            filter(Condition=="Normal") %>%  
            filter(Cancer_type==t) %>%
            group_by(across(all_of(c("Depth2")))) %>%
            summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
                    ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
                    TaskScoreSum = sum(TaskScore, na.rm=T),
                    TaskScoreMean = mean(TaskScore, na.rm=T))
        temp = degs_H[degs_H$Cancer_type%in%t,] %>% 
            group_by(across(all_of(c("Depth2")))) %>%
            summarize(SigUpTaskPerc = mean(adj.P.Val<0.05&logFC>0),
                      SigDownTaskPerc = mean(adj.P.Val<0.05&logFC<0)) %>%
            left_join(ActiveTaskSum) %>% na.omit()
        if (nrow(temp)>2) {
            n = temp[, c("SigDownTaskPerc", sum_statistic)] %>%
                na.omit() %>% nrow()
            temp1 = cor.test(temp$SigDownTaskPerc, temp[,sum_statistic,drop=T])
            temp1 = data.frame(cor_dn=temp1$estimate, p.value_dn=temp1$p.value, N_dn=n)
            n = temp[, c("SigUpTaskPerc", sum_statistic)] %>%
                na.omit() %>% nrow()
            temp2 = cor.test(temp$SigUpTaskPerc, temp[,sum_statistic,drop=T])
            temp2 = data.frame(cor_up=temp2$estimate, p.value_up=temp2$p.value, N_up=n)
            temp = cbind(temp1, temp2)
            temp$Cancer_type = t
            temp$Study="Others"
        } else {
            temp = data.frame(
            cor_dn = numeric(),
            p.value_dn = numeric(),
            cor_up = numeric(),
            p.value_up = numeric(),
            Cancer_type = character(),
            Study = character(),
            N_dn = numeric(),
            N_up = numeric()
            )
        }
        return(temp)
    }) %>% Reduce(rbind, .) %>%
        mutate(p.adj_up = p.adjust(p.value_up, method="fdr")) %>%
        mutate(p.adj_dn = p.adjust(p.value_dn, method="fdr")) %>%
        arrange(p.adj_dn)

    return(list(
        sigDiffCor_depth2Zhou2020 = sigDiffCor_depth2,
        sigDiffCor_depth2Others = sigDiffCor_depth2H))
}


CORwActiveTaskPerc = getCor(sum_statistic="ActiveTaskPerc") 
cors = CORwActiveTaskPerc$sigDiffCor_depth2Zhou2020 %>% 
    dplyr::select(Study, Cancer_type, cor_dn, p.value_dn, p.adj_dn, N_dn) %>%
    arrange(-cor_dn) %>% 
    setNames(c("Study", "Cancer_type", "cor", "p.value", "p.adj", "N")) %>%
    mutate(Diff.Sys.="down") %>%
    full_join(
        CORwActiveTaskPerc$sigDiffCor_depth2Zhou2020 %>%
        dplyr::select(Study, Cancer_type, cor_up, p.value_up, p.adj_up, N_up) %>%
        arrange(-cor_up) %>% 
        setNames(c("Study", "Cancer_type","cor", "p.value", "p.adj", "N")) %>%
        mutate(Diff.Sys.="up") 
    ) %>%
    full_join(
        CORwActiveTaskPerc$sigDiffCor_depth2Others %>%
        dplyr::select(Study, Cancer_type, cor_dn, p.value_dn, p.adj_dn, N_dn) %>%
        arrange(-cor_dn) %>% 
        setNames(c("Study", "Cancer_type", "cor", "p.value", "p.adj", "N")) %>%
        mutate(Diff.Sys.="down") %>%
        full_join(
            CORwActiveTaskPerc$sigDiffCor_depth2Others %>%
            dplyr::select(Study, Cancer_type, cor_up, p.value_up, p.adj_up, N_up) %>%
            arrange(-cor_up) %>% 
            setNames(c("Study", "Cancer_type", "cor", "p.value", "p.adj", "N")) %>%
            mutate(Diff.Sys.="up") 
        ) 
    ) %>% mutate(cor=round(cor, 2)) %>% arrange(-cor) %>%
    mutate(Significance = ifelse(p.adj<0.1, "Sig", "nonSig"))

cors_Z = cors %>% filter(Study=="Zhou2020") %>% arrange(-cor) %>%
    mutate(Sig.diff.task=Diff.Sys.)
cors_Z$Cancer_type = factor(cors_Z$Cancer_type, 
    levels=rev(cors_Z$Cancer_type %>% unique()))
cors_Z$Significance = factor(cors_Z$Significance, levels=c("Sig", "nonSig"))
pdf("../doc/FiguresTables/correlation/CorrelationBetweenNormalTissueActiveTaskPercentageAndDifferentialTaskNr_depth2_Zhou2020.pdf", h=3.5, w=5.5)
ggplot(cors_Z, aes(x=Cancer_type, y=cor)) +
    geom_point(aes(shape=Significance, color=Sig.diff.task), size=5) +
    #facet_wrap(~Study, ncol=2) + 
    scale_shape_manual(values = c("Sig" = 20, "nonSig" = 18)) +
    ggtitle("Cor: normal meta. & diff. perc.") +
    scale_color_manual(values=c("up"="darkred", "down"="grey")) +
    theme_bw(base_size = 15) + 
    # scale_size(trans="reverse") +
    guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
    theme(
        # axis.title = element_blank(),
        strip.text.x = element_text(size = 14),
        axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) +
    labs(x="", y="Correlation",
         caption=paste0("p.adj Cutoff = 0.1, p.adj Method = fdr"))
dev.off()

cors_H = cors %>% filter(Study=="Others") %>% arrange(-cor) %>%
    mutate(Sig.diff.task=Diff.Sys.)
cors_H$Cancer_type = factor(cors_H$Cancer_type, 
    levels=rev(cors_H$Cancer_type %>% unique()))
cors_H$Significance = factor(cors_H$Significance, levels=c("Sig", "nonSig"))
pdf("../doc/FiguresTables/correlation/CorrelationBetweenNormalTissueActiveTaskPercentageAndDifferentialTaskNr_depth2_Hu2025.pdf", h=3.5, w=5.5)
ggplot(cors_H, aes(x=Cancer_type, y=cor)) +
    geom_point(aes(shape=Significance, color=Sig.diff.task), size=5) +
    #facet_wrap(~Study, ncol=2) + 
    scale_shape_manual(values = c("Sig" = 20, "nonSig" = 18)) +
    ggtitle("Cor: normal meta. & diff. perc.") +
    scale_color_manual(values=c("up"="darkred", "down"="grey")) +
    theme_bw(base_size = 15) + 
    # scale_size(trans="reverse") +
    guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
    theme(
        # axis.title = element_blank(),
        strip.text.x = element_text(size = 14),
        axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) +
    labs(x="", y="Correlation",
         caption=paste0("p.adj Cutoff = 0.1, p.adj Method = fdr"))
dev.off()

openxlsx::write.xlsx(rbind(cors_Z, cors_H) %>% select(-Diff.Sys.), 
  file=paste0("../doc/FiguresTables/correlation//CorrelationBetweenNormalTissueActiveTaskPercentageAndDifferentialTaskNr_depth2_Cancer_type.xlsx"))
  
```


# Correlation between the nr of up/downregulated tasks and the Active task percentage of different subsystems/systems in normal tissues.

```{r}
taskInfo = read.csv2("../results/CellFie_Hu2025/CellFieOut/ESCC/taskInfo.csv", 
                     sep=",", stringsAsFactors = FALSE) %>%
                     setNames(c("Depth3.ID", "Depth3", "Depth1", "Depth2"))
DEG_files_H = "../results/CellFie_Hu2025/Limma/TaskDepth3/Limma_differentialAnalysis_result_Depth3.xlsx"
degs_H = lapply(openxlsx::getSheetNames(DEG_files_H), function(sheet) {
  openxlsx::read.xlsx(DEG_files_H, sheet = sheet) %>% 
    mutate(Cancer_type=sheet) %>%
    left_join(taskInfo %>% mutate(Task=Depth3))
}) %>% Reduce(rbind, .)

DEG_files_Z = "../results/CellFie_Zhou2020/Limma/TaskDepth3/Limma_differentialAnalysis_result_Depth3.xlsx"
degs_Z = lapply(openxlsx::getSheetNames(DEG_files_Z), function(sheet) {
  openxlsx::read.xlsx(DEG_files_Z, sheet = sheet) %>% 
    mutate(Cancer_type=sheet) %>%
    left_join(taskInfo %>% mutate(Task=Depth3))
}) %>% Reduce(rbind, .)

getCor = function(sum_statistic="ActiveTaskPerc") {
    # Zhou 2020 
    ## Depth2
    tasks = unique(cf5$Depth2)
    sigDiffCor_depth2 = lapply(tasks, function(t) {
        ActiveTaskSum = cf5 %>% mutate(Sample_ID = Sample) %>% dplyr::select(-Sample) %>%
            left_join(Zhou2020_meta) %>%
            filter(Sample_type=="Normal") %>%  
            filter(Depth2==t) %>%
            group_by(across(all_of(c("Cancer_type")))) %>%
            summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
                    ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
                    TaskScoreSum = sum(TaskScore, na.rm=T),
                    TaskScoreMean = mean(TaskScore, na.rm=T))     
        temp = degs_Z[degs_Z$Depth2%in%t,] %>% 
            group_by(across(all_of(c("Cancer_type")))) %>%
            summarize(SigUpTaskPerc = mean(adj.P.Val<0.05&logFC>0),
                      SigDownTaskPerc = mean(adj.P.Val<0.05&logFC<0)) %>%
            left_join(ActiveTaskSum) %>% na.omit()
        t = gsub("METABOLISM", "M.", t)
        if (nrow(temp)>2) {
            temp1 = cor.test(temp$SigDownTaskPerc, temp[,sum_statistic,drop=T])
            temp1 = data.frame(cor_dn=temp1$estimate, p.value_dn=temp1$p.value)
            temp2 = cor.test(temp$SigUpTaskPerc, temp[,sum_statistic,drop=T])
            temp2 = data.frame(cor_up=temp2$estimate, p.value_up=temp2$p.value)
            temp = cbind(temp1, temp2)
            temp$Subsystem = t
            temp$Study="Zhou2020"
        } else {
            temp = data.frame(
            cor_dn = numeric(),
            p.value_dn = numeric(),
            cor_up = numeric(),
            p.value_up = numeric(),
            Subsystem = character(),
            Study = character()
            )
        }
        return(temp)
    }) %>% Reduce(rbind, .) %>%
        mutate(p.adj_up = p.adjust(p.value_up, method="BH")) %>%
        mutate(p.adj_dn = p.adjust(p.value_dn, method="BH")) %>%
        arrange(p.adj_dn)

    ## Depth1
    tasks = unique(cf5$Depth1)
    sigDiffCor_depth1 = lapply(tasks, function(t) {
        ActiveTaskSum = cf5 %>% mutate(Sample_ID = Sample) %>% dplyr::select(-Sample) %>%
            left_join(Zhou2020_meta) %>%
            filter(Sample_type=="Normal") %>%  
            filter(Depth1==t) %>%
            group_by(across(all_of(c("Cancer_type")))) %>%
            summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
                    ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
                    TaskScoreSum = sum(TaskScore, na.rm=T),
                    TaskScoreMean = mean(TaskScore, na.rm=T))
        temp = degs_Z[degs_Z$Depth1%in%t,] %>% 
            group_by(across(all_of(c("Cancer_type")))) %>%
            summarize(SigUpTaskPerc = mean(adj.P.Val<0.05&logFC>0),
                      SigDownTaskPerc = mean(adj.P.Val<0.05&logFC<0)) %>%
            left_join(ActiveTaskSum) %>% na.omit()
        t = gsub("METABOLISM", "M.", t)
        if (nrow(temp)>2) {
            temp1 = cor.test(temp$SigDownTaskPerc, temp[,sum_statistic,drop=T])
            temp1 = data.frame(cor_dn=temp1$estimate, p.value_dn=temp1$p.value)
            temp2 = cor.test(temp$SigUpTaskPerc, temp[,sum_statistic,drop=T])
            temp2 = data.frame(cor_up=temp2$estimate, p.value_up=temp2$p.value)
            temp = cbind(temp1, temp2)
            temp$System = t
            temp$Study="Zhou2020"
        } else {
            temp = data.frame(
            cor_dn = numeric(),
            p.value_dn = numeric(),
            cor_up = numeric(),
            p.value_up = numeric(),
            System = character(),
            Study = character()
            )
        }
        return(temp)
    }) %>% Reduce(rbind, .) %>%
        mutate(p.adj_up = p.adjust(p.value_up, method="BH")) %>%
        mutate(p.adj_dn = p.adjust(p.value_dn, method="BH")) %>%
        arrange(p.adj_dn)


    # Hu 2025
    ## Depth2
    tasks = unique(cfs$Depth2)
    sigDiffCor_depth2H = lapply(tasks, function(t) {
        ActiveTaskSum = cfs %>% mutate(Sample_ID = Sample) %>% dplyr::select(-Sample) %>%
            left_join(Hu2025_meta) %>%
            filter(Sample_type=="Normal") %>%  
            filter(Depth2==t) %>% 
            group_by(across(all_of(c("Cancer_type")))) %>%
            summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
                    ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
                    TaskScoreSum = sum(TaskScore, na.rm=T),
                    TaskScoreMean = mean(TaskScore, na.rm=T))
        temp = degs_H[degs_H$Depth2%in%t,] %>% 
            group_by(across(all_of(c("Cancer_type")))) %>%
            summarize(SigUpTaskPerc = mean(adj.P.Val<0.05&logFC>0),
                      SigDownTaskPerc = mean(adj.P.Val<0.05&logFC<0)) %>%
            left_join(ActiveTaskSum) %>% na.omit()
        t = gsub("METABOLISM", "M.", t)
        if (nrow(temp)>2) {
            temp1 = cor.test(temp$SigDownTaskPerc, temp[,sum_statistic,drop=T])
            temp1 = data.frame(cor_dn=temp1$estimate, p.value_dn=temp1$p.value)
            temp2 = cor.test(temp$SigUpTaskPerc, temp[,sum_statistic,drop=T])
            temp2 = data.frame(cor_up=temp2$estimate, p.value_up=temp2$p.value)
            temp = cbind(temp1, temp2)
            temp$Subsystem = t
            temp$Study="Others"
        } else {
            temp = data.frame(
            cor_dn = numeric(),
            p.value_dn = numeric(),
            cor_up = numeric(),
            p.value_up = numeric(),
            Subsystem = character(),
            Study = character()
            )
        }
        return(temp)
    }) %>% Reduce(rbind, .) %>%
        mutate(p.adj_up = p.adjust(p.value_up, method="BH")) %>%
        mutate(p.adj_dn = p.adjust(p.value_dn, method="BH")) %>%
        arrange(p.adj_dn)

    ## Depth1
    tasks = unique(cfs$Depth1)
    sigDiffCor_depth1H = lapply(tasks, function(t) {
        ActiveTaskSum = cfs %>% mutate(Sample_ID = Sample) %>% dplyr::select(-Sample) %>%
            left_join(Hu2025_meta) %>%
            filter(Sample_type=="Normal") %>%  
            filter(Depth1==t) %>% 
            group_by(across(all_of(c("Cancer_type")))) %>%
            summarize(ActiveTaskNr = sum(BinaryTaskScore, na.rm=T),
                    ActiveTaskPerc = mean(BinaryTaskScore, na.rm=T),
                    TaskScoreSum = sum(TaskScore, na.rm=T),
                    TaskScoreMean = mean(TaskScore, na.rm=T))
        temp = degs_H[degs_H$Depth1%in%t,] %>% 
            group_by(across(all_of(c("Cancer_type")))) %>%
            summarize(SigUpTaskPerc = mean(adj.P.Val<0.05&logFC>0),
                      SigDownTaskPerc = mean(adj.P.Val<0.05&logFC<0)) %>%
            left_join(ActiveTaskSum) %>% na.omit()
        t = gsub("METABOLISM", "M.", t)
        if (nrow(temp)>2) {
            temp1 = cor.test(temp$SigDownTaskPerc, temp[,sum_statistic,drop=T])
            temp1 = data.frame(cor_dn=temp1$estimate, p.value_dn=temp1$p.value)
            temp2 = cor.test(temp$SigUpTaskPerc, temp[,sum_statistic,drop=T])
            temp2 = data.frame(cor_up=temp2$estimate, p.value_up=temp2$p.value)
            temp = cbind(temp1, temp2)
            temp$System = t
            temp$Study="Others"
        } else {
            temp = data.frame(
            cor_dn = numeric(),
            p.value_dn = numeric(),
            cor_up = numeric(),
            p.value_up = numeric(),
            System = character(),
            Study = character()
            )
        }
        return(temp)
    }) %>% Reduce(rbind, .) %>%
        mutate(p.adj_up = p.adjust(p.value_up, method="BH")) %>%
        mutate(p.adj_dn = p.adjust(p.value_dn, method="BH")) %>%
        arrange(p.adj_dn)

    return(list(
                sigDiffCor_depth2Zhou2020 = sigDiffCor_depth2,
                sigDiffCor_depth1Zhou2020 = sigDiffCor_depth1,
                sigDiffCor_depth2Others = sigDiffCor_depth2H,
                sigDiffCor_depth1Others = sigDiffCor_depth1H))
}

dir.create("../doc/FiguresTables/correlation", recursive=T, showWarnings=F)
CORwActiveTaskPerc = getCor(sum_statistic="ActiveTaskPerc")

openxlsx::write.xlsx(CORwActiveTaskPerc, 
  file=paste0("../doc/FiguresTables/correlation//CorrelationBetweenNormalTissueActiveTaskPercentageAndDifferentialTaskNr.xlsx"))

```

## Depth1_systems

```{r}
taskInfo = read.csv2("../results/CellFie_Hu2025/CellFieOut/ESCC/taskInfo.csv", 
                     sep=",", stringsAsFactors = FALSE) %>%
                     setNames(c("Depth3.ID", "Depth3", "Depth1", "Depth2"))

cors = CORwActiveTaskPerc$sigDiffCor_depth1Zhou2020 %>% 
    filter(p.value_dn<0.05) %>% 
    dplyr::select(Study, System, cor_dn, p.value_dn, p.adj_dn) %>%
    arrange(-cor_dn) %>% 
    setNames(c("Study", "System", "cor", "p.value", "p.adj")) %>%
    mutate(Diff.Sys.="down") %>%
    full_join(
        CORwActiveTaskPerc$sigDiffCor_depth1Zhou2020 %>%
        filter(p.adj_up<0.1) %>% 
        dplyr::select(Study, System, cor_up, p.value_up, p.adj_up) %>%
        arrange(-cor_up) %>% 
        setNames(c("Study", "System","cor", "p.value", "p.adj")) %>%
        mutate(Diff.Sys.="up") 
    ) %>%
full_join(
    CORwActiveTaskPerc$sigDiffCor_depth1Others %>%
    filter(p.value_dn<0.05) %>% 
    dplyr::select(Study, System, cor_dn, p.value_dn, p.adj_dn) %>%
    arrange(-cor_dn) %>% 
    setNames(c("Study", "System", "cor", "p.value", "p.adj")) %>%
    mutate(Diff.Sys.="down") %>%
    full_join(
        CORwActiveTaskPerc$sigDiffCor_depth1Others %>%
        filter(p.adj_up<0.1) %>% 
        dplyr::select(Study, System, cor_up, p.value_up, p.adj_up) %>%
        arrange(-cor_up) %>% 
        setNames(c("Study", "System", "cor", "p.value", "p.adj")) %>%
        mutate(Diff.Sys.="up") 
    ) 
) %>% mutate(cor=round(cor, 2)) %>% arrange(-cor) 

cors$System = factor(cors$System, 
    levels=rev(unique(c(cors%>%filter(Diff.Sys.=="down")%>%pull(System),cors$System))))
pdf("../doc/FiguresTables/correlation/CorrelationBetweenNormalTissueActiveTaskPercentageAndDifferentialTaskNr_depth1.pdf", h=3.5, w=7.5)
ggplot(cors, aes(x=cor, y=System)) +
    geom_point(aes(size=p.adj, shape=Diff.Sys., color=Diff.Sys.)) +
    facet_wrap(~Study, ncol=2) + 
    scale_shape_manual(values = c("up" = 20, "down" = 18)) +
    ggtitle("Cor: normal meta. activity & diff. nr.") +
    scale_color_manual(values=c("up"="grey", "down"="orange")) +
    theme_bw(base_size = 15) + 
    scale_size(trans="reverse") +
    guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
    theme(
        # axis.title = element_blank(),
        strip.text.x = element_text(size = 14),
        axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) +
    labs(y="", x="Correlation",
         caption=paste0("p.value Cutoff = 0.05, p.adj Method = BH"))
dev.off()

cors

```

## Depth2
```{r}
cors = CORwActiveTaskPerc$sigDiffCor_depth2Zhou2020 %>%
    mutate(Study="Zhou2020") %>% 
    filter(p.value_dn<0.05) %>% 
    dplyr::select(Study, Subsystem, cor_dn, p.value_dn, p.adj_dn) %>%
    arrange(-cor_dn) %>% 
    setNames(c("Study", "Subsystem", "cor", "p.value", "p.adj")) %>%
    mutate(Diff.Sys.="down")  %>%
    full_join(
        CORwActiveTaskPerc$sigDiffCor_depth2Zhou2020 %>%
        mutate(Study="Zhou2020") %>% 
        filter(p.adj_up<0.05) %>% 
        dplyr::select(Study, Subsystem, cor_up, p.value_up, p.adj_up) %>%
        arrange(-cor_up) %>% 
        setNames(c("Study", "Subsystem", "cor", "p.value", "p.adj")) %>%
        mutate(Diff.Sys.="up") 
    ) %>%
full_join(
    CORwActiveTaskPerc$sigDiffCor_depth2Others %>%
    mutate(Study="Others") %>% 
    filter(p.value_dn<0.1) %>% 
    dplyr::select(Study, Subsystem, cor_dn, p.value_dn, p.adj_dn) %>%
    arrange(-cor_dn) %>% 
    setNames(c("Study", "Subsystem", "cor", "p.value", "p.adj")) %>%
    mutate(Diff.Sys.="down") %>%
    full_join(
        CORwActiveTaskPerc$sigDiffCor_depth2Others %>%
        mutate(Study="Others") %>% 
        filter(p.adj_up<0.1) %>% 
        dplyr::select(Study, Subsystem, cor_up, p.value_up, p.adj_up) %>%
        arrange(-cor_up) %>% 
        setNames(c("Study", "Subsystem", "cor", "p.value", "p.adj")) %>%
        mutate(Diff.Sys.="up") 
    ) 
) %>% mutate(cor=round(cor, 2)) %>% arrange(-cor) %>%
    left_join(taskInfo%>%mutate(Subsystem=gsub("METABOLISM","M.",Depth2))) %>%
    mutate(System=gsub(" METABOLISM","",Depth1)) %>%
    mutate(Depth1=stringr::str_sub(Depth1, 1, 3)) 

cors$Subsystem = factor(cors$Subsystem, 
    levels=rev(unique(c(cors%>%filter(Diff.Sys.=="down")%>%pull(Subsystem),cors$Subsystem))))
pdf("../doc/FiguresTables/correlation/CorrelationBetweenNormalTissueActiveTaskPercentageAndDifferentialTaskNr_depth2.pdf", 
    h=12, w=8)
ggplot(cors, aes(x=cor, y=Subsystem)) +
    geom_point(aes(size=p.adj, shape=Diff.Sys., color=Diff.Sys.)) +
    # facet_wrap(~Study, ncol=2) + 
    facet_nested(Depth1~Study, scales = "free_y", space = "free") +
    scale_shape_manual(values = c("up" = 20, "down" = 18)) +
    ggtitle("Cor: normal meta. activity & diff. nr.") +
    scale_color_manual(values=c("up"="grey", "down"="orange")) +
    theme_bw(base_size = 15) + 
    scale_size(trans="reverse") +
    guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
    theme(
        # axis.title = element_blank(),
        strip.text.y = element_text(size = 8),
        axis.text.x = element_text(angle = 90, hjust=0.5, vjust=0.5)) +
    labs(y="", x="Correlation",
         caption=paste0("p.value Cutoff = 0.05, p.adj Method = BH"))
dev.off()

pdf("../doc/FiguresTables/correlation/boxplots.pdf", w=4, h=3.8)
temp = cors[cors$Study=="Others",] %>% arrange(Depth1)
temp$Depth1 = factor(temp$Depth1, levels=unique(temp$Depth1))
ggplot(temp, aes(fill=Diff.Sys., y=cor, x=System)) + 
    # geom_bar(position="dodge", stat="identity") +
    geom_boxplot(width=0.5,position = position_dodge(width=0.7)) +
    theme_bw(base_size = 15) +
    ylab("Correlation") +
    xlab("System") +
    labs(fill="Regulation") +
    theme(
        # axis.title = element_blank(),
        # strip.text.y = element_text(size = 10, angle = 90, hjust=0.5),
        axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
    # geom_bar(position="dodge", stat="identity")
dev.off()

```

## summary

### Depth2
```{r}
CORwActiveTaskPerc$sigDiffCor_depth2Zhou2020 %>% filter(p.value_dn<0.05) %>% pull(cor_dn) %>% na.omit() %>% min()
# [1] 0.6231626
CORwActiveTaskPerc$sigDiffCor_depth2Zhou2020 %>% filter(p.value_dn<0.05) %>% pull(cor_dn) %>% na.omit() %>% max()
# [1] 0.8763726
CORwActiveTaskPerc$sigDiffCor_depth2Zhou2020 %>% filter(p.value_dn<0.05) %>% pull(cor_dn) %>% na.omit() %>% mean()
# [1] 0.7495821
```

```{r}
CORwActiveTaskPerc$sigDiffCor_depth2Zhou2020 %>% filter(p.value_up<0.05) %>% pull(cor_up) %>% na.omit() %>% min()
# [1] -0.8615612
CORwActiveTaskPerc$sigDiffCor_depth2Zhou2020 %>% filter(p.value_up<0.05) %>% pull(cor_up) %>% na.omit() %>% max()
# [1] -0.622255
CORwActiveTaskPerc$sigDiffCor_depth2Zhou2020 %>% filter(p.value_up<0.05) %>% pull(cor_up) %>% na.omit() %>% mean()
# [1] -0.7343217
```

```{r}
CORwActiveTaskPerc$sigDiffCor_depth2Others %>% filter(p.value_dn<0.05) %>% pull(cor_dn) %>% na.omit() %>% min()
# [1] 0.5996359
CORwActiveTaskPerc$sigDiffCor_depth2Others %>% filter(p.value_dn<0.05) %>% pull(cor_dn) %>% na.omit() %>% max()
# [1] 0.9202862
CORwActiveTaskPerc$sigDiffCor_depth2Others %>% filter(p.value_dn<0.05) %>% pull(cor_dn) %>% na.omit() %>% mean()
# [1] 0.7718389
```

```{r}
CORwActiveTaskPerc$sigDiffCor_depth2Others %>% filter(p.value_up<0.05) %>% pull(cor_up) %>% na.omit() %>% min()
# [1] -0.9141435
CORwActiveTaskPerc$sigDiffCor_depth2Others %>% filter(p.value_up<0.05) %>% pull(cor_up) %>% na.omit() %>% max()
# [1] -0.5906396
CORwActiveTaskPerc$sigDiffCor_depth2Others %>% filter(p.value_up<0.05) %>% pull(cor_up) %>% na.omit() %>% mean()
# [1] -0.7399185
```

### Depth1
```{r}
CORwActiveTaskPerc$sigDiffCor_depth1Zhou2020 %>% filter(p.value_dn<0.05) %>% pull(cor_dn) %>% na.omit() %>% min()
# [1] 0.6824061
CORwActiveTaskPerc$sigDiffCor_depth1Zhou2020 %>% filter(p.value_dn<0.05) %>% pull(cor_dn) %>% na.omit() %>% max()
# [1] 0.7423764
CORwActiveTaskPerc$sigDiffCor_depth1Zhou2020 %>% filter(p.value_dn<0.05) %>% pull(cor_dn) %>% na.omit() %>% mean()
# [1] 0.7123912
```

```{r}
CORwActiveTaskPerc$sigDiffCor_depth1Zhou2020 %>% filter(p.value_up<0.05) %>% pull(cor_up) %>% na.omit() %>% min()
CORwActiveTaskPerc$sigDiffCor_depth1Zhou2020 %>% filter(p.value_up<0.05) %>% pull(cor_up) %>% na.omit() %>% max()
CORwActiveTaskPerc$sigDiffCor_depth1Zhou2020 %>% filter(p.value_up<0.05) %>% pull(cor_up) %>% na.omit() %>% mean()
```

```{r}
CORwActiveTaskPerc$sigDiffCor_depth1Others %>% filter(p.value_dn<0.05) %>% pull(cor_dn) %>% na.omit() %>% min()
# [1] 0.721767
CORwActiveTaskPerc$sigDiffCor_depth1Others %>% filter(p.value_dn<0.05) %>% pull(cor_dn) %>% na.omit() %>% max()
# [1] 0.8759472
CORwActiveTaskPerc$sigDiffCor_depth1Others %>% filter(p.value_dn<0.05) %>% pull(cor_dn) %>% na.omit() %>% mean()
# [1] 0.797708
```

```{r}
CORwActiveTaskPerc$sigDiffCor_depth1Others %>% filter(p.value_up<0.05) %>% pull(cor_up) %>% na.omit() %>% min()
# [1]  -0.8032965
CORwActiveTaskPerc$sigDiffCor_depth1Others %>% filter(p.value_up<0.05) %>% pull(cor_up) %>% na.omit() %>% max()
# [1] -0.6233187
CORwActiveTaskPerc$sigDiffCor_depth1Others %>% filter(p.value_up<0.05) %>% pull(cor_up) %>% na.omit() %>% mean()
# [1] -0.747471
```

# Limma table

```{r}
library(dplyr)
limma_results = list.files(path="../results/CellFie_Zhou2020/Limma/", pattern="\\.xlsx$", 
                           full.names = TRUE, recursive = TRUE)
Zhou2020 = lapply(limma_results, function(file) {
  sheets = readxl::excel_sheets(file)
  all_sheets = lapply(sheets, function(sheet) {
    c = readxl::read_excel(file, sheet = sheet)
    c$Cancer_type = sheet
    c
  }) %>% Reduce(rbind, .)
})


limma_results2 = list.files(path="../results/CellFie_Hu2025/Limma/", pattern="\\.xlsx$", 
                           full.names = TRUE, recursive = TRUE)
Hu2025 = lapply(limma_results2, function(file) {
  sheets = readxl::excel_sheets(file)
  all_sheets = lapply(sheets, function(sheet) {
    c = readxl::read_excel(file, sheet = sheet)
    c$Cancer_type = sheet
    c
  }) %>% Reduce(rbind, .)
})

openxlsx::write.xlsx(c(setNames(Zhou2020, paste0("Zhou2020_Depth", 1:3)), 
                       setNames(Hu2025, paste0("Others_Depth", 1:3))),
  file=paste0("../doc/FiguresTables/LimmaResults.xlsx"))

```




```{r}
output = capture.output(sessionInfo())
writeLines(output, "../doc/FiguresTables/sessionInfo.txt")
sessionInfo()
```

