---
title: "Hub gene survial analysis"
format: html
---


```{r}
# BiocManager::install("mygene")
# library(mygene)
library(dplyr)
library(parseMetab)

```

```{r}
outdir = "../results/SurvivalAnalysis_TCGA_hubGenes/"
dir.create(outdir, recursive=T, showWarnings=F)
```

# Import data

## Import normalized TCGA data

```{r}
# Get normalized TCGA gene expression data
NdatTCGA = 
    read.csv2("../results/TCGA/Limma_GSVA/NormalizedCount.csv", 
    sep="\t")
NdatTCGA = NdatTCGA %>% tibble::column_to_rownames(var="Features") %>%
    tibble::rownames_to_column(var="symbol")
NdatTCGA[,2:ncol(NdatTCGA)] = 
    lapply(NdatTCGA[,2:ncol(NdatTCGA)], function(x) as.numeric(x))
databaseDIR="../results/TCGA/KEGG_metabolism/"
metab_db_table = 
    readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS"))
TCGA_meta2 = 
    openxlsx::read.xlsx("../results/Summary/Sample_summary_all.xlsx", sheet="TCGA")

```

## Import hub genes

```{r}
library(openxlsx)
# Load the workbook structure
hub.file = "../results/FeatureDifferential_Network/Network/Hubness.xlsx"
sheet_names <- loadWorkbook(hub.file) %>% names()
# Define the exact target terms
targets <- c("Nucleotide.metabolism", "Glycan.biosynthesis.and.metabolism")
# Use sapply to check each sheet, returning a logical vector
sheets2keep.index <- sapply(sheet_names, function(sheet) {
    dat = openxlsx::read.xlsx(hub.file, sheet=sheet)
    colnames(dat)[2] %in% targets
})
sheets2keep = sheet_names[sheets2keep.index]
hubgenes = sapply(sheets2keep, function(sheet){
    dat = openxlsx::read.xlsx(hub.file, sheet=sheet)
    setNames(dat[,2,drop=TRUE], dat[,1,drop=TRUE]) %>%
        sort(.,decreasing=T) %>% 
        head(3) %>% names()
}) %>% as.character()

```

# Survival analysis

## Select cancer types that have >50 samples with survival data

```{r}
dirs = list.files(path = "../data/TCGA/", full.names = TRUE)  # list files/folders in current directory
dirs = dirs[file.info(dirs)$isdir] %>% gsub(".*/", "", .)              # keep only directories
# We will only analyze cancer that have more than 50 samples with available data
cancers2analyze = sapply(dirs, function(c) {
    meta = openxlsx::read.xlsx(paste0("../data/TCGA/",c,"/Clinical_",c,".xlsx"))
    meta0 = openxlsx::read.xlsx(paste0("../data/TCGA/",c,"/Meta_",c,".xlsx"))
    meta = meta0[, c("cases.submitter_id", "sample.submitter_id", "sample_type")] %>% 
        setNames(c("submitter_id", "sample.submitter_id", "sample_type")) %>%
        dplyr::right_join(meta, by="submitter_id")
    meta$sample.submitter_id = gsub("-", ".", meta$sample.submitter_id)
    meta = meta %>% filter(sample_type == "Primary Tumor")
    if (c %in% TCGA_meta2$Cancer_type){
        if(nrow(meta[!(is.na(meta$days_to_death)),])>50) {
            return(c)
        } else {return(NA)}
    } else {return(NA)}
}) %>% as.character()
cancers2analyze = 
    cancers2analyze[!is.na(cancers2analyze)] %>% as.character()

```


## Define the functions for survial analysis

```{r}
library(survival)
library(survminer)
library(dplyr)

survivalAnalysis2 = function(surv_df, ct, hubgene) {
    if (nrow(surv_df) > 10) {
        # 1. Categorize using a Median Split
        median_val <- median(surv_df$Gene_Expression)
        df_clean <- surv_df %>%
            mutate(Gene_Group = ifelse(Gene_Expression >= median_val, "High", "Low")) %>%
            mutate(Gene_Group = factor(Gene_Group, levels = c("Low", "High")))
        
        # 2. Fit the Kaplan-Meier curve using the categorized groups
        km_fit <- survfit(Surv(Time, Event) ~ Gene_Group, data = df_clean)
        
        # 3. Plot a publication-quality curve
        pdf(paste0(outdir, "ggsurvplot_", ct, "_", hubgene, ".pdf"), width = 8, height = 7)
        print(ggsurvplot(
            km_fit,
            data = df_clean,
            pval = TRUE,               
            conf.int = TRUE,  
            conf.int.style = "step",         
            risk.table = TRUE,         
            palette = c("#2E9FDF", "#E7B800"), 
            legend.labs = c("Low Expression", "High Expression"),
            xlab = "Time (Days)",      
            ylab = "Survival Probability",
            ggtheme = theme_minimal()  
        ))
        dev.off()

        # 4. Fit Univariate Cox Model
        uni_cox <- coxph(Surv(Time, Event) ~ Gene_Expression, data = df_clean)
        cat("\n=== Univariate Cox Summary for", ct, "-", hubgene, "===\n")
        print(summary(uni_cox)) # Force output to show in console log

        # 5. DYNAMIC FORMULA BUILDER: Replaces the 8 error-prone if-statements perfectly
        potential_confounders <- c("Age", "Stage", "Gender")
        existing_confounders  <- potential_confounders[potential_confounders %in% colnames(surv_df)]
        all_formula_features  <- c("Gene_Expression", existing_confounders)
        
        survival_formula <- as.formula(
            paste("Surv(Time, Event) ~", paste(all_formula_features, collapse = " + "))
        )

        # 6. Fit Multivariate Cox Model
        multi_cox <- tryCatch({
            coxph(survival_formula, data = df_clean)
        }, error = function(e) { NULL })

        # 7. Check if model exists and successfully converged
        if (!is.null(multi_cox)) {
            cat("\n=== Multivariate Cox Summary for", ct, "-", hubgene, "===\n")
            print(summary(multi_cox)) # Force output to show in console log
            
            # Extract coefficients and check for NA or massive/infinite numbers
            model_coefs <- summary(multi_cox)$coefficients
            has_infinite_coef <- any(is.na(model_coefs[, "se(coef)"])) || 
                                any(model_coefs[, "se(coef)"] > 100) || 
                                any(abs(model_coefs[, "coef"]) > 20)
            
            # Only plot if the model is mathematically stable
            if (!has_infinite_coef) {
                
                # Forest Plot (Removed duplicate file entry block)
                pdf(paste0(outdir, "Forestplot_", ct, "_", hubgene, ".pdf"), width = 8, height = 5)
                print(ggforest(multi_cox, data = df_clean))
                dev.off()

                # Calculate Schoenfeld residuals
                ph_test <- tryCatch({ cox.zph(multi_cox) }, error = function(e) { NULL })

                if (!is.null(ph_test)) {
                    cat("\n--- Proportional Hazards Test ---\n")
                    print(ph_test)

                    # Plot the residuals (Fixed: Added explicit print() statement wrapper)
                    pdf(paste0(outdir, "ggcoxzph_", ct, "_", hubgene, ".pdf"), width = 7, height = 7)
                    print(ggcoxzph(ph_test))
                    dev.off()
                } else {
                    ph_test <- NA
                }
                
            } else {
                message(paste("⚠️ Skipped Forestplot for", ct, "-", hubgene, ": Model has infinite coefficients."))
                ph_test = NA
            }
        } else {
            message(paste("❌ Skipped Forestplot for", ct, "-", hubgene, ": Model failed to fit."))
            ph_test = NULL
        }
    } else {
        km_fit = NULL
        uni_cox = NULL
        multi_cox = NULL
        ph_test = NULL
    }

    return(list(km_fit = km_fit, uni_cox = uni_cox, multi_cox = multi_cox, ph_test = ph_test))
}


```


## Suvival analysis

```{r}

survival_rslt_list = lapply(hubgenes, function(hubgene){
    lapply(cancers2analyze, function(ct) {
        samples = TCGA_meta2 %>% 
                filter(Cancer_type==ct) %>% 
                pull(Sample) %>% intersect(., colnames(NdatTCGA))
        meta = openxlsx::read.xlsx(paste0("../data/TCGA/",ct,"/Clinical_",ct,".xlsx"))
        meta0 = openxlsx::read.xlsx(paste0("../data/TCGA/",ct,"/Meta_",ct,".xlsx"))
        meta = meta0[, c("cases.submitter_id", "sample.submitter_id", "sample_type")] %>% 
            setNames(c("submitter_id", "sample.submitter_id", "sample_type")) %>%
            dplyr::right_join(meta, by="submitter_id")
        meta$sample.submitter_id = gsub("-", ".", meta$sample.submitter_id)
        meta = meta %>% filter(sample_type == "Primary Tumor") %>% 
            filter(sample.submitter_id %in% samples)
        Time = 
            ifelse(!is.na(meta$days_to_death), 
            meta$days_to_death, meta$days_to_last_follow_up) %>% 
            setNames(meta$sample.submitter_id )
        # Event: 1 if dead, 0 if alive
        Event = ifelse(meta$vital_status == "Dead", 1, 0) %>% 
            setNames(meta$sample.submitter_id)
        Age = meta$age_at_diagnosis %>% as.numeric() %>% setNames(meta$sample.submitter_id)
        Gender = meta$gender %>% setNames(meta$sample.submitter_id)
        Stage = meta$ajcc_pathologic_stage %>% 
            gsub("Stage IA", "Stage I", .) %>%
            gsub("Stage IIA", "Stage II", .) %>%
            gsub("Stage IIB", "Stage II", .) %>%
            gsub("Stage IIIA|Stage IIIB|Stage IIIC", "Stage III", .) %>%
            setNames(meta$sample.submitter_id)
        surv_df = data.frame(
            Sample = samples,
            Time = Time[samples] %>% as.numeric(), 
            Event = Event[samples] %>% as.numeric(),
            Gene_Expression = NdatTCGA %>% 
                tibble::column_to_rownames(var="symbol") %>% 
                .[hubgene, samples] %>% as.numeric(),
            Age = Age[samples] %>% as.numeric() %>% log10(),
            Gender = Gender[samples] %>% as.factor(),
            Stage = Stage[samples] %>% as.factor()
        ) %>% na.omit() 
        rownames(surv_df) = NULL
        surv_df = surv_df %>% tibble::column_to_rownames(var="Sample")

        # 1. Check Gender: Drop it if any group has 0 events, 0 survivors, or < 2 patients
        if ("Gender" %in% colnames(surv_df)) {
            gender_table <- table(surv_df$Gender, surv_df$Event)
            
            # Check if any gender group has zero events (0) or zero survivors (1), or fewer than 2 total patients
            if (any(gender_table == 0) || any(rowSums(gender_table) < 2) || length(unique(surv_df$Gender)) < 2) {
                surv_df <- surv_df %>% dplyr::select(-Gender)
            }
        }

        # 2. Check Stage: Drop it if any group has 0 events, 0 survivors, or < 2 patients
        if ("Stage" %in% colnames(surv_df)) {
            stage_table <- table(surv_df$Stage, surv_df$Event)
            
            # Check if any stage group has zero events (0) or zero survivors (1), or fewer than 2 total patients
            if (any(stage_table == 0) || any(rowSums(stage_table) < 2) || length(unique(surv_df$Stage)) < 2) {
                surv_df <- surv_df %>% dplyr::select(-Stage)
            }
        }

        survivalAnalysis2(surv_df, ct, hubgene) 

    }) %>% setNames(cancers2analyze)
}) %>% setNames(hubgenes)

saveRDS(survival_rslt_list, paste0(outdir, "survival_rslt_list.RDS"))

# Extract multi_cox p values:
lapply(hubgenes, function(hubgene) {
    lapply(survival_rslt_list[[hubgene]], function(ct) {
    
    # 1. Ensure the model exists, is a valid coxph object, and is not NA/NULL
    if (!is.null(ct$multi_cox) && inherits(ct$multi_cox, "coxph")) {
        
        coef_matrix <- summary(ct$multi_cox)$coefficients
        
        # 2. Double-check that "Gene_Expression" is actually a row in the matrix
        if ("Gene_Expression" %in% rownames(coef_matrix)) {
        return(coef_matrix["Gene_Expression", "Pr(>|z|)"])
        }
    }
    
    # Return NA if the model failed or didn't contain the gene variable
    return(NA) 
    })
}) %>% setNames(hubgenes)




```

$DTYMK
$DTYMK$BRCA
[1] 0.1058096

$DTYMK$COAD
[1] 0.956502

$DTYMK$ESCA
[1] 0.5045361

$DTYMK$HNSC
[1] 0.8771243

$DTYMK$KIRC
[1] 0.2790758

$DTYMK$LIHC
[1] 0.2223341

$DTYMK$LUAD
[1] 0.2948976

$DTYMK$LUSC
[1] 0.7690232

$DTYMK$PAAD
[1] NA

$DTYMK$UCEC
[1] NA


$RRM2
$RRM2$BRCA
[1] 0.409178

$RRM2$COAD
[1] 0.6789857

$RRM2$ESCA
[1] 0.7010074

$RRM2$HNSC
[1] 0.6304727

$RRM2$KIRC
[1] 0.118267

$RRM2$LIHC
[1] 0.6428563

$RRM2$LUAD
[1] 0.2604951

$RRM2$LUSC
[1] 0.6400252

$RRM2$PAAD
[1] NA

$RRM2$UCEC
[1] NA


$NME1
$NME1$BRCA
[1] 0.5928409

$NME1$COAD
[1] 0.6536146

$NME1$ESCA
[1] 0.9980014

$NME1$HNSC
[1] 0.9617346

$NME1$KIRC
[1] 0.1638014

$NME1$LIHC
[1] 0.3071921

$NME1$LUAD
[1] 0.3364658

$NME1$LUSC
[1] 0.1387894

$NME1$PAAD
[1] NA

$NME1$UCEC
[1] NA


$RPN2
$RPN2$BRCA
[1] 0.369054

$RPN2$COAD
[1] 0.9917923

$RPN2$ESCA
[1] 0.4008399

$RPN2$HNSC
[1] 0.01578119

$RPN2$KIRC
[1] 0.1164738

$RPN2$LIHC
[1] 0.5672467

$RPN2$LUAD
[1] 0.4687461

$RPN2$LUSC
[1] 0.1589894

$RPN2$PAAD
[1] NA

$RPN2$UCEC
[1] NA


$ALG3
$ALG3$BRCA
[1] 0.9596745

$ALG3$COAD
[1] 0.4888238

$ALG3$ESCA
[1] 0.8461447

$ALG3$HNSC
[1] 0.8926775

$ALG3$KIRC
[1] 0.2321092

$ALG3$LIHC
[1] 0.1229703

$ALG3$LUAD
[1] 0.006883037

$ALG3$LUSC
[1] 0.5451583

$ALG3$PAAD
[1] NA

$ALG3$UCEC
[1] NA


$DDOST
$DDOST$BRCA
[1] 0.6954913

$DDOST$COAD
[1] 0.4949864

$DDOST$ESCA
[1] 0.9071666

$DDOST$HNSC
[1] 0.1971339

$DDOST$KIRC
[1] 0.4660783

$DDOST$LIHC
[1] 0.4728688

$DDOST$LUAD
[1] 0.2780172

$DDOST$LUSC
[1] 0.2456013

$DDOST$PAAD
[1] NA

$DDOST$UCEC
[1] NA



```{r}
out = capture.output(sessionInfo())
writeLines(out, paste0(outdir, "/sessionInfo.txt"))

```
