---
title: "TCGA"
author: Yuan Li (yuan.li@nbis.se)
format: html
---




```{r}
# BiocManager::install("TCGAbiolinks")
library(TCGAbiolinks)
library(SummarizedExperiment)
library(dplyr)
library(tidyr)
library(ggplot2)
library(parseMetab)
# remove.packages("parseMetab")
# detach("package:KinaseDownstream", unload = TRUE)
# remotes::install_github("Maj18/parseMetab", dependencies = TRUE)
# library(parseMetab)
```

```{r}
outdir = "../results/TCGA_sensitity_purity/"
dir.create(outdir, recursive=T, showWarnings=F)
```

# Download/import data

```{r}
# Download TCGA purity data
# Fetch the built-in purity dataset
data("Tumor.purity")
# Convert it to a data frame for easier manipulation
purity_df = as.data.frame(Tumor.purity)
head(purity_df)
# Export the entire purity table to your working directory
write.csv(purity_df, 
    file = paste0(outdir, "/TCGA_tumor_purity_data.csv"), 
    row.names = FALSE)

# Import paired data meta
paired_data = readRDS("../results/TCGA/Summary/paired_data.RDS")

# Add purity into meta
purity_df = purity_df %>% 
    mutate(sample.submitter_id = Sample.ID) %>%
    dplyr::select(-Sample.ID)
head(purity_df)
meta_data = lapply(names(paired_data), function(cancer) {
    p = paired_data[[cancer]]
    p = p %>% left_join(purity_df) 
    p$CPE = gsub(",", ".", as.character(p$CPE))
    p$CPE = as.numeric(p$CPE)
    p = p %>% mutate(CPE = replace_na(CPE, 0))
    patientsNOpurityInfo = p[p$CPE==0&p$sample_type=="Primary Tumor", "cases.submitter_id"]
    p %>% filter(!cases.submitter_id %in% patientsNOpurityInfo)
}) %>% setNames(names(paired_data))
# Remove empty items in the list:
meta_data = meta_data[sapply(meta_data, function(p) nrow(p)>0)]
lapply(meta_data, function(p) {
    p %>% pull(sample_type) %>% table()
})
saveRDS(meta_data, paste0(outdir, "Meta_data.RDS"))
# rm(paired_data)
# gc()

# Paired data
paired_data_count = 
    readRDS("../results/TCGA/paired_data_count.RDS") %>%
    .[names(meta_data)]

```


# Import KEGG metabolic pathways

```{r}
databaseDIR="../results/TCGA/KEGG_metabolism/"
kegg_metab_dbs = 
    list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
        kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))
kegg_metab_db = kegg_metab_dbs$kegg_metab_db
names(kegg_metab_dbs)
```


# Paired differential analysis on sample GSVA scores, controllign for CPE (tumor purity)

```{r}
GSVA_limma_rslt = lapply(names(paired_data_count), function(c) {
    print(c)
    dat = paired_data_count[[c]] 
    rownames(dat) = NULL
    dat = dat %>% as.data.frame() %>%
        tibble::column_to_rownames(var="gene_name") %>% na.omit()
    meta = meta_data[[c]] %>% as.data.frame() %>%
        mutate(Sample = sample.submitter_id) %>%
        mutate(Group = sample_type) %>%
        mutate(Patient = cases.submitter_id)
    meta$Group = factor(meta$Group, levels=c("Solid Tissue Normal", "Primary Tumor"))
    paste0(outdir, "/Limma_GSVA/", c, "/") %>%
        dir.create(., recursive=TRUE, showWarnings=FALSE)
    GSVAlimmaTest(dat, paired=TRUE, Feature.filtering=TRUE,
        currentCovariate="CPE", meta=meta, paired_variable="Patient", 
        kegg_metab_db=kegg_metab_db, 
        OUTDIR=paste0(outdir, "/Limma_GSVA/", c, "/"))
}) %>% setNames(names(paired_data_count))

dir.create(paste0(outdir, "/Limma_GSVA"), recursive=T, showWarnings=F)
saveRDS(GSVA_limma_rslt, 
    paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
# GSVA_limma_rslt = readRDS(paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))

# Make a circular packing plot to show the number of pathways in each class
getCirPackingPlot_GSVA(
    kegg_metab_db_table = 
        kegg_metab_dbs$kegg_metab_db_table %>% 
        filter(gs_name %in% (sapply(GSVA_limma_rslt, function(g) {
            rownames(g$gsva)
        }) %>% unlist() %>% as.character() %>% unique())),
    outdir=outdir)

```


## Visualize differential results

```{r}
library(dplyr)
library(stringr)
rlst_list = lapply(GSVA_limma_rslt, function(l) l$rslt_of_interest)
makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA", nested=T, 
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA_tops", nested=T, top=TRUE,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())


makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.1, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA", nested=T, 
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.1, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA_tops", nested=T, top=TRUE,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())


makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA2", nested=T, 
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique(),
        color_low="blue", color_high="red")

makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA_tops2", nested=T, top=TRUE,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique(),
        color_low="blue", color_high="red")

```


### Visualize the number of significant pathways

```{r}
visualizeSigNr(OUTDIR=paste0(outdir, "/Limma_GSVA/"), 
    GSVA_limma_rslt = GSVA_limma_rslt)

names(GSVA_limma_rslt)
sapply(GSVA_limma_rslt, function(c) {
    c$rslt_of_interest %>% filter(adj.P.Val<0.05) %>% nrow()
})

```


### Visualize differential effect

#### Differential effect boxplot by class

```{r}
GSVA_limma_rslt = 
    readRDS(paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
databaseDIR="../results/TCGA/KEGG_metabolism/"
kegg_metab_dbs = 
    list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
    kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))

diffEffectBoxplot_bySystem_GSVA(rlst_list = GSVA_limma_rslt, 
    taskHierarchy=
    kegg_metab_dbs$kegg_metab_db_table %>% select(gs_name, class) %>% unique(), 
    OUT_DIR=paste0(outdir, "/Limma_GSVA"), 
    weight.effect.by.gene=TRUE,
    w=6.2, h=4.2, title="TCGA")

```

#### Differential effect boxplot by cancer

```{r}
diffEffectBoxplot_byCancer_GSVA(
  rlst_list = GSVA_limma_rslt, 
  OUT_DIR=paste0(outdir, "/Limma_GSVA"), w=2.75, h=3.0, title="TCGA")
```


# Summarize GSVA scores

## Visualize pathway activity GSVA scores

### Per sample

Here, we will normalize all cancer type together and then calculate GSVA scores together to make it relatively more comparable across cancer types
```{r}
# Run GSVA all together 
paste0(outdir, "/Limma_GSVA/") %>% 
    dir.create(., recursive=TRUE, showWarnings=FALSE)
dat = paired_data_count %>% Reduce(function(x,y) full_join(x,y,by="gene_name"), .)
rownames(dat) = NULL
dat = dat %>% as.data.frame() %>%
    tibble::column_to_rownames(var="gene_name") %>% na.omit()
    
meta = lapply(names(meta_data), function(c) {
    meta_data[[c]] %>% as.data.frame() %>% 
            mutate(cancer_type=c) %>% 
            select(cancer_type, sample.submitter_id, 
                    sample_type, cases.submitter_id, 
                    CPE)
    }) %>% Reduce(rbind, .) %>% 
    mutate(Sample = sample.submitter_id) %>%
    mutate(Group = sample_type) %>%
    mutate(Patient = cases.submitter_id)
meta$Group = factor(meta$Group, levels=c("Solid Tissue Normal", "Primary Tumor"))

GSVA_limma_rslt_all = GSVAlimmaTest(dat, paired=TRUE, Feature.filtering=TRUE,
    currentCovariate="CPE", meta=meta, paired_variable="Patient", 
    kegg_metab_db=kegg_metab_db, 
    OUTDIR=paste0(outdir, "/Limma_GSVA/"))

GSVA_limma_rslt_gsva_all = lapply(names(meta_data), function(c) {
    GSVA_limma_rslt_all$gsva[,meta_data[[c]]$sample.submitter_id]
}) %>% setNames(names(meta_data))

visualizeGSVSscoresGroup(
    GSVA_limma_rslt_gsva = GSVA_limma_rslt_gsva_all, 
    paired_data = meta_data, 
    OUTDIR = paste0(outdir, "/Summary/")
)

```

### Per metabolic class

```{r}
# paired_data = paired_data[names(GSVA_limma_rslt)]
visualizeGSVSscoresClass(GSVA_limma_rslt, 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table, 
    paired_data = meta_data, 
    OUTDIR = paste0(outdir, "/Summary/"))
```

##  Visualize pathway sizes

```{r}
genes4GSVA = read.delim(
    file = paste0(outdir, "/Limma_GSVA/NormalizedCount.csv"),
    header = TRUE,
    stringsAsFactors = FALSE
    ) %>% pull(Features) %>% unique()

visualizeGSsizeClass(
    GSVA_limma_rslt_gsva=lapply(GSVA_limma_rslt, function(r) r$gsva), 
    genes4GSVA = genes4GSVA, 
    kegg_metab_db = kegg_metab_dbs$kegg_metab_db, 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table,
    OUTDIR=paste0(outdir, "/Summary/"))
```


```{r}
save.image(paste0(outdir, "/TCGA.RData"))
# load(paste0(outdir, "/TCGA.RData"))
out = capture.output(sessionInfo())
writeLines(out, paste0(outdir, "/sessionInfo.txt"))

```


