---
title: "TCGA"
author: Yuan Li (yuan.li@nbis.se)
format: html
---


```{r}
#  #' @param GSVA_limma_rslt_gsva GSVA_limma_rslt_gsva A named list of GSVA result matrices (or data frames). Each list
#  #'   element should correspond to a group (the list names indicate the groups, e.g. cancer
#  #'   types) and contain KEGG pathway identifiers as row names and sample submitter IDs as
#  #'   column names.
# #' @param rslt_of_interest
# #' @param paired_data
# #' @param sample
# #' @param OUTDIR
# #' 
# getNormalVStcorPlot = function(
#         GSVA_limma_rslt_gsva, rslt_of_interest, paired_data,
#         sample="normal", OUTDIR) {
#     dat.list = lapply(seq_along(GSVA_limma_rslt_gsva), function(j) {
#         s = names(GSVA_limma_rslt_gsva)[j]
#         if (sample=="normal") {
#             normal_samples = paired_data[[s]] %>%
#                 filter(sample_type=="Solid Tissue Normal") %>% pull(sample.submitter_id)
#             gsva = GSVA_limma_rslt_gsva[[s]] %>% .[, normal_samples] ###
#         } else if (sample=="all") {
#             gsva = GSVA_limma_rslt_gsva[[s]]
#         } else {stop("parameter sample should be either ´normal´ or ´all´")}
#         pathway_t = setNames(rslt_of_interest[[j]]$t, rownames(rslt_of_interest[[j]])) ###
#         temp = data.frame(
#             means = (rowSums(gsva, na.rm=TRUE)/ncol(gsva)) %>% as.numeric(),
#             Cancer_type = names(GSVA_limma_rslt_gsva)[j],
#             effect_t = pathway_t[rownames(gsva)] %>% as.numeric())
#         temp
#     }) %>% setNames(names(GSVA_limma_rslt_gsva))

#     dat = lapply(seq_along(dat.list), function(k) {
#         temp = dat.list[[k]]
#         n = nrow(temp)
#         temp = cor.test(temp$effect_t, temp$means)
#         temp = data.frame(cor=temp$estimate, p.value=temp$p.value)
#         temp$Cancer_type = names(dat.list)[k]
#         temp$Study="TCGA"
#         temp$SampleSize = n
#         temp
#     }) %>% Reduce(rbind, .) %>%
#         mutate(p.adj = p.adjust(p.value, method="BH")) %>%
#         arrange(p.adj) %>%
#         mutate(Regulation = ifelse(p.adj<0.05, "Sig", "No"))

#     dat$Cancer_type = factor(dat$Cancer_type, levels=dat$Cancer_type)
#     pdf(paste0(OUTDIR, "/Cor_ScoreVSt_TCGA_", sample,".pdf"), h=3.85, w=5.5)
#         print(ggplot2::ggplot(dat, ggplot2::aes(x=Cancer_type, y=cor)) +
#             ggplot2::geom_point(aes(shape=Regulation, color=cor), size=5) +
#             ggplot2::ggtitle(paste0("corr(",sample,", effect)")) +
#             ggplot2::scale_size(trans="reverse") +
#             ggplot2::scale_color_gradient2(low="steelblue",
#                             mid="snow2", high="firebrick", midpoint=0) +
#             ggplot2::theme_bw(base_size = 15) +
#             ggplot2::guides(alpha="none", size=ggplot2::guide_legend(override.aes=list(shape=21))) +
#             ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0)) +
#             ggplot2::labs(X="Cancer_type", y="Pearson correlation",
#                 caption=paste0("n = ", min(dat$SampelSize), "-",
#                 max(dat$SampelSize), ", p.adj Method = BH")))
#     dev.off()
# }



# getNormalExprVStcorPlot = function(normalizedExpr_all, kegg_metab_db,
#         GSVA_limma_rslt_gsva, rslt_of_interest, paired_data, OUTDIR) {
#     dat.list = lapply(seq_along(GSVA_limma_rslt_gsva), function(j) {
#         s = names(GSVA_limma_rslt_gsva)[j]
#         normal_samples = paired_data[[s]] %>%
#             filter(sample_type=="Solid Tissue Normal") %>% pull(sample.submitter_id)
#         normExpr = normalizedExpr_all[, intersect(normal_samples, colnames(normalizedExpr_all))]
#         pathways = rownames(GSVA_limma_rslt_gsva[[j]])
#         Pexpr = lapply(pathways, function(p) {
#             Pgenes = intersect(kegg_metab_db[[p]], rownames(normalizedExpr_all))
#             dat = data.frame(Pmedians = normExpr[Pgenes,] %>% as.matrix() %>%
#                 matrixStats::colMedians(.,na.rm=T)) %>%
#                 t()
#             rownames(dat) = p
#             dat
#         }) %>% Reduce(rbind, .)
#         pathway_t = setNames(rslt_of_interest[[j]]$t, rownames(rslt_of_interest[[j]])) ###
#         temp = data.frame(
#             means = (rowSums(Pexpr, na.rm=TRUE)/ncol(Pexpr)) %>% as.numeric(),
#             Cancer_type = names(GSVA_limma_rslt_gsva)[j],
#             effect_t = pathway_t[rownames(Pexpr)] %>% as.numeric())
#         temp
#     }) %>% setNames(names(GSVA_limma_rslt_gsva))

#     dat = lapply(seq_along(dat.list), function(k) {
#         temp = dat.list[[k]]
#         n = nrow(temp)
#         temp = cor.test(temp$effect_t, temp$means)
#         temp = data.frame(cor=temp$estimate, p.value=temp$p.value)
#         temp$Cancer_type = names(dat.list)[k]
#         temp$Study="TCGA"
#         temp$SampleSize = n
#         temp
#     }) %>% Reduce(rbind, .) %>%
#         mutate(p.adj = p.adjust(p.value, method="BH")) %>%
#         arrange(p.adj) %>%
#         mutate(Regulation = ifelse(p.adj<0.05, "Sig", "No"))

#     dat$Cancer_type = factor(dat$Cancer_type, levels=dat$Cancer_type)
#     pdf(paste0(OUTDIR, "/Cor_pathwExprVSt_TCGA.pdf"), h=3.85, w=5.5)
#         print(ggplot2::ggplot(dat, ggplot2::aes(x=Cancer_type, y=cor)) +
#             ggplot2::geom_point(aes(shape=Regulation, color=cor), size=5) +
#             ggplot2::ggtitle("corr(NormalExpr., effect)") +
#             ggplot2::scale_size(trans="reverse") +
#             ggplot2::scale_color_gradient2(low="steelblue",
#                             mid="snow2", high="firebrick", midpoint=0) +
#             ggplot2::theme_bw(base_size = 15) +
#             ggplot2::guides(alpha="none", size=ggplot2::guide_legend(override.aes=list(shape=21))) +
#             ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0)) +
#             ggplot2::labs(X="Cancer_type", y="Pearson correlation",
#                 caption=paste0("n = ", min(dat$SampelSize), "-",
#                 max(dat$SampelSize), ", p.adj Method = BH")))
#     dev.off()
# }



# #' @param rlst_list a list of limma differential analysis output
# #' @param taskHierarchy a table at least with two columns: Depth and ´Depth1´
# #' 
# makeLimmaDotplot_TCGA = function(rlst_list, adj.P.Val.cutoff, w=NULL, h=NULL, OUTDIRV, 
#         Depth="Depth3", lowfigure=F, title = "Dysregulated pathways_TCGA", 
#         suffix="TCGA", taskHierarchy=cfs, nested=T, top=FALSE,
#         color_low="skyblue", color_high="purple"){
#     taskHierarchy$Task = taskHierarchy[[Depth]]
#     rlst_long = lapply(seq_along(rlst_list), function(i) {
#         rlst = rlst_list[[i]]
#         rlst$Cancer = names(rlst_list)[i]
#         rlst$Regulation = 
#             ifelse(rlst$logFC>0&rlst$adj.P.Val<adj.P.Val.cutoff, "Up", 
#             ifelse (rlst$logFC<0&rlst$adj.P.Val<adj.P.Val.cutoff, "Down", "NonSig"))
#         rlst %>% tibble::rownames_to_column(var="Task") %>% 
#             filter(adj.P.Val<adj.P.Val.cutoff)
#     }) %>% Reduce(rbind, .) %>%
#         dplyr::left_join(taskHierarchy) %>%
#             mutate(logFC=round(logFC, 2)) %>%
#             mutate(Class = gsub("Carbohydrate metabolism","CAR",class)) %>%
#             mutate(Class = gsub("Lipid metabolism", "LIP", Class)) %>%
#             mutate(Class = gsub("Metabolism of cofactors and vitamins", "COF", Class)) %>%
#             mutate(Class = gsub("Energy metabolism", "ENE", Class)) %>%
#             mutate(Class = gsub("Amino acid metabolism", "AMI", Class)) %>%
#             mutate(Class = gsub("Nucleotide metabolism", "NUC", Class)) %>%
#             mutate(Class = gsub("Biosynthesis of other secondary metabolites", "OSM", Class)) %>%
#             mutate(Class = gsub("Metabolism of other amino acids", "OAA", Class)) %>%
#             mutate(Class = gsub("Glycan biosynthesis and metabolism", "GLY", Class)) %>%
#             mutate(Class = gsub("Metabolism of terpenoids and polyketides", "TER", Class)) %>%
#             mutate(Class = gsub("Xenobiotics biodegradation and metabolism" , "XEN", Class))

#     if (top) {
#         temp = rlst_long %>% dplyr::select(Task, Cancer, Regulation) %>% 
#             unique() %>% filter(Regulation %in% c("Up", "Down")) %>%
#             pull(Task) %>% table()
#         tops = names(temp)[temp > median(temp)]
#         rlst_long = rlst_long %>% filter(Task %in% tops)
#     }
#     rlst_long$t[rlst_long$t>5] = 5
#     rlst_long$t[rlst_long$t< -5] = -5
  
#     p = ggplot2::ggplot(data=rlst_long, mapping=ggplot2::aes_string(x="Cancer", y="Task")) +
#         ggplot2::geom_point(ggplot2::aes(size=adj.P.Val, color=t, shape=Regulation))
#     if (nested) p = p + ggh4x::facet_nested(Class~., scales = "free_y", space = "free")
#     p = p + ggplot2::scale_shape_manual(values = c("Up" = 18, "Down" = 20)) +
#         ggplot2::ggtitle(title) +
#         ggplot2::ylab("") +  ggplot2::labs(size="adj.P.Val") +
#         ggplot2::scale_color_gradient2(low=color_low, mid="snow2", high=color_high, midpoint=0) +
#         # ggplot2::scale_color_gradient2(low="blue", mid="snow2", high="red", midpoint=0) +
#         ggplot2::scale_size(trans="reverse") +
#         ggplot2::theme_bw(base_size = 15) + 
#         ggplot2::scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 1000)) +
#         ggplot2::guides(alpha="none", size=ggplot2::guide_legend(override.aes=list(shape=21))) +
#         ggplot2::theme(
#             axis.title =  ggplot2::element_blank(),
#             strip.text.y = ggplot2::element_text(size = 10, angle = 90, hjust=0.5),
#             axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0)) +
#         ggplot2::labs(caption=paste0("adj.P.ValCutoff = ",adj.P.Val.cutoff,", pAdjustMethod = BH"))

#     if (lowfigure) {
#       p = p + ggplot2::theme(
#               axis.title = ggplot2::element_blank(),
#               legend.position = "bottom",
#               legend.box = "vertical",                  # Stack legends vertically
#               legend.box.just = "left",                 # Center the whole legend box
#               legend.text = ggplot2::element_text(size = 7),
#               legend.title = ggplot2::element_text(size = 11),
#               legend.spacing.y = ggplot2::unit(0, "cm") # Reduce spacing between legend rows
#             ) +
#             ggplot2::guides(
#               fill  = ggplot2::guide_legend(ncol=1, title.position = "top", byrow = F),
#               shape = ggplot2::guide_legend(title.position = "left"),
#               size  = ggplot2::guide_legend(title.position = "left"))
#     }
#     if (is.null(h)|is.null(w)) {
#         h = ((unique(rlst_long$gs_name) %>% length())/6) + 3
#         w = (nchar(unique(rlst_long$gs_name)) %>% max())/11 + 3.5
#     }
#     pdf(paste0(OUTDIRV,"/Dotplot_all_adj.P.ValCutoff",
#         adj.P.Val.cutoff,"_",suffix,".pdf"),w=w,h=h)
#         show(p)
#     dev.off()
# }
```

```{r}
# BiocManager::install("TCGAbiolinks")
library(TCGAbiolinks)
library(SummarizedExperiment)
library(dplyr)
library(dplyr)
library(ggplot2)
library(parseMetab)
# remove.packages("parseMetab")
# detach("package:KinaseDownstream", unload = TRUE)
# remotes::install_github("Maj18/parseMetab", dependencies = TRUE)
# library(parseMetab)
```

# Download data

```{r}
# # Check how many TCGA cancer datasets do we have
# projects = getGDCprojects()
# tcga_projects = projects[grep("^TCGA-", projects$project_id), ]
# tcga_projects %>% openxlsx::write.xlsx("../data/TCGA/TCGA_projects.xlsx")
# cancer_types = tcga_projects$tumor
# cancer_types
# #  [1] "BRCA" "UCEC" "LAML" "CHOL" "ACC"  "DLBC" "CESC" "ESCA" "KICH" "HNSC"
# # [11] "COAD" "BLCA" "KIRP" "GBM"  "MESO" "SARC" "PCPG" "READ" "LUAD" "LUSC"
# # [21] "PAAD" "LIHC" "KIRC" "LGG"  "SKCM" "PRAD" "STAD" "OV"   "THYM" "UCS" 
# # [31] "UVM"  "TGCT" "THCA"

# # Loop through each cancer type
# for (cancer in setdiff(cancer_types,c("BLCA", "KIRP", "LGG"))) {
#     cat("Downloading data for:", cancer, "\n")
#     cancer_path = paste0("../data/TCGA/", cancer,"/")
#     if (!dir.exists(cancer_path)) {
#         # Query
#         query = GDCquery(project = paste0("TCGA-", cancer),
#                         data.category = "Transcriptome Profiling",
#                         data.type = "Gene Expression Quantification",
#                         workflow.type = "STAR - Counts",
#                         sample.type = c("Primary Tumor", "Solid Tissue Normal"))
#         if (!is.null(query)) {
#             results = getResults(query)           
#             results %>% openxlsx::write.xlsx(.,paste0("Meta_", cancer,".xlsx"))
#             clinical = GDCquery_clinic(project = paste0("TCGA-", cancer), type = "clinical") 
#             clinical[,sapply(clinical, function(col) is.atomic(col) && !is.list(col))] %>%
#                 openxlsx::write.xlsx(., paste0("Clinical_", cancer,".xlsx"))
            
#             # Download
#             GDCdownload(query, files.per.chunk = 100, method = "api")

#             # Move the output folder and result to cancer_path
#             cancer_path %>% dir.create(., recursive=T, showWarnings=F)
#             from_files = list.files("./", pattern="*.tar.gz|*.txt|*xlsx", full.names = TRUE)
#             lapply(from_files, function(from) {
#                 file.rename(from = from,
#                         to   = paste0(cancer_path, "/", from))
#             })
#             file.rename(from = "GDCdata", to   = paste0(cancer_path, "/GDCdata"))
#         }
#     }
# }

# # TCGA-BLCA problematic

```

```{r}
# library(TCGAbiolinks)
# library(SummarizedExperiment)

# # Check how many TCGA cancer datasets do we have
# projects = getGDCprojects()
# tcga_projects = projects[grep("^TCGA-", projects$project_id), ]
# tcga_projects %>% openxlsx::write.xlsx("../data/TCGA/TCGA_projects.xlsx")
# cancer_types = tcga_projects$tumor
# cancer_types
# #  [1] "BRCA" "UCEC" "LAML" "CHOL" "ACC"  "DLBC" "CESC" "ESCA" "KICH" "HNSC"
# # [11] "COAD" "BLCA" "KIRP" "GBM"  "MESO" "SARC" "PCPG" "READ" "LUAD" "LUSC"
# # [21] "PAAD" "LIHC" "KIRC" "LGG"  "SKCM" "PRAD" "STAD" "OV"   "THYM" "UCS" 
# # [31] "UVM"  "TGCT" "THCA"

# # # Loop through each cancer type
# for (cancer in setdiff(cancer_types,c("BLCA", "KIRP", "LGG"))) {
#     query = GDCquery(project = paste0("TCGA-", cancer),
#                     data.category = "Transcriptome Profiling",
#                     data.type = "Gene Expression Quantification",
#                     workflow.type = "STAR - Counts",
#                     sample.type = c("Primary Tumor", "Solid Tissue Normal"))
#     # Download and prepare
#     # GDCdownload(query)
#     data = GDCprepare(query)
#     # Extract count matrix and metadata
#     counts = assay(data)
#     metadata = colData(data)
#     saveRDS(data, paste0("../data/TCGA/", cancer, "/TCGA_UCEC.RDS"))
#     saveRDS(counts,paste0("../data/TCGA/", cancer, "/counts.RDS"))
#     saveRDS(metadata, paste0("../data/TCGA/", cancer, "/metadat.RDS"))
# }

```

# Data checking

```{r}
outdir = "../results/TCGA"
dir.create(outdir, recursive=T, showWarnings=F)
```

## Only take paired data
```{r}
INDIR = "../data/TCGA/"
ctypes = list.dirs(INDIR, recursive=FALSE, full.name=F)
ctypes
paired_data = lapply(ctypes, function(ctype) {
    meta = openxlsx::read.xlsx(paste0(INDIR, ctype, "/Meta_", ctype, ".xlsx")) %>% 
        full_join(
            openxlsx::read.xlsx(paste0(INDIR, ctype, "/Clinical_", ctype, ".xlsx")) %>%
            select(-created_datetime),
        by=c("cases.submitter_id" = "submitter_id"))  # Here we have patient IDs
    cases.submitter_id2keep = 
        intersect(meta$cases.submitter_id[meta$sample_type=="Primary Tumor"],
        meta$cases.submitter_id[meta$sample_type=="Solid Tissue Normal"]) %>%
        na.omit()
    if (length(cases.submitter_id2keep)>3) {
        return(meta %>% filter(cases.submitter_id %in% cases.submitter_id2keep))
    } else {return(NA)}
}) %>% setNames(ctypes) %>% .[!is.na(.)]

lapply(names(paired_data), function(c) {
    temp = paired_data[[c]]
    temp = temp[temp$cases.submitter_id %in% 
        (which(table(temp$cases.submitter_id)>2) %>% names()), ]
    as.data.frame(temp) %>% count(sample_type, cases.submitter_id)
}) %>% setNames(names(paired_data))

print("If there are more than one tumor data file per patient, pick the first one")
paired_data = lapply(paired_data, function(cancer) {
    cancer[!duplicated(paste0(cancer$sample_type,cancer$cases.submitter_id)),]
})
# Check 
lapply(names(paired_data), function(c) {
    temp = paired_data[[c]]
    temp = temp[temp$cases.submitter_id %in% 
        (which(table(temp$cases.submitter_id)>2) %>% names()),]
    as.data.frame(temp) %>% count(sample_type, cases.submitter_id)
}) %>% setNames(names(paired_data))

saveRDS(paired_data, paste0(outdir, "/Summary/paired_data.RDS"))
# paired_data = readRDS("../results/TCGA/Summary/paired_data.RDS")

```

## Circular barplot to present sample size

```{r}
sampleSizeCircularBarplot = function(paired_data, OUTDIR) {
    sample_sizes = lapply(names(paired_data), function(cancer) {
        dat = paired_data[[cancer]]
        count = dat %>% select(sample_type, sample.submitter_id) %>% 
            unique() %>% group_by(sample_type) %>% 
            summarise(Count=n()) %>%
            mutate(sample_type=gsub("Primary Tumor", "Tumor", sample_type)) %>%
            mutate(sample_type=gsub("Solid Tissue Normal", "Normal", sample_type)) %>%
            mutate(Cancer_type=cancer)
        count = count %>% left_join(
                count %>% group_by(Cancer_type) %>%
                summarise(total=sum(Count))) %>%
                mutate(Cancer_type = paste0(Cancer_type, "\n(", total, ")"))
    }) %>% Reduce(rbind, .)

    TCGA_summary = sample_sizes %>%
        ggplot(., aes(x=Cancer_type, y=Count, fill=sample_type)) +
        geom_bar(stat="identity", show.legend=FALSE) +
        theme_minimal() +
        theme(axis.text.y = element_blank(),
                axis.ticks = element_blank(),
                axis.title = element_blank(),
                plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")) + 
        coord_polar() +
        ggtitle("TCGA") 
    pdf(paste0(outdir, "/Summary/Sample_size_circular_barplot.pdf"), 
        w=9, h=4.5)
        print(cowplot::plot_grid(TCGA_summary, TCGA_summary))
    dev.off()
    return(sample_sizes)
}

paste0(outdir, "/Summary/") %>% dir.create(.,recursive=T, showWarnings=F)
sample_sizes = sampleSizeCircularBarplot(
    paired_data, 
    OUTDIR=paste0(outdir, "/Summary/"))
saveRDS(sample_sizes, paste0(outdir, "/Summary/sample_sizes.RDS"))
# sample_sizes = readRDS(paste0(outdir, "/Summary/sample_sizes.RDS"))
```

## Import raw data overview

```{r}
# Import the data:
paired_data_count = lapply(names(paired_data), function(cancer) {
    folder = paste0("../data/TCGA/", cancer, "/GDCdata/TCGA-", 
        cancer, "/Transcriptome_Profiling/Gene_Expression_Quantification/")
    p = paired_data[[cancer]]
    files = paste0(folder,"/", p$file_id, "/", p$file_name)
    Samples = p$sample.submitter_id
    count_table = lapply(1:length(files), function(i) {
        file = files[[i]]
        expr = read.delim(file, skip=1, header=TRUE)
        expr$gene_id = sub("\\..*$", "", expr$gene_id)
        expr %>% as.data.frame() %>% 
            select(gene_id, gene_name, gene_type, unstranded) %>%
            setNames(c("gene_id", "gene_name", "gene_type", Samples[[i]]))
    })

    count_table

}) %>% setNames(names(paired_data))

# Combine all samples to the same table for each cancer type
paired_data_count = lapply(paired_data_count, function(cancer){
    temp2 = lapply(1:length(cancer), function(i) {
        sample = cancer[[i]]
        sample = sample[!duplicated(sample$gene_name), ]
        rownames(sample) = NULL
        temp = sample %>% as.data.frame() %>% 
            tibble::column_to_rownames(var="gene_name") 
        temp = temp[, 3, drop=FALSE]
        temp = temp[!duplicated(paired_data_count[[1]][[1]]$gene_name), ,drop=FALSE]
        if (i==1) {
            temp = temp %>% as.data.frame() %>% 
                tibble::rownames_to_column(var="gene_name")
        } else {rownames(temp)=NULL}
        temp
    }) %>% Reduce(cbind, .)
    temp2[5:nrow(temp2), ]
})

names(paired_data_count)
head(paired_data[[1]])

saveRDS(paired_data_count, paste0(outdir, "/paired_data_count.RDS"))
# outdir = "../results/TCGA"
# paired_data_count = readRDS(paste0(outdir, "/paired_data_count.RDS"))

```

# Import KEGG metabolic pathways

```{r}
# # Get KEGG metabolic pathways
# paste0(outdir, "/KEGG_metabolism/") %>% 
#     dir.create(., recursive=TRUE, showWarnings=FALSE)

# kegg_metab_dbs = 
#     getdb_metabolism(databaseDIR=paste0(outdir, "/KEGG_metabolism/"))

outdir = "../results/TCGA"
databaseDIR=paste0(outdir, "/KEGG_metabolism/")
kegg_metab_dbs = 
list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
    kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))
kegg_metab_db = kegg_metab_dbs$kegg_metab_db
names(kegg_metab_dbs)
# kegg_metab_dbs$kegg_metab_db_table$gene_symbol %>% unique() %>% length()

```


# Paired differential analysis on sample GSVA scores

```{r}
GSVA_limma_rslt = lapply(names(paired_data_count), function(c) {
    print(c)
    dat = paired_data_count[[c]] 
    rownames(dat) = NULL
    dat = dat %>% as.data.frame() %>%
        tibble::column_to_rownames(var="gene_name") %>% na.omit()
    meta = paired_data[[c]] %>% as.data.frame() %>%
        mutate(Sample = sample.submitter_id) %>%
        mutate(Group = sample_type) %>%
        mutate(Patient = cases.submitter_id)
    meta$Group = factor(meta$Group, levels=c("Solid Tissue Normal", "Primary Tumor"))
    paste0(outdir, "/Limma_GSVA/", c, "/") %>%
        dir.create(., recursive=TRUE, showWarnings=FALSE)
    GSVAlimmaTest(dat, paired=TRUE, Feature.filtering=TRUE,
        currentCovariate=NULL, meta=meta, paired_variable="Patient", 
        kegg_metab_db=kegg_metab_db, 
        OUTDIR=paste0(outdir, "/Limma_GSVA/", c, "/"))
}) %>% setNames(names(paired_data_count))

saveRDS(GSVA_limma_rslt, 
    paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))

# GSVA_limma_rslt = readRDS(paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))

# Make a circular packing plot to show the number of pathways in each class
getCirPackingPlot_GSVA(
    kegg_metab_db_table = 
        kegg_metab_dbs$kegg_metab_db_table %>% 
        filter(gs_name %in% (sapply(GSVA_limma_rslt, function(g) {
            rownames(g$gsva)
        }) %>% unlist() %>% as.character() %>% unique())),
    outdir=outdir)

```


## Visualize differential results

```{r}
library(dplyr)
library(stringr)
rlst_list = lapply(GSVA_limma_rslt, function(l) l$rslt_of_interest)
makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA", nested=T, 
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA_tops", nested=T, top=TRUE,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA2", nested=T, 
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique(),
        color_low="blue", color_high="red")

makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA_tops2", nested=T, top=TRUE,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique(),
        color_low="blue", color_high="red")

```


### Visualize the number of significant pathways

```{r}
visualizeSigNr(OUTDIR=paste0(outdir, "/Limma_GSVA/"), 
    GSVA_limma_rslt = GSVA_limma_rslt)

names(GSVA_limma_rslt)
sapply(GSVA_limma_rslt, function(c) {
    c$rslt_of_interest %>% filter(adj.P.Val<0.05) %>% nrow()
})

# BRCA CHOL COAD ESCA HNSC KICH KIRC KIRP LIHC LUAD LUSC PAAD READ UCEC 
#   64   65   66   14   61   56   69   53   65   66   62    0   39   52 
```

```{r}
# Check samples:
sapply(GSVA_limma_rslt, function(c)
    names(c$gsva)) %>% unlist() %>%
    as.character() %>% unique()
```


### Visualize differential effect

#### Differential effect boxplot by class

```{r}
outdir = "../results/TCGA"
GSVA_limma_rslt = 
    readRDS(paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
databaseDIR=paste0(outdir, "/KEGG_metabolism/")
kegg_metab_dbs = 
    list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
    kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))

diffEffectBoxplot_bySystem_GSVA(rlst_list = GSVA_limma_rslt, 
    taskHierarchy=
    kegg_metab_dbs$kegg_metab_db_table %>% select(gs_name, class) %>% unique(), 
    OUT_DIR=paste0(outdir, "/Limma_GSVA"), 
    weight.effect.by.gene=TRUE,
    w=6.2, h=4.2, title="TCGA")

```

#### Differential effect boxplot by cancer

```{r}
diffEffectBoxplot_byCancer_GSVA(
  rlst_list = GSVA_limma_rslt, 
  OUT_DIR=paste0(outdir, "/Limma_GSVA"), w=2.75, h=3.0, title="TCGA")
```


# Summarize GSVA scores

## Visualize pathway activity GSVA scores

### Per sample

Here, we will normalize all cancer type together and then calculate GSVA scores together to make it relatively more comparable across cancer types
```{r}
# Run GSVA all together 
paste0(outdir, "/Limma_GSVA/") %>% 
    dir.create(., recursive=TRUE, showWarnings=FALSE)
dat = paired_data_count %>% Reduce(function(x,y) full_join(x,y,by="gene_name"), .)
rownames(dat) = NULL
dat = dat %>% as.data.frame() %>%
    tibble::column_to_rownames(var="gene_name") %>% na.omit()
    
meta = lapply(names(paired_data), function(c) {
    paired_data[[c]] %>% as.data.frame() %>% 
            mutate(cancer_type=c) %>% 
            select(cancer_type, sample.submitter_id, sample_type, cases.submitter_id)
    }) %>% Reduce(rbind, .) %>% 
    mutate(Sample = sample.submitter_id) %>%
    mutate(Group = sample_type) %>%
    mutate(Patient = cases.submitter_id)
meta$Group = factor(meta$Group, levels=c("Solid Tissue Normal", "Primary Tumor"))

GSVA_limma_rslt_all = GSVAlimmaTest(dat, paired=TRUE, Feature.filtering=TRUE,
    currentCovariate=NULL, meta=meta, paired_variable="Patient", 
    kegg_metab_db=kegg_metab_db, 
    OUTDIR=paste0(outdir, "/Limma_GSVA/"))

GSVA_limma_rslt_gsva_all = lapply(names(paired_data), function(c) {
    GSVA_limma_rslt_all$gsva[,paired_data[[c]]$sample.submitter_id]
}) %>% setNames(names(paired_data))

visualizeGSVSscoresGroup(
    GSVA_limma_rslt_gsva = GSVA_limma_rslt_gsva_all, 
    paired_data = paired_data, 
    OUTDIR = paste0(outdir, "/Summary/")
)

```

### Per metabolic class

```{r}
visualizeGSVSscoresClass(GSVA_limma_rslt, 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table, 
    paired_data = paired_data, 
    OUTDIR = paste0(outdir, "/Summary/"))
```

##  Visualize pathway sizes
```{r}
genes4GSVA = read.delim(
    file = paste0(outdir, "/Limma_GSVA/NormalizedCount.csv"),
    header = TRUE,
    stringsAsFactors = FALSE
    ) %>% pull(Features) %>% unique()

visualizeGSsizeClass(
    GSVA_limma_rslt_gsva=lapply(GSVA_limma_rslt, function(r) r$gsva), 
    genes4GSVA = genes4GSVA, 
    kegg_metab_db = kegg_metab_dbs$kegg_metab_db, 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table,
    OUTDIR=paste0(outdir, "/Summary/"))
```


# Visualize correlations

## Correlation between normal pathway GSVA scores and differential effect size

```{r}
# dir.create(paste0(outdir, 
#     "/correlation_scoreVSt/"), recursive=T, showWarnings=F)
# rslt_of_interest = lapply(GSVA_limma_rslt, function(g) g$rslt_of_interest)
# # Here the normal sample GSVA scores from different cancer types were calculated together
# getNormalVStcorPlot(
#         GSVA_limma_rslt_gsva=GSVA_limma_rslt_gsva_all,  
#         rslt_of_interest, paired_data, 
#         OUTDIR=paste0(outdir, "/correlation_scoreVSt/"),
#         sample="normal")

```

## Correlation between normal pathway GSVA scores and differential effect size

```{r}
# # Here the normal sample GSVA scores from different cancer types were calculated together
# getNormalVStcorPlot(
#         GSVA_limma_rslt_gsva=GSVA_limma_rslt_gsva_all,  
#         rslt_of_interest, paired_data, 
#         OUTDIR=paste0(outdir, "/correlation_scoreVSt/"),
#         sample="all")

```

## Correlation between normal pathway gene expr (median) and differential effect size

```{r}
# normalizedExpr_all = read.delim(
#     file = paste0(outdir, "/Limma_GSVA/NormalizedCount.csv"),
#     header = TRUE,
#     stringsAsFactors = FALSE
#     ) %>% tibble::column_to_rownames(var="Features")
# colnames(normalizedExpr_all) = gsub("[.]", "-", colnames(normalizedExpr_all))
# getNormalExprVStcorPlot(normalizedExpr_all, 
#     kegg_metab_db=kegg_metab_dbs$kegg_metab_db,
#     GSVA_limma_rslt_gsva=lapply(GSVA_limma_rslt, function(r) r$gsva), 
#     rslt_of_interest=lapply(GSVA_limma_rslt, function(r) r$rslt_of_interest), 
#     paired_data, 
#     OUTDIR = "../results/TCGA/Limma_GSVA/")
```


```{r}
save.image(paste0(outdir, "/TCGA.RData"))
# load(paste0(outdir, "/TCGA.RData"))
out = capture.output(sessionInfo())
writeLines(out, paste0(outdir, "/sessionInfo.txt"))

```


