---
title: Cancer metabolism
subtitle: 
author: Yuan Li
toc: true
number-sections: true
format: 
  html:
    #embed-resources: true #This works from command line not from Rstudio
    code-fold: true
    code-tools: true
    code-link: true
    theme: pandoc
    df-print: paged
    html-math-method: katex #use katex for math text
  pdf:
    code-line-numbers: true
    geometry: 
      - top=30mm
      - left=30mm
  docx: default
language:
  code-summary: Click to show code
editor_options:
  chunk_output_type: inline
editor: visual
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 7,
                      results    = "hold",
                      message    = FALSE,
                      cache.lazy = FALSE,
                      warning    = FALSE)

```

# 1. Import and Process the data

```{r}
library(ggplot2)
library(dplyr)
```

```{r}
library(dplyr)
Pdat = 
  readxl::read_xlsx("../data/Zhou2020/13045_2020_1013_MOESM3_ESM*.xlsx") %>%
  mutate(genes = gsub(";.*", "", `Gene Group`)) %>%
  select(genes, everything(), -`Gene Group`, -Index, -`Protein Group`)
dim(Pdat)
```

# Check whether the data has been normalized or not

```{r}
#' @import ggplot2
overviewPlot = function(data, variable2plot="TotalAbundance", plot="barplot", 
                        unit="Sample", log=F, ylim=F) {
  if (variable2plot=="TotalAbundance") {
    temp = colSums(data, na.rm=T)
    if (log) temp=log2(temp+0.0001)
  } else if (variable2plot=="Nr.MissingValues") {
    temp = apply(data, 2, function(col) {
            sum(is.na(col))})
  } else if (variable2plot=="Nr.NonMissingValues") {
    temp = apply(data, 2, function(col) {
            sum(!is.na(col))})
  } else {
    temp = colSums(data, na.rm=T)
    if (log) temp=log2(temp+0.0001)
  }
  if (variable2plot %in% c("TotalAbundance","Nr.MissingValues","Nr.NonMissingValues")) {
    temp2 = data.frame(varaible2plot=as.numeric(temp),
                       Sample=names(temp),
                       Group=sub("_([^_]+)$", "", names(temp)))
    if (log) {
      variable2plot = paste0(variable2plot, "_log2")
    }
    colnames(temp2)[1] = variable2plot
  }
  if (plot == "boxplot") {
    if (unit=="Sample") {
      if (log) {
        temp3 = tidyr::gather(log2(data+0.0001), key="Sample", value="Abundance")
        temp3$Group = gsub("-.*","",temp3$Sample) %>% gsub(".*_","",.) %>%
          gsub("N", "Normal",.) %>% gsub("T","Tumor",.)
      } else {
        temp3 = tidyr::gather(data, key="Sample", value="Abundance")
        temp3$Group = sub("-.*","",temp3$Sample) %>% gsub(".*_","",.) %>%
          gsub("N", "Normal",.) %>% gsub("T","Tumor",.)
      }
      if (log) {
        v2plot = "Abundance_log2"
        colnames(temp3)[2]=v2plot} else {v2plot="Abundance"}
      p = ggplot2::ggplot(temp3, ggplot2::aes_string(x="Sample", y=v2plot)) +
        ggplot2::geom_boxplot(ggplot2::aes(fill=Group),
                    outlier.shape = 16,
                    outlier.size = 1,
                    outlier.colour = "grey") +
        ggplot2::labs(x="", y=v2plot) 
    } else {
      p = ggplot2::ggplot(temp2, ggplot2::aes_string(x="Group", y=variable2plot)) +
        ggplot2::geom_boxplot(ggplot2::aes(fill=Group), 
                     outlier.shape = 16,
                     outlier.size = 1,
                     outlier.colour = "grey") +
        ggplot2::labs(x="Sample", y=variable2plot)     
    }
  } else if (plot == "barplot") {
    p = ggplot2::ggplot(temp2, ggplot2::aes_string(x="Sample", y=variable2plot)) +
      ggplot2::geom_bar(aes(fill=Group), stat = "identity") +
      ggplot2::theme(legend.position = "none") 
  }
  if (ylim) {p = p + ylim(-250, 250)}
  p + ggplot2::theme_bw() +
    ggplot2::theme(axis.text.x=ggplot2::element_text(angle=270, vjust=0.5,hjust=1)) +
    ggplot2::theme(legend.position = "none") +
    ggtitle("Zhou2020_all")
}
OUTDIR = "../results/CellFie_Zhou2020/"
QCDIR = paste0(OUTDIR, "/data/QC/")
dir.create(QCDIR, recursive=T, showWarnings=F)
pdf(paste0(QCDIR, "RawDataBoxplot.pdf"), h=4, w=30)
overviewPlot(Pdat[,2:ncol(Pdat)], 
             variable2plot="TotalAbundance", plot="boxplot", 
             unit="Sample", log=T, ylim=F)
dev.off()

# hist(log2(Pdat[,24,drop=T]), breaks=500, plot=F)
# min(log2(Pdat[,24,drop=T]))
# log2(Pdat[,24,drop=T])[log2(Pdat[,24,drop=T]) < 6]
# # This data has been imputed
# # Median normalization
# ## 1. Calculate the median for each sample
# sample_medians = apply(Pdat[,2:ncol(Pdat)], 2, median, na.rm = TRUE)
# ## 2. Determine a target median. 
# ### Often, this is chosen as the global median of the sample medians.
# global_median = median(sample_medians, na.rm = TRUE)
# ## 3. Calculate scaling factors for each sample as: scaling_factor = global_median / sample_median
# scaling_factors = global_median / sample_medians
# ## 4. Normalize the data by multiplying each column by its respective scaling factor.
# ### We use the sweep() function to apply the multiplication across columns.
# Median_normalized = sweep(Pdat[,2:ncol(Pdat)], 2, scaling_factors, FUN = "*") %>% log2()
# overviewPlot(Median_normalized, 
#              variable2plot="TotalAbundance", plot="boxplot", 
#              unit="Sample", log=F, ylim=F)
# rm(sample_medians, global_median, scaling_factors)
# gc()
# median(log2(Pdat[,24,drop=T]))
# min(log2(Pdat[,24,drop=T]))
# max(log2(Pdat[,24,drop=T]))
# pdf("test.pdf")
# hist(log2(Pdat[,24,drop=T]), breaks=100)
# dev.off()

# We will need to convert the protein uniprot ID to ENTREZ ID
# BiocManager::install("org.Hs.eg.db")

```

```{r}
library(AnnotationDbi)
library(org.Hs.eg.db)
library(dplyr)
Pdat2 = AnnotationDbi::select(
  org.Hs.eg.db,
  keys = Pdat$genes,
  columns = c("ENTREZID"),
  keytype = "SYMBOL"
) %>% na.omit() %>%
  .[!duplicated(.$SYMBOL),] %>%
  setNames(c("genes", "ENTREZID")) %>%
  right_join(Pdat) %>%
  na.omit() %>%
  dplyr::select(-genes)
names(Pdat2)[1] = "genes"
# length(unique(Pdat2$SYMBOL))
# length(unique(Pdat2$ENTREZID))
dim(Pdat2)
# In total 232 samples

# Log transform the data
Pdat2[,2:ncol(Pdat2)] = log2(Pdat2[,2:ncol(Pdat2)])

# Save the data table
OUTDIR = "../results/CellFie_Zhou2020/"
INDIR = paste0(OUTDIR, "data/")
dir.create(INDIR, recursive = T, showWarnings = F)
# Save the input data to a csv file
write.table(Pdat2, paste0(INDIR, "/Zhou2020/CancerProteomics.csv"), sep=",", 
            quote=F, row.names=F)

```

# 2. Run CellFie

> I then sumbit the input file to the [genePatten webserver](https://cloud.genepattern.org/gp/pages/index.jsf), and run it there

-   Data="PsoPproteomics.csv",
-   Sample.Number="148",
-   Reference.Model= Human_iHsa \# "human.1",
-   Thresholding.Approach="local",
-   Percentile.Or.Value="percentile",
-   Local.Threshold.Type="minmaxmean",
-   Local.Lower.Bound="25",
-   Local.Upper.Bound="75"
-   Uses gene-specific thresholds, customized per gene (based on its expression distribution)
-   For each gene, CellFie computes its 25th and 75th percentile and considers it ON if expression \> mean of \[25th, 75th\].

# 3. Analyze CellFie output

```{r}
cf = read.csv2(paste0(OUTDIR, "/CellFieOutput/detailScoring.csv"), 
               sep=",", stringsAsFactors = FALSE) 
```

## Convert complex number to only keep the real part

```{r}
complexCol = grep("TaskScore_|ExpressionScoreEssentialRxnsTask_", colnames(cf))
for (i in complexCol) {
  cf[, i] = gsub("\\+\\-", "-", cf[, i])
  # x_fixed <- gsub("\\-\\+", "-", x_fixed)
  # x_fixed <- gsub("\\+\\+", "+", x_fixed)
  # x_fixed <- gsub("\\-\\-", "+", x_fixed)
  cf[, i] = Re(as.complex(cf[, i]))
}
```

## Import taskInfo.csv

```{r}
taskInfo = read.csv2(paste0(OUTDIR, "/CellFieOutput/taskInfo.csv"), 
                     sep=",", stringsAsFactors = FALSE) %>%
                     setNames(c("Depth3.ID", "Depth3", "Depth1", "Depth2"))

colnames(taskInfo)
cat(paste0("The number of tasks at Depth3 are ", length(unique(taskInfo$Depth3))))
cat(paste0("The number of tasks at Depth2 are ", length(unique(taskInfo$Depth2))))
cat(paste0("The number of tasks at Depth1 are ", length(unique(taskInfo$Depth1))))
# The number of tasks at Depth3 are 195
# The number of tasks at Depth2 are 73
# The number of tasks at Depth1 are 7
taskInfo %>% dplyr::count(Depth2, Depth1)
taskInfo %>%  dplyr::count(Depth3, Depth1)

# Summarize task nr
TaskNr1 = taskInfo %>% group_by(Depth1) %>%
  summarise(Depth2 = n_distinct(Depth2))
TaskNr2 = taskInfo %>% group_by(Depth1) %>%
  summarise(Depth3 = n_distinct(Depth3))
TaskNr = left_join(TaskNr1, TaskNr2)

TableNames = list(TaskNr) %>% setNames("TaskHierarchy_summary")
openxlsx::write.xlsx(TableNames, 
  file=paste0(OUTDIR,"Summary/TaskHierarchy_summary.xlsx"))
```

## Reshuffle the table to have tasks in the rows and samples in the columns

```{r}
Samples = gsub(".*_", "", colnames(cf)) %>% unique()
cf2 =lapply(Samples, function(S) {
  temp = cf[, grep(paste0(S, "$"), colnames(cf))]
  colnames(temp) = gsub("_.*", "", colnames(temp))
  temp
}) %>% Reduce(rbind, .) %>% 
  mutate(Depth3.ID = TaskID) %>%
  dplyr::select(Depth3.ID, everything()) %>% 
  filter(TaskScore!=-1) # remove unscorable rows (-1)
```

## Add in task depth information

```{r}
cf3 = right_join(taskInfo, cf2)
```

## Add in Sample information

```{r}
SampleInfo = data.frame(Sample = colnames(Pdat2)[-1],
             SampleID = 1:length(colnames(Pdat2)[-1]))
cf4 = right_join(SampleInfo, cf3) %>%
      mutate(Group=gsub("[-].*", "", Sample)) %>%
      dplyr::select(Group, everything()) 

write.table(cf4, paste0(OUTDIR, "/CellFieOut/detailScoring_new.csv"), sep=",", 
            quote=F, row.names=F)

cf5 = dplyr::select(cf4, -EssentialRxnsTask, -ExpressionScoreEssentialRxnsTask,
            -GeneAssociatedToEssentialRxnsTask, -GeneExpressionValue) %>%
            unique()
cf5$ActiveTaskScore = cf5$TaskScore*cf5$BinaryTaskScore
saveRDS(cf5, paste0(OUTDIR, "/Summary/cf5.RDS"))
# cf5 = readRDS(paste0(OUTDIR, "/Summary/cf5.RDS"))

```

## Sum up task scores to higher level categories

```{r}
sumUpTaskScores = function(Depth,ncol=32,h=12,w=30) {
    cols_to_group = c("Sample", Depth)
    Depths = cf5 %>%
        group_by(across(all_of(cols_to_group))) %>%
        summarise(
            MeanTaskScore = mean(TaskScore, na.rm = TRUE),
            MeanActiveTaskScore = mean(ActiveTaskScore, na.rm = TRUE),
            SumTaskScore = sum(TaskScore, na.rm = TRUE),
            SumActiveTaskScore = sum(ActiveTaskScore, na.rm = TRUE),
            FractionOfActiveTasks = mean(BinaryTaskScore==1, na.rm = TRUE),
            NrOfActiveTasks = sum(BinaryTaskScore==1, na.rm = TRUE)
        ) %>%  mutate(Group=gsub("[-].*", "", Sample)) %>%
        dplyr::select(Group, everything()) 
    Depths[,Depth] = factor(Depths[,Depth,drop=T], 
                    levels=unique(Depths[,Depth,drop=T]))
    library(ggplot2)
    margin(1, 1, 1, 1, unit = "cm")
    P = ggplot(Depths, aes_string(x=Depth, y="SumTaskScore", fill="Sample")) +
        geom_bar(stat="identity", position="dodge") +
        facet_wrap(~ Group, ncol=ncol) +
        coord_flip() + xlab("") +
        theme_bw(base_size = 12) +
        theme(legend.position = "none",
        axis.text.x = element_text(angle=90, hjust=1))
    dir.create(paste0(OUTDIR, "/Summary/"), 
        recursive = T, showWarnings = F) 
    pdf(paste0(OUTDIR, "Summary/TaskScores_", Depth,".pdf"),h=h,w=w)
    show(P)
    dev.off()

    L = Depths$Group %>% unique() %>% length()
    Depths$Condition = gsub(".*_", "", Depths$Group) %>% 
      gsub("N","Normal",.) %>% gsub("T", "Tumor", .)
    P = ggplot(Depths, aes_string(y=Depth, x="Sample")) +
        geom_point(shape=18, aes(size=abs(FractionOfActiveTasks), color=Condition)) +
        labs(size = "FractionOfActiveTasks") + 
        scale_color_manual(values=c("Tumor"="darkorange", "Normal"="darkturquoise")) +
        xlab("") + 
        ggtitle(paste0("Fraction of active tasks_",Depth)) + 
        theme_bw(base_size = 12) +
        theme( # legend.position = "none",
          strip.text.x = element_text(size = 14),
          axis.text.x = element_text(angle = 90, hjust=1, vjust=0))
    pdf(paste0(OUTDIR, "/Summary/BinaryTaskScores_", Depth,".pdf"),h=h,w=w)
      show(P)
    dev.off()

    return(Depths)
}

### Task depth1
Depth="Depth1"
Depth1 = sumUpTaskScores(Depth=Depth,ncol=32,h=3.5, w=43) %>% 
  mutate(TaskScore=MeanTaskScore)
saveRDS(Depth1, paste0(OUTDIR, "/Summary/Depth1.RDS")) ###############

### Task depth2
Depth="Depth2"
Depth2 = sumUpTaskScores(Depth=Depth,ncol=32,h=13, w=43)%>%
  mutate(TaskScore=MeanTaskScore)
saveRDS(Depth2, paste0(OUTDIR, "/Summary/Depth2.RDS")) ###############

### Task depth3
Depth = "Depth3"
L = cf5$Group %>% unique() %>% length()
cf5$Condition = gsub(".*_", "", cf5$Group) %>% 
      gsub("N","Normal",.) %>% gsub("T", "Tumor", .)
P = ggplot(cf5, aes_string(y=Depth, x="Sample")) +
    geom_point(shape=18, aes(size=BinaryTaskScore, color=Condition)) +
    labs(size = "BinaryTaskScore") + 
    scale_color_manual(values=c("Tumor"="darkorange", "Normal"="darkturquoise")) +
    xlab("") + 
    ggtitle(paste0("Binary Task Score_", Depth)) + 
    theme_bw(base_size = 12) +
    theme( # legend.position = "none",
      strip.text.x = element_text(size = 14),
      axis.text.x = element_text(angle = 90, hjust=1, vjust=0))
pdf(paste0(OUTDIR, "/Summary/BinaryTaskScores_", Depth,".pdf"),h=26,w=43)
  show(P)
dev.off()

TableNames = list(
  Depth3 = cf5,
  Depth2 = Depth2,
  Depth1 = Depth1)
openxlsx::write.xlsx(TableNames, 
    file=paste0(OUTDIR,"/Summary/TaskScores_differentDepths.xlsx"))

```

## Summarize BinaryTaskScore to higher-level categories

```{r}
# plotActive = function(Depth, ncol=1, h=12, w=15) {
#     library(dplyr)
#     cols_to_group <- c("Group", Depth)
#     cf5_summary = cf5 %>%
#         group_by(across(all_of(cols_to_group))) %>%
#         summarise(Active=sum(BinaryTaskScore==1,na.rm=TRUE)) %>%
#         arrange(Active)
#     cf5_summary[,Depth] = factor(cf5_summary[,Depth,drop=T],
#                          levels=unique(cf5_summary[,Depth,drop=T]))
#     cf5_summary$Cancer = gsub("_.*", "", cf5_summary$Group)
#     cf5_summary$Condition = gsub(".*_", "", cf5_summary$Group) %>%
#       gsub("N","Normal",.) %>% gsub("T","Tumor",.) %>% 
#       factor(., levels=c("Tumor", "Normal"))
#     library(ggplot2)
#     TableNames = list(cf5_summary) %>% setNames(Depth)
#     openxlsx::write.xlsx(TableNames, 
#         file=paste0(OUTDIR,"Summary/BinaryTaskScores_",Depth,".xlsx"))
#     L = cf5_summary$Group %>% unique() %>% length()
#     P = ggplot(cf5_summary, aes_string(y=Depth, x="Group")) +
#         geom_point(shape=18, aes(size=abs(Active), color=Condition)) +
#         labs(size = "Active.Tasks") + 
#         scale_color_manual(values=c("Tumor"="darkorange", "Normal"="darkturquoise")) +
#         # guides(fill = guide_legend(ncol = ncol)) +
#         # scale_fill_gradientn(colors = 
#         #   grDevices::colors()[grep('gr(a|e)y', grDevices::colors(), invert = T)]) +
#         # guides(fill = guide_legend(override.aes = list(alpha = 0.5))) +
#         # coord_flip()  +
#         # facet_wrap(~Group, ncol=L) + 
#         # scale_fill_gradient2(
#         #   high = munsell::mnsl("5B 7/8"),
#         #   low = munsell::mnsl("5Y 7/8")
#         # ) +
#         xlab("") + 
#         ggtitle(paste0("Active task count_",Depth)) + 
#         theme_bw(base_size = 12) +
#         theme( # legend.position = "none",
#           strip.text.x = element_text(size = 14),
#           axis.text.x = element_text(angle = 90, hjust=1, vjust=0))
#     dir.create(paste0(OUTDIR, "/Summary/"), recursive = T, showWarnings = F) 
#     pdf(paste0(OUTDIR, "Summary/BinaryTaskScores_",Depth,".pdf"), h=h, w=w)
#     show(P)
#     dev.off()   
# }

# ### Task depth 3
# Depth = "Depth3"
# plotActive(Depth, ncol=1, h=26, w=17)
# ### Task depth 2
# Depth = "Depth2"
# plotActive(Depth, ncol=1, h=12, w=10)
# ### Task depth 1
# Depth = "Depth1"
# plotActive(Depth, ncol=8, h=3, w=8.8)

```

# Limma differential analysis of Task scores between groups

## Sort out meta data

```{r}
meta = 
    readxl::read_xlsx("../data/Zhou2020/13045_2020_1013_MOESM2_ESM*.xlsx",
    sheet="Patient_clinical charateristics") 
colnames(meta) = gsub(" ", "_", colnames(meta))
meta = meta %>% mutate(Sample=Tumor_case_ID,
        # Cancer_type=Cancer_type,
        `Tumor_size_(cm)`=`Tumor_size_(cm)`,
        Grade=Grade,
        Histologic_type=Histologic_type,
        Tumor_Stage_Pathological=Tumor_Stage_Pathological,
        Path_Stage_Primary_Tumor_pT=Path_Stage_Primary_Tumor_pT,
        Path_Stage_Lymph_Nodes_pN=Path_Stage_Lymph_Nodes_pN) %>% 
        dplyr::select(-Normal_case_ID, -Tumor_case_ID) %>%
        rbind(
            meta %>% 
            mutate(Sample=Normal_case_ID,
                   # Cancer_type = NA,
                   `Tumor_size_(cm)`= NA,
                   Grade=NA,
                   Histologic_type=NA,
                   Tumor_Stage_Pathological=NA,
                   Path_Stage_Primary_Tumor_pT=NA,
                   Path_Stage_Lymph_Nodes_pN=NA
        ) %>%
        dplyr::select(-Normal_case_ID, -Tumor_case_ID)) %>%
        .[!is.na(.[,"Sample"]), ] %>%
        mutate(Group = gsub("[-].*", "", Sample))
meta = meta %>% filter(Sample %in% cf5$Sample) %>% 
  mutate(Condition=gsub(".*_","",Group)%>%
    gsub("T","Tumor",.)%>%gsub("N","Normal",.)) %>%
  mutate(Cancer_type=gsub("ESCA", "ESCC", Cancer_type)) %>%
  mutate(Cancer_type=gsub("LUSC", "SCLC", Cancer_type)) %>%
  mutate(Cancer_type=gsub("LUSQ", "LUSC", Cancer_type))
TableNames = list(
  MetaData = meta,
  SampleSummary = meta %>% dplyr::count(Cancer_type, Condition))
openxlsx::write.xlsx(TableNames, 
    file=paste0(OUTDIR,"/data/Zhou2020/meta.xlsx"))
saveRDS(meta, paste0(OUTDIR,"/data/Zhou2020/meta.RDS")) ###############
#head(meta)
rm(TableNames)
gc()

```

## Set up the function

```{r}
invisible(capture.output({library(limma)}))
invisible(capture.output({library(tibble)}))
#' @param dat the dataset needs to have features as rownames
#' @param meta meta data needs to have at least two columns named "Sample", "Group", for the moment, we only support Group with two categories
#'
limmaTest = function(dat, paired=TRUE, currentCovariate=NULL, 
                     checkCovariate=FALSE, meta=meta,
                     paired_variable="Patient") {
  meta = filter(meta, Sample%in%colnames(dat))
  # Make sure the sample are in the same order in both meta and dat:
  dat = dat[, meta$Sample]
  meta = meta %>% mutate(across(where(is.factor), droplevels))
  # print("When running differential analysis, we only keep features that have at least 2 values in each group...")
  featureskeep = apply(!is.na(dat), 1, function(row) {
      A=row
      B=sub("-.*", "", colnames(dat))
      sum(aggregate(A~B, data=data.frame(A=A,B=B), sum)[,"A"]>1)>1
  })
  dat = dat[featureskeep, ]
  #print(dim(dat))
  #if (!is.null(currentCovariate)) {currentCovariate=NULL}
  baseFormula = ~ Group
  if (is.null(currentCovariate)) {
    currentFormula = baseFormula
  } else {
    currentFormula = as.formula(paste0("~ ", currentCovariate, " + Group"))}
  
  design = model.matrix(currentFormula, data=meta) # each column is for coefficient
  group.categories = meta$Group %>% unique() %>% length()
  colnames(design) = gsub("Group", "", colnames(design))
  
  if (paired) {
    #Fit with correlated arrays
    dupcor = duplicateCorrelation(dat, design, 
                            block=meta[,paired_variable,drop=T])
    fit = lmFit(dat, design, block=meta[,paired_variable,drop=T],
                correlation=dupcor$consensus)
  } else {
    fit = lmFit(dat, design)
  }
  #fit3 = eBayes(fit, robust=TRUE)
  #print("Check whether we should include a factor in the model by checking for how many metabolites this factor is significant in the model: ")
  if (checkCovariate) {
    fit.con_co = eBayes(fit)
    rlst_co = lapply(2:(ncol(design)-group.categories+1), function(i) {
        topTable(fit.con_co, n=Inf, coef=colnames(design)[i])}) %>%
        setNames(colnames(design)[2:(ncol(design)-group.categories+1)])
    #n refers to the number of top-ranked genes to be returned.
  } else {rlst_co = NA}
  
  fit.con = eBayes(fit, robust = TRUE)
  rlst_interest = topTable(fit.con, n=Inf, coef=levels(meta$Group)[2]) %>%
      arrange(-t) 
  
  return(list(rslt_of_interest=rlst_interest, rslt_co=rlst_co, data=dat, fitted.model=fit))
}

#We have so many meta data, which ones to add?
##One method is to fit a model to the biological groups of interest with one addition factor to observe whether the factor has substantial influence on gene expression (such that many genes are detected as differentially expressed for that factor). Repeat this for subsequenct factors to determine the factors that should be included into the final model. 
globalAnalaysis = function(data_matrix=data_matrix_tissue, paired = FALSE, 
                           checkCovariates = c(FALSE, rep(TRUE, 6)), 
                           currentCovariates=currentCovariates, 
                           groups = groups, meta=meta_tissue,
                           varianceCutoff = 0, paired_variable="Patient") {
  sub = data_matrix[, c(grep(groups[1], colnames(data_matrix), value=T), 
                        grep(groups[2], colnames(data_matrix), value=T))]
  ## Remove features with low variance
  invisible(capture.output({library(matrixStats)}))
  keep = rowSds(sub%>%as.matrix(), na.rm = T) > varianceCutoff
  sub = sub[keep, ]
  rlst = lapply(seq_along(currentCovariates), function(i) {
    print(currentCovariates[i])
    currentCovariate = currentCovariates[i]
    checkCovariate = checkCovariates[i]
    limmaTest(dat = sub, paired=paired, 
              currentCovariate=currentCovariate, 
              checkCovariate=checkCovariate,
              meta=meta, paired_variable=paired_variable)
  }) %>% setNames(paste0("rlst_", currentCovariates))
  return(rlst)
}

modelSummary = function(rlst, Pcut, logFCcut) {
  print("Check whether to include a factor into the model by checking the number of features for which the factor is significant when added in the model: ")
  lapply(seq_along(rlst), function(j) {
    r=rlst[[j]]
    print(names(rlst)[j])
    if (names(rlst)[j]!="rlst_NULL") {
      print(sum((lapply(r$rslt_co, function(c) {
        c$adj.P.Val<Pcut
      }) %>% Reduce("+", .))>0, na.rm=T))
    }
  })

  print("Check how many features are significant in different models: ")
  lapply(seq_along(rlst), function(k) {
    r=rlst[[k]]
    print(names(rlst)[k])
    print(sum(r$rslt_of_interest$adj.P.Val<Pcut&abs(r$rslt_of_interest$logFC)>logFCcut, na.rm=T))
  })  
  
  print("Check how many significant features are shared when adding in an extra factor in comparison to only sample group is included in the model: ")
  sigFeatures = lapply(seq_along(rlst), function(k) {
    r=rlst[[k]]
    print(names(rlst)[k])
    rownames(r$rslt_of_interest)[r$rslt_of_interest$adj.P.Val<Pcut&abs(r$rslt_of_interest$logFC)>logFCcut]
  }) %>% setNames(currentCovariates)
  lapply(2:length(sigFeatures), function(l) {
    print(intersect(sigFeatures[[1]], sigFeatures[[l]]) %>% length())
  })
  return(sigFeatures)
}
```

## Limma_Task Depth3 (task level)

```{r}
runCellFie = function(depth, cf5, meta, OUTDIRV, w=9.5, height=23) {
    dir.create(OUTDIRV, recursive = TRUE, showWarnings = FALSE)
    rownames(cf5) = NULL
    cf5_Depth = cf5[, c(depth, "Sample", "TaskScore")] %>% 
      tidyr::spread(key="Sample", value="TaskScore") %>% 
      as.data.frame() %>%
      tibble::column_to_rownames(var=depth) %>% 
      .[, intersect(colnames(.), meta$Sample)]

    meta$Group = factor(meta$Group, 
        levels=c(grep("_N",meta$Group,value=T),grep("_T",meta$Group,value=T)) %>% 
        unique())

    cancerType = unique(meta$Cancer_type)
    nonPaired = c("PRAD", "OV", "GBM")
    rlst_list_paired = lapply(setdiff(cancerType,nonPaired), function(ct) {
        paired = TRUE
        checkCovariate = FALSE
        currentCovariate = "NULL"
        meta_sub = meta[meta$Cancer_type==ct,]
        sub = cf5_Depth[, meta_sub$Sample]
        ## Remove features with 0 variance
        invisible(capture.output({library(matrixStats)}))
        keep = rowSds(sub%>%as.matrix(), na.rm = T) > 0
        sub = sub[keep, ]
        ## Paired analysis with only group in the model
        rlst = limmaTest(dat=sub, paired=paired, 
            currentCovariate=currentCovariate, 
            checkCovariate=checkCovariate,
            meta=meta_sub, 
            paired_variable="Pair_ID")
        sub = sub[rlst$rslt_of_interest %>%
                tibble::rownames_to_column(var="Task")%>%
                arrange(logFC) %>% pull(Task) %>% rev(), ] 
        colAnn = data.frame(Group = colnames(sub) %>% gsub("[-].*", "", .)) 
        rownames(colAnn) = colnames(sub)
        L = ncol(sub)
        pheatmap::pheatmap(sub,
                        show_colnames=TRUE,
                        # labels_col=rep("", ncol(sub)),
                        color=viridisLite::magma(50),
                        breaks=seq(-2,2,len=50),
                        border_color="black",
                        scale="row",
                        annotation_col=colAnn,
                        # annotation_colors=ann_colors,
                        main="Task Scores",
                        cluster_cols = FALSE,
                        cluster_rows = FALSE,
                        filename=paste0(OUTDIRV,"/TaskScore_heatmap_",ct,"_wSampleNames.pdf"), 
                        width=w+0.14*L, height=height+1.5)
        pheatmap::pheatmap(sub,
                        # show_colnames=TRUE,
                        labels_col=rep("", ncol(sub)),
                        color=viridisLite::magma(50),
                        breaks=seq(-2,2,len=50),
                        border_color="black",
                        scale="row",
                        annotation_col=colAnn,
                        # annotation_colors=ann_colors,
                        main="Task Scores",
                        cluster_cols = FALSE,
                        cluster_rows = FALSE,
                        filename=paste0(OUTDIRV,"/TaskScore_heatmap_",ct,".pdf"), 
                        width=w+0.14*L, height=height)
        rlst
    }) %>% setNames(setdiff(cancerType,nonPaired))

    cts = c("PRAD", "OV")
    rlst_list_unpaired = lapply(cts, function(ct) {
        paired = FALSE
        checkCovariate = FALSE
        currentCovariate = "NULL"
        meta_sub = meta[meta$Cancer_type==ct,]
        sub = cf5_Depth[, meta_sub$Sample]
        ## Remove features with 0 variance
        invisible(capture.output({library(matrixStats)}))
        keep = rowSds(sub%>%as.matrix(), na.rm = T) > 0
        sub = sub[keep, ]
        ## Unpaired analysis with only group in the model
        rlst = limmaTest(dat = sub, paired=paired, 
            currentCovariate=currentCovariate, 
            checkCovariate=checkCovariate,
            meta = meta_sub, 
            paired_variable="Pair_ID")
        sub = sub[rlst$rslt_of_interest %>%
                tibble::rownames_to_column(var="Task")%>%
                arrange(logFC) %>% pull(Task) %>% rev(), ] 
        colAnn = data.frame(Group = colnames(sub) %>% gsub("[-].*", "", .)) 
        rownames(colAnn) = colnames(sub)
        L = ncol(sub)
        pheatmap::pheatmap(sub,
                        show_colnames=TRUE,
                        # labels_col=rep("", ncol(sub)),
                        color=viridisLite::magma(50),
                        breaks=seq(-2,2,len=50),
                        border_color="black",
                        scale="row",
                        annotation_col=colAnn,
                        # annotation_colors=ann_colors,
                        main="Task Scores",
                        cluster_cols = FALSE,
                        cluster_rows = FALSE,
                        filename=paste0(OUTDIRV,"/TaskScore_heatmap_",ct,"_wSampleNames.pdf"), 
                        width=w+0.14*L, height=height+1.5)
        pheatmap::pheatmap(sub,
                        # show_colnames=TRUE,
                        labels_col=rep("", ncol(sub)),
                        color=viridisLite::magma(50),
                        breaks=seq(-2,2,len=50),
                        border_color="black",
                        scale="row",
                        annotation_col=colAnn,
                        # annotation_colors=ann_colors,
                        main="Task Scores",
                        cluster_cols = FALSE,
                        cluster_rows = FALSE,
                        filename=paste0(OUTDIRV,"/TaskScore_heatmap_",ct,".pdf"), 
                        width=w+0.14*L, height=height)
        rlst
    }) %>% setNames(cts)

    #here are only cancer samples for this kind of cancer, so can't compare
    ct = "GBM"

    ### Save the limma results
    #Combine all limma results together
    rlst_list = c(rlst_list_paired, rlst_list_unpaired)
    saveRDS(rlst_list, paste0(OUTDIRV, "/result.RDS"))
    TableNames = lapply(names(rlst_list), function(ct) {
        rlst_list[[ct]]$rslt_of_interest %>% 
        tibble::rownames_to_column(var="Task")
    }) %>% setNames(names(rlst_list))
    openxlsx::write.xlsx(TableNames, 
        file=paste0(OUTDIRV, "/Limma_differentialAnalysis_result_Depth3.xlsx"))
    return(list(cf5_Depth=cf5_Depth, rlst_list= rlst_list))
}

depth = "Depth3"
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")
result3 = runCellFie(depth, 
  cf5%>%mutate(TaskScore=log2(TaskScore+0.001)), 
  meta, OUTDIRV, w=9.5, height=23)
cf5_Depth3 = result3$cf5_Depth
rlst_list3 = result3$rlst_list
```

### Upset plot

```{r}
makeUpset = function(rlst_list, adj.P.Val.cutoff, OUTDIRO, h=8, w=18) {
  sig.feature.list = c(lapply(names(rlst_list), function(ct) {
          rlst_list[[ct]]$rslt_of_interest %>% 
          tibble::rownames_to_column(var="Task") %>%
          filter(adj.P.Val<adj.P.Val.cutoff&logFC>0) %>% pull(Task)
      }) %>% setNames(paste0(names(rlst_list), "_up")),
      lapply(names(rlst_list), function(ct) {
          rlst_list[[ct]]$rslt_of_interest %>% 
          tibble::rownames_to_column(var="Task") %>%
          filter(adj.P.Val<adj.P.Val.cutoff&logFC<0) %>% pull(Task)
      }) %>% setNames(paste0(names(rlst_list), "_down")))
  m = sig.feature.list %>% make_comb_mat()
  Orders = names(m)
  pdf(paste0(OUTDIRO, "/Sig.Diff.features_upset.pdf"), h=h, w=w)
      print(UpSet(m,
              top_annotation = upset_top_annotation(m, add_numbers = TRUE),
              right_annotation = upset_right_annotation(m, add_numbers = TRUE),
              comb_order = rev(order(comb_size(m))),
              set_order = Orders,
              column_title = paste0("Significant features (adj.P.Val<0.05)")))
  dev.off()
  graphics.off()
  saveRDS(sig.feature.list, 
      paste0(OUTDIRO, "/Sig.Diff.features_list.RDS"))
  return(sig.feature.list)
}

library(ComplexHeatmap)
OUTDIRO = paste0(OUTDIRV, "/Upset_overlapping/")
dir.create(OUTDIRO, recursive=T, showWarnings = F)
adj.P.Val.cutoff = 0.05
sig.feature.list3 = makeUpset(rlst_list=rlst_list3, 
  adj.P.Val.cutoff, OUTDIRO, h=8, w=18)
```

### Heatmap of all sample scores together

```{r}
makeTaskScoreHeatmap = function(sig.feature.list, cf5_Depth, OUTDIRV, w=9.5,h=23) {
    ups = sig.feature.list[grep("_up", names(sig.feature.list),value=T)] %>% 
        unlist() %>% table() %>%
        .[order(.,decreasing = T)] %>% names()
    downs = sig.feature.list[grep("_down", names(sig.feature.list),value=T)] %>% 
        unlist() %>% table() %>%
        .[order(.,decreasing = T)] %>% names()
    colAnn = data.frame(
      Group = colnames(cf5_Depth) %>% gsub("[-].*", "", .) %>% 
          gsub(".*_", "", .) %>% gsub("T", "Tumor", .) %>%
          gsub("N", "Normal", .) %>% factor(levels=c("Normal", "Tumor")),
      Cancer = colnames(cf5_Depth) %>% gsub("[-].*", "", .) %>%
          gsub("_.*", "", .))
    colAnn$Cancer = factor(colAnn$Cancer, levels = unique(colAnn$Cancer))
    rownames(colAnn) = colnames(cf5_Depth)
    L = ncol(cf5_Depth)
    cf5_Depth_2 = cf5_Depth[c(ups,downs)%>%unique(),]
    pheatmap::pheatmap(cf5_Depth_2,
        show_colnames=TRUE,
        color=viridisLite::magma(50),
        breaks=seq(-2,2,len=50),
        border_color="black",
        scale="row",
        annotation_col=colAnn,
        # annotation_colors=ann_colors,
        main="Task Scores",
        cluster_cols = TRUE,
        cluster_rows = TRUE,
        filename=paste0(OUTDIRV,"/TaskScore_heatmap_all_wSampleNames.pdf"), 
        width=w+0.14*L, height=h)
}

makeTaskScoreHeatmap(sig.feature.list=sig.feature.list3, cf5_Depth=cf5_Depth3, OUTDIRV, w=9.5,h=23)
```

### Dotplot showing limma results all together

```{r}
library(dplyr)
makeLimmaDotplot = function(rlst_list, sig.feature.list, 
  adj.P.Val.cutoff, w=15, h=30, OUTDIRV, lowfigure=F){
    ups = sig.feature.list[grep("_up", names(sig.feature.list),value=T)] %>% 
      unlist() %>% table() %>%
      .[order(.,decreasing = T)] %>% names()
    downs = sig.feature.list[grep("_down", names(sig.feature.list),value=T)] %>% 
      unlist() %>% table() %>%
      .[order(.,decreasing = T)] %>% names()

    # OUTDIR = "../results/CellFie/"
    # OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")
    # rlst_list = readRDS(paste0(OUTDIRV, "/result.RDS"))
    # names(rlst_list)
    rlst_list$OV$rslt_of_interest %>% colnames()
    rlst_long = lapply(seq_along(rlst_list), function(i) {
      rlst = rlst_list[[i]]$rslt_of_interest
      rlst$Cancer = names(rlst_list)[i]
      rlst$Regulation = ifelse(rlst$logFC>0&rlst$adj.P.Val<adj.P.Val.cutoff, "Up", 
        ifelse (rlst$logFC<0&rlst$adj.P.Val<adj.P.Val.cutoff, "Down", "NonSig"))
      rlst %>% tibble::rownames_to_column(var="Task") %>% 
        filter(Task %in% c(ups, downs)) %>%
        filter(adj.P.Val<adj.P.Val.cutoff)
    }) %>% Reduce(rbind, .) 

    rlst_long$Task = factor(rlst_long$Task,
      levels=rev(c(ups, downs) %>% unique()))
    rlst_long$t[rlst_long$t>5] = 5
    rlst_long$t[rlst_long$t< -5] = -5
    # table(rlst_long$Cancer)
    library(ggplot2)
    p = ggplot(data=rlst_long, mapping=aes_string(x="Cancer", y="Task")) +
        geom_point(aes(size=adj.P.Val, shape=Regulation, color=logFC)) +
        # facet_wrap(~Cancer, ncol=length(rlst_list)) + 
        scale_shape_manual(values = c("Up" = 18, "Down" = 20)) +
        ggtitle("Limma_task") +
        xlab("Limma_t") +
        ylab("") + # labs(fill="adj.P.Val") +
        scale_color_gradient2(low="blue", mid="snow2", high="red", midpoint=0) +
        theme_bw(base_size = 15) + 
        scale_size(trans="reverse") +
        guides(alpha="none", size=guide_legend(override.aes=list(shape=21))) +
        theme(strip.text.x = element_text(size = 14),
              axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) +
        labs(caption=paste0("adj.P.ValCutoff = ",adj.P.Val.cutoff,", pAdjustMethod = BH"))
    if (lowfigure) {
      p = p + theme(
              legend.position = "bottom",
              legend.box = "vertical",           # Stack legends vertically
              legend.box.just = "center",        # Center the whole legend box
              legend.text = element_text(size = 10),
              legend.title = element_text(size = 11),
              legend.spacing.y = unit(0.3, "cm") # Reduce spacing between legend rows
            ) +
            guides(
              fill  = guide_legend(nrow = 3, title.position = "top"),
              shape = guide_legend(nrow = 3, title.position = "top"),
              size  = guide_legend(nrow = 3, title.position = "top")
            )
    }

    pdf(paste0(OUTDIRV,"/Dotplot_all_adj.P.ValCutoff",adj.P.Val.cutoff,".pdf"),w=w,h=h)
    show(p)
    dev.off()
}

adj.P.Val.cutoff = 0.05
makeLimmaDotplot(rlst_list=rlst_list3, sig.feature.list=sig.feature.list3, 
  adj.P.Val.cutoff, w=15, h=30, OUTDIRV)

adj.P.Val.cutoff = 0.1
makeLimmaDotplot(rlst_list=rlst_list3, sig.feature.list=sig.feature.list3, 
  adj.P.Val.cutoff, w=15, h=30, OUTDIRV)

adj.P.Val.cutoff = 1
makeLimmaDotplot(rlst_list=rlst_list3, sig.feature.list=sig.feature.list3, 
  adj.P.Val.cutoff, w=15, h=30, OUTDIRV)
```

### Boxplot/Barplot

#### Showing group wise TaskScore

```{r}
# Depth3
cf5 = cf5 %>% dplyr::left_join(meta) %>% mutate(Cancer=Cancer_type) %>%
  mutate(Group=paste0(Cancer_type,"_",Condition))
boxplot_D3 = ggplot(cf5, aes(x=Group, y=TaskScore, fill=Cancer)) +
  geom_boxplot(position=position_dodge()) +
  facet_wrap(~Depth3, ncol=10, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
pdf(paste0(OUTDIR,"/Summary/Boxplots_TaskScores_depth3.pdf"),w=62,h=48)
  show(boxplot_D3)
dev.off()

# Depth2
Depth2 = Depth2 %>% dplyr::left_join(meta) %>% mutate(Cancer=Cancer_type) %>%
  mutate(Group=paste0(Cancer_type,"_",Condition))
boxplot_D2 = ggplot(Depth2, aes(x=Group, y=SumTaskScore, fill=Cancer)) +
  geom_boxplot(position=position_dodge()) +
  facet_wrap(~Depth2, ncol=10, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
pdf(paste0(OUTDIR,"/Summary/Boxplots_TaskScoreSums_depth2.pdf"),w=50,h=24)
  show(boxplot_D2)
dev.off()
boxplot_D2_2 = ggplot(Depth2, aes(x=Group, y=MeanTaskScore, fill=Cancer)) +
  geom_boxplot(position=position_dodge()) +
  facet_wrap(~Depth2, ncol=10, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
pdf(paste0(OUTDIR,"/Summary/Boxplots_TaskScoreMeans_depth2.pdf"),w=50,h=24)
  show(boxplot_D2_2)
dev.off()

# Depth1
Depth1 = Depth1 %>% dplyr::left_join(meta) %>% mutate(Cancer=Cancer_type) %>%
  mutate(Group=paste0(Cancer_type,"_",Condition))
boxplot_D1 = ggplot(Depth1, aes(x=Group, y=SumTaskScore, fill=Cancer)) +
  geom_boxplot(position=position_dodge()) +
  facet_wrap(~Depth1, ncol=4, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
pdf(paste0(OUTDIR,"/Summary/Boxplots_TaskScoreSums_depth1.pdf"),w=21,h=7)
  show(boxplot_D1)
dev.off()
boxplot_D1_2 = ggplot(Depth1, aes(x=Group, y=MeanTaskScore, fill=Cancer)) +
  geom_boxplot(position=position_dodge()) +
  facet_wrap(~Depth1, ncol=4, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
pdf(paste0(OUTDIR,"/Summary/Boxplots_TaskScoreMeans_depth1.pdf"),w=21,h=7)
  show(boxplot_D1_2)
dev.off()

```

#### Showing group-wise BinaryTaskScore

```{r}
# Depth3
cols_to_group = c("Group", "Depth3")
Depth3B = cf5 %>%
    group_by(across(all_of(cols_to_group))) %>%
    summarise(
        NrOfActiveTasks = sum(BinaryTaskScore==1, na.rm = TRUE)
    ) %>%  mutate(Cancer=gsub("_.*", "", Group)) %>%
    dplyr::select(Group, Cancer, everything()) 

boxplot_DB3 = ggplot(Depth3B, aes(x=Group, y=NrOfActiveTasks, fill=Cancer)) +
  geom_bar(stat="identity", position=position_dodge()) +
  facet_wrap(~Depth3, ncol=10, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
pdf(paste0(OUTDIR,"/Summary/Boxplots_NrOfActiveTasks_depth3.pdf"),w=62,h=48)
  show(boxplot_DB3)
dev.off()

# Depth2
cols_to_group = c("Group", "Depth2")
Depth2B = cf5 %>%
    group_by(across(all_of(cols_to_group))) %>%
    summarise(
        NrOfActiveTasks = sum(BinaryTaskScore==1, na.rm = TRUE)
    ) %>%  mutate(Cancer=gsub("_.*", "", Group)) %>%
    dplyr::select(Group, Cancer, everything()) 

boxplot_DB2 = ggplot(Depth2B, aes(x=Group, y=NrOfActiveTasks, fill=Cancer)) +
  geom_bar(stat="identity", position=position_dodge()) +
  facet_wrap(~Depth2, ncol=10, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
pdf(paste0(OUTDIR,"/Summary/Boxplots_NrOfActiveTasks_depth2.pdf"),w=50,h=24)
  show(boxplot_DB2)
dev.off()

# Depth1
cols_to_group = c("Group", "Depth1")
Depth1B = cf5 %>%
    group_by(across(all_of(cols_to_group))) %>%
    summarise(
        NrOfActiveTasks = sum(BinaryTaskScore==1, na.rm = TRUE)
    ) %>%  mutate(Cancer=gsub("_.*", "", Group)) %>%
    dplyr::select(Group, Cancer, everything()) 

boxplot_DB1 = ggplot(Depth1B, aes(x=Group, y=NrOfActiveTasks, fill=Cancer)) +
  geom_bar(stat="identity", position=position_dodge()) +
  facet_wrap(~Depth1, ncol=4, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
pdf(paste0(OUTDIR,"/Summary/Boxplots_NrOfActiveTasks_depth1.pdf"),w=21,h=7)
  show(boxplot_DB1)
dev.off()

```

#### Showing group-wise TaskScore\*BinaryTaskScore

```{r}
# Depth3
boxplot_D3 = ggplot(cf5, aes(x=Group, y=ActiveTaskScore, fill=Cancer)) +
  geom_boxplot(position=position_dodge()) +
  facet_wrap(~Depth3, ncol=10, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
pdf(paste0(OUTDIR,"/Summary/Boxplots_ActiveTaskScores_depth3.pdf"),w=62,h=48)
  show(boxplot_D3)
dev.off()

# Depth2
boxplot_D2 = ggplot(Depth2, aes(x=Group, y=SumActiveTaskScore, fill=Cancer)) +
  geom_boxplot(position=position_dodge()) +
  facet_wrap(~Depth2, ncol=10, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
pdf(paste0(OUTDIR,"/Summary/Boxplots_ActiveTaskScoreSums_depth2.pdf"),w=50,h=24)
  show(boxplot_D2)
dev.off()
boxplot_D2_2 = ggplot(Depth2, aes(x=Group, y=MeanActiveTaskScore, fill=Cancer)) +
  geom_boxplot(position=position_dodge()) +
  facet_wrap(~Depth2, ncol=10, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
pdf(paste0(OUTDIR,"/Summary/Boxplots_ActiveTaskScoreMeans_depth2.pdf"),w=50,h=24)
  show(boxplot_D2_2)
dev.off()

# Depth1
boxplot_D1 = ggplot(Depth1, aes(x=Group, y=SumActiveTaskScore, fill=Cancer)) +
  geom_boxplot(position=position_dodge()) +
  facet_wrap(~Depth1, ncol=4, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
pdf(paste0(OUTDIR,"/Summary/Boxplots_ActiveTaskScoreSums_depth1.pdf"),w=21,h=7)
  show(boxplot_D1)
dev.off()
boxplot_D1_2 = ggplot(Depth1, aes(x=Group, y=MeanActiveTaskScore, fill=Cancer)) +
  geom_boxplot(position=position_dodge()) +
  facet_wrap(~Depth1, ncol=4, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
pdf(paste0(OUTDIR,"/Summary/Boxplots_ActiveTaskScoreMeans_depth1.pdf"),w=21,h=7)
  show(boxplot_D1_2)
dev.off()

```

## Limma_Task Depth2 (task level)

```{r}
depth = "Depth2"
OUTDIRV2 = paste0(OUTDIR, "/Limma/TaskDepth2/")
result2 = runCellFie(depth, cf5=Depth2%>%mutate(TaskScore=log2(TaskScore+0.001)), 
  meta, OUTDIRV=OUTDIRV2, w=5.5, height=11)
cf5_Depth2 = result2$cf5_Depth
rlst_list2 = result2$rlst_list
```

### Upset plot

```{r}
library(ComplexHeatmap)
OUTDIRO2 = paste0(OUTDIRV2, "/Upset_overlapping/")
dir.create(OUTDIRO2, recursive=T, showWarnings = F)
adj.P.Val.cutoff = 0.05
sig.feature.list2 = makeUpset(rlst_list=rlst_list2, 
  adj.P.Val.cutoff, OUTDIRO=OUTDIRO2, h=6.5, w=11)
```

### Heatmap of all sample scores together

```{r}
makeTaskScoreHeatmap(sig.feature.list=sig.feature.list2, 
  cf5_Depth=cf5_Depth2, OUTDIRV=OUTDIRV2, w=7,h=16)
```

### Dotplot showing limma results all together

```{r}
adj.P.Val.cutoff = 0.05
makeLimmaDotplot(rlst_list=rlst_list2, sig.feature.list=sig.feature.list2, 
  adj.P.Val.cutoff, w=9, h=15, OUTDIRV=OUTDIRV2)

adj.P.Val.cutoff = 0.1
makeLimmaDotplot(rlst_list=rlst_list2, sig.feature.list=sig.feature.list2, 
  adj.P.Val.cutoff, w=9, h=15, OUTDIRV=OUTDIRV2)

adj.P.Val.cutoff = 1
makeLimmaDotplot(rlst_list=rlst_list2, sig.feature.list=sig.feature.list2, 
  adj.P.Val.cutoff, w=9, h=15, OUTDIRV=OUTDIRV2)
```

## Limma_Task Depth1 (task level)

```{r}
depth = "Depth1"
OUTDIRV1 = paste0(OUTDIR, "/Limma/TaskDepth1/")
result1 = runCellFie(depth, cf5=Depth1%>%mutate(TaskScore=log2(TaskScore+0.001)), 
  meta, OUTDIRV=OUTDIRV1, w=5.5, height=2)
cf5_Depth1 = result1$cf5_Depth
rlst_list1 = result1$rlst_list

```

### Upset plot

```{r}
library(ComplexHeatmap)
OUTDIRO1 = paste0(OUTDIRV1, "/Upset_overlapping/")
dir.create(OUTDIRO1, recursive=T, showWarnings = F)
adj.P.Val.cutoff = 0.05
sig.feature.list1 = makeUpset(rlst_list=rlst_list1, 
  adj.P.Val.cutoff, OUTDIRO=OUTDIRO1, h=6, w=3)
```

### Heatmap of all sample scores together

```{r}
makeTaskScoreHeatmap(sig.feature.list=sig.feature.list1, 
  cf5_Depth=cf5_Depth1, OUTDIRV=OUTDIRV1, w=7,h=4)
```

### Dotplot showing limma results all together

```{r}
adj.P.Val.cutoff = 0.05
makeLimmaDotplot(rlst_list=rlst_list1, sig.feature.list=sig.feature.list1, 
  adj.P.Val.cutoff, w=6, h=7, OUTDIRV=OUTDIRV1, lowfigure=T)

adj.P.Val.cutoff = 0.1
makeLimmaDotplot(rlst_list=rlst_list1, sig.feature.list=sig.feature.list1, 
  adj.P.Val.cutoff, w=6, h=7, OUTDIRV=OUTDIRV1, lowfigure=T)

adj.P.Val.cutoff = 1
makeLimmaDotplot(rlst_list=rlst_list1, sig.feature.list=sig.feature.list1, 
  adj.P.Val.cutoff, w=6, h=7, OUTDIRV=OUTDIRV1, lowfigure=T)
```

# Volcano plot

```{r}
library(ggplot2)
doVolcano = function(resi, NAME, adj.P.Val.cutoff=0.05, logFC.cutoff=0.3, lfcShrink=F) {
  selectedCols = c("Task", "logFC", "P.Value", "adj.P.Val", "Depth")
  # resi$Protein = gsub("sp[|]", "", resi$Protein) %>%
  #   sub("[|].*", "", .)
  toExport = resi[, selectedCols]
  toExport = toExport[order(toExport$P.Value),]

  tmp = as.data.frame(toExport) %>%
    filter(!is.na(adj.P.Val)) %>%
    mutate(sig = ifelse(adj.P.Val<adj.P.Val.cutoff&logFC>logFC.cutoff, "up", 
    ifelse(adj.P.Val<adj.P.Val.cutoff&(logFC< -logFC.cutoff), "down", "nonSig")))
  tmp$PointSize = ifelse(tmp$Depth == "Depth3", 1,
                    ifelse(tmp$Depth == "Depth2", 2, 3))
  tmpSignif = tmp %>%
    filter(tmp$logFC>quantile(abs(tmp$logFC),0.98)|(tmp$logFC< -quantile(abs(tmp$logFC),0.98))) %>%
    filter(sig!="nonSig")
  
  p = ggplot(data = tmp, aes(x=logFC, y=-log10(adj.P.Val), col=sig)) + 
    geom_point(alpha = 0.5, aes(size=PointSize)) + 
    xlab("Log2 Fold Change") +
    ggrepel::geom_text_repel(data = tmpSignif, 
      aes(x=logFC, y=-log10(adj.P.Val), label = Task), 
      color="cornsilk4") +
    geom_hline(yintercept = -log10(0.05), lty = 2) +
    geom_vline(xintercept = c(-0, 0), lty = 2) + 
    scale_colour_manual(values = c("down"="blue", "nonSig"="black", "up"="red")) + 
    theme(legend.position="none") +
    ggtitle(ifelse(lfcShrink, paste0(NAME, "_lfcShrink"), paste0(NAME, "")))
  return(p)
}

library(ggplot2)
adj.P.Val.cutoff = 0.05
plots = lapply(names(result1$rlst_list), function(cancer) {
  dat2plot = rbind(result1$rlst_list[[cancer]]$rslt_of_interest %>% 
                      tibble::rownames_to_column(var="Task") %>%
                      #filter(adj.P.Val<adj.P.Val.cutoff) %>%
                      mutate(Depth = "Depth1"),
                   result2$rlst_list[[cancer]]$rslt_of_interest %>% 
                      tibble::rownames_to_column(var="Task") %>%
                      #filter(adj.P.Val<adj.P.Val.cutoff)%>%
                      mutate(Depth = "Depth2"),
                   result3$rlst_list[[cancer]]$rslt_of_interest%>% 
                      tibble::rownames_to_column(var="Task") %>%
                      #filter(adj.P.Val<adj.P.Val.cutoff)%>%
                      mutate(Depth = "Depth3"))
  #print(head(dat2plot))
  p = doVolcano(resi=dat2plot, NAME=cancer, adj.P.Val.cutoff=0.05, 
                logFC.cutoff=0, lfcShrink=F)
  p
}) 

pdf(paste0(OUTDIR, "/Limma/Volcano_plots.pdf"),h=4,w=6)
  plots
dev.off()

pdf(paste0(OUTDIR, "/Limma/Volcano_plots2.pdf"),h=20,w=18)
  cowplot::plot_grid(plots[[1]],plots[[2]],plots[[3]],plots[[4]], 
                     plots[[5]],plots[[6]],plots[[7]],plots[[8]],
                     plots[[9]],plots[[10]],plots[[11]],plots[[12]],
                     plots[[13]],plots[[14]],plots[[15]],
                     ncol=3,nrow=5)
dev.off()

```

# Network

```{r}
TaskNetworks = function(pair, logFCcutoff=0, sig.cutoff=0.05, suffix, height, width) {
  combined_limmaTable = 
    Reduce(rbind,list(result1$rlst_list[[pair]]$rslt_of_interest, 
                      result2$rlst_list[[pair]]$rslt_of_interest,
                      result3$rlst_list[[pair]]$rslt_of_interest)) %>%
    as.data.frame()  
  
  library(igraph)
  # Define edges (from parent to child)
  edges = cf5 %>% dplyr::select(Depth1, Depth2) %>% setNames(c("", "")) %>%
    rbind(cf5 %>% dplyr::select(Depth2, Depth3) %>% setNames(c("", ""))) %>%
    unique() %>% as.matrix()
  
  g = graph_from_edgelist(edges, directed = TRUE)
  V(g)$color = ifelse(V(g)$name%in%cf5$Depth1, adjustcolor("coral4", alpha.f = 0.7), 
                      ifelse(V(g)$name%in%cf5$Depth2, "grey" ,"skyblue"))
  
  label_up = combined_limmaTable %>% filter(logFC>0) %>% rownames(.)
  label_dn = combined_limmaTable %>% filter(logFC<0) %>% rownames(.)
  label.colors = ifelse(V(g)$name%in%label_up, "darkred", 
                        ifelse(V(g)$name%in%label_dn, "darkblue", "black"))
  
  labelsig = combined_limmaTable %>% 
    filter(abs(logFC)>logFCcutoff&adj.P.Val<sig.cutoff) %>% rownames(.)
  edge.colors = ifelse(V(g)$name%in%labelsig,  adjustcolor("orange", alpha.f=1), "white")
  
  V(g)$name[V(g)$name%in%labelsig] = paste0(V(g)$name[V(g)$name%in%labelsig], "*")
  
  sizes = rep(5, length(V(g)))
  
  set.seed(808)
  plot(g, 
       layout = layout_with_fr(g), 
       vertex.label = V(g)$name,
       vertex.frame.color = edge.colors,
       vertex.size = sizes,
       vertex.label.color = label.colors,
       vertex.label.cex = 1.2, 
       edge.arrow.size = 0.5,
       vertex.label.font = 2,
       vertex.frame.width=6)
  paste0(OUTDIR,"/Network/") %>% dir.create(recursive=T, showWarnings=F)
  pdf(paste0(OUTDIR,"/Network/NetworksOfTaskDepths_",pair,"_sig.cutoff", 
      suffix, ".pdf"),h=height,w=width)
  set.seed(808)
  plot(g, 
       layout = layout_with_fr(g), 
       # vertex.label = V(g)$name,
       vertex.frame.color = edge.colors,
       vertex.size = sizes,
       vertex.label.color = label.colors,
       vertex.label.cex = 1.2, 
       edge.arrow.size = 0.5,
       vertex.label.font = 2,
       vertex.frame.width=6,
       main=paste0(pair,"_", suffix))
  dev.off()
  return(g)
}

Pairs = names(result1$rlst_list)
gs = lapply(Pairs, function(pair) TaskNetworks(pair, suffix="adj.P.valueCutoff0.05", 
    height=40, width=40))
gs = lapply(Pairs, function(pair) TaskNetworks(pair, suffix="adj.P.valueCutoff0.05_larger", 
    height=80, width=80))

# # install.packages("visNetwork")
# library(visNetwork)
# # Convert to data frames
# nodes = data.frame(id = V(gs[[1]])$name) # %>%
# #   mutate(group = ifelse(V(gs[[1]])$name%in%cf5$Depth1, "Depth1", 
# #                       ifelse(V(gs[[1]])$name%in%cf5$Depth2, "Depth2",
# #                       ifelse(V(gs[[1]])$name%in%cf5$Depth3, "Depth3", "NA"))))
# edges = as_data_frame(gs[[1]])
# # Create interactive plot
# vis = visNetwork(nodes, edges) %>% visLegend() %>% 
#   visOptions(manipulation = TRUE)
# visSave(vis, file = "network.html")

```

# Classification

## Based on ENERGY METABOLISM

```{r}
rownames(cf5) = NULL
depth = "Depth3"
EnergyD3 = cf5 %>% filter(Depth1 == "ENERGY METABOLISM") %>%
  .[, c(depth, "Sample", "TaskScore")] %>% 
  tidyr::spread(key="Sample", value="TaskScore") %>% 
  as.data.frame() %>%
  tibble::column_to_rownames(var=depth) %>% 
  .[, intersect(colnames(.), meta$Sample)]

EnergyD3[,1:3]

c_pca = prcomp(t(na.omit(EnergyD3)),
               scale=TRUE, center = TRUE)
frac_var = function(x) x^2/sum(x^2)
var_explained = c_pca$sdev %>% 
        tibble::as_tibble() %>% 
        frac_var() %>% 
        round(digits = 2) %>%
        pull(value)
OUTDIRC = paste0(OUTDIR, "/Classification/")
dir.create(OUTDIRC, recursive=T, showWarnings=F)
pdf(paste0(OUTDIRC,"ENERGYMETABOLISM_PCA.pdf"))
plot(1:length(var_explained), var_explained, type = "b", pch = 19,
     xlab = "Number of PCs",
     ylab = "Variance explained",
     main = "Elbow Method for Determining Optimal PCs")
dev.off()

# ggplot(as.data.frame(c_pca$x), aes(PC1, PC2)) + geom_point() +
#   ggrepel::geom_label_repel(aes(label=rownames(c_pca$x)))

print("Clustering ...")
# BiocManager::install("ConsensusClusterPlus")
library(ConsensusClusterPlus)
results = ConsensusClusterPlus(
  t(scale(t(EnergyD3))),         # data matrix (rows = features, cols = samples)
  maxK = 6,        # maximum number of clusters (K) to evaluate
  reps = 1000,     # number of resampling iterations
  pItem = 0.8,     # proportion of samples per iteration
  pFeature = 1,    # proportion of features per iteration
  clusterAlg = "km",              # "hc" (hierarchical) or "km" (k-means) or "pam"
  # distance = "pearson",           # distance metric: "pearson", "euclidean", etc.
  # innerLinkage = "average",       # linkage for subsample clustering
  # finalLinkage = "average",       # linkage for consensus clustering
  seed = 1234,                    # for reproducibility
  title = "ConsensusCluster",     # output prefix
  plot = "png",                   # output image format (“png”, “pdf”, or NULL to suppress)
  writeTable = TRUE               # write consensus matrices and class assignments
)

# compute quantitative stability scores:
icl = calcICL(results)
# This returns clusterConsensus (mean pairwise consensus within each cluster) and itemConsensus (per-sample stability), enabling you to identify poorly supported assignments 

met = lapply(2:6, function(i) {
  test = icl$itemConsensus %>% filter(k==i) %>%  
    arrange(desc(itemConsensus))
  test = test[!duplicated(test$item), ]
  print(mean(test$itemConsensus))
}) %>% setNames(2:6)
sapply(seq_along(met), function(i) {
  print(paste0("Mean of item consensus indices across all samples for k==", 
               names(met)[i], ": ", met[i]))})
max = (2:6)[which.max(met)]
print(paste0("\n The k with the maximum of mean Item-consensus index across all samples is ", max))

test = icl$itemConsensus %>% filter(k==max) %>% 
    mutate(degree = ifelse(itemConsensus>0.4, 1, 0)) %>% filter(degree>0) %>% 
  arrange(desc(itemConsensus))
test2 = test[!duplicated(test$item), ]
mapping2 = setNames(test2$cluster, test2$item)
pca_mat = c_pca$x %>% as.data.frame()
pca_mat$Cluster = as.factor(mapping2[rownames(pca_mat)])
pca_mat$Cancer = gsub("_.*", "", rownames(pca_mat))
pca_mat$Condition = gsub("[-].*", "", rownames(pca_mat)) %>%
  gsub(".*_", "", .)
mapping3 = setNames(meta$Tissue, meta$Sample)
pca_mat$Tissue = mapping3[rownames(pca_mat)]
pca_mat = pca_mat %>% cbind(t(EnergyD3)[rownames(pca_mat), ])

saveRDS(results, paste0(OUTDIRC, "ENERGYMETABOLISM_Classification_results.RDS"))
saveRDS(pca_mat, paste0(OUTDIRC, "ENERGYMETABOLISM_pca_mat.RDS"))
file.rename("ConsensusCluster", paste0(OUTDIRC,"/ENERGYMETABOLISM_ConsensusCluster"))

library(RColorBrewer)
mycolors = c(brewer.pal(name="Dark2", n = 8), brewer.pal(name="Paired", n = 8))
pdf(paste0(OUTDIRC, "ENERGYMETABOLISM_Classification_results.pdf"),h=30,w=14)
p01 = ggplot(pca_mat, aes(PC1, PC2)) + 
  geom_point(aes(shape=Cluster, col=Cancer)) +
  theme_bw() +
  scale_color_manual(values = mycolors)
  # ggrepel::geom_label_repel(aes(label=PatID))
p02 = ggplot(pca_mat, aes(PC1, PC2)) + 
  geom_point(aes(shape=Condition, col=Cancer)) +
  theme_bw() +
  scale_color_manual(values = mycolors)
  # ggrepel::geom_label_repel(aes(label=PatID))
p1 = ggplot(pca_mat, aes(PC1, PC2)) + 
  geom_point(aes(shape=Cluster, col=Tissue)) +
  theme_bw() +
  scale_color_manual(values = mycolors)
  # ggrepel::geom_label_repel(aes(label=PatID))
p2 = ggplot(pca_mat, aes(PC1, PC2)) + 
  geom_point(aes(shape=Condition, col=Tissue)) +
  theme_bw() +
  scale_color_manual(values = mycolors)
  # ggrepel::geom_label_repel(aes(label=PatID))
p3 = ggplot(pca_mat, aes(PC1, PC2)) + 
  geom_point(aes(shape=Condition, col=scale(`ATP generation from glucose (hypoxic conditions) - glycolysis`))) +
  theme_bw() +
  scale_color_gradient2(low="blue", mid="snow2", high="red", midpoint=0)
  # scale_color_manual(values = mycolors)
  # ggrepel::geom_label_repel(aes(label=PatID))
p4 = ggplot(pca_mat, aes(PC1, PC2)) + 
  geom_point(aes(shape=Condition, col=scale(`Krebs cycle - NADH generation`))) +
  theme_bw() +
  scale_color_gradient2(low="blue", mid="snow2", high="red", midpoint=0)
  # scale_color_manual(values = mycolors)
  # ggrepel::geom_label_repel(aes(label=PatID))
p5 = ggplot(pca_mat, aes(PC1, PC2)) + 
  geom_point(aes(shape=Cluster, col=scale(`Krebs cycle - oxidative decarboxylation of pyruvate`))) +
  theme_bw() +
  scale_color_gradient2(low="blue", mid="snow2", high="red", midpoint=0)
  # scale_color_manual(values = mycolors)
  # ggrepel::geom_label_repel(aes(label=PatID))
p6 = ggplot(pca_mat, aes(PC1, PC2)) + 
  geom_point(aes(shape=Cluster, col=scale(`Presence of the thioredoxin system through the thioredoxin reductase activity`))) +
  theme_bw() +
  scale_color_gradient2(low="blue", mid="snow2", high="red", midpoint=0)
  # scale_color_manual(values = mycolors)
  # ggrepel::geom_label_repel(aes(label=PatID))
library(ggpubr)
ggarrange(    
  ggarrange(p01,p02, ncol = 2, labels = c("G", "H")),                                             # First row with scatter plot
  ggarrange(p1, p2, ncol = 2, labels = c("A", "B")),
  p3,p4,p5,p6, # Second row with box and dot plots
  nrow = 6, 
  labels = c("C", "D", "E", "F")                                        # Labels of the scatter plot
  ) 
dev.off()

temp = pca_mat %>% tibble::rownames_to_column(var="Sample") %>%
  dplyr::left_join(meta) %>% 
  mutate(Condition=gsub("N", "Normal", Condition)) %>%
  mutate(Condition=gsub("T", "Tumor", Condition)) %>%
  mutate(Group = paste0(Cancer_type, "_", Condition))
TableNames = list(Classification_result = temp)
openxlsx::write.xlsx(TableNames, 
    file=paste0(OUTDIRC, "/ENERGYMETABOLISM_Classification_result.xlsx"))
pdf(paste0(OUTDIRC, "ENERGYMETABOLISM_Classification_results_stackedBarplot.pdf"), h=2.5, w=6)
temp %>% dplyr::count(Cluster, Group) %>% mutate(SampleNrPercent = n) %>%
  mutate(Cancer=gsub("_.*", "", Group)) %>%
  mutate(Condition=gsub(".*_", "", Group)) %>%
  ggplot(., aes(x=Group, y=SampleNrPercent, fill=Cluster)) +
  geom_bar(stat="identity", position="fill") +
  # facet_wrap(~Cluster, ncol=2, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
dev.off()

```

## Based on KREBS CYCLE

```{r}
rownames(cf5) = NULL
depth = "Depth3"
EnergyD2 = cf5 %>% filter(Depth2 == "KREBS CYCLE") %>%
  .[, c(depth, "Sample", "TaskScore")] %>% 
  tidyr::spread(key="Sample", value="TaskScore") %>% 
  as.data.frame() %>%
  tibble::column_to_rownames(var=depth) %>% 
  .[, intersect(colnames(.), meta$Sample)]

print("Clustering ...")
# BiocManager::install("ConsensusClusterPlus")
library(ConsensusClusterPlus)
results = ConsensusClusterPlus(
  t(scale(t(EnergyD2))),         # data matrix (rows = features, cols = samples)
  maxK = 6,        # maximum number of clusters (K) to evaluate
  reps = 1000,     # number of resampling iterations
  pItem = 0.8,     # proportion of samples per iteration
  pFeature = 1,    # proportion of features per iteration
  clusterAlg = "km",              # "hc" (hierarchical) or "km" (k-means) or "pam"
  # distance = "pearson",           # distance metric: "pearson", "euclidean", etc.
  # innerLinkage = "average",       # linkage for subsample clustering
  # finalLinkage = "average",       # linkage for consensus clustering
  seed = 1234,                    # for reproducibility
  title = "ConsensusCluster",     # output prefix
  plot = "png",                   # output image format (“png”, “pdf”, or NULL to suppress)
  writeTable = TRUE               # write consensus matrices and class assignments
)

# compute quantitative stability scores:
icl = calcICL(results)
# This returns clusterConsensus (mean pairwise consensus within each cluster) and itemConsensus (per-sample stability), enabling you to identify poorly supported assignments 

met = lapply(2:6, function(i) {
  test = icl$itemConsensus %>% filter(k==i) %>%  
    arrange(desc(itemConsensus))
  test = test[!duplicated(test$item), ]
  print(mean(test$itemConsensus))
}) %>% setNames(2:6)
sapply(seq_along(met), function(i) {
  print(paste0("Mean of item consensus indices across all samples for k==", 
               names(met)[i], ": ", met[i]))})
max = (2:6)[which.max(met)]
print(paste0("\n The k with the maximum of mean Item-consensus index across all samples is ", max))

test = icl$itemConsensus %>% filter(k==max) %>% 
    mutate(degree = ifelse(itemConsensus>0.4, 1, 0)) %>% filter(degree>0) %>% 
  arrange(desc(itemConsensus))
test2 = test[!duplicated(test$item), ]
mapping2 = setNames(test2$cluster, test2$item)
pca_mat = t(EnergyD2) %>% as.data.frame() %>%
  tibble::rownames_to_column(var="Sample") %>%
  mutate(Cluster=as.factor(mapping2[Sample])) %>%
  dplyr::left_join(meta) %>%
  mutate(Condition = gsub(".*_", "", Group)%>% gsub("N","Normal",.)%>%gsub("T","Tumor",.))

saveRDS(results, paste0(OUTDIRC, "KREBSCYCLE_Classification_results.RDS"))
saveRDS(pca_mat, paste0(OUTDIRC, "KREBSCYCLE_pca_mat.RDS"))
file.rename("ConsensusCluster", paste0(OUTDIRC,"/KREBSCYCLE_ConsensusCluster"))


library(RColorBrewer)
mycolors = c(brewer.pal(name="Dark2", n = 8), brewer.pal(name="Paired", n = 8))
pdf(paste0(OUTDIRC, "KREBSCYCLE_Classification_results.pdf"),h=10,w=14)
p01 = ggplot(pca_mat, aes(x=`Krebs cycle - NADH generation`, y=`Krebs cycle - oxidative decarboxylation of pyruvate`)) + 
  geom_point(aes(shape=Cluster, col=Cancer_type)) +
  theme_bw() +
  scale_color_manual(values = mycolors)

p02 = ggplot(pca_mat, aes(`Krebs cycle - NADH generation`, 
  `Krebs cycle - oxidative decarboxylation of pyruvate`)) + 
  geom_point(aes(shape=Condition, col=Cancer_type)) +
  theme_bw() +
  scale_color_manual(values = mycolors)
  # ggrepel::geom_label_repel(aes(label=PatID))
p1 = ggplot(pca_mat, aes(`Krebs cycle - NADH generation`, 
  `Krebs cycle - oxidative decarboxylation of pyruvate`)) + 
  geom_point(aes(shape=Cluster, col=Tissue)) +
  theme_bw() +
  scale_color_manual(values = mycolors)
  # ggrepel::geom_label_repel(aes(label=PatID))
p2 = ggplot(pca_mat, aes(`Krebs cycle - NADH generation`, 
  `Krebs cycle - oxidative decarboxylation of pyruvate`)) + 
  geom_point(aes(shape=Condition, col=Tissue)) +
  theme_bw() +
  scale_color_manual(values = mycolors)
  # ggrepel::geom_label_repel(aes(label=PatID))

library(ggpubr)
ggarrange(    
  ggarrange(p01,p02, ncol = 2, labels = c("G", "H")),                                             # First row with scatter plot
  ggarrange(p1, p2, ncol = 2, labels = c("A", "B")),
  nrow = 2                                      # Labels of the scatter plot
  ) 
dev.off()

temp = pca_mat %>% tibble::rownames_to_column(var="Sample") %>%
  dplyr::left_join(meta) %>% 
  mutate(Condition=gsub("N", "Normal", Condition)) %>%
  mutate(Condition=gsub("T", "Tumor", Condition)) %>%
  mutate(Group = paste0(Cancer_type, "_", Condition))
TableNames = list(Classification_result = temp)
openxlsx::write.xlsx(TableNames, 
    file=paste0(OUTDIRC, "/KREBSCYCLE_Classification_result.xlsx"))
pdf(paste0(OUTDIRC, "KREBSCYCLE_Classification_results_stackedBarplot.pdf"), h=2.5, w=6)
temp %>% dplyr::count(Cluster, Group) %>% mutate(SampleNrPercent = n) %>%
  mutate(Cancer=gsub("_.*", "", Group)) %>%
  mutate(Condition=gsub(".*_", "", Group)) %>%
  ggplot(., aes(x=Group, y=SampleNrPercent, fill=Cluster)) +
  geom_bar(stat="identity", position="fill") +
  # facet_wrap(~Cluster, ncol=2, scales="free") +
  theme_bw() +
  theme(strip.text.x = element_text(size = 14),
    axis.text.x = element_text(angle = 90, hjust=1, vjust=0)) 
dev.off()

```

# SessionInfo

```{r}
out = capture.output(
  sessionInfo()
)
writeLines(out, paste0(OUTDIR,"/sessionInfo.txt"))
out
```
