---
title: Cancer metabolism
subtitle: 
author: Yuan Li
toc: true
number-sections: true
format: 
  html:
    #embed-resources: true #This works from command line not from Rstudio
    code-fold: true
    code-tools: true
    code-link: true
    theme: pandoc
    df-print: paged
    html-math-method: katex #use katex for math text
  pdf:
    code-line-numbers: true
    geometry: 
      - top=30mm
      - left=30mm
  docx: default
language:
  code-summary: Click to show code
editor_options:
  chunk_output_type: inline
editor: visual
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 7,
                      results    = "hold",
                      message    = FALSE,
                      cache.lazy = FALSE,
                      warning    = FALSE)

```

# Import and Prepare the data for CellFie

```{r}
# remove.packages("parseMetab")
# detach("package:parseMetab", unload = TRUE)
# remotes::install_github("Maj18/parseMetab", dependencies = TRUE)
library(parseMetab)
library(ggplot2)
library(dplyr)
library(parseMetab)
```

```{r}
Pdat = 
  readxl::read_xlsx("../data/Zhou2020/13045_2020_1013_MOESM3_ESM*.xlsx") %>%
  mutate(genes = gsub(";.*", "", `Gene Group`)) %>%
  select(genes, everything(), -`Gene Group`, -Index, -`Protein Group`)
dim(Pdat)
```

## Check whether the data has been normalized or not

```{r}
#' @import ggplot2
overviewPlot = function(data, variable2plot="TotalAbundance", plot="barplot", 
                        unit="Sample", log=F, ylim=F) {
  if (variable2plot=="TotalAbundance") {
    temp = colSums(data, na.rm=T)
    if (log) temp=log2(temp+0.0001)
  } else if (variable2plot=="Nr.MissingValues") {
    temp = apply(data, 2, function(col) {
            sum(is.na(col))})
  } else if (variable2plot=="Nr.NonMissingValues") {
    temp = apply(data, 2, function(col) {
            sum(!is.na(col))})
  } else {
    temp = colSums(data, na.rm=T)
    if (log) temp=log2(temp+0.0001)
  }
  if (variable2plot %in% c("TotalAbundance","Nr.MissingValues","Nr.NonMissingValues")) {
    temp2 = data.frame(varaible2plot=as.numeric(temp),
                       Sample=names(temp),
                       Group=sub("_([^_]+)$", "", names(temp)))
    if (log) {
      variable2plot = paste0(variable2plot, "_log2")
    }
    colnames(temp2)[1] = variable2plot
  }
  if (plot == "boxplot") {
    if (unit=="Sample") {
      if (log) {
        temp3 = tidyr::gather(log2(data+0.0001), key="Sample", value="Abundance")
        temp3$Group = gsub("-.*","",temp3$Sample) %>% gsub(".*_","",.) %>%
          gsub("N", "Normal",.) %>% gsub("T","Tumor",.)
      } else {
        temp3 = tidyr::gather(data, key="Sample", value="Abundance")
        temp3$Group = sub("-.*","",temp3$Sample) %>% gsub(".*_","",.) %>%
          gsub("N", "Normal",.) %>% gsub("T","Tumor",.)
      }
      if (log) {
        v2plot = "Abundance_log2"
        colnames(temp3)[2]=v2plot} else {v2plot="Abundance"}
      p = ggplot2::ggplot(temp3, ggplot2::aes_string(x="Sample", y=v2plot)) +
        ggplot2::geom_boxplot(ggplot2::aes(fill=Group),
                    outlier.shape = 16,
                    outlier.size = 1,
                    outlier.colour = "grey") +
        ggplot2::labs(x="", y=v2plot) 
    } else {
      p = ggplot2::ggplot(temp2, ggplot2::aes_string(x="Group", y=variable2plot)) +
        ggplot2::geom_boxplot(ggplot2::aes(fill=Group), 
                     outlier.shape = 16,
                     outlier.size = 1,
                     outlier.colour = "grey") +
        ggplot2::labs(x="Sample", y=variable2plot)     
    }
  } else if (plot == "barplot") {
    p = ggplot2::ggplot(temp2, ggplot2::aes_string(x="Sample", y=variable2plot)) +
      ggplot2::geom_bar(aes(fill=Group), stat = "identity") +
      ggplot2::theme(legend.position = "none") 
  }
  if (ylim) {p = p + ylim(-250, 250)}
  p + ggplot2::theme_bw() +
    ggplot2::theme(axis.text.x=ggplot2::element_text(angle=270, vjust=0.5,hjust=1)) +
    ggplot2::theme(legend.position = "none") +
    ggtitle("Zhou2020_all")
}
OUTDIR = "../results/CellFie_Zhou2020/"
QCDIR = paste0(OUTDIR, "/data/QC/")
dir.create(QCDIR, recursive=T, showWarnings=F)
pdf(paste0(QCDIR, "RawDataBoxplot.pdf"), h=4, w=30)
overviewPlot(Pdat[,2:ncol(Pdat)], 
             variable2plot="TotalAbundance", plot="boxplot", 
             unit="Sample", log=T, ylim=F)
dev.off()

```

```{r}
library(AnnotationDbi)
library(org.Hs.eg.db)
Pdat2 = AnnotationDbi::select(
  org.Hs.eg.db,
  keys = Pdat$genes,
  columns = c("ENTREZID"),
  keytype = "SYMBOL"
) %>% na.omit() %>%
  .[!duplicated(.$SYMBOL),] %>%
  setNames(c("genes", "ENTREZID")) %>%
  right_join(Pdat) %>%
  na.omit() %>%
  dplyr::select(-genes)
names(Pdat2)[1] = "genes"
dim(Pdat2)
# In total 232 samples

# Log transform the data
Pdat2[,2:ncol(Pdat2)] = log2(Pdat2[,2:ncol(Pdat2)])

# Save the data table
OUTDIR = "../results/CellFie_Zhou2020/"
INDIR = paste0(OUTDIR, "data/")
dir.create(INDIR, recursive = T, showWarnings = F)
# Save the input data to a csv file
write.table(Pdat2, paste0(INDIR, "/Zhou2020/CancerProteomics.csv"), sep=",", 
            quote=F, row.names=F)
saveRDS(Pdat2, paste0(OUTDIR, "/Pdat2.RDS"))
# Pdat2 = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
SampleNames = setdiff(colnames(Pdat2), "genes")
rm(Pdat, Pdat2)
gc()

```

# Run CellFie

  Submit the non-negative normalized input file to the [genePatten webserver](https://cloud.genepattern.org/gp/pages/index.jsf), and run it with the following settings:

  -   Data="PsoPproteomics.csv",
  -   Sample.Number=,
  -   Reference.Model= Human_iHsa,
  -   Thresholding.Approach="local",
  -   Percentile.Or.Value="percentile",
  -   Local.Threshold.Type="minmaxmean",
  -   Local.Lower.Bound="25",
  -   Local.Upper.Bound="75"
  -   Uses gene-specific thresholds, customized per gene (based on its expression distribution)
  -   For each gene, CellFie computes its 25th and 75th percentile and considers it ON if expression > mean of [25th, 75th].

# Analyze CellFie output

## Sort out meta data

```{r}
meta = 
    readxl::read_xlsx("../data/Zhou2020/13045_2020_1013_MOESM2_ESM*.xlsx",
    sheet="Patient_clinical charateristics") 
colnames(meta) = gsub(" ", "_", colnames(meta))
meta = meta %>% mutate(Sample=Tumor_case_ID,
        `Tumor_size_(cm)`=`Tumor_size_(cm)`,
        Grade=Grade,
        Histologic_type=Histologic_type,
        Tumor_Stage_Pathological=Tumor_Stage_Pathological,
        Path_Stage_Primary_Tumor_pT=Path_Stage_Primary_Tumor_pT,
        Path_Stage_Lymph_Nodes_pN=Path_Stage_Lymph_Nodes_pN) %>% 
        dplyr::select(-Normal_case_ID, -Tumor_case_ID) %>%
        rbind(
            meta %>% 
            mutate(Sample=Normal_case_ID,
                   # Cancer_type = NA,
                   `Tumor_size_(cm)`= NA,
                   Grade=NA,
                   Histologic_type=NA,
                   Tumor_Stage_Pathological=NA,
                   Path_Stage_Primary_Tumor_pT=NA,
                   Path_Stage_Lymph_Nodes_pN=NA
        ) %>%
        dplyr::select(-Normal_case_ID, -Tumor_case_ID)) %>%
        .[!is.na(.[,"Sample"]), ] %>%
        mutate(Group = gsub("[-].*", "", Sample))
meta = meta %>% mutate(Condition=gsub(".*_","",Group)%>%
    gsub("T","Tumor",.)%>%gsub("N","Normal",.)) %>%
    mutate(Cancer_type=gsub("ESCA", "ESCC", Cancer_type)) %>%
    mutate(Cancer_type=gsub("LUSC", "SCLC", Cancer_type)) %>%
    mutate(Cancer_type=gsub("LUSQ", "LUSC", Cancer_type)) %>%
    mutate(Cancer_Condition=paste0(Cancer_type, "_", Condition))

meta = meta %>% filter(Sample %in% SampleNames)
meta$Batch = "batch1"
meta$Patient = meta$Pair_ID
meta$Patient[is.na(meta$Patient)] = 1:sum(is.na(meta$Patient))
paired = names(table(na.omit(meta$Pair_ID)))[table(na.omit(meta$Pair_ID))==2]
meta$Paired = ifelse(meta$Pair_ID%in%paired, "Paired", "No")

openxlsx::write.xlsx(list(
  MetaData = meta,
  SampleSummary = meta %>% dplyr::count(Cancer_type, Condition)), 
    file=paste0(OUTDIR,"/data/Zhou2020/meta.xlsx"))

saveRDS(meta, paste0(OUTDIR,"/data/Zhou2020/meta.RDS")) 
# meta = radRDS(paste0(OUTDIR,"/data/Zhou2020/meta.RDS"))
gc()

```

## Data summary

### Sample size summary

```{r}
meta = readRDS(paste0(OUTDIR,"/data/Zhou2020/meta.RDS"))
SampleSizeCircularBarplot(meta, 
  title="Zhou2020", OUT_DIR=paste0(OUTDIR, "/Summary/"), h=4, w=4.5)

```

## Process CellFie output

```{r}
cfs = processCellFieOutput(combined=TRUE,
    outdir = paste0(OUTDIR, "/CellFieOut/"), 
    SampleNames, meta=meta, samples2keep=meta$Sample)
saveRDS(cfs, paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
# cfs = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
rm(cfs)
gc()


# # Boxplot of task size (unique gene nr) summary
# plotTaskSizes(taskInfo = lapply(list("all"), function(cancer) {
#     read.csv(paste0(OUTDIR, "/CellFieOut/", cancer, "/detailScoring_new.csv"),
#           header = TRUE, sep="\t", stringsAsFactors = FALSE) 
#   }) %>% Reduce(rbind,.), 
#               OUTDIR = paste0(OUTDIR, "/CellFieOut/../Summary/"))

```

### Metabolic activity boxplot by cancer

```{r}
meta = readRDS(paste0(OUTDIR,"/data/Zhou2020/meta.RDS"))
cfs = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
ActivityBoxplot_byCancer(meta=meta, 
  cf=cfs, OUT_DIR=paste0(OUTDIR, "/Summary/"), title="Zhou2020")

```

### Metabolic activity boxplot by system

```{r}
cfs = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
ActivityBoxplot_bySystem(cf=cfs%>%mutate(Depth1=gsub("METABOLISM","M.", Depth1)), 
  OUT_DIR=paste0(OUTDIR, "/Summary/"), w=2, h=3.5, title="Zhou2020")

```

## Limma analysis (paired)

```{r}
cfs = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
meta = readRDS(paste0(OUTDIR,"/data/Zhou2020/meta.RDS"))
depth = "Depth3"
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")

result3 = runLimmaCellFie(depth, 
  cfs=cfs, 
  meta=meta, OUTDIRV, w=9.5, height=23)
saveRDS(result3, paste0(OUTDIRV, "/result3.RDS"))
# result3 = readRDS(paste0(OUTDIRV, "/result3.RDS"))

adj.P.Val.cutoff = 0.1
makeTaskScoreHeatmap_CellFie(rlst_list=result3$rlst_list, 
  cf5_Depth=result3$cf5_Depth, OUTDIRV=OUTDIRV, w=11, h=26,
  adj.P.Val.cutoff, meta=meta)

makeLimmaDotplot_CellFie(
  rlst_list=result3$rlst_list,
  adj.P.Val.cutoff, w=14.5, # h=19,  
  OUTDIRV=OUTDIRV, Depth=depth, lowfigure=F, title = "", 
  suffix="Zhou2020_tasks_all", taskHierarchy=cfs, nested=T, filter=F)

makeLimmaDotplot_CellFie(
  rlst_list=result3$rlst_list,
  adj.P.Val.cutoff, w=12.5, # h=13.5,  
  OUTDIRV=OUTDIRV, Depth=depth, lowfigure=F, title = "", 
  suffix="Zhou2020_tasks", taskHierarchy=cfs, nested=T, filter=T,
  probs_cutoff=0.65)

rm(result3)
gc()

```

```{r}
cfs = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
meta = readRDS(paste0(OUTDIR,"/data/Zhou2020/meta.RDS"))
depth = "Depth2"
Depth2 = readRDS(paste0(OUTDIR, "Summary/TaskScores_", depth,".RDS"))
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth2/")
result2 = runLimmaCellFie(depth, 
  cfs=Depth2, 
  meta=meta, OUTDIRV, w=5.2, height=15)
saveRDS(result2, paste0(OUTDIRV, "/result2.RDS"))
# result2 = readRDS(paste0(OUTDIRV, "/result2.RDS"))

adj.P.Val.cutoff = 0.1
makeTaskScoreHeatmap_CellFie(rlst_list=result2$rlst_list, 
  cf5_Depth=result2$cf5_Depth, OUTDIRV=OUTDIRV, w=11, h=17,
  adj.P.Val.cutoff, meta=meta)

makeLimmaDotplot_CellFie(
  rlst_list=result2$rlst_list,
  adj.P.Val.cutoff, w=9, # h=19,  
  OUTDIRV=OUTDIRV, Depth=depth, lowfigure=F, title = "", 
  suffix="Zhou2020_tasks_all", taskHierarchy=cfs, nested=T, filter=F)

rm(result2)
gc()

```

```{r}
cfs = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
meta = readRDS(paste0(OUTDIR,"/data/Zhou2020/meta.RDS"))
depth = "Depth1"
Depth1 = readRDS(paste0(OUTDIR, "Summary/TaskScores_", depth,".RDS"))
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth1/")
result1 = runLimmaCellFie(depth, 
  cfs=Depth1, 
  meta=meta, OUTDIRV, w=4.7, height=3.3)
saveRDS(result1, paste0(OUTDIRV, "/result1.RDS"))
# result2 = readRDS(paste0(OUTDIRV, "/result2.RDS"))

adj.P.Val.cutoff = 0.1
makeTaskScoreHeatmap_CellFie(rlst_list=result1$rlst_list, 
  cf5_Depth=result1$cf5_Depth, OUTDIRV=OUTDIRV, w=9.5, h=4.5,
  adj.P.Val.cutoff, meta=meta)

makeLimmaDotplot_CellFie(
  rlst_list=result1$rlst_list,
  adj.P.Val.cutoff, h=3.2, w=7.5, #w=9, # h=19,  
  OUTDIRV=OUTDIRV, Depth=depth, lowfigure=F, title = "", 
  suffix="Zhou2020_tasks_all", taskHierarchy=cfs, nested=F, filter=F)

makeLimmaDotplot_CellFie(
  rlst_list=result1$rlst_list,
  adj.P.Val.cutoff, h=4.5, w=6, #w=9, # h=19,  
  OUTDIRV=OUTDIRV, Depth=depth, lowfigure=T, title = "", 
  suffix="Zhou2020_tasks_all2", taskHierarchy=cfs, nested=F, filter=F)


rm(result1)
gc()

```


### Volcano plots

```{r}
adj.P.Val.cutoff = 0.1
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth1/")
result1 = readRDS(paste0(OUTDIRV, "result1.RDS"))
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth2/")
result2 = readRDS(paste0(OUTDIRV, "result2.RDS"))
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")
result3 = readRDS(paste0(OUTDIRV, "/result3.RDS"))
plots = lapply(names(result1$rlst_list), function(cancer) {
  dat2plot = rbind(result1$rlst_list[[cancer]]$rslt_of_interest %>% 
                      tibble::rownames_to_column(var="Task") %>%
                      #filter(adj.P.Val<adj.P.Val.cutoff) %>%
                      mutate(Depth = "Depth1"),
                   result2$rlst_list[[cancer]]$rslt_of_interest %>% 
                      tibble::rownames_to_column(var="Task") %>%
                      #filter(adj.P.Val<adj.P.Val.cutoff)%>%
                      mutate(Depth = "Depth2"),
                   result3$rlst_list[[cancer]]$rslt_of_interest%>% 
                      tibble::rownames_to_column(var="Task") %>%
                      #filter(adj.P.Val<adj.P.Val.cutoff)%>%
                      mutate(Depth = "Depth3"))
  #print(head(dat2plot))
  p = doVolcano_CellFie(resi=dat2plot, NAME=cancer, adj.P.Val.cutoff=0.05, 
                logFC.cutoff=0, lfcShrink=F)
  p
}) 

pdf(paste0(OUTDIR, "/Limma/Volcano_plots.pdf"),h=4,w=6)
  plots
dev.off()

pdf(paste0(OUTDIR, "/Limma/Volcano_plots2.pdf"),h=20,w=18)
  cowplot::plot_grid(plots[[1]],plots[[2]],plots[[3]],plots[[4]], 
                     plots[[5]],plots[[6]],plots[[7]],plots[[8]],
                     plots[[9]],plots[[10]],plots[[11]],plots[[12]],
                     ncol=3,nrow=5)
dev.off()
rm(plots)
rm (result1, result2, restul3)
gc()
```

### Limma result summary

#### Sig task nr barplot_by cancer

```{r}
OUTDIR = "../results/CellFie_Zhou2020/"
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")
result3 = readRDS(paste0(OUTDIRV, "/result3.RDS"))
rlst_list = result3$rlst_list
SigNrBarplot(rlst_list, 
  sig.cutoff=0.1, w=2.75, h=3.0, OUT_DIR=paste0(OUTDIR, "/Limma/"))
```

#### Differential effect boxplot by system

```{r}
OUTDIR = "../results/CellFie_Zhou2020/"
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")
result3 = readRDS(paste0(OUTDIRV, "/result3.RDS"))
cfs = readRDS(paste0(OUTDIR, "/CellFieOut/cfs.RDS"))
diffEffectBoxplot_bySystem(rlst_list = result3$rlst_list, 
  cf=cfs%>%mutate(Depth1=gsub("METABOLISM","M.", Depth1)), 
  OUT_DIR=paste0(OUTDIR, "/Limma/"), 
  w=2, h=3.5, title="Zhou2020", sig.cutoff=0.05)

```

#### Differential effect boxplot by cancer

```{r}
OUTDIR = "../results/CellFie_Zhou2020/"
OUTDIRV = paste0(OUTDIR, "/Limma/TaskDepth3/")
result3 = readRDS(paste0(OUTDIRV, "/result3.RDS"))
diffEffectBoxplot_byCancer(
  rlst_list = result3$rlst_list, 
  OUT_DIR=paste0(OUTDIR, "/Limma/"), w=2.75, h=3.0, title="Zhou2020")
```

# SessionInfo
```{r}
out = capture.output(
  sessionInfo()
)
writeLines(out, paste0(OUTDIR,"/sessionInfo.txt"))
out
```

