---
title: "Zhou2020_GSVA"
author: Yuan Li (yuan.li@nbis.se)
format: html
---


```{r}
# BiocManager::install("TCGAbiolinks")
library(TCGAbiolinks)
library(SummarizedExperiment)
library(dplyr)
library(dplyr)
library(ggplot2)
library(parseMetab)
# remove.packages("parseMetab")
# detach("package:KinaseDownstream", unload = TRUE)
# remotes::install_github("Maj18/parseMetab", dependencies = TRUE)
# library(parseMetab)
```


# Data checking

```{r}
outdir = "../results/Zhou2020_GSVA/"
dir.create(outdir, recursive=T, showWarnings=F)
```

## Only take paired data
```{r}
Pdat = 
  readxl::read_xlsx("../data/Zhou2020/13045_2020_1013_MOESM3_ESM*.xlsx") %>%
  mutate(genes = gsub(";.*", "", `Gene Group`)) %>%
  select(genes, everything(), -`Gene Group`, -Index, -`Protein Group`)
dim(Pdat)
meta = readRDS("../results/CellFie_Zhou2020/data/Zhou2020/meta.RDS")
print("Only use paired data ...")
samples2keep = meta %>%
    filter(meta$Pair_ID %in% names(table(meta$Pair_ID))[table(meta$Pair_ID)>1]) %>%
    pull(Sample)
Pdat3 = Pdat[, c("genes", samples2keep)]
meta2 = meta[meta$Sample%in%samples2keep, ]
dim(Pdat3)
rm(Pdat, meta)
gc()

```



# Import KEGG metabolic pathways

```{r}
# # Get KEGG metabolic pathways
# paste0(outdir, "/KEGG_metabolism/") %>% 
#     dir.create(., recursive=TRUE, showWarnings=FALSE)

# kegg_metab_dbs = 
#     getdb_metabolism(databaseDIR=paste0(outdir, "/KEGG_metabolism/"))

databaseDIR="../results/TCGA/KEGG_metabolism/"
kegg_metab_dbs = 
list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
    kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))
kegg_metab_db = kegg_metab_dbs$kegg_metab_db
names(kegg_metab_dbs)
# kegg_metab_dbs$kegg_metab_db_table$gene_symbol %>% unique() %>% length()

```


# Paired differential analysis on sample GSVA scores

```{r}
head(meta2)
GSVA_limma_rslt = lapply(meta2$Cancer_type %>% unique(), function(c) {
    print(c)
    meta_sub = meta2 %>% filter(Cancer_type==c)
    samples = meta_sub$Sample
    temp = Pdat3[, 2:ncol(Pdat3)] %>% rowSums() 
    Pdat3 = Pdat3[temp %>% order(decreasing=T), ]
    sub = Pdat3[!duplicated(Pdat3$genes), c("genes", samples)] 
    rownames(sub) = NULL
    sub = sub %>% tibble::column_to_rownames(var="genes") %>% na.omit()

    meta_sub$Group = gsub(".*_T", "Primary Tumor", meta_sub$Group)
    meta_sub$Group = gsub(".*_N", "Solid Tissue Normal", meta_sub$Group)
    meta_sub$Group = factor(meta_sub$Group, levels=c("Solid Tissue Normal", "Primary Tumor"))
    if (length(unique(meta_sub$Group))>1) {
        paste0(outdir, "/Limma_GSVA/", c, "/") %>%
            dir.create(., recursive=TRUE, showWarnings=FALSE)
        rslt = GSVAlimmaTest(dat=sub, paired=TRUE, Feature.filtering=TRUE,
            currentCovariate=NULL, meta=meta_sub, paired_variable="Patient", 
            kegg_metab_db=kegg_metab_db, 
            OUTDIR=paste0(outdir, "/Limma_GSVA/", c, "/"))
        return(rslt)
    } else {return(NA)}
}) %>% setNames(meta2$Cancer_type %>% unique())

# Remove NA
GSVA_limma_rslt = GSVA_limma_rslt[!is.na(GSVA_limma_rslt)]
# Remove NULL
GSVA_limma_rslt = GSVA_limma_rslt[
    sapply(GSVA_limma_rslt, function(r) !is.null(r$rslt_of_interest)) %>%
    unlist()]
saveRDS(GSVA_limma_rslt, 
    paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
# GSVA_limma_rslt = readRDS(paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))

# Make a circular packing plot to show the number of pathways in each class
getCirPackingPlot_GSVA(
    kegg_metab_db_table = 
        kegg_metab_dbs$kegg_metab_db_table %>% 
        filter(gs_name %in% (sapply(GSVA_limma_rslt, function(g) {
            rownames(g$gsva)
        }) %>% unlist() %>% as.character() %>% unique())),
    outdir=outdir)

```


## Visualize differential results

```{r}
library(dplyr)
library(stringr)
rlst_list = lapply(GSVA_limma_rslt, function(l) as.data.frame(l$rslt_of_interest))
adj.P.Val.cutoff = 0.1
makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=adj.P.Val.cutoff, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA", nested=T, 
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=adj.P.Val.cutoff, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA_tops", nested=T, top=TRUE,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=adj.P.Val.cutoff, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA2", nested=T, 
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique(),
        color_low="blue", color_high="red")

makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=adj.P.Val.cutoff, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA_tops2", nested=T, top=TRUE,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique(),
        color_low="blue", color_high="red")

```


### Visualize the number of significant pathways

```{r}
visualizeSigNr(OUTDIR=paste0(outdir, "/Limma_GSVA/"), 
    GSVA_limma_rslt = GSVA_limma_rslt)

names(GSVA_limma_rslt)
sapply(GSVA_limma_rslt, function(c) {
    c$rslt_of_interest %>% filter(adj.P.Val<0.05) %>% nrow()
})

# BRCA CHOL COAD ESCA HNSC KICH KIRC KIRP LIHC LUAD LUSC PAAD READ UCEC 
#   64   65   66   14   61   56   69   53   65   66   62    0   39   52 
```

```{r}
# Check samples:
sapply(GSVA_limma_rslt, function(c)
    names(c$gsva)) %>% unlist() %>%
    as.character() %>% unique()
```


### Visualize differential effect

#### Differential effect boxplot by class

```{r}

GSVA_limma_rslt = 
    readRDS(paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
databaseDIR=paste0(outdir, "/KEGG_metabolism/")
kegg_metab_dbs = 
    list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
    kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))

diffEffectBoxplot_bySystem_GSVA(rlst_list = GSVA_limma_rslt, 
    taskHierarchy=
    kegg_metab_dbs$kegg_metab_db_table %>% select(gs_name, class) %>% unique(), 
    OUT_DIR=paste0(outdir, "/Limma_GSVA"), 
    weight.effect.by.gene=TRUE,
    w=6.2, h=4.2, title="TCGA")

```

#### Differential effect boxplot by cancer

```{r}
diffEffectBoxplot_byCancer_GSVA(
  rlst_list = GSVA_limma_rslt, 
  OUT_DIR=paste0(outdir, "/Limma_GSVA"), w=2.75, h=3.0, title="TCGA")
```


# Summarize GSVA scores

## Visualize pathway activity GSVA scores

### Per sample

```{r}
GSVA_limma_rslt_gsva = lapply(GSVA_limma_rslt, function(r) r$gsva)
visualizeGSVSscoresGroup(
    GSVA_limma_rslt_gsva = GSVA_limma_rslt_gsva, 
    paired_data = Pdat3, 
    OUTDIR = paste0(outdir, "/Summary/"),
    plot_healthy_control = FALSE
)

```

### Per metabolic class

```{r}
visualizeGSVSscoresClass(GSVA_limma_rslt, 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table, 
    paired_data = Pdat3, 
    OUTDIR = paste0(outdir, "/Summary/"))
```

##  Visualize pathway sizes
```{r}
genes4GSVA = read.delim(
    file = paste0(outdir, "/Limma_GSVA/UCEC/NormalizedCount.csv"),
    header = TRUE,
    stringsAsFactors = FALSE
    ) %>% pull(Features) %>% unique()

visualizeGSsizeClass(
    GSVA_limma_rslt_gsva=lapply(GSVA_limma_rslt, function(r) r$gsva), 
    genes4GSVA = genes4GSVA, 
    kegg_metab_db = kegg_metab_dbs$kegg_metab_db, 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table,
    OUTDIR=paste0(outdir, "/Summary/"))
```


```{r}
save.image(paste0(outdir, "/TCGA.RData"))
# load(paste0(outdir, "/TCGA.RData"))
out = capture.output(sessionInfo())
writeLines(out, paste0(outdir, "/sessionInfo.txt"))

```

