---
title: "Untitled"
format: html
---


```{r}
BiocManager::install("mygene")
library(mygene)
```

```{r}
outDirHu = "../results/GlycanDifferential/Hu/"
dir.create(outDirHu, recursive=TRUE, showWarnings=FALSE)
outDirZhou = "../results/GlycanDifferential/Zhou/"
dir.create(outDirZhou, recursive=TRUE, showWarnings=FALSE)
```
# Extract the GLYCAN METABOLISM genes
```{r}
#' Convert gene entrez ID to gene symbol using mygene, then extract the GLYCAN METABOLISM.
#' 
#' @param taskInfo2 A cellfie output tabel generated with the processCellFieOutput() function. It should have a column named GeneAssociatedToEssentialRxnsTask.
#' @examples 
#' # getGlycan(taskInfo2)
#' 
getGlycan = function(taskInfo2) {
    entrezID = unique(taskInfo2$GeneAssociatedToEssentialRxnsTask)
    res = queryMany(
        entrezID,
        scopes = "entrezgene",
        fields = "symbol",
        species = "human"
    )
    mapping = setNames(res$symbol, res$query)
    taskInfo2$GeneAssociatedToEssentialRxnsTask_symbol = 
        mapping[as.character(taskInfo2$GeneAssociatedToEssentialRxnsTask)]
    head(taskInfo2)
    glycan = taskInfo2 %>% filter(Depth1=="GLYCAN METABOLISM")
    return(glycan)
}

# Hu2025
OUTDIR = "../results/CellFie_Hu2025/"
taskInfo2 = readRDS(paste0(OUTDIR, "/CellFieOut/taskInfo.RDS"))
glycanHu = getGlycan(taskInfo2)
glycanHu = glycanHu %>% unique()
saveRDS(glycanHu, paste0(outDirHu, "/glycanTaskInfoHu.RDS"))
openxlsx::write.xlsx(list(Hu2025=glycanHu),
    paste0(outDirHu, "glycanTaskInfoHu.xlsx"))

rm(taskInfo2, OUTDIR)
gc()

# Zhou2020
OUTDIR = "../results/CellFie_Zhou2020/"
taskInfo2 = 
    read.csv(paste0(OUTDIR, "/CellFieOut/all/detailScoring_new.csv"), sep="\t", header=TRUE)
head(taskInfo2)
glycanZhou = getGlycan(taskInfo2)
glycanZhou = glycanZhou %>% unique()
saveRDS(glycanZhou, paste0(outDirZhou, "/glycanTaskInfoZhou.RDS"))
openxlsx::write.xlsx(list(Zhou2020=glycanZhou),
    paste0(outDirZhou, "glycanTaskInfoZhou.xlsx"))
head(glycanZhou)
rm(taskInfo2, OUTDIR)
gc()

```

# Limma trend differential analysis per comparing pair

## Import the normalized protein abundance data

```{r}
OUTDIR = "../results/CellFie_Hu2025/"
NdatHu = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
metaHu = readRDS(paste0(OUTDIR, "/data/Hu2025/Meta_hu2025.RDS"))
saveRDS(metaHu, paste0(outDirHu, "/metaHu.RDS"))
rm(OUTDIR)
head(NdatHu)
head(metaHu)

OUTDIR = "../results/CellFie_Zhou2020/"
NdatZhou = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
metaZhou = readRDS(paste0(OUTDIR,"/data/Zhou2020/meta.RDS"))
saveRDS(metaZhou, paste0(outDirZhou, "/metaZhou.RDS"))
rm(OUTDIR)
head(NdatZhou)
head(metaZhou)
```

## Limma differential analysis

```{r}
# Check the Hu2025 data
metaHu %>% filter(Paired==Paired) %>% 
    group_by(Cancer_type, Condition) %>% 
    summarize(count=n()) %>% print(n=Inf)
```
| Tumor_code | Condition | count |
|------------|-----------|-------|
| COAD       | Normal    | 100   |
| COAD       | Tumor     | 97    |
| ESCC       | Normal    | 124   |
| ESCC       | Tumor     | 124   |
| GBM        | Normal    | 10    |
| GBM        | Tumor     | 99    |
| HNSC       | Normal    | 63    |
| HNSC       | Tumor     | 108   |
| KIRC       | Normal    | 84    |
| KIRC       | Tumor     | 92    |
| LIHC       | Normal    | 165   |
| LIHC       | Tumor     | 165   |
| LUAD       | Normal    | 101   |
| LUAD       | Tumor     | 110   |
| LUSC       | Normal    | 100   |
| LUSC       | Tumor     | 108   |
| OV         | Normal    | 20    |
| OV         | Tumor     | 83    |
| PAAD       | Normal    | 67    |
| PAAD       | Tumor     | 140   |
| STAD       | Normal    | 80    |
| STAD       | Tumor     | 80    |
| UCEC       | Normal    | 25    |
| UCEC       | Tumor     | 70    |

```{r}
# Check the Zhou2020 data
metaZhou %>% filter(Paired==Paired) %>% 
    group_by(Cancer_type, Condition) %>% 
    summarize(count=n()) %>% print(n=Inf)
```
| Cancer_type | Condition | count |
|-------------|-----------|-------|
| BLCA        | Normal    | 11    |
| BLCA        | Tumor     | 11    |
| BRCA        | Normal    | 5     |
| BRCA        | Tumor     | 5     |
| COAD        | Normal    | 11    |
| COAD        | Tumor     | 11    |
| ESCC        | Normal    | 6     |
| ESCC        | Tumor     | 7     |
| GBM         | Tumor     | 12    |
| HNSC        | Normal    | 5     |
| HNSC        | Tumor     | 6     |
| KIRC        | Normal    | 10    |
| KIRC        | Tumor     | 9     |
| LIHC        | Normal    | 8     |
| LIHC        | Tumor     | 8     |
| LUAD        | Normal    | 8     |
| LUAD        | Tumor     | 9     |
| LUSC        | Normal    | 7     |
| LUSC        | Tumor     | 8     |
| OV          | Normal    | 6     |
| OV          | Tumor     | 7     |
| PAAD        | Normal    | 5     |
| PAAD        | Tumor     | 8     |
| PRAD        | Normal    | 7     |
| PRAD        | Tumor     | 7     |
| SCLC        | Normal    | 6     |
| SCLC        | Tumor     | 6     |
| STAD        | Normal    | 7     |
| STAD        | Tumor     | 8     |
| UCEC        | Normal    | 4     |
| UCEC        | Tumor     | 4     |

```{r}
OUTDIR = "../results/CellFie_Hu2025/"
NdatHu = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
glycanHu = readRDS(paste0(outDirHu, "/glycanTaskInfoHu.RDS"))
metaHu = readRDS(paste0(outDirHu, "/metaHu.RDS"))

head(NdatHu)
head(metaHu)
sum(metaHu$Sample %in% colnames(NdatHu))
sum(colnames(NdatHu) %in% metaHu$Sample)

NdatHu = NdatHu[!duplicated(NdatHu$genes),]
rownames(NdatHu) = NULL
NdatHu = NdatHu %>% tibble::column_to_rownames(var="genes")
NdatHu[] = lapply(NdatHu, as.numeric)
sharedGenes = 
    intersect(as.character(unique(glycanHu$GeneAssociatedToEssentialRxnsTask)), rownames(NdatHu))
patientPair = which(table(metaHu$Patient)==2) %>% names()
limma_rslt_listHu = 
    lapply(unique(metaHu$Cancer_type), function(c) {
        print(c)
        meta_sub = metaHu %>% filter(Patient%in%patientPair) %>% filter(Cancer_type==c)
        if ((nrow(meta_sub)>4) & length(unique(meta_sub$Condition))==2) {
            limmaTest_CellFie(
                dat=NdatHu[sharedGenes,], 
                paired=TRUE, 
                currentCovariate=NULL, 
                checkCovariate=FALSE, 
                meta=meta_sub, 
                paired_variable="Patient")  
        } else {NA}
    }) %>%setNames(unique(metaHu$Cancer_type))
limma_rslt_listHu = limma_rslt_listHu[!is.na(limma_rslt_listHu)]
rm(NdatHu, metaHu, sharedGenes, OUTDIR, patientPair)
gc()

names(limma_rslt_listHu)
mappingHu = setNames(glycanHu$GeneAssociatedToEssentialRxnsTask_symbol,
    glycanHu$GeneAssociatedToEssentialRxnsTask)
limma_rslt_listHu = lapply(limma_rslt_listHu, function(x) {
        x$rslt_of_interest %>% 
        tibble::rownames_to_column(var="EntrezID")%>%
        mutate(Feature=as.character(mappingHu[EntrezID])) %>%
        dplyr::select(Feature, everything())})
saveRDS(limma_rslt_listHu, paste0(outDirHu, "limma_rslt_listHu.RDS"))
openxlsx::write.xlsx(
    limma_rslt_listHu, 
    file = paste0(outDirHu, "limma_reslut.xlsx"))

```

```{r}
OUTDIR = "../results/CellFie_Zhou2020/"
NdatZhou = readRDS(paste0(OUTDIR, "/Pdat2.RDS"))
glycanZhou = readRDS(paste0(outDirZhou, "/glycanTaskInfoZhou.RDS"))
metaZhou = readRDS(paste0(outDirZhou, "/metaZhou.RDS"))
# lapply(unique(metaZhou$Cancer_type), function(c) {
#     metaZhou %>% filter(Paired==Paired) %>% filter(Cancer_type==c) %>% dim()  
# }) 

head(NdatZhou)
head(metaZhou)
sum(metaZhou$Sample %in% colnames(NdatZhou))
sum(colnames(NdatZhou) %in% metaZhou$Sample)

NdatZhou = NdatZhou[!duplicated(NdatZhou$genes),]
rownames(NdatZhou) = NULL
NdatZhou = NdatZhou %>% tibble::column_to_rownames(var="genes")
NdatZhou[] = lapply(NdatZhou, as.numeric)
sharedGenes = 
    intersect(as.character(unique(glycanZhou$GeneAssociatedToEssentialRxnsTask)), rownames(NdatZhou))
patientPair = which(table(metaZhou$Patient)==2) %>% names()
limma_rslt_listZhou = 
    lapply(unique(metaZhou$Cancer_type), function(c) {
        print(c)
        meta_sub = metaZhou %>% filter(Patient%in%patientPair) %>% filter(Cancer_type==c)
        if ((nrow(meta_sub)>4) & length(unique(meta_sub$Condition))==2) {
            limmaTest_CellFie(
                dat=NdatZhou[sharedGenes,], 
                paired=TRUE, 
                currentCovariate=NULL, 
                checkCovariate=FALSE, 
                meta=meta_sub, 
                paired_variable="Patient")  
        } else {NA}
    }) %>%setNames(unique(metaZhou$Cancer_type))
limma_rslt_listZhou = limma_rslt_listZhou[!is.na(limma_rslt_listZhou)]
rm(NdatZhou, metaZhou, sharedGenes, OUTDIR, patientPair)
gc()

mappingZhou = 
    setNames(glycanZhou$GeneAssociatedToEssentialRxnsTask_symbol,
    glycanZhou$GeneAssociatedToEssentialRxnsTask)
limma_rslt_listZhou = lapply(limma_rslt_listZhou,function(x) {
        x$rslt_of_interest %>% 
        tibble::rownames_to_column(var="EntrezID")%>%
        mutate(Feature=as.character(mappingZhou[EntrezID])) %>%
        dplyr::select(Feature, everything())})
saveRDS(limma_rslt_listZhou, paste0(outDirZhou, "limma_rslt_listZhou.RDS"))
openxlsx::write.xlsx(
    limma_rslt_listZhou, 
    file = paste0(outDirZhou, "limma_reslut.xlsx"))

```

# Visualize differential results (all cancers together)

```{r}
visualizedDiffGenes = function(
    limma_rslt=limma_rslt_listZhou, 
    outdir=outDirZhou, 
    taskInfo=lipidZhou,
    h=20, w=8,
    max_logFC=3) {

    logFCdat = lapply(names(limma_rslt), function(l) {
        lr = limma_rslt[[l]] %>% dplyr::select(Feature, logFC) %>%
            setNames(c("Feature", l))
    }) %>% Reduce(dplyr::full_join, .) %>%
        tibble::column_to_rownames(var="Feature")
    logFCdat = logFCdat[rev(order(rowSums(logFCdat))), ]

    padj_dat = lapply(names(limma_rslt), function(l) {
        lr = limma_rslt[[l]] %>% dplyr::select(Feature, adj.P.Val) %>%
            setNames(c("Feature", l))
    }) %>% Reduce(dplyr::full_join, .) %>%
        tibble::column_to_rownames(var="Feature")
    padj_dat = padj_dat[rownames(logFCdat), ]

    sig_mat = matrix("", nrow=nrow(logFCdat), ncol=ncol(logFCdat))
    sig_mat[padj_dat<0.05] = "*"

    ann_row = taskInfo %>% dplyr::select(Depth3, GeneAssociatedToEssentialRxnsTask_symbol) %>%
        na.omit() %>% unique() %>% 
        filter(GeneAssociatedToEssentialRxnsTask_symbol %in% as.character(rownames(logFCdat))) %>%
        mutate(value = 1) %>%  # mark presence with 1
        tidyr::pivot_wider(names_from = Depth3, 
            values_from = value,
            values_fill = list(value = 0)) %>%
        tibble::column_to_rownames(var="GeneAssociatedToEssentialRxnsTask_symbol")
    ann_row = ann_row[rownames(logFCdat), ]
    ann_colors = 
        replicate(ncol(ann_row), c("0" = "snow2", "1" = "#d95f02"), 
        simplify = FALSE) %>% setNames(colnames(ann_row))
    
    logFCdat[abs(logFCdat)>max_logFC] = max_logFC
    max_abs = max(abs(logFCdat), na.rm = TRUE)
    # Define breaks centered at 0
    breaks = seq(-max_abs, max_abs, length.out = 101)
    p1 = pheatmap::pheatmap(logFCdat,
            color = colorRampPalette(c("blue", "white", "red"))(100),
            breaks = breaks,
            # annotation_col = ann_row,   # top bar
            display_numbers = sig_mat,
            annotation_row = ann_row,
            annotation_colors = ann_colors,
            show_rownames=TRUE, 
            show_colnames=TRUE,
            fontsize = 10,
            border_color = "white",
            annotation_legend = FALSE,
            main=paste0("logFC (*: padj<0.05)"))
    pdf(paste0(outdir, "/logFC_tumorVSnormal.pdf"),
        h=h, w=w)
    print(p1)
    dev.off()
}

visualizedDiffGenes(limma_rslt=limma_rslt_listZhou, 
    outdir=outDirZhou, 
    taskInfo=glycanZhou,
    h=20, w=7.3,
    max_logFC=5)

visualizedDiffGenes(limma_rslt=limma_rslt_listHu, 
    outdir=outDirHu, 
    taskInfo=glycanHu,
    h=8, w=6)

```


# Network, hub genes


