---
title: "TCGA"
author: Yuan Li (yuan.li@nbis.se)
format: html
---



```{r}
# BiocManager::install("TCGAbiolinks")
library(TCGAbiolinks)
library(SummarizedExperiment)
library(dplyr)
library(dplyr)
library(ggplot2)
library(parseMetab)
# remove.packages("parseMetab")
# detach("package:KinaseDownstream", unload = TRUE)
# remotes::install_github("Maj18/parseMetab", dependencies = TRUE)
# library(parseMetab)
```

# Download SMTdb cancer slice data

```{bash}
#!/bin/bash

(
  echo -e "cancer\ttissue\tpaper\tpaper_html\tdatasetID\tspot_num\tcell_num\tbac_num\tslice_type\tsample"
  jq -r '
    .data[] | [
      .cancer,
      .tissue,
      .paper,
      .paper_html,
      .datasetID,
      .spot_num,
      .cell_num,
      .bac_num,
      .slice_type,
      .sample
    ] | @tsv
  ' ../data/SMTdb/download_file.json
) > ../data/SMTdb/output.tsv

# First get the header to know column numbers
header=$(head -1 ../data/SMTdb/output.tsv)

# Example columns based on your JSON (adjust if different):
# cancer = column 1
# datasetID = column 5
# slice_type = column 9

# List of files to download
# "sc_celltype_file.txt", "sc_gene_exp_count.txt"
files=("st_cluster.txt" "st_neb.txt" "st_dev_rctd.txt" "st_gene_exp_count.txt")

tail -n +2 ../data/SMTdb/output.tsv | awk -F '\t' '{ print $1 "\t" $5 }' | while IFS=$'\t' read -r cancer dataset_id; do
  mkdir -p "../data/SMTdb/$cancer"/${dataset_id}

  for file in "../data/SMTdb/${files[@]}"; do
    output_file="../data/SMTdb/${cancer}/${dataset_id}/${file}"
    if [ -f "$output_file" ]; then
      echo "File $output_file already exists, skipping download."
    else
      echo "Downloading $file for $dataset_id into folder $cancer"
      curl -L -o "$output_file" "http://bio-bigdata.hrbmu.edu.cn/SMTdb/downloadCopy.jsp?filename=${file}&loc=${dataset_id}"
    fi
  done

done

```

# Process the data

```{r}
INDIR = "../data/SMTdb/"
meta = read.table(
  paste0(INDIR, "/output.tsv"),
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE,
  fill = TRUE,
  quote = "",
  comment.char = ""
)

# We will only keep cancer slices that have >5 cancer slices:
print("We will only keep cancer slices that have >5 cancer slices...")
cancers2keep = meta %>% filter(slice_type=="cancer") %>%
    group_by(cancer) %>% 
    summarise(count = n()) %>%
    filter(count > 5) %>% pull(cancer)
meta = meta %>% 
    filter(cancer %in% cancers2keep) %>% filter(slice_type=="cancer")
OUTDIR = "../results/SMTdb/"
OUTDIR %>% dir.create(recursive=T, showWarnings=FALSE)
write.table(meta, paste0(OUTDIR, "/meta.csv"),
                    quote=F, sep="\t", row.names=F)
saveRDS(meta, paste0(OUTDIR, "/meta.RDS"))
# meta = readRDS(paste0(OUTDIR, "/meta.RDS"))

```

## Import data

```{r}
#' Import SMTdb files
#' 
#' Import SMTdb files for all cancer types and all slices.
#' 
#' @param INDIR Directory where the SMTdb files are placed, it shoud be organized as <Cancer_type>/<Sample_ID>/*.txt.
#' @param pattern What files to import, e.g. "st_neb.txt".
#' 
#' @return A list of list of tables:
#'  The first level of lists contains all cancer types.
#'  The second level of lists contains all slices within each cancer type.
#'
#' @examples 
#' # importFiles_SMTdb(INDIR, pattern="st_gene_exp_count.txt")
#' 
importFiles_SMTdb = function(INDIR, pattern="st_gene_exp_count.txt", meta) {
    cancers = meta$cancer %>% unique()
    # cancers = list.dirs(INDIR, full.names=FALSE, recursive=FALSE)
    lapply(cancers, function(c) {
        cancer_dir = paste0(INDIR, "/", c, "/")
        # file_paths = 
        #     list.files(cancer_dir, 
        #     pattern=pattern, full.names=TRUE, recursive=TRUE)
        # file_samples = list.dirs(cancer_dir, full.names=F, recursive=FALSE)
        if (pattern=="st_neb.txt") {
            file_samples = 
                meta %>% 
                filter(cancer==c&slice_type=="cancer") %>%
                pull(datasetID) %>% unique()
        } else {
            file_samples = 
                meta %>% filter(cancer==c) %>% pull(datasetID) %>% unique()
        }
        file_paths = paste0(INDIR, "/", c, "/", file_samples, "/", pattern)

        lapply(file_paths, function(file) {
            tryCatch(read.table(file, header = TRUE, 
                sep = "\t", stringsAsFactors = FALSE, row.names = NULL) %>%
            as.data.frame(), error=function(e) NA)
        }) %>% setNames(file_samples)
    }) %>% setNames(cancers)
}

INDIR = "../data/SMTdb/"
SpatialNeighborhood = importFiles_SMTdb(INDIR, pattern="st_neb.txt", meta=meta)
STexpr_all = importFiles_SMTdb(INDIR, pattern="st_gene_exp_count.txt", meta=meta)

# Save the objects
OUTDIR = "../results/SMTdb/"
OUTDIR %>% dir.create(., recursive=TRUE, showWarnings=FALSE)
saveRDS(SpatialNeighborhood, paste0(OUTDIR, "/SpatialNeighborhood.RDS"))
# SpatialNeighborhood = readRDS(paste0(OUTDIR, "/SpatialNeighborhood.RDS"))
saveRDS(STexpr_all, paste0(OUTDIR, "/STexpr_all.RDS"))
# STexpr_all = readRDS(paste0(OUTDIR, "/STexpr_all.RDS"))
lapply(STexpr_all, function(STexpr) names(STexpr))
lapply(SpatialNeighborhood, function(STexpr) names(STexpr))
lapply(SpatialNeighborhood, function(x) length(x))
lapply(STexpr_all, function(x) length(x))

# Get ST data for only cancer
STexpr_cancer = lapply(names(STexpr_all), function(c) {
    cancer_samples = names(SpatialNeighborhood[[c]])
    STexpr_all[[c]][cancer_samples]
}) %>% setNames(names(STexpr_all))
saveRDS(STexpr_cancer, paste0(OUTDIR, "/STexpr_cancer.RDS"))
lapply(STexpr_cancer, function(STexpr) names(STexpr))

rm(STexpr_all)
gc()

```

## Make pseudo-bulk per slice

### Malignant spots on cacner slices

```{r}
#' Make pseudobulk data from spatial transcriptomics data.
#' 
#' Aggregate the gene count across all Malignant/Non-malignant spots for each cancer slice.
#' 
#' @param SpatialNeighborhood The spatial neighborhood data got with importFiles_SMTdb().
#' @param STexpr The spatial transcriptomcis expression data got with importFiles_SMTdb().
#' 
#' @return A list of pseudobulk RNA counts, with each component represent a cancer type.
#' 
#' @examples 
#' # getPseudoBulk(SpatialNeighborhood, STexpr)
#' 
getPseudoBulk_malignant = function(SpatialNeighborhood, STexpr) {
    lapply(names(SpatialNeighborhood), function(c) {
        print(c)
        sp_nei = SpatialNeighborhood[[c]]
        st_expr = STexpr[[c]]
        lapply(names(sp_nei), function(sample) {
            sp_nei_sample = sp_nei[[sample]] %>% na.omit() 
            st_expr_sample = st_expr[[sample]] %>% na.omit() %>%
                tibble::column_to_rownames(var="row.names")
            st_expr_sample = st_expr_sample[rowSums(st_expr_sample, na.rm=T)>1,]
            st_expr_sample = st_expr_sample[, colSums(st_expr_sample, na.rm=T)>0]
            Malignant_spots = sp_nei_sample %>% 
                filter(Location%in%c("Malignant", "Mal")) %>% pull(row.names) %>%
                gsub("-", ".", .) %>%
                intersect(., colnames(st_expr_sample))
            Non_malignant_spots = sp_nei_sample %>% 
                filter(Location%in%c("Non-malignant", "nMal")) %>% pull(row.names) %>%
                gsub("-", ".", .) %>%
                intersect(., colnames(st_expr_sample))
                # we only keep cancers with at least 3 slices
            if (length(Malignant_spots)>=3) {
                Malignant =  rowSums(st_expr_sample[, Malignant_spots], na.rm=T)} else {
                    Malignant = NA
                }
                # we only keep cancers with at least 3 slices
            if (length(Non_malignant_spots)>=3) {
                Non_malignant =  rowSums(st_expr_sample[, Non_malignant_spots], na.rm=T)} else {
                    Non_malignant = NA
                }
            data.frame(Feature=rownames(st_expr_sample),
                Malignant =  Malignant,
                Non_malignant = Non_malignant
            ) %>% setNames(c("Feature", paste0(c("Malignant", "nonMalignant"), "_", sample)))
        }) %>% Reduce(full_join, .)
    }) %>% setNames(names(SpatialNeighborhood))
}

# 
pseudoBulk_cancer = getPseudoBulk_malignant(SpatialNeighborhood, STexpr_cancer)
pseudoBulk_cancer = pseudoBulk_cancer[!is.na(pseudoBulk_cancer)]
saveRDS(pseudoBulk_cancer, paste0(OUTDIR, "/pseudoBulk_cancer.RDS"))
# pseudoBulk = readRDS(paste0(OUTDIR, "/pseudoBulk_cancer.RDS"))
openxlsx::write.xlsx(pseudoBulk_cancer, 
    file=paste0(OUTDIR, "/PseudoCount_filtered_cancer.xlsx"))

```

### Normal slices

```{r}
#' Generate pseudobulk data from spatial transcriptomics data
#' 
#' @param STexpr This is a list of lists of tables, each innernlist represents a cancer type, each table represents a slice, and each table should have features as rownames and spots as column names.
#' @examples 
#' # getPseudoBulk_normal(STexpr_normal)
#' 
getPseudoBulk_normal = function(STexpr) {
    lapply(names(STexpr), function(c) {
        st_expr = STexpr[[c]]
        lapply(names(st_expr), function(s) {
            expr = st_expr[[s]]
            data.frame(
                Feature = rownames(expr),
                Normal =  rowSums(expr, na.rm=T)
            ) %>% 
            setNames(c("Feature", paste0("Normal_", s)))
        }) %>% Reduce(full_join, .)
    }) %>% setNames(names(STexpr))
}

normal_slices = meta %>% filter(slice_type=="normal") %>% pull(datasetID)
STexpr_normal = lapply(STexpr_all, function(c) {
    dats = c[intersect(normal_slices, names(c))] 
    lapply(dats, function(dat) {
        dat %>%
        as.data.frame() %>%
        tibble::column_to_rownames(var="row.names")
    }) 
})

pseudoBulk_normal = getPseudoBulk_normal(STexpr_normal)
colnames(pseudoBulk_normal[[1]])

lapply(names(pseudoBulk_cancer), function(c) {
    # grep("^Malignant_",names(pseudoBulk_cancer[[c]]), value=TRUE)
    length(names(pseudoBulk_cancer[[c]]))
})

lapply(names(pseudoBulk_cancer), function(c) {
    length(names(pseudoBulk_normal[[c]]))
})

```



# GSVA metabolic differential analysis

## Get KEGG metabolic pathways

```{r}
# Get KEGG metabolic pathways
paste0(OUTDIR, "/KEGG_metabolism/") %>% 
    dir.create(., recursive=TRUE, showWarnings=FALSE)

kegg_metab_dbs = 
    getdb_metabolism(databaseDIR=paste0(OUTDIR, "/KEGG_metabolism/"))

kegg_metab_db = kegg_metab_dbs$kegg_metab_db

# databaseDIR=paste0(OUTDIR, "/KEGG_metabolism/")
# kegg_metab_dbs = 
# list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
#     kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))

# Make a circular packing plot to show the number of pathways in each class
getCirPackingPlot_GSVA(
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table,
    outdir = OUTDIR)

```

## Paired differential analysis on sample GSVA scores

```{r}
OUTDIR = "../results/SMTdb/"
GSVA_limma_rslt = lapply(names(pseudoBulk), function(c) {
    print(c)
    dat = pseudoBulk[[c]] 
    rownames(dat) = NULL
    dat = dat %>% as.data.frame() %>%
        tibble::column_to_rownames(var="Feature") %>% na.omit()
    meta = data.frame(Slice = colnames(dat) %>% 
                gsub("nonMalignant_", "", .) %>%
                gsub("Malignant_", "", .),
            Group = gsub("_.*", "", colnames(dat)),
            Sample = colnames(dat))
    meta$Group = factor(meta$Group, levels=c("nonMalignant", "Malignant")) # Reference first!
    paste0(OUTDIR, "/Limma_GSVA/", c, "/") %>%
        dir.create(., recursive=TRUE, showWarnings=FALSE)
    GSVAlimmaTest(dat, paired=TRUE, Feature.filtering=TRUE,
        currentCovariate=NULL, meta=meta, paired_variable="Slice", 
        kegg_metab_db=kegg_metab_dbs$kegg_metab_db, 
        OUTDIR=paste0(OUTDIR, "/Limma_GSVA/", c, "/"),
        min.count=5)
}) %>% setNames(names(pseudoBulk))

GSVA_limma_rslt = 
    GSVA_limma_rslt[sapply(GSVA_limma_rslt, function(g) {is.data.frame(g$gsva)})]

saveRDS(GSVA_limma_rslt, 
    paste0(OUTDIR, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))

# GSVA_limma_rslt = readRDS(paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
```

## Visualize differential results

```{r}
rlst_list = lapply(GSVA_limma_rslt, function(l) l$rslt_of_interest)
makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.2, w=8, h=8, 
        OUTDIRV = paste0(OUTDIR, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dys.Path_SMTdb", 
        suffix="SMTdb", nested=T,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

# temp = as.data.frame(GSVA_limma_rslt[["OSCC"]]$gsva)["Linoleic acid metabolism",] 
# temp[, grep("nonMalignant", colnames(temp))]  %>% as.numeric() %>% mean()
# temp[, !grepl("nonMalignant", colnames(temp))]  %>% as.numeric() %>% mean()                    

```










```{r}
stop
```

```{r}
outdir = "../results/TCGA"
dir.create(outdir, recursive=T, showWarnings=F)
```

## Only take paired data
```{r}
INDIR = "../data/TCGA/"
ctypes = list.dirs(INDIR, recursive=FALSE, full.name=F)
ctypes
paired_data = lapply(ctypes, function(ctype) {
    meta = openxlsx::read.xlsx(paste0(INDIR, ctype, "/Meta_", ctype, ".xlsx")) %>% 
        full_join(
            openxlsx::read.xlsx(paste0(INDIR, ctype, "/Clinical_", ctype, ".xlsx")) %>%
            select(-created_datetime),
        by=c("cases.submitter_id" = "submitter_id"))  # Here we have patient IDs
    cases.submitter_id2keep = 
        intersect(meta$cases.submitter_id[meta$sample_type=="Primary Tumor"],
        meta$cases.submitter_id[meta$sample_type=="Solid Tissue Normal"]) %>%
        na.omit()
    if (length(cases.submitter_id2keep)>3) {
        return(meta %>% filter(cases.submitter_id %in% cases.submitter_id2keep))
    } else {return(NA)}
}) %>% setNames(ctypes) %>% .[!is.na(.)]

lapply(names(paired_data), function(c) {
    temp = paired_data[[c]]
    temp = temp[temp$cases.submitter_id %in% 
        (which(table(temp$cases.submitter_id)>2) %>% names()),]
    as.data.frame(temp) %>% count(sample_type, cases.submitter_id)
}) %>% setNames(names(paired_data))

print("If there are more than one tumor data file per patient, pick the first one")
paired_data = lapply(paired_data, function(cancer) {
    cancer[!duplicated(paste0(cancer$sample_type,cancer$cases.submitter_id)),]
})
# Check 
lapply(names(paired_data), function(c) {
    temp = paired_data[[c]]
    temp = temp[temp$cases.submitter_id %in% 
        (which(table(temp$cases.submitter_id)>2) %>% names()),]
    as.data.frame(temp) %>% count(sample_type, cases.submitter_id)
}) %>% setNames(names(paired_data))

```

## Circular barplot to present sample size

```{r}
sampleSizeCircularBarplot = function(paired_data, OUTDIR) {
    sample_sizes = lapply(names(paired_data), function(cancer) {
        dat = paired_data[[cancer]]
        count = dat %>% select(sample_type, sample.submitter_id) %>% 
            unique() %>% group_by(sample_type) %>% 
            summarise(Count=n()) %>%
            mutate(sample_type=gsub("Primary Tumor", "Tumor", sample_type)) %>%
            mutate(sample_type=gsub("Solid Tissue Normal", "Normal", sample_type)) %>%
            mutate(Cancer_type=cancer)
        count = count %>% left_join(
                count %>% group_by(Cancer_type) %>%
                summarise(total=sum(Count))) %>%
                mutate(Cancer_type = paste0(Cancer_type, "\n(", total, ")"))
    }) %>% Reduce(rbind, .)

    TCGA_summary = sample_sizes %>%
    ggplot(., aes(x=Cancer_type, y=Count, fill=sample_type)) +
    geom_bar(stat="identity", show.legend=FALSE) +
    theme_minimal() +
    theme(axis.text.y = element_blank(),
            axis.ticks = element_blank(),
            axis.title = element_blank(),
            plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")) + 
    coord_polar() +
    ggtitle("TCGA") 
    pdf(paste0(outdir, "/Summary/Sample_size_circular_barplot.pdf"), 
        w=9, h=4.5)
        print(cowplot::plot_grid(TCGA_summary, TCGA_summary))
    dev.off()
}

paste0(outdir, "/Summary/") %>% dir.create(.,recursive=T, showWarnings=F)
sampleSizeCircularBarplot(
    paired_data, 
    OUTDIR=paste0(outdir, "/Summary/"))

```

## Import raw data overview

```{r}
# Import the data:
paired_data_count = lapply(names(paired_data), function(cancer) {
    folder = paste0("../data/TCGA/", cancer, "/GDCdata/TCGA-", 
        cancer, "/Transcriptome_Profiling/Gene_Expression_Quantification/")
    p = paired_data[[cancer]]
    files = paste0(folder,"/", p$file_id, "/", p$file_name)
    Samples = p$sample.submitter_id
    count_table = lapply(1:length(files), function(i) {
        file = files[[i]]
        expr = read.delim(file, skip=1, header=TRUE)
        expr$gene_id = sub("\\..*$", "", expr$gene_id)
        expr %>% as.data.frame() %>% 
            select(gene_id, gene_name, gene_type, unstranded) %>%
            setNames(c("gene_id", "gene_name", "gene_type", Samples[[i]]))
    })

    count_table

}) %>% setNames(names(paired_data))

# Combine all samples to the same table for each cancer type
paired_data_count = lapply(paired_data_count, function(cancer){
    temp2 = lapply(1:length(cancer), function(i) {
        sample = cancer[[i]]
        sample = sample[!duplicated(sample$gene_name), ]
        rownames(sample) = NULL
        temp = sample %>% as.data.frame() %>% 
            tibble::column_to_rownames(var="gene_name") 
        temp = temp[, 3, drop=FALSE]
        temp = temp[!duplicated(paired_data_count[[1]][[1]]$gene_name), ,drop=FALSE]
        if (i==1) {
            temp = temp %>% as.data.frame() %>% 
                tibble::rownames_to_column(var="gene_name")
        } else {rownames(temp)=NULL}
        temp
    }) %>% Reduce(cbind, .)
    temp2[5:nrow(temp2), ]
})

names(paired_data_count)
head(paired_data[[1]])

saveRDS(paired_data_count, paste0(outdir, "/paired_data_count.RDS"))

```

# Import KEGG metabolic pathways

```{r}
# Get KEGG metabolic pathways
paste0(outdir, "/KEGG_metabolism/") %>% 
    dir.create(., recursive=TRUE, showWarnings=FALSE)

kegg_metab_dbs = 
    getdb_metabolism(databaseDIR=paste0(outdir, "/KEGG_metabolism/"))

kegg_metab_db = kegg_metab_dbs$kegg_metab_db

# databaseDIR=paste0(outdir, "/KEGG_metabolism/")
# kegg_metab_dbs = 
# list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
#     kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))

# Make a circular packing plot to show the number of pathways in each class
getCirPackingPlot_GSVA(
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table)

```


# Paired differential analysis on sample GSVA scores

```{r}
GSVA_limma_rslt = lapply(names(paired_data_count), function(cancer) {
    print(cancer)
    dat = paired_data_count[[cancer]] 
    rownames(dat) = NULL
    dat = dat %>% as.data.frame() %>%
        tibble::column_to_rownames(var="gene_name") %>% na.omit()
    meta = paired_data[[cancer]] %>% as.data.frame() %>%
        mutate(Sample = sample.submitter_id) %>%
        mutate(Group = sample_type) %>%
        mutate(Patient = cases.submitter_id)
    meta$Group = factor(meta$Group, levels=c("Solid Tissue Normal", "Primary Tumor"))
    paste0(outdir, "/Limma_GSVA/", cancer, "/") %>%
        dir.create(., recursive=TRUE, showWarnings=FALSE)
    GSVAlimmaTest(dat, paired=TRUE, Feature.filtering=TRUE,
        currentCovariate=NULL, meta=meta, paired_variable="Patient", 
        kegg_metab_db=kegg_metab_db, 
        OUTDIR=paste0(outdir, "/Limma_GSVA/", cancer, "/"))
}) %>% setNames(names(paired_data_count))

saveRDS(GSVA_limma_rslt, 
    paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))

# GSVA_limma_rslt = readRDS(paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
```

## Visualize differential results

```{r}
library(dplyr)
library(stringr)
rlst_list = lapply(GSVA_limma_rslt, function(l) l$rslt_of_interest)
makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA", nested=T, 
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA_tops", nested=T, top=TRUE,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())

makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA2", nested=T, 
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique(),
        color_low="blue", color_high="red")

makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA_tops2", nested=T, top=TRUE,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique(),
        color_low="blue", color_high="red")

```


### Visualize the number of significant pathways

```{r}
paste0(outdir, "/Limma_GSVA") %>% 
    dir.create(., recursive=T, showWarnings=FALSE)
visualizeSigNr(OUTDIR=paste0(outdir, "/Limma_GSVA"), 
                GSVA_limma_rslt)

```


### Visualize differential effect

#### Differential effect boxplot by class

```{r}
outdir = "../results/TCGA"
GSVA_limma_rslt = 
    readRDS(paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
databaseDIR=paste0(outdir, "/KEGG_metabolism/")
kegg_metab_dbs = 
    list(kegg_metab_db = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db.RDS")),
    kegg_metab_db_table = readRDS(paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS")))

diffEffectBoxplot_bySystem_GSVA(rlst_list = GSVA_limma_rslt, 
  taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                    select(gs_name, class) %>% unique(), 
  OUT_DIR=paste0(outdir, "/Limma_GSVA"), w=2, h=3.5, title="TCGA")

```

#### Differential effect boxplot by cancer

```{r}
diffEffectBoxplot_byCancer_GSVA(
  rlst_list = GSVA_limma_rslt, 
  OUT_DIR=paste0(outdir, "/Limma_GSVA"), w=2.75, h=3.0, title="TCGA")
```


# Summarize GSVA scores

## Visualize pathway activity GSVA scores

### Per sample

Here, we will normalize all cancer type together and then calculate GSVA scores together to make it relatively more comparable across cancer types
```{r}
# Run GSVA all together 
paste0(outdir, "/Limma_GSVA/") %>% 
    dir.create(., recursive=TRUE, showWarnings=FALSE)
dat = paired_data_count %>% Reduce(function(x,y) full_join(x,y,by="gene_name"), .)
rownames(dat) = NULL
dat = dat %>% as.data.frame() %>%
    tibble::column_to_rownames(var="gene_name") %>% na.omit()
    
meta = lapply(names(paired_data), function(c) {
        paired_data[[c]] %>% as.data.frame() %>% 
            mutate(cancer_type=c) %>% 
            select(cancer_type, sample.submitter_id, sample_type, cases.submitter_id)
    }) %>% Reduce(rbind, .) %>% 
    mutate(Sample = sample.submitter_id) %>%
    mutate(Group = sample_type) %>%
    mutate(Patient = cases.submitter_id)
meta$Group = factor(meta$Group, levels=c("Solid Tissue Normal", "Primary Tumor"))

GSVA_limma_rslt_all = GSVAlimmaTest(dat, paired=TRUE, Feature.filtering=TRUE,
        currentCovariate=NULL, meta=meta, paired_variable="Patient", 
        kegg_metab_db=kegg_metab_db, 
        OUTDIR=paste0(outdir, "/Limma_GSVA/"))

GSVA_limma_rslt_gsva_all = lapply(names(paired_data), function(c) {
    GSVA_limma_rslt_all$gsva[,paired_data[[c]]$sample.submitter_id]
}) %>% setNames(names(paired_data))

visualizeGSVSscoresGroup(
    GSVA_limma_rslt_gsva = GSVA_limma_rslt_gsva_all, 
    paired_data = paired_data, 
    OUTDIR = paste0(outdir, "/Summary/")
)

```

### Per metabolic class

```{r}
visualizeGSVSscoresClass(GSVA_limma_rslt, 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table, 
    paired_data, 
    OUTDIR = paste0(outdir, "/Summary/"))
```

##  Visualize pathway sizes
```{r}
genes4GSVA = read.delim(
    file = paste0(outdir, "/Limma_GSVA/NormalizedCount.csv"),
    header = TRUE,
    stringsAsFactors = FALSE
    ) %>% pull(Features) %>% unique()

visualizeGSsizeClass(
    GSVA_limma_rslt_gsva=lapply(GSVA_limma_rslt, function(r) r$gsva), 
    genes4GSVA = genes4GSVA, 
    kegg_metab_db = kegg_metab_dbs$kegg_metab_db, 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table,
    OUTDIR=paste0(outdir, "/Summary/"))
```


# Visualize correlations

## Correlation between normal pathway GSVA scores and differential effect size

```{r}
# dir.create(paste0(outdir, 
#     "/correlation_scoreVSt/"), recursive=T, showWarnings=F)
# rslt_of_interest = lapply(GSVA_limma_rslt, function(g) g$rslt_of_interest)
# # Here the normal sample GSVA scores from different cancer types were calculated together
# getNormalVStcorPlot(
#         GSVA_limma_rslt_gsva=GSVA_limma_rslt_gsva_all,  
#         rslt_of_interest, paired_data, 
#         OUTDIR=paste0(outdir, "/correlation_scoreVSt/"),
#         sample="normal")

```

## Correlation between normal pathway GSVA scores and differential effect size

```{r}
# # Here the normal sample GSVA scores from different cancer types were calculated together
# getNormalVStcorPlot(
#         GSVA_limma_rslt_gsva=GSVA_limma_rslt_gsva_all,  
#         rslt_of_interest, paired_data, 
#         OUTDIR=paste0(outdir, "/correlation_scoreVSt/"),
#         sample="all")

```

## Correlation between normal pathway gene expr (median) and differential effect size

```{r}
# normalizedExpr_all = read.delim(
#     file = paste0(outdir, "/Limma_GSVA/NormalizedCount.csv"),
#     header = TRUE,
#     stringsAsFactors = FALSE
#     ) %>% tibble::column_to_rownames(var="Features")
# colnames(normalizedExpr_all) = gsub("[.]", "-", colnames(normalizedExpr_all))
# getNormalExprVStcorPlot(normalizedExpr_all, 
#     kegg_metab_db=kegg_metab_dbs$kegg_metab_db,
#     GSVA_limma_rslt_gsva=lapply(GSVA_limma_rslt, function(r) r$gsva), 
#     rslt_of_interest=lapply(GSVA_limma_rslt, function(r) r$rslt_of_interest), 
#     paired_data, 
#     OUTDIR = "../results/TCGA/Limma_GSVA/")
```


```{r}
save.image(paste0(outdir, "/TCGA.RData"))
# load(paste0(outdir, "/TCGA.RData"))
out = capture.output(sessionInfo())
writeLines(out, paste0(outdir, "/sessionInfo.txt"))

```


```{r}
#install.packages("DiagrammeR")
library(DiagrammeR)

graph = grViz("
digraph nfcore_style_pipeline {
  graph [rankdir=TB, fontsize=12, splines=true]

  node [shape=box, style=filled, fillcolor=white, color=gray30, fontname=Arial, fontsize=10, penwidth=1]

  # Cluster for TFfootprint
  subgraph cluster_TFfootprint {
    label = 'TFfootprint (ATAC-seq)'
    style=rounded
    color=purple
    MERGEBAM [label='MERGEBAM']
    ATACORRECT [label='ATACORRECT']
    FOOTPRINTSCORES [label='FOOTPRINTSCORES']
    BINDETECT [label='BINDETECT']
    FOOTPRINTPLOT [label='FOOTPRINTPLOT']
    BINDINGHEATMAP [label='BINDINGHEATMAP']

    MERGEBAM -> ATACORRECT -> FOOTPRINTSCORES -> BINDETECT -> FOOTPRINTPLOT -> BINDINGHEATMAP
  }

  # Cluster for TFmotifEnrichment
  subgraph cluster_TFmotifEnrichment {
    label = 'TFmotifEnrichment (ATAC-seq)'
    style=rounded
    color=purple
    MONALISA [label='MONALISA']
  }

  # Cluster for TF activity
  subgraph cluster_TFactivity {
    label = 'TF activity (RNA-seq)'
    style=rounded
    color=orange
    DIFFTFACTIVITY [label='DIFFTFACTIVITY']
  }

  # Cluster for TF expression
  subgraph cluster_TFexpression {
    label = 'TF expression (RNA-seq)'
    style=rounded
    color=orange
    TFexpression
  }

  # Integration process
  node [shape=rectangle, style=filled, fillcolor=snow2, fontname=Arial, fontsize=11]
  INTEGRATION [label='INTEGRATION']

  # Connections to integration
  BINDINGHEATMAP -> INTEGRATION
  MONALISA -> INTEGRATION
  DIFFTFACTIVITY -> INTEGRATION
  TFexpression -> INTEGRATION

  # Layout the clusters left to right
  { rank = same; cluster_TFfootprint; cluster_TFmotifEnrichment; cluster_TFactivity; cluster_TFexpression}
}
")

htmlwidgets::saveWidget(graph, "diagram.html")

```