---
title: "TCGA"
author: Yuan Li (yuan.li@nbis.se)
format: html
---




# Define functions

```{r}
# invisible(capture.output({library(limma)}))
# invisible(capture.output({library(tibble)}))
#' @param dat the dataset needs to have features as rownames, samples as column names
#' @param meta meta data needs to have at least two columns named "Sample", "Group", for the moment, we only support Group with two categories
#'
GSVAlimmaTest = function(dat, paired=TRUE, Feature.filtering=TRUE,
        currentCovariate=NULL, meta=meta, paired_variable="Patient", 
        OUTDIR, kegg_metab_db) {
    meta = filter(meta, Sample%in%colnames(dat))
    # Make sure the sample are in the same order in both meta and dat:
    dat = dat[, meta$Sample]
    meta = meta %>% mutate(across(where(is.factor), droplevels))
    mapping = setNames(meta$Group, meta$Sample)
    print("Only keep features that have at least min.nr number of values that are >5 in each group...")
    min.nr = round(max((ncol(dat)/2)*0.15, 2), 0)
    if (Feature.filtering) {
        B = mapping[colnames(dat)] %>% as.character()
        featureskeep = apply(dat>5, 1, function(row) {
            A = row
            sum(aggregate(A~B, data=data.frame(A=A,B=B), sum, na.rm=T)[,"A"]>=min.nr)>1
        })
        dat = dat[featureskeep, ]
    }

    # library(limma)
    # conda install conda-forge::r-locfit
    # conda install bioconda::bioconductor-edger
    # BiocManager::install("edgeR")
    # library(edgeR)
    Group = factor(meta$Group)
    # Create a DEGList from the edgeR package
    dge = edgeR::DGEList(counts=dat)
    # Normalize the data
    dge = edgeR::calcNormFactors(dge) 
    # Note: calcNormFactors doesn’t normalize the data, it just calculates normalization factors for use downstream.
    pdf(paste0(OUTDIR, "/MDS_plot.pdf"))
        print(plotMDS(dge, col=as.numeric(Group)))
    dev.off()  

    #if (!is.null(currentCovariate)) {currentCovariate=NULL}
    baseFormula = ~ Group
    if (is.null(currentCovariate)) {
        currentFormula = baseFormula
    } else {
        currentFormula = as.formula(paste0("~ ", currentCovariate, " + Group"))}
    
    design = model.matrix(currentFormula, data=meta)
    colnames(design) = gsub("Group", "", colnames(design))
    dat2 = limma::voom(dge, design, plot=T)
    # What is voom doing?
    # Counts are transformed to log2 counts per million reads (CPM), where “per million reads” is defined based on the normalization factors we calculated earlier
    # A linear model is fitted to the log2 CPM for each gene, and the residuals are calculated
    # A smoothed curve is fitted to the sqrt(residual standard deviation) by average expression (see red line in plot above)
    # The smoothed curve is used to obtain weights for each gene and sample that are passed into limma along with the log2 CPMs
    # Save normalized count
    normalizedCount = dat2$E
    write.table(
        normalizedCount %>% as.data.frame() %>% 
                tibble::rownames_to_column(var="Features"), 
        paste0(OUTDIR, "/NormalizedCount.csv"),
        quote = F, row.names = F, col.names = T, sep="\t")  
    
    # Calcualte GSVA per sample
    # BiocManager::install("GSVA")
    # library(GSVA)
    # library(tibble)
    # library(dplyr)
    normalizedCount = as.matrix(normalizedCount)
    normalizedCount = matrix(as.numeric(normalizedCount), 
        nrow = nrow(normalizedCount), 
        ncol = ncol(normalizedCount), dimnames = dimnames(normalizedCount))
    param = GSVA::gsvaParam(normalizedCount, kegg_metab_db, minSize=5, maxSize=500)
    gsva_scores = GSVA::gsva(param, verbose=TRUE)
    write.table(
        gsva_scores %>% as.data.frame() %>% 
                tibble::rownames_to_column(var="Features"), 
        paste0(OUTDIR, "/GSVAscores.csv"),
        quote = F, row.names = F, col.names = T, sep="\t")  

    if (paired) {
        #Fit with correlated arrays
        dupcor = limma::duplicateCorrelation(gsva_scores, design, 
            block=meta[,paired_variable,drop=T])
        fit = limma::lmFit(gsva_scores, design, block=meta[,paired_variable,drop=T], 
            correlation=dupcor$consensus)
    } else { fit = limma::lmFit(gsva_scores, design) }

    # Limma trend to refine gene-level variance estimates
    gssizes = sapply(kegg_metab_db[rownames(fit$coefficients)], function(gs) {
        length(gs)
    })
    fit.con = limma::eBayes(fit, robust = TRUE, trend=gssizes)
    rlst_interest = 
        limma::topTable(fit.con, n=Inf, coef=levels(meta$Group)[2]) %>%
        arrange(-t) 
    
    return(list(rslt_of_interest=rlst_interest, gsva=gsva_scores, fitted.model=fit))  

}


# library(KEGGREST)
# install.packages("BiocManager")
# BiocManager::install("KEGGREST")

#' @import dplyr
#'
getdb_metabolism = function(databaseDIR) {
    # Get all human metabolic pathways
    query = KEGGREST::keggGet(c("hsa01100"))
    metab_pathways = names(query[[1]]$REL_PATHWAY)

    get_pathway_info = function(path_id) {
      info = KEGGREST::keggGet(path_id)[[1]]
      gene_ids = if (!is.null(info$GENE)) info$GENE[seq(1, length(info$GENE), 2)] else character()
      gene_symbols = if (!is.null(info$GENE)) info$GENE[seq(2, length(info$GENE), 2)]%>%
            strsplit(., ";") %>% sapply(., function(g) g[1]) else character()
      gene_infos = if (!is.null(info$GENE)) info$GENE[seq(2, length(info$GENE), 2)]%>%
            strsplit(., ";") %>% sapply(., function(g) g[2]) else character()   
      compound_ids = if (!is.null(info$COMPOUND)) names(info$COMPOUND) else character() 
      compound_Names = if (!is.null(info$COMPOUND)) as.character(info$COMPOUND) else character() 
      class = if (!is.null(info$CLASS)) as.character(info$CLASS) else character()      
      list(
        ID = as.character(info$ENTRY),
        pathway_name = info$NAME,
        genes = gene_ids,
        gene_symbols = gene_symbols,
        gene_infos = gene_infos,
        compounds = compound_ids,
        compound_Names = compound_Names,
        class = class
      )
    }

    metabolism_Pathways = 
        lapply(metab_pathways, function(path_id) get_pathway_info(path_id))

    # Remove empty pathways
    metabolism_Pathways %>% lapply(., function(p) {
        if(length(p$genes)!=0) {
            data.frame(
            ID = p$ID,
            pathway_name = p$pathway_name %>% gsub(" - Homo sapiens \\(human\\)", "", .),
            genes = paste(p$genes, collapse=";"),
            gene_symbols = paste(p$gene_symbols, collapse=";"),
            gene_infos = paste(p$gene_infos, collapse=";"),
            compounds = paste(p$compounds, collapse=";"),
            compound_Names = paste(p$compound_Names, collapse=";"),
            class = p$class)
    }}) %>% Filter(Negate(is.null), .) %>%
        Reduce(rbind, .) %>% 
        write.table(paste0(databaseDIR, "/KEGG_metabolism.csv"), 
            quote=F, sep="\t", row.names=F)

    # Create KEGG metabolic database
    kegg_metab = read.delim(
        paste0(outdir, "/KEGG_metabolism/KEGG_metabolism.csv"),
        header = TRUE, stringsAsFactors = FALSE) %>%
        mutate(class = gsub("Metabolism; ", "", class))
    kegg_metab_db = 
        stringr::str_split(kegg_metab[,"gene_symbols",drop=T], ";") %>%
        setNames(., kegg_metab$pathway_name)

    path_class = setNames(kegg_metab$class, kegg_metab$pathway_name)
    kegg_metab_db_table = lapply(seq_along(kegg_metab_db), function(i) {
        p = kegg_metab_db[[i]]
        data.frame(gs_name=rep(names(kegg_metab_db)[i],length(p)), 
            gene_symbol = p, class=path_class[names(kegg_metab_db)[i]])
        }) %>% Reduce(rbind, .) %>% tibble::as_tibble()
    saveRDS(kegg_metab_db, paste0(databaseDIR, "/KEGG_metabolism_db.RDS"))
    saveRDS(kegg_metab_db_table, paste0(databaseDIR, "/KEGG_metabolism_db_table.RDS"))

    return(list(kegg_metab_db=kegg_metab_db, kegg_metab_db_table=kegg_metab_db_table))

} 


# library(ggraph)
# library(igraph)
getCirPackingPlot = function(kegg_metab_db_table) {
    pathway_hierachy =  kegg_metab_db_table %>% mutate(Depth0="Metabolism") %>%
        mutate(Depth2=gs_name) %>% mutate(Depth1=class) %>% 
        dplyr::select(Depth0, Depth1, Depth2) %>% unique()
    edges = pathway_hierachy %>% .[,c("Depth0","Depth1")] %>%
        unique()%>%setNames(c("from", "to")) %>%
        rbind(pathway_hierachy[, c("Depth1","Depth2")] %>% 
            unique() %>% setNames(c("from", "to"))) 
    vertices = pathway_hierachy %>% .[, c("Depth0","Depth2")] %>%
        unique()%>%pull(Depth0)%>%table()%>%as.data.frame()%>%setNames(c("name", "size")) %>%
        rbind(pathway_hierachy[,c("Depth1","Depth2")]%>%unique()%>%pull(Depth1)%>%table()%>%
            as.data.frame()%>%setNames(c("name", "size"))) %>%
        rbind(pathway_hierachy %>% pull(Depth2) %>% unique() %>% table() %>%
            as.data.frame() %>% setNames(c("name","size"))) %>%
            mutate(shortName=gsub("etabolism",".",paste0(name,"(",size,")"))) %>%
            mutate(shortName=gsub("iosynthesis","iosyn.", shortName)) %>%
            mutate(shortName=gsub("secondary metabolites", "sec. mets.", shortName))
    vertices$shortName[!vertices$name%in%unique(pathway_hierachy$Depth1)] = NA
    mygraph = igraph::graph_from_data_frame(edges, vertices=vertices)       
    p = ggraph::ggraph(mygraph, layout = 'circlepack', weight=size) +
        ggraph::geom_node_circle(aes(col=depth)) +
        ggraph::geom_node_label(aes(label=shortName),repel=T,size=4,color="darkgreen") +
        ggplot2::theme_void() +
        viridis::scale_color_viridis() +
        ggplot2::theme(legend.position="FALSE") +
        ggplot2::ggtitle("                                           TCGA")

    pdf(paste0(outdir, "/KEGG_metabolism/TaskSummary.pdf"), h=5, w=9.7)
        print(cowplot::plot_grid(p,p,ncol=2))
    dev.off()
}

#' @param GSVA_limma_rslt_gsva a list of gsva result, the names of the lsit should indicate groups, each dataset has features as rownames
#' @param paired_data a list of metadata, which should correspond to GSVA_limma_rslt_gsva, each metadata table should at least have two columns: ´sample_type´ and ´sample.submitter_id´
#' 
visualizeGSVSscoresGroup = function(GSVA_limma_rslt_gsva, paired_data, OUTDIR) {
    # GSVA scores for all samples
    dat = lapply(seq_along(GSVA_limma_rslt_gsva), function(i) {
        data.frame(
            means = (
                rowSums(GSVA_limma_rslt_gsva[[i]],na.rm=TRUE)/ncol(GSVA_limma_rslt_gsva[[i]])) %>% 
                    as.numeric(), 
            cancer_type = names(GSVA_limma_rslt_gsva)[i])
        }) %>% Reduce(rbind, .) 
    cancer_type_order = dat %>% group_by(cancer_type) %>% 
            summarise(Means=mean(means)) %>% 
            arrange(Means) %>% pull(cancer_type) %>% unique()
    P1 = dat %>% mutate(cancer_type = factor(cancer_type, levels=cancer_type_order)) %>%
            ggplot2::ggplot(., ggplot2::aes(x=cancer_type, y=means)) + 
            ggplot2::geom_boxplot(fill="green4") +
            ggplot2::labs(title = "TCGA",
                x="Cancer_type", y="Mean Metab. Act.")+
            ggplot2::theme_minimal() +
            ggplto2::theme(legend.position="none",
                axis.text.x = element_text(angle = 90, hjust=1, vjust=0))

    # GSVA scores for healthy control samples only
    dat = lapply(seq_along(GSVA_limma_rslt_gsva), function(j) {
        normal_samples = paired_data[[j]] %>% 
            filter(sample_type=="Solid Tissue Normal") %>% pull(sample.submitter_id)
        gsva = GSVA_limma_rslt_gsva[[j]] %>% .[, normal_samples]
        data.frame(
            means = (rowSums(gsva, na.rm=TRUE)/ncol(gsva)) %>% as.numeric(), 
            cancer_type = names(GSVA_limma_rslt_gsva)[j])
        }) %>% Reduce(rbind, .) 
    cancer_type_order = dat %>% group_by(cancer_type) %>% 
            summarise(Means=mean(means)) %>% 
            arrange(Means) %>% pull(cancer_type) %>% unique()
    P2 = dat %>% mutate(cancer_type = factor(cancer_type, levels=cancer_type_order)) %>%
            ggplot2::ggplot(., ggplot2::aes(x=cancer_type, y=means)) +
            ggplot2::geom_boxplot(fill="green4") +
            ggplot2::labs(title = "TCGA",
                x="Cancer_type", y="Mean Norm. Metab. Act.")+
            ggplot2::theme_minimal() +
            ggplot2::theme(legend.position="none",
                axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0))

    pdf(paste0(OUTDIR, "/GSVA_samples.pdf"), w=5.5, h=3.0)
        print(cowplot::plot_grid(P1, P2))
    dev.off()

}

#' @param GSVA_limma_rslt_gsva a list of gsva tables
#' @param genes4GSVA a vector of genes used for calculating GSVA scores
#' @param kegg_metab_db a list of pathways showing their gene members
#' @param kegg_metab_db_table a table showing the pathways (´gs_name´) and their classes (´class´)
#' 
visualizeGSsizeClass = function(
        GSVA_limma_rslt_gsva, genes4GSVA, 
        kegg_metab_db, kegg_metab_db_table,
        OUTDIR) {
    pathway_class_mapping = 
        setNames(kegg_metab_db_table$class, kegg_metab_db_table$gs_name)
    P = lapply(GSVA_limma_rslt_gsva, function(g) {
        # get pathways
        pathways = rownames(g)
        pathway_sizes = sapply(kegg_metab_db[pathways],function(d) {
            intersect(d, genes4GSVA) %>% length()
        })
        data.frame(Pathway = names(pathway_sizes),
                   Psize = as.numeric(pathway_sizes),
                   Class = pathway_class_mapping[names(pathway_sizes)])
    }) %>% Reduce(rbind, .) %>%
        mutate(Class = gsub("etabolism",".", Class)) %>%
        mutate(Class = gsub("thesis", ".", Class)) %>%
        mutate(Class = gsub("secondary metabolites", "sec. mets.", Class)) %>%
        mutate(Class = gsub("biodegradation", "biodegrad.", Class)) %>%
        ggplot2::ggplot(., ggplot2::aes(x=Class, y=Psize)) +
        ggplot2::geom_boxplot(fill="green4") +
        ggplot2::labs(# title = "TCGA",
            x="", y="Pathway size")+
        ggplot2::theme_minimal() +
        ggplot2::theme(legend.position="none",
            axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0))
    pdf(paste0(OUTDIR, "/Pathway_size_class.pdf"), w=2.5, h=3.5)
        print(P)
    dev.off()
}


visualizeGSVSscoresClass = function(GSVA_limma_rslt, kegg_metab_db_table, paired_data, OUTDIR) {
    dat = lapply(seq_along(GSVA_limma_rslt), function(i) {
        s = names(GSVA_limma_rslt)[i]
        pathway_means = rowSums(GSVA_limma_rslt[[i]]$gsva,na.rm=TRUE)/ncol(GSVA_limma_rslt[[i]]$gsva)
        kegg_metab_db_table = 
            kegg_metab_db_table %>% dplyr::select(gs_name, class) %>%
            unique() 
        kegg_metab_db_mapping = setNames(kegg_metab_db_table$class, kegg_metab_db_table$gs_name)
        data.frame(Pathway = names(pathway_means), 
                    Means = as.numeric(pathway_means), 
                    Class = kegg_metab_db_mapping[names(pathway_means)]) %>%
                    mutate(Class = gsub("etabolism",".", Class)) %>%
                    mutate(Class = gsub("thesis", ".", Class)) %>%
                    mutate(Class = gsub("secondary metabolites", "sec. mets.", Class)) %>%
                    mutate(Class = gsub("biodegradation", "biodegrad.", Class))
        }) %>% Reduce(rbind, .) 
        Class_order = dat %>% group_by(Class) %>% 
            summarise(Means=mean(Means)) %>% arrange(Means) %>% pull(Class) %>% unique()
    P1 = dat %>% mutate(Class = factor(Class, levels=Class_order)) %>%
            ggplot2::ggplot(., ggplot2::aes(x=Class, y=Means)) +
            ggplot2::geom_boxplot(fill="green4") +
            ggplot2::labs(# title = "TCGA",
                x="", y="Mean Metab. Act.")+
            ggplot2::theme_minimal() +
            ggplot2::theme(legend.position="none",
                axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0))

    dat = lapply(seq_along(GSVA_limma_rslt), function(j) {
        s = names(GSVA_limma_rslt)[j]
        normal_samples = paired_data[[s]] %>% 
            filter(sample_type=="Solid Tissue Normal") %>% pull(sample.submitter_id)
        gsva = GSVA_limma_rslt[[s]]$gsva %>% .[, normal_samples]
        pathway_means = rowSums(gsva,na.rm=TRUE)/ncol(gsva)
        kegg_metab_db_table = 
            kegg_metab_dbs$kegg_metab_db_table %>% dplyr::select(gs_name, class) %>%
            unique() 
        kegg_metab_db_mapping = setNames(kegg_metab_db_table$class, kegg_metab_db_table$gs_name)
        data.frame(Pathway = names(pathway_means), 
                    Means = as.numeric(pathway_means), 
                    Class = kegg_metab_db_mapping[names(pathway_means)]) %>%
                    mutate(Class = gsub("etabolism",".",Class)) %>%
                    mutate(Class = gsub("thesis", ".", Class)) %>%
                    mutate(Class = gsub("secondary metabolites", "sec. mets.", Class)) %>%
                    mutate(Class = gsub("biodegradation", "biodegrad.", Class))
        }) %>% Reduce(rbind, .) 
        Class_order = dat %>% group_by(Class) %>% 
            summarise(Means=mean(Means)) %>% arrange(Means) %>% pull(Class) %>% unique()
    P2 = dat %>% mutate(Class = factor(Class, levels=Class_order)) %>%
            ggplot2::ggplot(., ggplot2::aes(x=Class, y=Means)) +
            ggplot2::geom_boxplot(fill="green4") +
            ggplot2::labs(# title = "TCGA",
                x="", y="Mean Norm. Metab.")+
            ggplot2::theme_minimal() +
            ggplot2::theme(legend.position="none",
                axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0))

    pdf(paste0(OUTDIR, "/GSVA_pathway_class.pdf"), w=5.5, h=3.5)
        cowplot::plot_grid(P1, P2)
    dev.off()
}


# library(tidyr)
visualizeSigNr = function(OUTDIR, GSVA_limma_rslt) {
    pdf(paste0(OUTDIR, "/Sig_pathway_nr.pdf"), w=3, h=3.5)
        sig_nr = lapply(names(GSVA_limma_rslt), function(cancer) {
            rslt = GSVA_limma_rslt[[cancer]]$rslt_of_interest %>% 
                filter(adj.P.Val<0.05) 
            data.frame(up = rslt %>% filter(logFC>0) %>% nrow(),
                down = -(rslt %>% filter(logFC<0) %>% nrow())) %>%
                t() %>% as.data.frame() %>% setNames(cancer) 
            }) %>% Reduce(cbind, .) %>% 
                tibble::rownames_to_column(var="Regulation") %>%
                tidyr::gather(key="Cancer_type", value="Count", -Regulation) %>%
                arrange(-Count) %>%
                mutate(Cancer_type=factor(Cancer_type,levels=unique(Cancer_type))) 
        print(ggplot2::ggplot(sig_nr, ggplot2::aes(x=Cancer_type, y=Count, fill=Regulation)) +
            ggplot2::geom_col() +
            ggplot2::scale_fill_manual(values=c("up"="firebrick","down"="steelblue")) +
            ggplot2::geom_hline(yintercept=0, color="black") +
            ggplot2::labs(title = "Sig. diff. systems",
                x="Cancer_type", y="Sig. Count")+
            ggplot2::theme_minimal() +
            ggplot2::theme(legend.position="none",
                axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0)))
    dev.off()
}

getNormalVStcorPlot = function(
        GSVA_limma_rslt_gsva, rslt_of_interest, paired_data, 
        sample="normal", OUTDIR) {
    dat.list = lapply(seq_along(GSVA_limma_rslt_gsva), function(j) {
        s = names(GSVA_limma_rslt_gsva)[j]
        if (sample=="normal") {
            normal_samples = paired_data[[s]] %>% 
                filter(sample_type=="Solid Tissue Normal") %>% pull(sample.submitter_id)
            gsva = GSVA_limma_rslt_gsva[[s]] %>% .[, normal_samples] ###
        } else if (sample=="all") {
            gsva = GSVA_limma_rslt_gsva[[s]]
        } else {stop("parameter sample should be either ´normal´ or ´all´")}
        pathway_t = setNames(rslt_of_interest[[j]]$t, rownames(rslt_of_interest[[j]])) ###
        temp = data.frame(
            means = (rowSums(gsva, na.rm=TRUE)/ncol(gsva)) %>% as.numeric(), 
            cancer_type = names(GSVA_limma_rslt_gsva)[j],
            effect_t = pathway_t[rownames(gsva)] %>% as.numeric())
        temp
    }) %>% setNames(names(GSVA_limma_rslt_gsva))

    dat = lapply(seq_along(dat.list), function(k) {
        temp = dat.list[[k]]
        n = nrow(temp)
        temp = cor.test(temp$effect_t, temp$means)
        temp = data.frame(cor=temp$estimate, p.value=temp$p.value)
        temp$Cancer_type = names(dat.list)[k]
        temp$Study="TCGA"
        temp$SampleSize = n
        temp
    }) %>% Reduce(rbind, .) %>%
        mutate(p.adj = p.adjust(p.value, method="BH")) %>%
        arrange(p.adj) %>%
        mutate(Regulation = ifelse(p.adj<0.05, "Sig", "No"))

    dat$Cancer_type = factor(dat$Cancer_type, levels=dat$Cancer_type)
    pdf(paste0(OUTDIR, "/Cor_ScoreVSt_TCGA_", sample,".pdf"), h=3.85, w=5.5)
        print(ggplot2::ggplot(dat, ggplot2::aes(x=Cancer_type, y=cor)) +
            ggplot2::geom_point(aes(shape=Regulation, color=cor), size=5) +
            ggplot2::ggtitle(paste0("corr(",sample,", effect)")) +
            ggplot2::scale_size(trans="reverse") +
            ggplot2::scale_color_gradient2(low="steelblue", 
                            mid="snow2", high="firebrick", midpoint=0) +
            ggplot2::theme_bw(base_size = 15) + 
            ggplot2::guides(alpha="none", size=ggplot2::guide_legend(override.aes=list(shape=21))) +
            ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0)) +
            ggplot2::labs(X="Cancer_type", y="Pearson correlation",
                caption=paste0("n = ", min(dat$SampelSize), "-", 
                max(dat$SampelSize), ", p.adj Method = BH")))
    dev.off()
}


getNormalExprVStcorPlot = function(normalizedExpr_all, kegg_metab_db,
        GSVA_limma_rslt_gsva, rslt_of_interest, paired_data, OUTDIR) {
    dat.list = lapply(seq_along(GSVA_limma_rslt_gsva), function(j) {
        s = names(GSVA_limma_rslt_gsva)[j]
        normal_samples = paired_data[[s]] %>% 
            filter(sample_type=="Solid Tissue Normal") %>% pull(sample.submitter_id)
        normExpr = normalizedExpr_all[, intersect(normal_samples, colnames(normalizedExpr_all))]
        pathways = rownames(GSVA_limma_rslt_gsva[[j]])
        Pexpr = lapply(pathways, function(p) {
            Pgenes = intersect(kegg_metab_db[[p]], rownames(normalizedExpr_all))
            dat = data.frame(Pmedians = normExpr[Pgenes,] %>% as.matrix() %>%
                matrixStats::colMedians(.,na.rm=T)) %>%
                t()
            rownames(dat) = p
            dat
        }) %>% Reduce(rbind, .)
        pathway_t = setNames(rslt_of_interest[[j]]$t, rownames(rslt_of_interest[[j]])) ###
        temp = data.frame(
            means = (rowSums(Pexpr, na.rm=TRUE)/ncol(Pexpr)) %>% as.numeric(), 
            cancer_type = names(GSVA_limma_rslt_gsva)[j],
            effect_t = pathway_t[rownames(Pexpr)] %>% as.numeric())
        temp
    }) %>% setNames(names(GSVA_limma_rslt_gsva))

    dat = lapply(seq_along(dat.list), function(k) {
        temp = dat.list[[k]]
        n = nrow(temp)
        temp = cor.test(temp$effect_t, temp$means)
        temp = data.frame(cor=temp$estimate, p.value=temp$p.value)
        temp$Cancer_type = names(dat.list)[k]
        temp$Study="TCGA"
        temp$SampleSize = n
        temp
    }) %>% Reduce(rbind, .) %>%
        mutate(p.adj = p.adjust(p.value, method="BH")) %>%
        arrange(p.adj) %>%
        mutate(Regulation = ifelse(p.adj<0.05, "Sig", "No"))

    dat$Cancer_type = factor(dat$Cancer_type, levels=dat$Cancer_type)
    pdf(paste0(OUTDIR, "/Cor_pathwExprVSt_TCGA.pdf"), h=3.85, w=5.5)
        print(ggplot2::ggplot(dat, ggplot2::aes(x=Cancer_type, y=cor)) +
            ggplot2::geom_point(aes(shape=Regulation, color=cor), size=5) +
            ggplot2::ggtitle("corr(NormalExpr., effect)") +
            ggplot2::scale_size(trans="reverse") +
            ggplot2::scale_color_gradient2(low="steelblue", 
                            mid="snow2", high="firebrick", midpoint=0) +
            ggplot2::theme_bw(base_size = 15) + 
            ggplot2::guides(alpha="none", size=ggplot2::guide_legend(override.aes=list(shape=21))) +
            ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0)) +
            ggplot2::labs(X="Cancer_type", y="Pearson correlation",
                caption=paste0("n = ", min(dat$SampelSize), "-", 
                max(dat$SampelSize), ", p.adj Method = BH")))
    dev.off()
}


#' @param rlst_list a list of limma differential analysis output
#' @param taskHierarchy a table at least with two columns: Depth and ´Depth1´
#' 
makeLimmaDotplot_TCGA = function(rlst_list, adj.P.Val.cutoff, w=NULL, h=NULL, OUTDIRV, 
        Depth="Depth3", lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA", taskHierarchy=cfs, nested=T, top=FALSE,
        color_low="skyblue", color_high="purple"){
    taskHierarchy$Task = taskHierarchy[[Depth]]
    rlst_long = lapply(seq_along(rlst_list), function(i) {
        rlst = rlst_list[[i]]
        rlst$Cancer = names(rlst_list)[i]
        rlst$Regulation = 
            ifelse(rlst$logFC>0&rlst$adj.P.Val<adj.P.Val.cutoff, "Up", 
            ifelse (rlst$logFC<0&rlst$adj.P.Val<adj.P.Val.cutoff, "Down", "NonSig"))
        rlst %>% tibble::rownames_to_column(var="Task") %>% 
            filter(adj.P.Val<adj.P.Val.cutoff)
    }) %>% Reduce(rbind, .) %>%
        dplyr::left_join(taskHierarchy) %>%
            mutate(logFC=round(logFC, 2)) %>%
            mutate(Class = gsub("Carbohydrate metabolism","CAR",class)) %>%
            mutate(Class = gsub("Lipid metabolism", "LIP", Class)) %>%
            mutate(Class = gsub("Metabolism of cofactors and vitamins", "COF", Class)) %>%
            mutate(Class = gsub("Energy metabolism", "ENE", Class)) %>%
            mutate(Class = gsub("Amino acid metabolism", "AMI", Class)) %>%
            mutate(Class = gsub("Nucleotide metabolism", "NUC", Class)) %>%
            mutate(Class = gsub("Biosynthesis of other secondary metabolites", "OSM", Class)) %>%
            mutate(Class = gsub("Metabolism of other amino acids", "OAA", Class)) %>%
            mutate(Class = gsub("Glycan biosynthesis and metabolism", "GLY", Class)) %>%
            mutate(Class = gsub("Metabolism of terpenoids and polyketides", "TER", Class)) %>%
            mutate(Class = gsub("Xenobiotics biodegradation and metabolism" , "XEN", Class))

    if (top) {
        temp = rlst_long %>% dplyr::select(Task, Cancer, Regulation) %>% 
            unique() %>% filter(Regulation %in% c("Up", "Down")) %>%
            pull(Task) %>% table()
        tops = names(temp)[temp > median(temp)]
        rlst_long = rlst_long %>% filter(Task %in% tops)
    }
    #rlst_long$Task = factor(rlst_long$Task, levels=rev(c(ups, downs) %>% unique()))
    rlst_long$t[rlst_long$t>10] = 10
    rlst_long$t[rlst_long$t< -10] = -10
  
    library(ggplot2)
    library(ggh4x) 
    p = ggplot2::ggplot(data=rlst_long, mapping=ggplot2::aes_string(x="Cancer", y="Task")) +
        ggplot2::geom_point(ggplot2::aes(size=adj.P.Val, color=t, shape=Regulation))
    if (nested) p = p + ggh4x::facet_nested(Class~., scales = "free_y", space = "free")
    p = p + ggplot2::scale_shape_manual(values = c("Up" = 18, "Down" = 20)) +
        ggplot2::ggtitle(title) +
        ggplot2::ylab("") +  ggplot2::labs(size="adj.P.Val") +
        ggplot2::scale_color_gradient2(low=color_low, mid="snow2", high=color_high, midpoint=0) +
        # ggplot2::scale_color_gradient2(low="blue", mid="snow2", high="red", midpoint=0) +
        ggplot2::scale_size(trans="reverse") +
        ggplot2::theme_bw(base_size = 15) + 
        ggplot2::scale_y_discrete(labels = function(x) stringr::str_wrap(x, width = 1000)) +
        ggplot2::guides(alpha="none", size=ggplot2::guide_legend(override.aes=list(shape=21))) +
        ggplot2::theme(
            axis.title =  ggplot2::element_blank(),
            strip.text.y = ggplot2::element_text(size = 10, angle = 90, hjust=0.5),
            axis.text.x = ggplot2::element_text(angle = 90, hjust=1, vjust=0)) +
        ggplot2::labs(caption=paste0("adj.P.ValCutoff = ",adj.P.Val.cutoff,", pAdjustMethod = BH"))

    if (lowfigure) {
      p = p + ggplot2::theme(
              axis.title = ggplot2::element_blank(),
              legend.position = "bottom",
              legend.box = "vertical",         # Stack legends vertically
              legend.box.just = "left",        # Center the whole legend box
              legend.text = ggplot2::element_text(size = 7),
              legend.title = ggplot2::element_text(size = 11),
              legend.spacing.y = ggplot2::unit(0, "cm") # Reduce spacing between legend rows
            ) +
            ggplot2::guides(
              fill  = ggplot2::guide_legend(ncol=1, title.position = "top", byrow = F),
              shape = ggplot2::guide_legend(title.position = "left"),
              size  = ggplot2::guide_legend(title.position = "left"))
    }
    if (is.null(h)|is.null(w)) {
        h = ((unique(rlst_long$gs_name) %>% length())/6) + 3
        w = (nchar(unique(rlst_long$gs_name)) %>% max())/11 + 3.5
    }
    pdf(paste0(OUTDIRV,"/Dotplot_all_adj.P.ValCutoff",adj.P.Val.cutoff,"_",suffix,".pdf"),w=w,h=h)
        show(p)
    dev.off()
}


```

```{r}
# BiocManager::install("TCGAbiolinks")
library(TCGAbiolinks)
library(SummarizedExperiment)
library(dplyr)
library(dplyr)
library(ggplot2)
```

# Download data

```{r}
# # Check how many TCGA cancer datasets do we have
# projects = getGDCprojects()
# tcga_projects = projects[grep("^TCGA-", projects$project_id), ]
# tcga_projects %>% openxlsx::write.xlsx("../data/TCGA/TCGA_projects.xlsx")
# cancer_types = tcga_projects$tumor
# cancer_types
# #  [1] "BRCA" "UCEC" "LAML" "CHOL" "ACC"  "DLBC" "CESC" "ESCA" "KICH" "HNSC"
# # [11] "COAD" "BLCA" "KIRP" "GBM"  "MESO" "SARC" "PCPG" "READ" "LUAD" "LUSC"
# # [21] "PAAD" "LIHC" "KIRC" "LGG"  "SKCM" "PRAD" "STAD" "OV"   "THYM" "UCS" 
# # [31] "UVM"  "TGCT" "THCA"

# # Loop through each cancer type
# for (cancer in setdiff(cancer_types,c("BLCA", "KIRP", "LGG"))) {
#     cat("Downloading data for:", cancer, "\n")
#     cancer_path = paste0("../data/TCGA/", cancer,"/")
#     if (!dir.exists(cancer_path)) {
#         # Query
#         query = GDCquery(project = paste0("TCGA-", cancer),
#                         data.category = "Transcriptome Profiling",
#                         data.type = "Gene Expression Quantification",
#                         workflow.type = "STAR - Counts",
#                         sample.type = c("Primary Tumor", "Solid Tissue Normal"))
#         if (!is.null(query)) {
#             results = getResults(query)           
#             results %>% openxlsx::write.xlsx(.,paste0("Meta_", cancer,".xlsx"))
#             clinical = GDCquery_clinic(project = paste0("TCGA-", cancer), type = "clinical") 
#             clinical[,sapply(clinical, function(col) is.atomic(col) && !is.list(col))] %>%
#                 openxlsx::write.xlsx(., paste0("Clinical_", cancer,".xlsx"))
            
#             # Download
#             GDCdownload(query, files.per.chunk = 100, method = "api")

#             # Move the output folder and result to cancer_path
#             cancer_path %>% dir.create(., recursive=T, showWarnings=F)
#             from_files = list.files("./", pattern="*.tar.gz|*.txt|*xlsx", full.names = TRUE)
#             lapply(from_files, function(from) {
#                 file.rename(from = from,
#                         to   = paste0(cancer_path, "/", from))
#             })
#             file.rename(from = "GDCdata", to   = paste0(cancer_path, "/GDCdata"))
#         }
#     }
# }

# # TCGA-BLCA problematic

```

```{r}
# library(TCGAbiolinks)
# library(SummarizedExperiment)

# # Check how many TCGA cancer datasets do we have
# projects = getGDCprojects()
# tcga_projects = projects[grep("^TCGA-", projects$project_id), ]
# tcga_projects %>% openxlsx::write.xlsx("../data/TCGA/TCGA_projects.xlsx")
# cancer_types = tcga_projects$tumor
# cancer_types
# #  [1] "BRCA" "UCEC" "LAML" "CHOL" "ACC"  "DLBC" "CESC" "ESCA" "KICH" "HNSC"
# # [11] "COAD" "BLCA" "KIRP" "GBM"  "MESO" "SARC" "PCPG" "READ" "LUAD" "LUSC"
# # [21] "PAAD" "LIHC" "KIRC" "LGG"  "SKCM" "PRAD" "STAD" "OV"   "THYM" "UCS" 
# # [31] "UVM"  "TGCT" "THCA"

# # # Loop through each cancer type
# for (cancer in setdiff(cancer_types,c("BLCA", "KIRP", "LGG"))) {
#     query = GDCquery(project = paste0("TCGA-", cancer),
#                     data.category = "Transcriptome Profiling",
#                     data.type = "Gene Expression Quantification",
#                     workflow.type = "STAR - Counts",
#                     sample.type = c("Primary Tumor", "Solid Tissue Normal"))
#     # Download and prepare
#     # GDCdownload(query)
#     data = GDCprepare(query)
#     # Extract count matrix and metadata
#     counts = assay(data)
#     metadata = colData(data)
#     saveRDS(data, paste0("../data/TCGA/", cancer, "/TCGA_UCEC.RDS"))
#     saveRDS(counts,paste0("../data/TCGA/", cancer, "/counts.RDS"))
#     saveRDS(metadata, paste0("../data/TCGA/", cancer, "/metadat.RDS"))
# }

```

# Data checking

```{r}
outdir = "../results/TCGA"
dir.create(outdir, recursive=T, showWarnings=F)
```

## Only take paired data
```{r}
INDIR = "../data/TCGA/"
ctypes = list.dirs(INDIR, recursive=FALSE, full.name=F)
ctypes
paired_data = lapply(ctypes, function(ctype) {
    meta = openxlsx::read.xlsx(paste0(INDIR, ctype, "/Meta_", ctype, ".xlsx")) %>% 
        full_join(
            openxlsx::read.xlsx(paste0(INDIR, ctype, "/Clinical_", ctype, ".xlsx")) %>%
            select(-created_datetime),
        by=c("cases.submitter_id" = "submitter_id"))  # Here we have patient IDs
    cases.submitter_id2keep = 
        intersect(meta$cases.submitter_id[meta$sample_type=="Primary Tumor"],
        meta$cases.submitter_id[meta$sample_type=="Solid Tissue Normal"]) %>%
        na.omit()
    if (length(cases.submitter_id2keep)>3) {
        return(meta %>% filter(cases.submitter_id %in% cases.submitter_id2keep))
    } else {return(NA)}
}) %>% setNames(ctypes) %>% .[!is.na(.)]

lapply(names(paired_data), function(c) {
    temp = paired_data[[c]]
    temp = temp[temp$cases.submitter_id %in% 
        (which(table(temp$cases.submitter_id)>2) %>% names()),]
    as.data.frame(temp) %>% count(sample_type, cases.submitter_id)
}) %>% setNames(names(paired_data))

print("If there are more than one tumor data file per patient, pick the first one")
paired_data = lapply(paired_data, function(cancer) {
    cancer[!duplicated(paste0(cancer$sample_type,cancer$cases.submitter_id)),]
})
# Check 
lapply(names(paired_data), function(c) {
    temp = paired_data[[c]]
    temp = temp[temp$cases.submitter_id %in% 
        (which(table(temp$cases.submitter_id)>2) %>% names()),]
    as.data.frame(temp) %>% count(sample_type, cases.submitter_id)
}) %>% setNames(names(paired_data))

```

## Circular barplot to present sample size

```{r}
sampleSizeCircularBarplot = function(paired_data, OUTDIR) {
    sample_sizes = lapply(names(paired_data), function(cancer) {
        dat = paired_data[[cancer]]
        count = dat %>% select(sample_type, sample.submitter_id) %>% 
            unique() %>% group_by(sample_type) %>% 
            summarise(Count=n()) %>%
            mutate(sample_type=gsub("Primary Tumor", "Tumor", sample_type)) %>%
            mutate(sample_type=gsub("Solid Tissue Normal", "Normal", sample_type)) %>%
            mutate(Cancer_type=cancer)
        count = count %>% left_join(
                count %>% group_by(Cancer_type) %>%
                summarise(total=sum(Count))) %>%
                mutate(Cancer_type = paste0(Cancer_type, "\n(", total, ")"))
    }) %>% Reduce(rbind, .)

    TCGA_summary = sample_sizes %>%
    ggplot(., aes(x=Cancer_type, y=Count, fill=sample_type)) +
    geom_bar(stat="identity", show.legend=FALSE) +
    theme_minimal() +
    theme(axis.text.y = element_blank(),
            axis.ticks = element_blank(),
            axis.title = element_blank(),
            plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")) + 
    coord_polar() +
    ggtitle("TCGA") 
    pdf(paste0(outdir, "/Summary/Sample_size_circular_barplot.pdf"), 
        w=9, h=4.5)
        print(cowplot::plot_grid(TCGA_summary, TCGA_summary))
    dev.off()
}

paste0(outdir, "/Summary/") %>% dir.create(.,recursive=T, showWarnings=F)
sampleSizeCircularBarplot(
    paired_data, 
    OUTDIR=paste0(outdir, "/Summary/"))

```

## Import raw data overview

```{r}
# Import the data:
paired_data_count = lapply(names(paired_data), function(cancer) {
    folder = paste0("../data/TCGA/", cancer, "/GDCdata/TCGA-", 
        cancer, "/Transcriptome_Profiling/Gene_Expression_Quantification/")
    p = paired_data[[cancer]]
    files = paste0(folder,"/", p$file_id, "/", p$file_name)
    Samples = p$sample.submitter_id
    count_table = lapply(1:length(files), function(i) {
        file = files[[i]]
        expr = read.delim(file, skip=1, header=TRUE)
        expr$gene_id = sub("\\..*$", "", expr$gene_id)
        expr %>% as.data.frame() %>% 
            select(gene_id, gene_name, gene_type, unstranded) %>%
            setNames(c("gene_id", "gene_name", "gene_type", Samples[[i]]))
    })

    count_table

}) %>% setNames(names(paired_data))

# Combine all samples to the same table for each cancer type
paired_data_count = lapply(paired_data_count, function(cancer){
    temp2 = lapply(1:length(cancer), function(i) {
        sample = cancer[[i]]
        sample = sample[!duplicated(sample$gene_name), ]
        rownames(sample) = NULL
        temp = sample %>% as.data.frame() %>% 
            tibble::column_to_rownames(var="gene_name") 
        temp = temp[, 3, drop=FALSE]
        temp = temp[!duplicated(paired_data_count[[1]][[1]]$gene_name), ,drop=FALSE]
        if (i==1) {
            temp = temp %>% as.data.frame() %>% 
                tibble::rownames_to_column(var="gene_name")
        } else {rownames(temp)=NULL}
        temp
    }) %>% Reduce(cbind, .)
    temp2[5:nrow(temp2), ]
})

names(paired_data_count)
head(paired_data[[1]])

```

# Import KEGG metabolic pathways

```{r}
# Get KEGG metabolic pathways
paste0(outdir, "/KEGG_metabolism/") %>% 
    dir.create(., recursive=TRUE, showWarnings=FALSE)
kegg_metab_dbs = 
    getdb_metabolism(databaseDIR=paste0(outdir, "/KEGG_metabolism/"))
kegg_metab_db = kegg_metab_dbs$kegg_metab_db

# Make a circular packing plot to show the number of pathways in each class
getCirPackingPlot(
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table)

```


# Paired differential analysis on sample GSVA scores

```{r}
GSVA_limma_rslt = lapply(names(paired_data_count), function(cancer) {
    print(cancer)
    dat = paired_data_count[[cancer]] 
    rownames(dat) = NULL
    dat = dat %>% as.data.frame() %>%
        tibble::column_to_rownames(var="gene_name") %>% na.omit()
    meta = paired_data[[cancer]] %>% as.data.frame() %>%
        mutate(Sample = sample.submitter_id) %>%
        mutate(Group = sample_type) %>%
        mutate(Patient = cases.submitter_id)
    meta$Group = factor(meta$Group, levels=c("Solid Tissue Normal", "Primary Tumor"))
    paste0(outdir, "/Limma_GSVA/", cancer, "/") %>%
        dir.create(., recursive=TRUE, showWarnings=FALSE)
    GSVAlimmaTest(dat, paired=TRUE, Feature.filtering=TRUE,
        currentCovariate=NULL, meta=meta, paired_variable="Patient", 
        kegg_metab_db=kegg_metab_db, 
        OUTDIR=paste0(outdir, "/Limma_GSVA/", cancer, "/"))
}) %>% setNames(names(paired_data_count))

saveRDS(GSVA_limma_rslt, 
    paste0(outdir, "/Limma_GSVA/GSVAlimma_rslt_all.RDS"))
```

## Visualize differential results

```{r}
library(dplyr)
library(stringr)
rlst_list = lapply(GSVA_limma_rslt, function(l) l$rslt_of_interest)
makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA", nested=T, 
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())
makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA_tops", nested=T, top=TRUE,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique())


makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA2", nested=T, 
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique(),
        color_low="blue", color_high="red")
makeLimmaDotplot_TCGA(rlst_list, adj.P.Val.cutoff=0.05, w=NULL, h=NULL, 
        OUTDIRV = paste0(outdir, "/Limma_GSVA/"), Depth="gs_name", 
        lowfigure=F, title = "Dysregulated pathways_TCGA", 
        suffix="TCGA_tops2", nested=T, top=TRUE,
        taskHierarchy=kegg_metab_dbs$kegg_metab_db_table %>% 
                            select(gs_name, class) %>% unique(),
        color_low="blue", color_high="red")
```


## Visualize pathway activity GSVA scores

### Per sample

Here, we will normalize all cancer type together and then calculate GSVA scores together to make it relatively more comparable across cancer types
```{r}
# Run GSVA all together 
paste0(outdir, "/Limma_GSVA/") %>% 
    dir.create(., recursive=TRUE, showWarnings=FALSE)
dat = paired_data_count %>% Reduce(function(x,y) full_join(x,y,by="gene_name"), .)
rownames(dat) = NULL
dat = dat %>% as.data.frame() %>%
    tibble::column_to_rownames(var="gene_name") %>% na.omit()
meta = lapply(names(paired_data), function(c) {
        paired_data[[c]] %>% as.data.frame() %>% 
            mutate(cancer_type=c) %>% 
            select(cancer_type, sample.submitter_id, sample_type, cases.submitter_id)
    }) %>% Reduce(rbind, .) %>% 
    mutate(Sample = sample.submitter_id) %>%
    mutate(Group = sample_type) %>%
    mutate(Patient = cases.submitter_id)
meta$Group = factor(meta$Group, levels=c("Solid Tissue Normal", "Primary Tumor"))
GSVA_limma_rslt_all = GSVAlimmaTest(dat, paired=TRUE, Feature.filtering=TRUE,
        currentCovariate=NULL, meta=meta, paired_variable="Patient", 
        kegg_metab_db=kegg_metab_db, 
        OUTDIR=paste0(outdir, "/Limma_GSVA/"))
GSVA_limma_rslt_gsva_all = lapply(names(paired_data), function(c) {
    GSVA_limma_rslt_all$gsva[,paired_data[[c]]$sample.submitter_id]
}) %>% setNames(names(paired_data))

visualizeGSVSscoresGroup(
    GSVA_limma_rslt_gsva = GSVA_limma_rslt_gsva_all, 
    paired_data = paired_data, 
    OUTDIR = paste0(outdir, "/Limma_GSVA/")
)

```

### Per metabolic class

```{r}
visualizeGSVSscoresClass(GSVA_limma_rslt, 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table, 
    paired_data, 
    OUTDIR = paste0(outdir, "/Limma_GSVA/"))
```

##  Visualize pathway sizes
```{r}
genes4GSVA = read.delim(
    file = paste0(outdir, "/Limma_GSVA/NormalizedCount.csv"),
    header = TRUE,
    stringsAsFactors = FALSE
    ) %>% pull(Features) %>% unique()

visualizeGSsizeClass(
    GSVA_limma_rslt_gsva=lapply(GSVA_limma_rslt, function(r) r$gsva), 
    genes4GSVA = genes4GSVA, 
    kegg_metab_db = kegg_metab_dbs$kegg_metab_db, 
    kegg_metab_db_table = kegg_metab_dbs$kegg_metab_db_table,
    OUTDIR="../results/TCGA/Limma_GSVA/")
```


## Visualize the number of significant pathways

```{r}
paste0(outdir, "/Limma_GSVA") %>% 
    dir.create(., recursive=T, showWarnings=FALSE)
visualizeSigNr(OUTDIR=paste0(outdir, "/Limma_GSVA"), 
                GSVA_limma_rslt)

```

## Correlation between normal pathway GSVA scores and differential effect size

```{r}
dir.create(paste0(outdir, 
    "/correlation_scoreVSt/"), recursive=T, showWarnings=F)
rslt_of_interest = lapply(GSVA_limma_rslt, function(g) g$rslt_of_interest)
# Here the normal sample GSVA scores from different cancer types were calculated together
getNormalVStcorPlot(
        GSVA_limma_rslt_gsva=GSVA_limma_rslt_gsva_all,  
        rslt_of_interest, paired_data, 
        OUTDIR=paste0(outdir, "/correlation_scoreVSt/"),
        sample="normal")

```

## Correlation between normal pathway GSVA scores and differential effect size

```{r}
# Here the normal sample GSVA scores from different cancer types were calculated together
getNormalVStcorPlot(
        GSVA_limma_rslt_gsva=GSVA_limma_rslt_gsva_all,  
        rslt_of_interest, paired_data, 
        OUTDIR=paste0(outdir, "/correlation_scoreVSt/"),
        sample="all")

```

## Correlation between normal pathway gene expr (median) and differential effect size

```{r}
normalizedExpr_all = read.delim(
    file = paste0(outdir, "/Limma_GSVA/NormalizedCount.csv"),
    header = TRUE,
    stringsAsFactors = FALSE
    ) %>% tibble::column_to_rownames(var="Features")
colnames(normalizedExpr_all) = gsub("[.]", "-", colnames(normalizedExpr_all))
getNormalExprVStcorPlot(normalizedExpr_all, 
    kegg_metab_db=kegg_metab_dbs$kegg_metab_db,
    GSVA_limma_rslt_gsva=lapply(GSVA_limma_rslt, function(r) r$gsva), 
    rslt_of_interest=lapply(GSVA_limma_rslt, function(r) r$rslt_of_interest), 
    paired_data, 
    OUTDIR = "../results/TCGA/Limma_GSVA/")
```

```{r}
save.image(paste0(outdir, "/TCGA.RData"))
# load(paste0(outdir, "/TCGA.RData"))
out = capture.output(sessionInfo())
writeLines(out, paste0(outdir, "/sessionInfo.txt"))

```
