---
title: Detailed mapping of metabolism in plasma and skin in psoriasis
subtitle: "NBIS support {{< 7546 >}}"
author: Yuan Li
toc: true
number-sections: true
format: 
  html:
    #embed-resources: true #This works from command line not from Rstudio
    code-fold: true
    code-tools: true
    code-link: true
    theme: pandoc
    df-print: paged
    html-math-method: katex #use katex for math text
  pdf:
    code-line-numbers: true
    geometry: 
      - top=30mm
      - left=30mm
  docx: default
language:
  code-summary: Click to show code
editor_options:
  chunk_output_type: inline
editor: visual
---

```{r Setup, echo = FALSE}
knitr::opts_chunk$set(fig.width  = 7,
                      results    = "hold",
                      message    = FALSE,
                      cache.lazy = FALSE,
                      warning    = FALSE)

```

```{r,echo=FALSE,include=FALSE}
# project variables
rep_nbis_id <- "7546"
rep_report_version <- "1.0"
rep_request <- "Monica Torres"
rep_request_email <- "monica.barbara.torres@ki.se"
rep_pi <- "Jakob Wikström"
rep_pi_email <- "jakob.wikstrom@ki.se"
rep_org <- "KI"
rep_nbis <- "Yuan Li"
rep_nbis_email <- "yuan.li@nbis.se"
```

::: boxy
**NBIS ID:** `r rep_nbis_id`\
**Report Version:** `r rep_report_version`\
**Request by:** `r paste0(rep_request," (",rep_request_email,")")`\
**Principal Investigator:** `r paste0(rep_pi," (",rep_pi_email,")")`\
**Organisation:** `r rep_org`\
**NBIS Staff:** `r paste0(rep_nbis," (",rep_nbis_email,")")` **conda env:** TTN
:::

<br>

# Support Request

[7546](https://projects.nbis.se/issues/7546) Psoriasis is associated with an increased risk of systemic metabolic disorders, including diabetes, dyslipidemia, and obesity. The metabolic processes within psoriasis plaques, however, remain insufficiently understood. This project aims to elucidate the skin metabolism in psoriasis using comprehensive omics approaches. We plan to collect plasma, and both lesional and non-lesional skin biopsies from 50 psoriasis patients and 50 matched healthy controls (for age and sex), recruited via a social media campaign. These biopsies will be snap-frozen and analyzed using non-targeted mass spectrometry-based metabolomics and phosphoproteomics, while plasma samples will undergo non-targeted metabolomics. This extensive research is expected to yield a wealth of novel data, revealing and integrating dysregulated metabolites and phosphorylated proteins. Although we hypothesize that psoriatic lesional skin exhibits altered metabolism, characterized by potentially reduced mitochondrial activity and enhanced glycolytic processes, we anticipate uncovering unexpected findings that may pave the way for new research avenues. Moreover, by making the data publicly available, we aim to foster further exploration in this field by other researchers. In the long term, our project hopes to identify novel metabolic drug targets for psoriasis treatment.

Research questions and hypotheses:

Hypothesis: We hypothesize that, due to a hyperproliferative state in the epidermis, the skin in psoriasis has an altered metabolism, namely higher activity of glycolytic pathways in detriment of mitochondrial activity. However, as this is an explorative project, we expect to unravel other altered metabolic pathways that can lead to a better understanding of the underlying pathophysiology. Aim 1. Recruitment of study participants and harvesting of samples (already performed) Through a social media campaign, we recruited 50 patients with psoriasis and 50 healthy-matched volunteers (for sex and age). Two lesional and two non-lesional punch biopsies were harvested from psoriasis patients, and from healthy volunteers two non-lesional biopsies from the same anatomical location. For a systemic comparison, blood plasma was harvested. Samples were snap-frozen in liquid nitrogen and stored in a biobank freezer. This effort, in a short time, generated a very large biobank consisting of 300 skin samples and 100 plasma samples. The samples were then sent in one batch to BGI Genomics for further analysis, as described below. Aim 2. Untargeted metabolomics and phosphoproteomics analysis of skin biopsies and plasma (ongoing analysis) The samples collected in Aim 1 were examined by untargeted mass spectrometry-based metabolomics and phosphoproteomics. Raw data will undergo bioinformatic analysis to identify **differentially expressed metabolites and phosphoproteins as well as identify key bioenergetic pathways altered in psoriasis**. Comparison between skin and plasma will be performed. Advanced multi-omics integration of the metabolomics and phosphoproteomics data will also be made for a wider understanding of the altered metabolism. This analysis will generate a large unique dataset that will give novel insights into psoriasis skin metabolism.

Help needed:

Parts of the differential expression analysis in metabolomics and phosphoproteomics have already been provided by BGI Genomics, however, it would be important for us to optimize the analysis (enrichment analysis, motif analysis, etc) from the initial table with intensities, graphical representation and perform a multi-omic integration of the data, namely of plasma and skin metabolomics and phosphoproteomics, and with publicly available datasets. Lastly, we would like to do the analysis in a paired way (since every patient has a matched controll for age and sex), include clinical data in the analysis (such as BMI, smoking, inflammatory parammeters, etc.) and upload the data to public platforms before publication. There is a plan for parts of these tasks, but we would need help with the code and statistical analysis, as well as advice on which platforms to use.

conda env: "SMS7546"

In this tutorial, we will perform [**CellFie**](https://www.cell.com/cell-reports-methods/fulltext/S2667-2375(21)00085-0) analysis to look at which metabolic tasks have been turned on or off

# Prepare input phosphoproteomics data

The phosphoproteomics data has been preprocessed by the company (filtered and normalized), further filtered by me, and have missing values imputed.

```{r}
library(dplyr)
```

```{r}
Pdat = 
  readxl::read_xlsx("../results/PhosphoProteomics/Phosphoprotein/rawDataProcessing/NAimputation/Phosphoprotein_filtered_normalized_log_imputed.xlsx")
head(Pdat)
grep("HS", colnames(Pdat), value=T) %>% length()
grep("PL", colnames(Pdat), value=T) %>% length()
grep("PNL", colnames(Pdat), value=T) %>% length()
# We will need to convert the protein uniprot ID to ENTREZ ID
# Pdat2 = AnnotationDbi::select(
#   org.Hs.eg.db,
#   keys = Pdat$Protein,
#   columns = c("SYMBOL"),
#   keytype = "UNIPROT"
# ) %>% na.omit() %>%
#   .[!duplicated(.$UNIPROT),] %>%
#   setNames(c("Protein", "SYMBOL")) %>%
#   right_join(Pdat) %>%
#   na.omit() %>%
#   dplyr::select(-Protein)
# names(Pdat2)[1] = "genes"
# 
# OUTDIR = "../results/CellFie/"
# INDIR = paste0(OUTDIR, "data/")
# dir.create(INDIR, recursive = T, showWarnings = F)
# # Save the input data to a csv file
# write.table(Pdat2, paste0(INDIR, "PsoPproteomics.csv"), sep="\t", 
#             quote=F, row.names=F)

# We will need to convert the protein uniprot ID to ENTREZ ID
library(AnnotationDbi)
library(org.Hs.eg.db)
library(dplyr)
Pdat2 = AnnotationDbi::select(
  org.Hs.eg.db,
  keys = Pdat$Protein,
  columns = c("ENTREZID"),
  keytype = "UNIPROT"
) %>% na.omit() %>%
  .[!duplicated(.$UNIPROT),] %>%
  setNames(c("Protein", "ENTREZID")) %>%
  right_join(Pdat) %>%
  na.omit() %>%
  dplyr::select(-Protein)
names(Pdat2)[1] = "genes"

dim(Pdat2)
# [1] 1975  149
OUTDIR = "../results/CellFie/"
INDIR = paste0(OUTDIR, "data/")
dir.create(INDIR, recursive = T, showWarnings = F)
# Save the input data to a csv file
write.table(Pdat2, paste0(INDIR, "PsoPproteomics.csv"), sep=",", 
            quote=F, row.names=F)
```

> I then sumbit the input file to the [genePatten webserver](https://cloud.genepattern.org/gp/pages/index.jsf), and run it there

-   Data="PsoPproteomics.csv",
-   Sample.Number="148",
-   Reference.Model= Human_iHsa \# "human.1",
-   Thresholding.Approach="local",
-   Percentile.Or.Value="percentile",
-   Local.Threshold.Type="minmaxmean",
-   Local.Lower.Bound="25",
-   Local.Upper.Bound="75"
-   Uses gene-specific thresholds, customized per gene (based on its expression distribution)
-   For each gene, CellFie computes its 25th and 75th percentile and considers it ON if expression \> mean of \[25th, 75th\]

# Process CellFie output

```{r}
cf = read.csv2(paste0(OUTDIR, "/CellFieOut/detailScoring.csv"), 
               sep=",", stringsAsFactors = FALSE) 
  
head(cf)
# Re(cf$TaskScore_S11%>%as.complex())
# as.complex(cf$TaskScore_S1)[is.na(as.complex(cf$TaskScore_S1))]
# cf$TaskScore_S1[is.na(as.complex(cf$TaskScore_S1))]
# as.complex("1.43070924464163+-1.52372968707745e-16i")
# as.complex("-1+0i")

```

Convert complex number to only keep the real part

```{r}
# # table(cf$SampleID_S1)
# # table(cf$TaskID_S1)
# table(cf$TaskScore_S1)
# table(cf$BinaryTaskScore_S1)
# # table(cf$EssentialRxnsTask_S1)
# table(cf$ExpressionScoreEssentialRxnsTask_S1)
# # table(cf$GeneAssociatedToEssentialRxnsTask_S1)

complexCol = grep("TaskScore_|ExpressionScoreEssentialRxnsTask_", colnames(cf))
for (i in complexCol) {
  cf[, i] = gsub("\\+\\-", "-", cf[, i])
  # x_fixed <- gsub("\\-\\+", "-", x_fixed)
  # x_fixed <- gsub("\\+\\+", "+", x_fixed)
  # x_fixed <- gsub("\\-\\-", "+", x_fixed)
  cf[, i] = Re(as.complex(cf[, i]))
}

# #cf$TaskScore_S1[is.na(as.complex(cf$TaskScore_S1))]
# # table(cf$SampleID_S1)
# # table(cf$TaskID_S1)
# table(cf$TaskScore_S1)
# table(cf$BinaryTaskScore_S1)
# # table(cf$EssentialRxnsTask_S1)
# table(cf$ExpressionScoreEssentialRxnsTask_S1)
# # table(cf$GeneAssociatedToEssentialRxnsTask_S1)

head(cf)
```

## Import taskInfo.csv

```{r}
taskInfo = read.csv2(paste0(OUTDIR, "/CellFieOut/taskInfo.csv"), 
                     sep=",", stringsAsFactors = FALSE) %>%
                     setNames(c("Depth3.ID", "Depth3", "Depth1", "Depth2"))
head(taskInfo)
```

Reshuffle the table to have tasks in the rows and samples in the columns

```{r}
Samples = gsub(".*_", "", colnames(cf)) %>% unique()
cf2 =lapply(Samples, function(S) {
  temp = cf[, grep(paste0(S, "$"), colnames(cf))]
  colnames(temp) = gsub("_.*", "", colnames(temp))
  temp
}) %>% Reduce(rbind, .) %>% 
  mutate(Depth3.ID = TaskID) %>%
  dplyr::select(Depth3.ID, everything()) %>% 
  filter(TaskScore!=-1) # remove unscorable rows (-1)

# Add in task depth information
cf3 = right_join(taskInfo, cf2)
head(cf3)

# Add in Sample information
SampleInfo = data.frame(Sample = colnames(Pdat2)[-1],
             SampleID = 1:length(colnames(Pdat2)[-1]))
cf4 = right_join(SampleInfo, cf3) %>%
      mutate(Group=gsub("[-].*", "", Sample)) %>%
      dplyr::select(Group, everything()) 

write.table(cf4, paste0(OUTDIR, "/CellFieOut/detailScoring_new.csv"), sep=",", 
            quote=F, row.names=F)

cf5 = dplyr::select(cf4, -EssentialRxnsTask, -ExpressionScoreEssentialRxnsTask,
            -GeneAssociatedToEssentialRxnsTask, -GeneExpressionValue) %>%
            unique()

```

## Sum up task scores to higher level categories

### Task Depth1

```{r, fig.height=2, fig.width=6}
library(dplyr)
Depth1= cf5 %>%
  group_by(Sample, Depth1) %>%
  summarise(
    MeanTaskScore_Depth1 = mean(TaskScore, na.rm = TRUE),
    SumTaskScore_Depth1 = sum(TaskScore, na.rm = TRUE),
    FractionOfActiveTasks_Depth1 = mean(BinaryTaskScore==1, na.rm = TRUE)
  ) %>% mutate(Group=gsub("[-].*", "", Sample)) %>%
  #mutate(Depth1 = gsub("METABOLISM", "M.", Depth1)) %>%
  dplyr::select(Group, everything())

library(ggplot2)
margin(1, 1, 1, 1, unit = "cm")
P = ggplot(Depth1, aes(x = Depth1, y = MeanTaskScore_Depth1, fill = Sample)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ Group) +
  coord_flip()  +
  xlab("") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1))
P

dir.create(paste0(OUTDIR, "/Summary/"), recursive = T, showWarnings = F) 
pdf(paste0(OUTDIR, "Summary/TaskScores_Depth1.pdf"), h=2, w=6)
P
dev.off()
rm(P)
```

> Glycan metabolism: The specific subset of carbohydrate metabolism dealing with the biosynthesis, modification, and degradation of glycans --- the complex sugar chains attached to proteins and lipids.

**Glycan** plays key roles in:

-   Cell--cell communication
-   Protein folding
-   Immune recognition
-   Cell adhesion and signaling

> Carbohydrates metabolism: The broad set of biochemical pathways involved in the breakdown, synthesis, and utilization of carbohydrates (sugars).

### Task depth2

```{r, fig.height=9, fig.width=8}
library(dplyr)
Depth2 = cf5 %>%
  group_by(Sample, Depth2) %>%
  summarise(
    MeanTaskScore_Depth2 = mean(TaskScore, na.rm = TRUE),
    SumTaskScore_Depth2 = sum(TaskScore, na.rm = TRUE),
    FractionOfActiveTasks_Depth2 = mean(BinaryTaskScore==1, na.rm = TRUE)
  ) %>% mutate(Group=gsub("[-].*", "", Sample)) %>%
  # mutate(Depth2 = gsub("METABOLISM", "M.", Depth2)) %>%
  dplyr::select(Group, everything()) 
Depth2$Depth2 = factor(Depth2$Depth2, 
                            levels=unique(cf5$Depth2))

library(ggplot2)
margin(1, 1, 1, 1, unit = "cm")
P = ggplot(Depth2, aes(x = Depth2, y = MeanTaskScore_Depth2, fill = Sample)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ Group) +
  coord_flip()  +
  xlab("") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 0, hjust = 1))
P

dir.create(paste0(OUTDIR, "/Summary/"), recursive = T, showWarnings = F) 
pdf(paste0(OUTDIR, "Summary/TaskScores_Depth2.pdf"), h=9, w=8)
P
dev.off()
rm(P)

```

## Summarize BinaryTaskScore to higher-leve categories

### Task depth 1

```{r, fig.height=2.5, fig.width=7}
library(dplyr)
cf5_summary = cf5 %>%
  group_by(Group, Depth1) %>%
  summarise(Active=sum(BinaryTaskScore==1,na.rm=TRUE)) 

library(ggplot2)
P = ggplot(cf5_summary, aes(x=reorder(Depth1, Active), y=Active)) +
    geom_point(shape=21, aes(size=abs(Active), fill=Group)) +
    # facet_wrap(~ Group) +
    #geom_hline(yintercept = 0, linetype = "dashed", color = "snow3") +
    coord_flip()  +
    xlab("") + 
    ggtitle("Active task count") + 
    theme_bw(base_size = 12) #+ theme(legend.position = "none")
P
dir.create(paste0(OUTDIR, "/Summary/"), recursive = T, showWarnings = F) 
pdf(paste0(OUTDIR, "Summary/BinaryTaskScores_Depth1.pdf"), h=2.5, w=7)
P
dev.off()

rm(cf5_summary, P)

```

### Task depth 2

```{r, fig.height=9, fig.width=9}
library(dplyr)
cf5_summary = cf5 %>%
  group_by(Group, Depth2) %>%
  summarise(Active=sum(BinaryTaskScore==1,na.rm=TRUE)) 
cf5_summary$Depth2 = factor(cf5_summary$Depth2,
                            levels=unique(cf5$Depth2))

library(ggplot2)
P = ggplot(cf5_summary, aes(x=reorder(Depth2, Active), y=Active)) +
    geom_point(shape=21, aes(size=abs(Active), fill=Group)) +
    # facet_wrap(~ Group) +
    #geom_hline(yintercept = 0, linetype = "dashed", color = "snow3") +
    coord_flip()  +
    xlab("") + 
    ggtitle("Active task count_Depth2") + 
    theme_bw(base_size = 12) #+ theme(legend.position = "none")
P
dir.create(paste0(OUTDIR, "/Summary/"), recursive = T, showWarnings = F) 
pdf(paste0(OUTDIR, "Summary/BinaryTaskScores_Depth2.pdf"), h=9, w=9)
P
dev.off()

rm(cf5_summary, P)

```

# Limma differential analysis of Task scores between groups

Sort out meta data

```{r}
# Process meta data
library(readxl)
path_meta_all = "../data/250304 Meta data psoriasis study_updated.xlsx"
patient = invisible(read_excel(path_meta_all,
          sheet=1, col_names=T, na="", skip=0)) %>%
  setNames(c("SampleO", colnames(.)[2:ncol(.)])) %>%
  setNames(colnames(.) %>% gsub(" ", "_", .)) %>%
  setNames(colnames(.) %>% gsub(",|[(]|[)]", "", .))
patient$Smoking = gsub(" [(].*[)]$", "",  patient$Smoking)
patient$Exercise_num = factor(patient$Exercise, 
                levels=c("0", "1-2 per week", "3 or more per week")) %>%
                as.numeric()
patientL = dplyr::select(patient, -Skin_sample_non_lesional) %>%
  dplyr::rename(., c("Skin_sample"="Skin_sample_lesional")) %>%
  mutate(Sample = paste0(patient$SampleO, ".PL.M"))
patientNL = dplyr::select(patient, -Skin_sample_lesional) %>%
  dplyr::rename(., c("Skin_sample"="Skin_sample_non_lesional")) %>%
  mutate(Sample = paste0(patient$SampleO, ".PNL.M"))
healthControl = invisible(read_excel(path_meta_all,
                        sheet=2, col_names=T, na="", skip=0)) %>%
  setNames(c("SampleO", colnames(.)[2:ncol(.)])) %>%
  setNames(colnames(.) %>% gsub(" ", "_", .)) %>%
  setNames(colnames(.) %>% gsub(",|[(]|[)]", "", .)) 
healthControl$Sample = paste0(healthControl$SampleO, ".HS.M")
healthControl$Smoking = gsub(" [(].*[)]$", "", healthControl$Smoking)
healthControl$Exercise_num = factor(healthControl$Exercise, 
                levels=c("0", "1-2 per week", "3 or more per week")) %>%
                as.numeric()
meta_extra = Reduce(full_join, list(patientL, patientNL, healthControl))

temp = meta_extra$Sample %>% gsub(".M", "", .) %>% gsub("HC", "P", .) %>% 
  stringr::str_split_fixed(., "[.]", 2) 
temp = paste0(temp[,2], "-", temp[,1]) %>%
  sub("([A-Za-z])([0-9])", "\\1_\\2", .)
meta_extra = meta_extra %>% mutate(Samples=temp) %>% 
  dplyr::select(Samples, everything())
meta_extra$Alcohol = gsub(" ", "_", meta_extra$Alcohol) %>% 
  gsub("-", "_", .) %>% as.factor()
meta_extra$Exercise = gsub(" ", "_", meta_extra$Exercise) %>% 
  gsub("-", "_", .) %>% as.factor()
meta_extra$Group = sub("-.*", "", meta_extra$Samples) %>% 
  as.character() %>% factor(., levels=unique(.))
meta_extra$Patient = sub("[.].*", "", meta_extra$Sample) %>% 
  as.character() %>% factor(., levels=unique(.))
meta_extra$Heredity_for_psoriasis = as.factor(meta_extra$Heredity_for_psoriasis)
meta_extra$Sex = as.factor(meta_extra$Sex)
meta_extra$Alergies_asthma = as.factor(meta_extra$Alergies_asthma)
meta_extra$Smoking = 
  factor(ifelse(meta_extra$Smoking=="anao", "No", meta_extra$Smoking),
         levels = c("No", "Yes"))
meta_extra$Date_of_collection = 
  as.numeric(meta_extra$Date_of_collection) 
# this will be Unix timestamps, representing the number of seconds since Jan 1, 1970
meta_extra$Date_of_collection = 
  (meta_extra$Date_of_collection-min(meta_extra$Date_of_collection))/(60*60*24) 
# Convert Unix timestamps to time differences in days
meta_extra$Anatomical_category = as.factor(meta_extra$Anatomical_category)
meta_extra$Disease_category2 = 
  gsub("No diseases", "NoDiseases", meta_extra$Disease_category) %>%
  gsub(", ", "_", .) %>% gsub("/", "OR", .) %>% as.factor()
meta_extra = as.data.frame(meta_extra)
rownames(meta_extra) = meta_extra$Samples

Pairs = invisible(read_excel(path_meta_all,
  sheet=3, col_names=T, na="", skip=0)) %>%
  setNames(c("Sample", "Additional", "Anatomical_location_of_samples")) %>%
  dplyr::select(Sample, Additional, Anatomical_location_of_samples)
Pairs$Match = paste0("M", 1:nrow(Pairs))
Mapping_match = c(setNames(Pairs$Match, Pairs$Sample), 
                  setNames(Pairs$Match, Pairs$Additional))

meta_extra$Match = Mapping_match[meta_extra$Patient]
head(meta_extra)

```

Set up the function

```{r}
invisible(capture.output({library(limma)}))
invisible(capture.output({library(tibble)}))

#' @param dat the dataset needs to have features as rownames
#' @param meta meta data needs to have at least two columns named "Samples", "Group", for the moment, we only support Group with two categories
#'
limmaTest = function(dat, paired=TRUE, currentCovariate=NULL, 
                     checkCovariate=FALSE, meta=meta,
                     paired_variable="Patient") {
  meta = filter(meta, Samples%in%colnames(dat))
  # Make sure the samples are in the same order in both meta and dat:
  dat = dat[, meta$Samples]
  meta = meta %>% mutate(across(where(is.factor), droplevels))
  # print("When running differential analysis, we only keep features that have at least 2 values in each group...")
  featureskeep = apply(!is.na(dat), 1, function(row) {
      A=row
      B=sub("-.*", "", colnames(dat))
      sum(aggregate(A~B, data=data.frame(A=A,B=B), sum)[,"A"]>1)>1
  })
  dat = dat[featureskeep, ]
  # print(dim(dat))
  #if (!is.null(currentCovariate)) {currentCovariate=NULL}
  baseFormula = ~ Group
  if (is.null(currentCovariate)) {
    currentFormula = baseFormula
  } else {
    currentFormula = as.formula(paste0("~ ", currentCovariate, " + Group"))}
  
  design = model.matrix(currentFormula, data=meta) # each column is for coefficient
  group.categories = meta$Group %>% unique() %>% length()
  colnames(design) = gsub("Group", "", colnames(design))
  
  if (paired) {
    # Fit with correlated arrays
    dupcor = duplicateCorrelation(dat, design, 
                            block=meta[,paired_variable,drop=T])
    fit = lmFit(dat, design, block=meta[,paired_variable,drop=T],
                correlation=dupcor$consensus)
  } else {
    fit = lmFit(dat, design)
  }
  # fit3 = eBayes(fit, robust=TRUE)
  # print("Check whether we should include a factor in the model by checking for how many metabolites this factor is significant in the model: ")
  if (checkCovariate) {
    fit.con_co = eBayes(fit)
    rlst_co = lapply(2:(ncol(design)-group.categories+1), function(i) {
        topTable(fit.con_co, n=Inf, coef=colnames(design)[i])}) %>%
        setNames(colnames(design)[2:(ncol(design)-group.categories+1)])
    # n refers to the number of top-ranked genes to be returned.
  } else {rlst_co = NA}
  
  fit.con = eBayes(fit, robust = TRUE)
  rlst_interest = topTable(fit.con, n=Inf, coef=levels(meta$Group)[2]) %>%
      arrange(-t) 
  
  return(list(rslt_of_interest=rlst_interest, rslt_co=rlst_co, data=dat, fitted.model=fit))
}

# We have so many meta data, which ones to add?
## One method is to fit a model to the biological groups of interest with one addition factor to observe whether the factor has substantial influence on gene expression (such that many genes are detected as differentially expressed for that factor). Repeat this for subsequenct factors to determine the factors that should be included into the final model. 
globalAnalaysis = function(data_matrix=data_matrix_tissue, paired = FALSE, 
                           checkCovariates = c(FALSE, rep(TRUE, 6)), 
                           currentCovariates=currentCovariates, 
                           groups = groups, meta=meta_tissue,
                           varianceCutoff = 0, paired_variable="Patient") {
  sub = data_matrix[, c(grep(groups[1], colnames(data_matrix), value=T), 
                        grep(groups[2], colnames(data_matrix), value=T))]
  ## Remove features with low variance
  invisible(capture.output({library(matrixStats)}))
  keep = rowSds(sub%>%as.matrix(), na.rm = T) > varianceCutoff
  sub = sub[keep, ]
  rlst = lapply(seq_along(currentCovariates), function(i) {
    print(currentCovariates[i])
    currentCovariate = currentCovariates[i]
    checkCovariate = checkCovariates[i]
    limmaTest(dat = sub, paired=paired, 
              currentCovariate=currentCovariate, 
              checkCovariate=checkCovariate,
              meta=meta, paired_variable=paired_variable)
  }) %>% setNames(paste0("rlst_", currentCovariates))
  return(rlst)
}

modelSummary = function(rlst, Pcut, logFCcut) {
  print("Check whether to include a factor into the model by checking the number of features for which the factor is significant when added in the model: ")
  lapply(seq_along(rlst), function(j) {
    r=rlst[[j]]
    print(names(rlst)[j])
    if (names(rlst)[j]!="rlst_NULL") {
      print(sum((lapply(r$rslt_co, function(c) {
        c$adj.P.Val<Pcut
      }) %>% Reduce("+", .))>0, na.rm=T))
    }
  })
  
  print("Check how many features are significant in different models: ")
  lapply(seq_along(rlst), function(k) {
    r=rlst[[k]]
    print(names(rlst)[k])
    print(sum(r$rslt_of_interest$adj.P.Val<Pcut&abs(r$rslt_of_interest$logFC)>logFCcut, na.rm=T))
  })  
  
  print("Check how many significant features are shared when adding in an extra factor in comparison to only sample group is included in the model: ")
  sigFeatures = lapply(seq_along(rlst), function(k) {
    r=rlst[[k]]
    print(names(rlst)[k])
    rownames(r$rslt_of_interest)[r$rslt_of_interest$adj.P.Val<Pcut&abs(r$rslt_of_interest$logFC)>logFCcut]
  }) %>% setNames(currentCovariates)
  lapply(2:length(sigFeatures), function(l) {
    print(intersect(sigFeatures[[1]], sigFeatures[[l]]) %>% length())
  })
  return(sigFeatures)
}

```

## Task Depth3 (task level)

```{r}
depth = "Depth3"
cf5_Depth = cf5[, c(depth, "Sample", "TaskScore")] %>% 
  tidyr::spread(key="Sample", value="TaskScore") %>% 
  tibble::column_to_rownames(var=depth)
```

### PNLvsPL

```{r}
paired = TRUE
checkCovariate = FALSE
currentCovariate = "NULL"
sub_PLvsPNL = cf5_Depth[, c(grep("PL", colnames(cf5_Depth), value=T), 
                        c(grep("PNL", colnames(cf5_Depth), value=T)))]

## Remove features with low variance
invisible(capture.output({library(matrixStats)}))
keep = rowSds(sub_PLvsPNL%>%as.matrix(), na.rm = T) > 0
sub_PLvsPNL = sub_PLvsPNL[keep, ]
meta_extra$Group = factor(meta_extra$Group, levels=c("HS", "PNL", "PL"))
## Paired analysis with only group in the model
rlst_PNLvsPL3 = limmaTest(dat = sub_PLvsPNL, paired=paired, 
                 currentCovariate=currentCovariate, 
                 checkCovariate=checkCovariate,
                 meta = meta_extra, paired_variable="Patient")
Pcut = 0.05
logFCcut = 0.5
print("Check how many features are signficiant: ")
sum(rlst_PNLvsPL3$rslt_of_interest$adj.P.Val<Pcut&abs(rlst_PNLvsPL3$rslt_of_interest$logFC)>logFCcut, na.rm=T)

OUTDIR_LIMMA = paste0(OUTDIR, "/Limma_differentialAnalysis/Task_Depth3/")
dir.create(OUTDIR_LIMMA, recursive=T, showWarnings=F)

rm(paired, checkCovariate, currentCovariate, sub_PLvsPNL, keep, Pcut, logFCcut )
```

### HSvsPL

```{r}
currentCovariates = c("NULL", "Age","Sex","Anatomical_category","Date_of_collection",
                      "Smoking","BMI","Heredity_for_psoriasis", #"Disease_category2",
                      "Alergies_asthma","Alcohol","Exercise",
                      "Exercise_num")
checkCovariates = c(FALSE, rep(TRUE, length(currentCovariates)-1))
groups = c("PL", "HS") # Here the treat group needs to come first
paired = TRUE

meta_extra$Group = factor(meta_extra$Group, levels=c("HS", "PNL", "PL"))
rlst_HSvsPL3 = globalAnalaysis(data_matrix=cf5_Depth, 
                           paired = paired, 
                           checkCovariates = checkCovariates, 
                           currentCovariates=currentCovariates, 
                           groups = groups, meta = meta_extra, 
                           paired_variable="Match")
Pcut = 0.05
logFCcut = 0.5
sigFeatures = modelSummary(rlst=rlst_HSvsPL3, Pcut=Pcut, logFCcut=logFCcut)

rm(currentCovariates, checkCovariates, groups, paired, Pcut, logFCcut, sigFeatures)

```

```{r}
currentCovariates = c("Exercise")
checkCovariates = FALSE
groups = c("PL", "HS")
paired = T
meta_extra$Group = factor(meta_extra$Group, levels=c("HS", "PNL", "PL"))
rlst_HSvsPL_final3 = globalAnalaysis(data_matrix=cf5_Depth, 
                           paired = paired, 
                           checkCovariates = checkCovariates, 
                           currentCovariates=currentCovariates, 
                           groups = groups, meta = meta_extra, 
                           paired_variable="Match")

Pcut = 0.05
logFCcut = 0.5
# Check for how many features the sample group is significant (adj.P.Val<0.05 and logFC>0.3):
rlst_HSvsPL_final3$`rlst_Exercise`$rslt_of_interest %>%
  filter(abs(logFC)>logFCcut&adj.P.Val<Pcut) %>% nrow()

rm(currentCovariates, checkCovariates, groups, paired, Pcut, logFCcut)
```

### HSvsPNL

```{r}
currentCovariates = c("NULL", "Age","Sex","Anatomical_category","Date_of_collection",
                      "Smoking","BMI","Heredity_for_psoriasis", #"Disease_category2",
                      "Alergies_asthma","Alcohol","Exercise",
                      "Exercise_num")
checkCovariates = c(FALSE, rep(TRUE, length(currentCovariates)-1))
groups = c("PNL", "HS") # The treat group needs to come first, important!!!
paired = TRUE
meta_extra$Group = factor(meta_extra$Group, levels=c("HS", "PNL", "PL"))
rlst_HSvsPNL3 = globalAnalaysis(data_matrix=cf5_Depth, 
                           paired = paired, 
                           checkCovariate = checkCovariates, 
                           currentCovariates=currentCovariates, 
                           groups = groups, meta = meta_extra, 
                           paired_variable="Match")
Pcut = 0.05
logFCcut = 0.5
sigFeatures = modelSummary(rlst=rlst_HSvsPNL3, Pcut=Pcut, logFCcut=logFCcut)
rm(currentCovariates, checkCovariates, groups, paired, Pcut, logFCcut, sigFeatures)
```

```{r}
rlst_HSvsPNL_final3 = rlst_HSvsPNL3$rlst_NULL
Pcut = 0.05
logFCcut = 0.5
# Check for how many features the sample group is significant (adj.P.Val<0.05 and logFC>0.3):
rlst_HSvsPNL_final3$rslt_of_interest %>%
  filter(abs(logFC)>0.5&adj.P.Val<0.05) %>% nrow()

saveRDS(list(
  PNLvsPL = rlst_PNLvsPL3,
  HSvsPL = rlst_HSvsPL_final3$rlst_Date_of_collection,
  HSvsPN = rlst_HSvsPNL3), 
        paste0(OUTDIR_LIMMA, "/Limma_differentialAnalysis_result_original.RDS"))

rm(Pcut, logFCcut)
```

### Visualization

```{r, fig.height=8, fig.width=8}
library(ggplot2)
doVolcano = function(resi, NAME, adj.P.Val.cutoff=0.05, logFC.cutoff=0.3, lfcShrink=F) {
  selectedCols = c("Task", "logFC", "P.Value", "adj.P.Val")
  # resi$Protein = gsub("sp[|]", "", resi$Protein) %>%
  #   sub("[|].*", "", .)
  toExport = resi[, selectedCols]
  toExport = toExport[order(toExport$P.Value),]

  tmp = as.data.frame(toExport) %>%
    filter(!is.na(adj.P.Val)) %>%
    mutate(sig = ifelse(adj.P.Val<adj.P.Val.cutoff&logFC>logFC.cutoff, "up", 
    ifelse(adj.P.Val<adj.P.Val.cutoff&(logFC< -logFC.cutoff), "down", "nonSig")))
  tmpSignif = tmp %>%
    filter(tmp$logFC>quantile(abs(tmp$logFC),0.98)|(tmp$logFC< -quantile(abs(tmp$logFC),0.98))) %>%
    filter(sig!="nonSig")
  
  p = ggplot(data = tmp, aes(logFC, -log10(adj.P.Val), col = sig)) + 
    geom_point() + 
    xlab("Log2 Fold Change") +
    ggrepel::geom_text_repel(data = tmpSignif, aes(logFC, -log10(adj.P.Val), 
                                          label = Task), color="cornsilk4") +
    geom_hline(yintercept = -log10(0.05), lty = 2) +
    geom_vline(xintercept = c(-0.5, 0.5), lty = 2) + 
    scale_colour_manual(values = c("down"="blue", "nonSig"="black", "up"="red")) + 
    theme(legend.position="none") +
    ggtitle(ifelse(lfcShrink, paste0(NAME, "_lfcShrink"), paste0(NAME, "")))
  return(p)
}

limma_rslt3 = list(
  PNLvsPL = rlst_PNLvsPL3$rslt_of_interes %>%
    tibble::rownames_to_column(var="Task") %>%
    arrange(-logFC),
  HSvsPL = rlst_HSvsPL_final3$`rlst_Exercise`$rslt_of_interest %>%
    tibble::rownames_to_column(var="Task") %>% arrange(-logFC),
  HSvsPNL = rlst_HSvsPNL_final3$rslt_of_interest %>%
    tibble::rownames_to_column(var="Task") %>% arrange(-logFC))

saveRDS(limma_rslt3, 
        paste0(OUTDIR_LIMMA, "/Limma_differentialAnalysis_result.RDS"))
# limma_rslt=readRDS(paste0(OUTDIR_LIMMA, "/Limma_differentialAnalysis_result.RDS"))

NAMES = names(limma_rslt3)
for (i in seq_along(limma_rslt3)) {
  p = doVolcano(resi=limma_rslt3[[i]], NAME=NAMES[i], adj.P.Val.cutoff=0.05, 
                logFC.cutoff=0.5, lfcShrink=F)
  assign(paste0("P", i), p)
}

pdf(paste0(OUTDIR_LIMMA,"/Volcano_plots.pdf"),h=10,w=15)
  cowplot::plot_grid(P1, P2, P3, ncol=2, nrow=2)
dev.off()
cowplot::plot_grid(P1, P2, P3, ncol=2, nrow=2)

# Save Limma results
TableNames = list(PNLvsPL = limma_rslt3$PNLvsPL,
                  HSvsPL = limma_rslt3$HSvsPL,
                  HSvsPN = limma_rslt3$HSvsPNL)
openxlsx::write.xlsx(TableNames, 
             file=paste0(OUTDIR_LIMMA, "/Limma_differentialAnalysis_result.xlsx"))

rm(NAMES, TableNames)
```

```{r}
logFC.cutoff = 0.5
adj.P.Val.cutoff = 0.05
ups = lapply(limma_rslt3, function(rslt) {
  rslt %>% filter((rslt$logFC>logFC.cutoff)&(rslt$adj.P.Val<adj.P.Val.cutoff)) %>%
    pull(Task)
})

downs = lapply(limma_rslt3, function(rslt) {
  rslt %>% filter((rslt$logFC< -logFC.cutoff)&(rslt$adj.P.Val<adj.P.Val.cutoff)) %>%
    pull(Task)
})

source("../src/ggvenn.R")
pdf(paste0(OUTDIR_LIMMA,"/Venn.pdf"),h=3.5,w=8)
  theme = theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))
  cowplot::plot_grid(ggvenn(ups, show_percentage = FALSE, set_name_size = 4) + 
            labs(title = "Up-regulated")+ theme,
          ggvenn(downs, show_percentage = FALSE, 
       set_name_size = 4) + 
    labs(title = "Down-regulated")+ theme)
dev.off()
cowplot::plot_grid(ggvenn(ups, show_percentage = FALSE, set_name_size = 4) + 
            labs(title = "Up-regulated")+ theme,
          ggvenn(downs, show_percentage = FALSE, 
       set_name_size = 4) + 
    labs(title = "Down-regulated")+ theme)

rm(logFC.cutoff, adj.P.Val.cutoff, ups, downs)
```

```{r, fig.height=16, fig.width=19}
depth = "Depth3"
cf5_Depth = cf5[, c(depth, "Sample", "TaskScore")] %>% 
  tidyr::spread(key="Sample", value="TaskScore") %>% 
  tibble::column_to_rownames(var=depth)

invisible(capture.output({library(matrixStats)}))
keep = rowSds(cf5_Depth%>%as.matrix(), na.rm = T) > 0
cf5_Depth = cf5_Depth[keep, ]
cf5_Depth_2 = cf5_Depth[,c(grep("HS",colnames(cf5_Depth)),
                             grep("PL",colnames(cf5_Depth)),
                             grep("PNL",colnames(cf5_Depth)))]
colAnn = data.frame(Group = colnames(cf5_Depth) %>% gsub("[-].*", "", .)) 
rownames(colAnn) = colnames(cf5_Depth)
cf5_Depth_2 = cf5_Depth_2[
  c(limma_rslt3$HSvsPL%>%arrange(logFC)%>%pull(Task), 
  setdiff(rownames(cf5_Depth_2),limma_rslt3$HSvsPL%>%pull(Task))), ] 
# pdf(paste0(OUTDIR_V,"/CelltypeFractions_heatmap_all_padjCutoff",adj.P.Val_cutoff,".pdf"), 
#    h=7, w=7)
pheatmap::pheatmap(cf5_Depth_2,
                   color=viridisLite::magma(50),
                   breaks=seq(-2,2,len=50),
                   border_color="black",
                   scale="row",
                   annotation_col=colAnn,
                   # annotation_colors=ann_colors,
                   labels_col=rep("", ncol(cf5_Depth_2)),
                   main="Task Scores",
                   cluster_cols = FALSE,
                   cluster_rows = FALSE,  
                   gaps_col =  c(50,99,148))

# install.packages("Cairo")  # if not installed
library(Cairo)
# CairoPDF(paste0(OUTDIR_LIMMA,"/Heatmap.png"), width = 19, height = 16)
# pdf(paste0(OUTDIR_LIMMA,"/Heatmap.png"),h=16,w=19)
pheatmap::pheatmap(cf5_Depth_2,
                   color=viridisLite::magma(50),
                   breaks=seq(-2,2,len=50),
                   border_color="black",
                   scale="row",
                   annotation_col=colAnn,
                   # annotation_colors=ann_colors,
                   labels_col=rep("", ncol(cf5_Depth_2)),
                   main="Task Scores",
                   cluster_cols = FALSE,
                   cluster_rows = FALSE,  
                   gaps_col =  c(50,99,148),
                   filename=paste0(OUTDIR_LIMMA,"/Heatmap.png"), 
                   width=19, height=16)
# dev.off()

rm(keep, colAnn, cf5_Depth_2, cf5_Depth, depth)
```

## Task Depth2

```{r}
depth = "Depth2"
DepthWide = Depth2[, c(depth, "Sample", "MeanTaskScore_Depth2")] %>% 
  tidyr::spread(key="Sample", value="MeanTaskScore_Depth2") %>% 
  tibble::column_to_rownames(var=depth)
```

### PNLvsPL

```{r}
paired = TRUE
checkCovariate = FALSE
currentCovariate = "NULL"
sub_PLvsPNL = DepthWide[, c(grep("PL", colnames(DepthWide), value=T), 
                        c(grep("PNL", colnames(DepthWide), value=T)))]

## Remove features with low variance
invisible(capture.output({library(matrixStats)}))
keep = rowSds(sub_PLvsPNL%>%as.matrix(), na.rm = T) > 0
sub_PLvsPNL = sub_PLvsPNL[keep, ]
meta_extra$Group = factor(meta_extra$Group, levels=c("HS", "PNL", "PL"))
## Paired analysis with only group in the model
rlst_PNLvsPL2 = limmaTest(dat = sub_PLvsPNL, paired=paired, 
                 currentCovariate=currentCovariate, 
                 checkCovariate=checkCovariate,
                 meta = meta_extra, paired_variable="Patient")
Pcut = 0.05
logFCcut = 0.5
print("Check how many features are signficiant: ")
sum(rlst_PNLvsPL2$rslt_of_interest$adj.P.Val<Pcut&abs(rlst_PNLvsPL2$rslt_of_interest$logFC)>logFCcut, na.rm=T)

OUTDIR_LIMMA = paste0(OUTDIR, "/Limma_differentialAnalysis/TaskDepth2/")
dir.create(OUTDIR_LIMMA, recursive=T, showWarnings=F)

rm(paired, checkCovariate, currentCovariate, sub_PLvsPNL, keep, Pcut, logFCcut )
```

### HSvsPL

```{r}
currentCovariates = c("NULL", "Age","Sex","Anatomical_category","Date_of_collection",
                      "Smoking","BMI","Heredity_for_psoriasis", #"Disease_category2",
                      "Alergies_asthma","Alcohol","Exercise",
                      "Exercise_num")
checkCovariates = c(FALSE, rep(TRUE, length(currentCovariates)-1))
groups = c("PL", "HS") # Here the treat group needs to come first
paired = TRUE

meta_extra$Group = factor(meta_extra$Group, levels=c("HS", "PNL", "PL"))
rlst_HSvsPL2 = globalAnalaysis(data_matrix=DepthWide, 
                           paired = paired, 
                           checkCovariates = checkCovariates, 
                           currentCovariates=currentCovariates, 
                           groups = groups, meta = meta_extra, 
                           paired_variable="Match")
Pcut = 0.05
logFCcut = 0.5
sigFeatures = modelSummary(rlst=rlst_HSvsPL2, Pcut=Pcut, logFCcut=logFCcut)

rm(currentCovariates, checkCovariates, groups, paired, Pcut, logFCcut, sigFeatures)

```

```{r}
currentCovariates = c("Exercise")
checkCovariates = FALSE
groups = c("PL", "HS")
paired = T
meta_extra$Group = factor(meta_extra$Group, levels=c("HS", "PNL", "PL"))
rlst_HSvsPL_final2 = globalAnalaysis(data_matrix=DepthWide, 
                           paired = paired, 
                           checkCovariates = checkCovariates, 
                           currentCovariates=currentCovariates, 
                           groups = groups, meta = meta_extra, 
                           paired_variable="Match")

Pcut = 0.05
logFCcut = 0.5
# Check for how many features the sample group is significant (adj.P.Val<0.05 and logFC>0.3):
rlst_HSvsPL_final2$`rlst_Exercise`$rslt_of_interest %>%
  filter(abs(logFC)>logFCcut&adj.P.Val<Pcut) %>% nrow()

rm(currentCovariates, checkCovariates, groups, paired, Pcut, logFCcut)
```

### HSvsPNL

```{r}
currentCovariates = c("NULL", "Age","Sex","Anatomical_category","Date_of_collection",
                      "Smoking","BMI","Heredity_for_psoriasis", #"Disease_category2",
                      "Alergies_asthma","Alcohol","Exercise",
                      "Exercise_num")
checkCovariates = c(FALSE, rep(TRUE, length(currentCovariates)-1))
groups = c("PNL", "HS") # The treat group needs to come first, important!!!
paired = TRUE
meta_extra$Group = factor(meta_extra$Group, levels=c("HS", "PNL", "PL"))
rlst_HSvsPNL2 = globalAnalaysis(data_matrix=DepthWide, 
                           paired = paired, 
                           checkCovariate = checkCovariates, 
                           currentCovariates=currentCovariates, 
                           groups = groups, meta = meta_extra, 
                           paired_variable="Match")
Pcut = 0.05
logFCcut = 0.5
sigFeatures = modelSummary(rlst=rlst_HSvsPNL2, Pcut=Pcut, logFCcut=logFCcut)
rm(currentCovariates, checkCovariates, groups, paired, Pcut, logFCcut, sigFeatures)
```

```{r}
currentCovariates = c("Heredity_for_psoriasis")
checkCovariates = FALSE
groups = c("HS", "PNL")
paired = T
meta_extra$Group = factor(meta_extra$Group, levels=c("HS", "PNL", "PL"))
rlst_HSvsPNL_final2 = globalAnalaysis(data_matrix=DepthWide, 
                           paired = paired, 
                           checkCovariates = checkCovariates, 
                           currentCovariates=currentCovariates, 
                           groups = groups, meta = meta_extra, 
                           paired_variable="Match")
Pcut = 0.05
logFCcut = 0.5
# Check for how many features the sample group is significant (adj.P.Val<0.05 and logFC>0.3):
rlst_HSvsPNL_final2$rlst_Heredity_for_psoriasis$rslt_of_interest %>%
  filter(abs(logFC)>0.5&adj.P.Val<0.05) %>% nrow()

saveRDS(list(
  PNLvsPL = rlst_PNLvsPL2,
  HSvsPL = rlst_HSvsPL_final2$rlst_Date_of_collection,
  HSvsPN = rlst_HSvsPNL_final2$rlst_Heredity_for_psoriasis), 
        paste0(OUTDIR_LIMMA, "/Limma_differentialAnalysis_result_original.RDS"))

rm(Pcut, logFCcut)
```

### Visualization

```{r, fig.height=8, fig.width=8}
library(ggplot2)
limma_rslt2 = list(
  PNLvsPL = rlst_PNLvsPL2$rslt_of_interes %>%
    tibble::rownames_to_column(var="Task") %>%
    arrange(-logFC),
  HSvsPL = rlst_HSvsPL_final2$`rlst_Exercise`$rslt_of_interest %>%
    tibble::rownames_to_column(var="Task") %>% arrange(-logFC),
  HSvsPNL = rlst_HSvsPNL_final2$rlst_Heredity_for_psoriasis$rslt_of_interest %>%
    tibble::rownames_to_column(var="Task") %>% arrange(-logFC))

saveRDS(limma_rslt2, 
        paste0(OUTDIR_LIMMA, "/Limma_differentialAnalysis_result.RDS"))
# limma_rslt=readRDS(paste0(OUTDIR_LIMMA, "/Limma_differentialAnalysis_result.RDS"))

NAMES = names(limma_rslt2)
for (i in seq_along(limma_rslt2)) {
  p = doVolcano(resi=limma_rslt2[[i]], NAME=NAMES[i], adj.P.Val.cutoff=0.05, 
                logFC.cutoff=0.5, lfcShrink=F)
  assign(paste0("P", i), p)
}

pdf(paste0(OUTDIR_LIMMA,"/Volcano_plots.pdf"),h=10,w=15)
  cowplot::plot_grid(P1, P2, P3, ncol=2, nrow=2)
dev.off()
cowplot::plot_grid(P1, P2, P3, ncol=2, nrow=2)

# Save Limma results
TableNames = list(PNLvsPL = limma_rslt2$PNLvsPL,
                  HSvsPL = limma_rslt2$HSvsPL,
                  HSvsPN = limma_rslt2$HSvsPNL)
openxlsx::write.xlsx(TableNames, 
             file=paste0(OUTDIR_LIMMA, "/Limma_differentialAnalysis_result.xlsx"))

rm(NAMES, TableNames)
```

```{r}
logFC.cutoff = 0.5
adj.P.Val.cutoff = 0.05
ups = lapply(limma_rslt2, function(rslt) {
  rslt %>% filter((rslt$logFC>logFC.cutoff)&(rslt$adj.P.Val<adj.P.Val.cutoff)) %>%
    pull(Task)
})

downs = lapply(limma_rslt2, function(rslt) {
  rslt %>% filter((rslt$logFC< -logFC.cutoff)&(rslt$adj.P.Val<adj.P.Val.cutoff)) %>%
    pull(Task)
})

source("../src/ggvenn.R")
pdf(paste0(OUTDIR_LIMMA,"/Venn.pdf"),h=3.5,w=8)
  theme = theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))
  cowplot::plot_grid(ggvenn(ups, show_percentage = FALSE, set_name_size = 4) + 
            labs(title = "Up-regulated")+ theme,
          ggvenn(downs, show_percentage = FALSE, 
       set_name_size = 4) + 
    labs(title = "Down-regulated")+ theme)
dev.off()
cowplot::plot_grid(ggvenn(ups, show_percentage = FALSE, set_name_size = 4) + 
            labs(title = "Up-regulated")+ theme,
          ggvenn(downs, show_percentage = FALSE, 
       set_name_size = 4) + 
    labs(title = "Down-regulated")+ theme)

rm(logFC.cutoff, adj.P.Val.cutoff, ups, downs)
```

```{r, fig.height=10, fig.width=19}
depth = "Depth2"
DepthWide = Depth2[, c(depth, "Sample", "MeanTaskScore_Depth2")] %>% 
  tidyr::spread(key="Sample", value="MeanTaskScore_Depth2") %>% 
  tibble::column_to_rownames(var=depth)

invisible(capture.output({library(matrixStats)}))
keep = rowSds(DepthWide%>%as.matrix(), na.rm = T) > 0
DepthWide = DepthWide[keep, ]
DepthWide_2 = DepthWide[,c(grep("HS",colnames(DepthWide)),
                             grep("PL",colnames(DepthWide)),
                             grep("PNL",colnames(DepthWide)))]
colAnn = data.frame(Group = colnames(DepthWide) %>% gsub("[-].*", "", .)) 
rownames(colAnn) = colnames(DepthWide)
DepthWide_2 = DepthWide_2[
  c(limma_rslt2$HSvsPL%>%arrange(logFC)%>%pull(Task), 
  setdiff(rownames(DepthWide_2),limma_rslt2$HSvsPL%>%pull(Task))), ] 
# pdf(paste0(OUTDIR_V,"/CelltypeFractions_heatmap_all_padjCutoff",adj.P.Val_cutoff,".pdf"), 
#    h=7, w=7)
pheatmap::pheatmap(DepthWide_2,
                   color=viridisLite::magma(50),
                   breaks=seq(-2,2,len=50),
                   border_color="black",
                   scale="row",
                   annotation_col=colAnn,
                   # annotation_colors=ann_colors,
                   labels_col=rep("", ncol(DepthWide_2)),
                   main="Task Scores",
                   cluster_cols = FALSE,
                   cluster_rows = FALSE,  
                   gaps_col =  c(50,99,148))

library(Cairo)
# CairoPDF(paste0(OUTDIR_LIMMA,"/Heatmap.png"), width = 19, height = 10)
# pdf(paste0(OUTDIR_LIMMA,"/Heatmap.png"),h=16,w=19)
pheatmap::pheatmap(DepthWide_2,
                   color=viridisLite::magma(50),
                   breaks=seq(-2,2,len=50),
                   border_color="black",
                   scale="row",
                   annotation_col=colAnn,
                   # annotation_colors=ann_colors,
                   labels_col=rep("", ncol(DepthWide_2)),
                   main="Task Scores",
                   cluster_cols = FALSE,
                   cluster_rows = FALSE,  
                   gaps_col=c(50,99,148),
                   filename=paste0(OUTDIR_LIMMA,"/Heatmap.png"), 
                   width=19, height=10)
# dev.off()

rm(keep, colAnn, DepthWide_2, DepthWide)
```

## Task Depth1

```{r}
depth = "Depth1"
DepthWide = Depth1[, c(depth, "Sample", "MeanTaskScore_Depth1")] %>% 
  tidyr::spread(key="Sample", value="MeanTaskScore_Depth1") %>% 
  tibble::column_to_rownames(var=depth)
```

### PNLvsPL

```{r}
paired = TRUE
checkCovariate = FALSE
currentCovariate = "NULL"
sub_PLvsPNL = DepthWide[, c(grep("PL", colnames(DepthWide), value=T), 
                        c(grep("PNL", colnames(DepthWide), value=T)))]

## Remove features with low variance
invisible(capture.output({library(matrixStats)}))
keep = rowSds(sub_PLvsPNL%>%as.matrix(), na.rm = T) > 0
sub_PLvsPNL = sub_PLvsPNL[keep, ]
meta_extra$Group = factor(meta_extra$Group, levels=c("HS", "PNL", "PL"))
## Paired analysis with only group in the model
rlst_PNLvsPL1 = limmaTest(dat = sub_PLvsPNL, paired=paired, 
                 currentCovariate=currentCovariate, 
                 checkCovariate=checkCovariate,
                 meta = meta_extra, paired_variable="Patient")
Pcut = 0.05
logFCcut = 0.5
print("Check how many features are signficiant: ")
sum(rlst_PNLvsPL1$rslt_of_interest$adj.P.Val<Pcut&abs(rlst_PNLvsPL1$rslt_of_interest$logFC)>logFCcut, na.rm=T)

OUTDIR_LIMMA = paste0(OUTDIR, "/Limma_differentialAnalysis/TaskDepth1")
dir.create(OUTDIR_LIMMA, recursive=T, showWarnings=F)

rm(paired, checkCovariate, currentCovariate, sub_PLvsPNL, keep, Pcut, logFCcut )
```

### HSvsPL

```{r}
currentCovariates = c("NULL", "Age","Sex","Anatomical_category","Date_of_collection",
                      "Smoking","BMI","Heredity_for_psoriasis", #"Disease_category2",
                      "Alergies_asthma","Alcohol","Exercise",
                      "Exercise_num")
checkCovariates = c(FALSE, rep(TRUE, length(currentCovariates)-1))
groups = c("PL", "HS") # Here the treat group needs to come first
paired = TRUE

meta_extra$Group = factor(meta_extra$Group, levels=c("HS", "PNL", "PL"))
rlst_HSvsPL1 = globalAnalaysis(data_matrix=DepthWide, 
                           paired = paired, 
                           checkCovariates = checkCovariates, 
                           currentCovariates=currentCovariates, 
                           groups = groups, meta = meta_extra, 
                           paired_variable="Match")
Pcut = 0.05
logFCcut = 0.5
sigFeatures = modelSummary(rlst=rlst_HSvsPL1, Pcut=Pcut, logFCcut=logFCcut)

rm(currentCovariates, checkCovariates, groups, paired, Pcut, logFCcut, sigFeatures)

```

```{r}
rlst_HSvsPL_final1 = rlst_HSvsPL1$rlst_NULL
Pcut = 0.05
logFCcut = 0.5
# Check for how many features the sample group is significant (adj.P.Val<0.05 and logFC>0.3):
rlst_HSvsPL_final1$rslt_of_interest %>%
  filter(abs(logFC)>logFCcut&adj.P.Val<Pcut) %>% nrow()

rm(Pcut, logFCcut)
```

### HSvsPNL

```{r}
currentCovariates = c("NULL", "Age","Sex","Anatomical_category","Date_of_collection",
                      "Smoking","BMI","Heredity_for_psoriasis", #"Disease_category2",
                      "Alergies_asthma","Alcohol","Exercise",
                      "Exercise_num")
checkCovariates = c(FALSE, rep(TRUE, length(currentCovariates)-1))
groups = c("PNL", "HS") # The treat group needs to come first, important!!!
paired = TRUE
meta_extra$Group = factor(meta_extra$Group, levels=c("HS", "PNL", "PL"))
rlst_HSvsPNL1 = globalAnalaysis(data_matrix=DepthWide, 
                           paired = paired, 
                           checkCovariate = checkCovariates, 
                           currentCovariates=currentCovariates, 
                           groups = groups, meta = meta_extra, 
                           paired_variable="Match")
Pcut = 0.05
logFCcut = 0.5
sigFeatures = modelSummary(rlst=rlst_HSvsPNL1, Pcut=Pcut, logFCcut=logFCcut)
rm(currentCovariates, checkCovariates, groups, paired, Pcut, logFCcut, sigFeatures)
```

```{r}
rlst_HSvsPNL_final1 = rlst_HSvsPNL1$rlst_NULL
Pcut = 0.05
logFCcut = 0.5
# Check for how many features the sample group is significant (adj.P.Val<0.05 and logFC>0.3):
rlst_HSvsPNL_final1$rslt_of_interest %>%
  filter(abs(logFC)>0.5&adj.P.Val<0.05) %>% nrow()

saveRDS(list(
  PNLvsPL = rlst_PNLvsPL1,
  HSvsPL = rlst_HSvsPL1,
  HSvsPN = rlst_HSvsPNL1), 
        paste0(OUTDIR_LIMMA, "/Limma_differentialAnalysis_result_original.RDS"))

rm(Pcut, logFCcut)
```

### Visualization

```{r, fig.height=8, fig.width=8}
library(ggplot2)
limma_rslt1 = list(
  PNLvsPL = rlst_PNLvsPL1$rslt_of_interes %>%
    tibble::rownames_to_column(var="Task") %>%
    arrange(-logFC),
  HSvsPL = rlst_HSvsPL_final1$rslt_of_interest %>%
    tibble::rownames_to_column(var="Task") %>% arrange(-logFC),
  HSvsPNL = rlst_HSvsPNL_final1$rslt_of_interest %>%
    tibble::rownames_to_column(var="Task") %>% arrange(-logFC))

saveRDS(limma_rslt1, 
        paste0(OUTDIR_LIMMA, "/Limma_differentialAnalysis_result.RDS"))
# limma_rslt=readRDS(paste0(OUTDIR_LIMMA, "/Limma_differentialAnalysis_result.RDS"))

NAMES = names(limma_rslt1)
for (i in seq_along(limma_rslt1)) {
  p = doVolcano(resi=limma_rslt1[[i]], NAME=NAMES[i], adj.P.Val.cutoff=0.05, 
                logFC.cutoff=0.5, lfcShrink=F)
  assign(paste0("P", i), p)
}

pdf(paste0(OUTDIR_LIMMA,"/Volcano_plots.pdf"),h=10,w=15)
  cowplot::plot_grid(P1, P2, P3, ncol=2, nrow=2)
dev.off()
cowplot::plot_grid(P1, P2, P3, ncol=2, nrow=2)

# Save Limma results
TableNames = list(PNLvsPL = limma_rslt1$PNLvsPL,
                  HSvsPL = limma_rslt1$HSvsPL,
                  HSvsPN = limma_rslt1$HSvsPNL)
openxlsx::write.xlsx(TableNames, 
             file=paste0(OUTDIR_LIMMA, "/Limma_differentialAnalysis_result.xlsx"))

rm(NAMES, TableNames)
```

```{r}
logFC.cutoff = 0.5
adj.P.Val.cutoff = 0.05
ups = lapply(limma_rslt1, function(rslt) {
  rslt %>% filter((rslt$logFC>logFC.cutoff)&(rslt$adj.P.Val<adj.P.Val.cutoff)) %>%
    pull(Task)
})

downs = lapply(limma_rslt1, function(rslt) {
  rslt %>% filter((rslt$logFC< -logFC.cutoff)&(rslt$adj.P.Val<adj.P.Val.cutoff)) %>%
    pull(Task)
})

source("../src/ggvenn.R")
pdf(paste0(OUTDIR_LIMMA,"/Venn.pdf"),h=3.5,w=8)
  theme = theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))
  cowplot::plot_grid(ggvenn(ups, show_percentage = FALSE, set_name_size = 4) + 
            labs(title = "Up-regulated")+ theme,
          ggvenn(downs, show_percentage = FALSE, 
       set_name_size = 4) + 
    labs(title = "Down-regulated")+ theme)
dev.off()
cowplot::plot_grid(ggvenn(ups, show_percentage = FALSE, set_name_size = 4) + 
            labs(title = "Up-regulated")+ theme,
          ggvenn(downs, show_percentage = FALSE, 
       set_name_size = 4) + 
    labs(title = "Down-regulated")+ theme)

rm(logFC.cutoff, adj.P.Val.cutoff, ups, downs)
```

```{r, fig.height=2.5, fig.width=9}
depth = "Depth1"
DepthWide = Depth1[, c(depth, "Sample", "MeanTaskScore_Depth1")] %>% 
  tidyr::spread(key="Sample", value="MeanTaskScore_Depth1") %>% 
  tibble::column_to_rownames(var=depth)

invisible(capture.output({library(matrixStats)}))
keep = rowSds(DepthWide%>%as.matrix(), na.rm = T) > 0
DepthWide = DepthWide[keep, ]
DepthWide_2 = DepthWide[,c(grep("HS",colnames(DepthWide)),
                             grep("PL",colnames(DepthWide)),
                             grep("PNL",colnames(DepthWide)))]
colAnn = data.frame(Group = colnames(DepthWide) %>% gsub("[-].*", "", .)) 
rownames(colAnn) = colnames(DepthWide)
DepthWide_2 = DepthWide_2[
  c(limma_rslt1$HSvsPL%>%arrange(logFC)%>%pull(Task), 
  setdiff(rownames(DepthWide_2),limma_rslt1$HSvsPL%>%pull(Task))), ] 
# pdf(paste0(OUTDIR_V,"/CelltypeFractions_heatmap_all_padjCutoff",adj.P.Val_cutoff,".pdf"), 
#    h=7, w=7)
pheatmap::pheatmap(DepthWide_2,
                   color=viridisLite::magma(50),
                   breaks=seq(-2,2,len=50),
                   border_color="black",
                   scale="row",
                   annotation_col=colAnn,
                   # annotation_colors=ann_colors,
                   labels_col=rep("", ncol(DepthWide_2)),
                   main="Task Scores",
                   cluster_cols = FALSE,
                   cluster_rows = FALSE,  
                   gaps_col =  c(50,99,148))


# CairoPDF(paste0(OUTDIR_LIMMA,"/Heatmap.png"), width = 9, height = 2.5)
pheatmap::pheatmap(DepthWide_2,
                   color=viridisLite::magma(50),
                   breaks=seq(-2,2,len=50),
                   border_color="black",
                   scale="row",
                   annotation_col=colAnn,
                   # annotation_colors=ann_colors,
                   labels_col=rep("", ncol(DepthWide_2)),
                   main="Task Scores",
                   cluster_cols = FALSE,
                   cluster_rows = FALSE,  
                   gaps_col =  c(50,99,148),
                   filename=paste0(OUTDIR_LIMMA,"/Heatmap.png"), 
                   width=9, height=2.5)
# dev.off()


rm(keep, colAnn, DepthWide_2)
```

# The hierarchy of the tasks at different depths

```{r, fig.height=40, fig.width=40}
TaskNetworks = function(pair, logFCcutoff=0.5, sig.cutoff=0.05, suffix, height, width) {
  combined_limmaTable = 
    Reduce(rbind, list(limma_rslt1[[pair]], limma_rslt2[[pair]], limma_rslt3[[pair]])) %>%
    tibble::column_to_rownames(var="Task") 
  
  library(igraph)
  # Define edges (from parent to child)
  edges = cf5 %>% dplyr::select(Depth1, Depth2) %>% setNames(c("", "")) %>%
    rbind(cf5 %>% dplyr::select(Depth2, Depth3) %>% setNames(c("", ""))) %>%
    unique() %>% as.matrix
  
  g = graph_from_edgelist(edges, directed = TRUE)
  V(g)$color = ifelse(V(g)$name%in%cf5$Depth1, adjustcolor("coral4", alpha.f = 0.7), 
                      ifelse(V(g)$name%in%cf5$Depth2, "grey" ,"skyblue"))
  
  label_up = combined_limmaTable %>% filter(logFC>0) %>% rownames(.)
  label_dn = combined_limmaTable %>% filter(logFC<0) %>% rownames(.)
  label.colors = ifelse(V(g)$name%in%label_up, "darkred", 
                        ifelse(V(g)$name%in%label_dn, "darkblue", "black"))
  
  labelsig = combined_limmaTable %>% 
    filter(abs(logFC)>logFCcutoff&adj.P.Val<sig.cutoff) %>% rownames(.)
  edge.colors = ifelse(V(g)$name%in%labelsig,  adjustcolor("orange", alpha.f=1), "white")
  
  V(g)$name[V(g)$name%in%labelsig] = paste0(V(g)$name[V(g)$name%in%labelsig], "*")
  
  sizes = rep(5, length(V(g)))
  
  set.seed(808)
  plot(g, 
       layout = layout_with_fr(g), 
       vertex.label = V(g)$name,
       vertex.frame.color = edge.colors,
       vertex.size = sizes,
       vertex.label.color = label.colors,
       vertex.label.cex = 1.2, 
       edge.arrow.size = 0.5,
       vertex.label.font = 2,
       vertex.frame.width=6)
  
  pdf(paste0(OUTDIR,"/Network/NetworksOfTaskDepths_",pair,"_sig.cutoff", suffix, ".pdf"),h=height,w=width)
  set.seed(808)
  plot(g, 
       layout = layout_with_fr(g), 
       vertex.label = V(g)$name,
       vertex.frame.color = edge.colors,
       vertex.size = sizes,
       vertex.label.color = label.colors,
       vertex.label.cex = 1.2, 
       edge.arrow.size = 0.5,
       vertex.label.font = 2,
       vertex.frame.width=6,
       main=paste0(pair,"_", suffix))
  
  # g_tbl <- as_tbl_graph(g)
  # ggraph(g_tbl, layout = "fr") +
  #   geom_edge_link() +
  #   geom_node_point(size = 3) +
  #   geom_node_text(aes(label = name), repel = TRUE, size = 3)
  dev.off()
}

```

Sig tasks: logFCcutoff=0.5, adj.P.valueCutoff=0.05.

The following 3 networks are repectively for PNLvsPL, HSvsPL and HSvsPNL:

```{r, fig.height=40, fig.width=40}
Pairs = names(limma_rslt1)
lapply(Pairs, function(pair) TaskNetworks(pair, suffix="adj.P.valueCutoff0.05", height=40, width=40))
lapply(Pairs, function(pair) TaskNetworks(pair, suffix="adj.P.valueCutoff0.05_larger", height=60, width=60))

```

Sig tasks: logFCcutoff=0.5, adj.P.valueCutoff=0.1.

The following 3 networks are repectively for PNLvsPL, HSvsPL and HSvsPNL:

```{r, fig.height=40, fig.width=40}
Pairs = names(limma_rslt1)
lapply(Pairs, function(pair) TaskNetworks(pair,logFCcutoff=0.5,sig.cutoff=0.1,
                          suffix="adj.P.valueCutoff0.1", height=40, width=40))
lapply(Pairs, function(pair) TaskNetworks(pair,logFCcutoff=0.5,sig.cutoff=0.1,
                          suffix="adj.P.valueCutoff0.1_larger", height=60, width=60))

```

Sig tasks: logFCcutoff=0.5, adj.P.valueCutoff=0.2.

The following 3 networks are repectively for PNLvsPL, HSvsPL and HSvsPNL:

```{r, fig.height=40, fig.width=40}
Pairs = names(limma_rslt1)
lapply(Pairs, function(pair) TaskNetworks(pair,logFCcutoff=0.5,sig.cutoff=0.2,
                          suffix="adj.P.valueCutoff0.2", height=40, width=40))
lapply(Pairs, function(pair) TaskNetworks(pair,logFCcutoff=0.5,sig.cutoff=0.2,
                          suffix="adj.P.valueCutoff0.2_larger", height=60, width=60))

```

Sig tasks: logFCcutoff=0.5, adj.P.valueCutoff=0.

The following 3 networks are repectively for PNLvsPL, HSvsPL and HSvsPNL:

```{r, fig.height=40, fig.width=40}
Pairs = names(limma_rslt1)
lapply(Pairs, function(pair) TaskNetworks(pair,logFCcutoff=0.5,sig.cutoff=0.00000000001,
                                      suffix="adj.P.valueCutoff0", height=40, width=40))
lapply(Pairs, function(pair) TaskNetworks(pair,logFCcutoff=0.5,sig.cutoff=0.00000000001,
                                      suffix="adj.P.valueCutoff0_larger", height=60, width=60))

```

```{r, fig.height=40, fig.width=40}
out = capture.output(
print("In the network, coral circles are task categories at depth1, grey circles ar task categories at depth2, while skyblue circles are task categrories at depth 3. Red labels: upregulated, blue labels: downregulated, circles with orange frame are signficant tasks (abs(logFC)>0.5, adj.P.Val<0.05/0.1/0.2/0)")
  )
writeLines(out, paste0(OUTDIR,"/NetworksOfTaskDepths_README.txt"))
out

print("In the network, coral circles are task categories at depth1, grey circles are task categories at depth2, while skyblue circles are task categrories at depth 3. Red labels: upregulated, blue labels: downregulated, circles with orange frame are signficant tasks (abs(logFC)>0.5, adj.P.Val<0.05/0.1/0.2/0)")

```

# sessionInfo

```{r}
out = capture.output(sessionInfo())
writeLines(out, paste0(OUTDIR, "sessionInfo.txt"))

sessionInfo()
```

# Deliverables

# Method summary

The phosphoproteomics data has been processed, filtered and normalized by the company.

PCA analyses were performed using the prcomp() function based on the top 500 HVGs. In order to perform PCA analysis, we performed Bayesian PCA with 2 components to impute missing values using the pca() function from the pcaMethods package.

Univariate differential analysis was performed using linear models implemented in the limma package, testing each gene individually for differences between conditions PNLvsPL, HSvsPL and HSvsPNL. For differential analysis between PNLvsPL, paired analysis was performed. For differential analysis between healthy control subjects and patients (i.e. HSvsPL and HSvsPNL), samples were also paired based on their their locations, and additional factors were considered, to determine which factors that should be included into the final model, we fit a model to the biological groups of interest with one additional factor to observe whether the factor has substantial influence on gene expression (such that many genes are detected as differentially expressed for that factor), we repeated this process for Age, Sex, Date_of_collection, Anatomical_category, Smoking, BMI, Heredity_for_psoriasis, Alergies_asthma, Alcohol, Exercise, Exercise_num. Eventually, Date of collection was added into the final model for HSvsPL, no additional factor was included in the final model for HSvsPNL. Benjamini-Hochberg multiple testing corrections were performed to control the false discovery rate.

In Limma univariate differential analysis, phosphoproteins with logFC\>0.5&adj.P.Val\<0.05 were considered to be signficant for PNLvsPL, HSvsPL and HSvsPNL.

Gene set enrichment analysis (GSEA) was also performed using several databases (MsigDB_Hallmark, MsigDB_GO:BP, MsigDB_GO:CC, MsigDB_C2 (without CGP), MitcoCarta3.0_MitoPathway3.0, MsigDB_C2 subcategory CP:kegg_legacy, MsigDB_C2 subcategory CP:KEGG_MEDICUS) with the GSEA() function from the clusterProfiler package. The features were ranked based on logFC. fdr multiple testing corrections were performed to control the false discovery rate.

# Directory

## Practical info

## Software

# Data responsibility

The responsibility for data archiving lies with the PI of the project. We do not offer long-term storage or retrieval of data.

-   **NBIS & UPPNEX:** We kindly ask that you remove the files from UPPMAX/UPPNEX. The main storage at UPPNEX is optimized for high-speed and parallel access, which makes it expensive and not the right place for longer time archiving. Please consider others by not taking up the expensive space. Please note that UPPMAX is a resource separate from the Bioinformatics Platform, administered by the Swedish National Infrastructure for Computing (SNIC) and SNIC-specifc project rules apply to all projects hosted at UPPMAX.
-   **Sensitive data:** Please note that special considerations may apply to the human-derived legally considered sensitive personal data. These should be handled according to specific laws and regulations as outlined e.g. [here](http://nbis.se/support/human-data.html).
-   **Long-term backup:** We recommend asking your local IT for support with long-term data archiving. The [Data Office](https://www.scilifelab.se/data/) at SciLifeLab may be of help to discuss other options.

# Acknowledgments

If you are presenting the results in a paper, at a workshop or conference, we kindly ask you to acknowledge us.

-   NBIS staff are encouraged to be co-authors when this is merited in accordance to the ethical recommendations for authorship, e.g. [ICMJE recommendations](http://www.icmje.org/recommendations/browse/roles-and-responsibilities/defining-the-role-of-authors-and-contributors.html). If applicable, please include **Name, Surname, National Bioinformatics Infrastructure Sweden, Science for Life Laboratory, Further Affliations**, as co-author. In other cases, NBIS would be grateful if support by us is acknowledged in publications according to this example:

> "Support by NBIS (National Bioinformatics Infrastructure Sweden) is gratefully acknowledged."

-   **UPPMAX** kindly asks you to [acknowledge UPPMAX and NAISS](https://www.uppmax.uu.se/support/faq/general-miscellaneous-faq/acknowledging-uppmax-and-naiss).

-   **NGI:** For publications based on data from NGI Sweden, NGI and SciLifeLab should be acknowledged. Instructions are included in the NGI reports as described [here](https://ngisweden.scilifelab.se/resources/getting-started-at-ngi/).

# sessionInfo

```{r}
output = capture.output({
  sessionInfo()
})
writeLines(output, paste0(OUTDIR, "sessionInfo.txt"))
output
```

# Deliverables

# Directory

## Practical info

## Software

# Data responsibility

The responsibility for data archiving lies with the PI of the project. We do not offer long-term storage or retrieval of data.

-   **NBIS & UPPNEX:** We kindly ask that you remove the files from UPPMAX/UPPNEX. The main storage at UPPNEX is optimized for high-speed and parallel access, which makes it expensive and not the right place for longer time archiving. Please consider others by not taking up the expensive space. Please note that UPPMAX is a resource separate from the Bioinformatics Platform, administered by the Swedish National Infrastructure for Computing (SNIC) and SNIC-specifc project rules apply to all projects hosted at UPPMAX.
-   **Sensitive data:** Please note that special considerations may apply to the human-derived legally considered sensitive personal data. These should be handled according to specific laws and regulations as outlined e.g. [here](http://nbis.se/support/human-data.html).
-   **Long-term backup:** We recommend asking your local IT for support with long-term data archiving. The [Data Office](https://www.scilifelab.se/data/) at SciLifeLab may be of help to discuss other options.

# Acknowledgments

If you are presenting the results in a paper, at a workshop or conference, we kindly ask you to acknowledge us.

-   NBIS staff are encouraged to be co-authors when this is merited in accordance to the ethical recommendations for authorship, e.g. [ICMJE recommendations](http://www.icmje.org/recommendations/browse/roles-and-responsibilities/defining-the-role-of-authors-and-contributors.html). If applicable, please include **Name, Surname, National Bioinformatics Infrastructure Sweden, Science for Life Laboratory, Further Affliations**, as co-author. In other cases, NBIS would be grateful if support by us is acknowledged in publications according to this example:

> "Support by NBIS (National Bioinformatics Infrastructure Sweden) is gratefully acknowledged."

-   **UPPMAX** kindly asks you to [acknowledge UPPMAX and NAISS](https://www.uppmax.uu.se/support/faq/general-miscellaneous-faq/acknowledging-uppmax-and-naiss).

-   **NGI:** For publications based on data from NGI Sweden, NGI and SciLifeLab should be acknowledged. Instructions are included in the NGI reports as described [here](https://ngisweden.scilifelab.se/resources/getting-started-at-ngi/).
